---
title: "DADA2 utilizing DECIPHER for Taxonomy calling"
author: "Chris Trivedi & Liz Suter - based on [Mike Lee's DADA2 Workflow Example](https://astrobiomike.github.io/amplicon/dada2_workflow_ex#binder-available)"
date: "10 May, 2020"
output: 
  html_notebook:
    #toc: true
    #toc_float: true
editor_options: 
  chunk_output_type: console
---
---
<br>

# Prepping and getting the data ready

### Install and load packages

First let's install and load the `DADA2` and `DECIPHER` packages which we will be using for this tutorial.
```{r}
rm(list=ls())
# dada2
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("dada2")
# DECIPHER
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("DECIPHER")
```

```{r}
library(dada2)
library(DECIPHER)
```

### Install cutadapt

To trim our primers we will use `cutadapt`, let's install this in a conda environment. 

>**NOTE:** Make sure to run this and the following commands in the RStudio `console`.
```bash
conda install -c bioconda cutadapt
```

### Download the tutorial dataset

We will be using Mike Lee's tutorial dataset from his [Amplicon Analysis workflow](https://astrobiomike.github.io/amplicon/dada2_workflow_ex). He has this hosted on his FigShare. 
```bash
cd #<to working folder of your choice>
curl -L -o dada2_amplicon_ex_workflow.tar.gz https://ndownloader.figshare.com/files/15072638
tar -xzvf dada2_amplicon_ex_workflow.tar.gz
rm dada2_amplicon_ex_workflow.tar.gz
cd dada2_amplicon_ex_workflow/
```

### Get sample names
```bash
ls *_R1.fq | cut -f1 -d "_" > samples
```

### Removing primers using cutadapt
```bash
cutadapt --version # 2.3
cutadapt -a ^GTGCCAGCMGCCGCGGTAA...ATTAGAWACCCBDGTAGTCC \
    -A ^GGACTACHVGGGTWTCTAAT...TTACCGCGGCKGCTGGCAC \
    -m 215 -M 285 --discard-untrimmed \
    -o B1_sub_R1_trimmed.fq -p B1_sub_R2_trimmed.fq \
    B1_sub_R1.fq B1_sub_R2.fq
```

### Checking before and after
```bash
### R1 BEFORE TRIMMING PRIMERS
head -n 2 B1_sub_R1.fq
# @M02542:42:000000000-ABVHU:1:1101:8823:2303 1:N:0:3
# GTGCCAGCAGCCGCGGTAATACGTAGGGTGCGAGCGTTAATCGGAATTACTGGGCGTAAAGCGTGCGCAGGCGGTCTTGT
# AAGACAGAGGTGAAATCCCTGGGCTCAACCTAGGAATGGCCTTTGTGACTGCAAGGCTGGAGTGCGGCAGAGGGGGATGG
# AATTCCGCGTGTAGCAGTGAAATGCGTAGATATGCGGAGGAACACCGATGGCGAAGGCAGTCCCCTGGGCCTGCACTGAC
# GCTCATGCACGAAAGCGTGGGGAGCAAACAGGATTAGATACCCGGGTAGTCC
### R1 AFTER TRIMMING PRIMERS
head -n 2 B1_sub_R1_trimmed.fq
# @M02542:42:000000000-ABVHU:1:1101:8823:2303 1:N:0:3
# TACGTAGGGTGCGAGCGTTAATCGGAATTACTGGGCGTAAAGCGTGCGCAGGCGGTCTTGTAAGACAGAGGTGAAATCCC
# TGGGCTCAACCTAGGAATGGCCTTTGTGACTGCAAGGCTGGAGTGCGGCAGAGGGGGATGGAATTCCGCGTGTAGCAGTG
# AAATGCGTAGATATGCGGAGGAACACCGATGGCGAAGGCAGTCCCCTGGGCCTGCACTGACGCTCATGCACGAAAGCGTG
# GGGAGCAAACAGG
### R2 BEFORE TRIMMING PRIMERS
head -n 2 B1_sub_R2.fq
# @M02542:42:000000000-ABVHU:1:1101:8823:2303 2:N:0:3
# GGACTACCCGGGTATCTAATCCTGTTTGCTCCCCACGCTTTCGTGCATGAGCGTCAGTGCAGGCCCAGGGGACTGCCTTC
# GCCATCGGTGTTCCTCCGCATATCTACGCATTTCACTGCTACACGCGGAATTCCATCCCCCTCTGCCGCACTCCAGCCTT
# GCAGTCACAAAGGCCATTCCTAGGTTGAGCCCAGGGATTTCACCTCTGTCTTACAAGACCGCCTGCGCACGCTTTACGCC
# CAGTAATTCCGATTAACGCTCGCACCCTACGTATTACCGCGGCTGCTGGCACTCACACTC
### R2 AFTER TRIMMING PRIMERS
head -n 2 B1_sub_R2_trimmed.fq
# @M02542:42:000000000-ABVHU:1:1101:8823:2303 2:N:0:3
# CCTGTTTGCTCCCCACGCTTTCGTGCATGAGCGTCAGTGCAGGCCCAGGGGACTGCCTTCGCCATCGGTGTTCCTCCGCA
# TATCTACGCATTTCACTGCTACACGCGGAATTCCATCCCCCTCTGCCGCACTCCAGCCTTGCAGTCACAAAGGCCATTCC
# TAGGTTGAGCCCAGGGATTTCACCTCTGTCTTACAAGACCGCCTGCGCACGCTTTACGCCCAGTAATTCCGATTAACGCT
# CGCACCCTACGTA
```

### Use a loop to run cutadapt on all samples iteratively
```bash
for sample in $(cat samples)
do
    echo "On sample: $sample"
    
    cutadapt -a ^GTGCCAGCMGCCGCGGTAA...ATTAGAWACCCBDGTAGTCC \
    -A ^GGACTACHVGGGTWTCTAAT...TTACCGCGGCKGCTGGCAC \
    -m 215 -M 285 --discard-untrimmed \
    -o ${sample}_sub_R1_trimmed.fq.gz -p ${sample}_sub_R2_trimmed.fq.gz \
    ${sample}_sub_R1.fq ${sample}_sub_R2.fq \
    >> cutadapt_primer_trimming_stats.txt 2>&1
done
```

Let's see the output stats
```bash
paste samples <(grep "passing" cutadapt_primer_trimming_stats.txt | cut -f3 -d "(" | tr -d ")") <(grep "filtered" cutadapt_primer_trimming_stats.txt | cut -f3 -d "(" | tr -d ")")
```

>**NOTE:** Great, time to start processing with DADA2!
# DADA2

### Set our working directory and list our files
```{r}
setwd("~/dada2_amplicon_ex_workflow")
list.files() # make sure what we think is here is actually here
```

First let's set a few variables that we're going to use in the future. The first with all sample names, by scanning our "samples" file we made earlier.

In bash to get sample names from files in a folder:
ls > samples.txt
```{r}
# samples <- scan("samples", what="character")
# samples <- scan("qiime_import_16S/samples.txt", what="character") #get our sample names
# samples <- unique(substr(samples, 1, 11))
# samples <- head(samples, -1) #remove last element 'samples.txt'
```

The next two will contain the file names of all the forward and reverse reads
```{r}
#forward_reads <- paste0(samples, "_sub_R1_trimmed.fq.gz")
#reverse_reads <- paste0(samples, "_sub_R2_trimmed.fq.gz")

# forward_reads <- paste0(samples, "_00_L001_R1_001.fastq.qz")
# reverse_reads <- paste0(samples, "_00_L001_R2_001.fastq.qz")
```


Finally, we'll make variables containing the file names for the filtered forward and reverse reads.
```{r}
#filtered_forward_reads <- paste0(samples, "_sub_R1_filtered.fq.gz")
#filtered_reverse_reads <- paste0(samples, "_sub_R2_filtered.fq.gz")
```

Okay, now we're ready to begin looking at our data.

### Plotting the quality profiles
```{r}

names <- list.files() #get file names
names <- names[7:length(names)] #only keep the names of our sequences
forward_reads <- grep("_R1_", names, value=TRUE) #assign to forward_reads variable
reverse_reads <- grep("_R2_", names, value=TRUE) #assign to reverse_reads variable
filtered_forward_reads <- paste0(substr(forward_reads, 1, 26), "_filtered.fastq.gz") #assign names for output files of filtered reads
filtered_reverse_reads <- paste0(substr(reverse_reads, 1, 26), "_filtered.fastq.gz") #assign names for output files of filtered reads

plotQualityProfile(forward_reads)
plotQualityProfile(reverse_reads)

```

Alternatively we can just plot a few of the samples if we'd like
```{r}
plotQualityProfile(forward_reads[17:20])
```

Based on how these look we can determine how much to cut from each side.
```{r}
filtered_out <- filterAndTrim(forward_reads, filtered_forward_reads,
                reverse_reads, filtered_reverse_reads, maxEE=c(2,2),
                rm.phix=TRUE, minLen=175, truncLen=c(250,200))
```

Make sure you have the same number of samples you started with before filtering, and then check out the quality profiles again.
```{r}
class(filtered_out) # matrix
dim(filtered_out) # 20 2
filtered_out
plotQualityProfile(filtered_forward_reads)
plotQualityProfile(filtered_reverse_reads)
plotQualityProfile(filtered_reverse_reads[17:20])
```

>**NOTE** These numbers are slightly different than Mike's in his original workflow. This could be attributed to a different version of DADA2.

### Error profiling
```{r}
#err_forward_reads <- learnErrors(filtered_forward_reads)
err_forward_reads <- learnErrors(filtered_forward_reads, multithread=TRUE) # Use this if you're running outside of Binder
#err_reverse_reads <- learnErrors(filtered_reverse_reads)
err_reverse_reads <- learnErrors(filtered_reverse_reads, multithread=TRUE)
```

### Plotting the error profiles
```{r}
plotErrors(err_forward_reads, nominalQ=TRUE)
plotErrors(err_reverse_reads, nominalQ=TRUE)
```

>**NOTE:** As of the newest versions of DADA2 >=1.12 we don't have to do a separate dereplication step! DADA2 can now do this on the fly directly from the trimmed files. You can do it, but [according the the dev, it's not necessary](https://github.com/benjjneb/dada2/issues/752).
  
### Sample inference
```{r}
#dada_forward <- dada(filtered_forward_reads, err=err_forward_reads, pool="pseudo")
dada_forward <- dada(filtered_forward_reads, err=err_forward_reads, pool="pseudo", multithread=TRUE) # Use this if not running in Binder
#dada_reverse <- dada(filtered_reverse_reads, err=err_reverse_reads, pool="pseudo")
dada_reverse <- dada(filtered_reverse_reads, err=err_reverse_reads, pool="pseudo", multithread=TRUE)
```

### Merge your sample infered reads
```{r}
merged_amplicons <- mergePairs(dada_forward, filtered_forward_reads, dada_reverse, filtered_reverse_reads, trimOverhang=TRUE, minOverlap=170, verbose = TRUE)
# this object holds a lot of information that may be the first place you'd want to look if you want to start poking under the hood
class(merged_amplicons) # list
length(merged_amplicons) # 20 elements in this list, one for each of our samples
names(merged_amplicons) # the names() function gives us the name of each element of the list 
class(merged_amplicons$SRR12805327_00_L001_R1_001_filtered.fastq.gz) # each element of the list is a dataframe that can be accessed and manipulated like any ordinary dataframe
names(merged_amplicons$SRR12805327_00_L001_R1_001_filtered.fastq.gz) # the names() function on a dataframe gives you the column names
# "sequence"  "abundance" "forward"   "reverse"   "nmatch"    "nmismatch" "nindel"    "prefer"    "accept"
```

### Creating a sequence table
```{r}
seqtab <- makeSequenceTable(merged_amplicons)
class(seqtab) # matrix
dim(seqtab) # 41 2526
```

### Removing chimeras
```{r}
seqtab.nochim <- removeBimeraDenovo(seqtab, verbose=TRUE) 
# Identified 17 bimeras out of 2521 input sequences.
# though we only lost 17 sequences, we don't know if they held a lot in terms of abundance, this is one quick way to look at that
sum(seqtab.nochim)/sum(seqtab) 
# 0.9931372 
# good, we barely lost any in terms of abundance
```

# Summary of read counts through the pipeline
```{r}
# set a little function
getN <- function(x) sum(getUniques(x))
# making a little table
summary_tab <- data.frame(row.names=unique(substr(names, 1, 11)), dada2_input=filtered_out[,1],
                          filtered=filtered_out[,2], dada_f=sapply(dada_forward, getN),
                          dada_r=sapply(dada_reverse, getN), merged=sapply(merged_amplicons, getN),
                          nonchim=rowSums(seqtab.nochim),
                          final_perc_reads_retained=round(rowSums(seqtab.nochim)/filtered_out[,1]*100, 1))
summary_tab
```

### Assigning taxonomy using DECIPHER

>**Note:** DADA2 typically uses RDP's kmer-based method using SILVA for its classification database. Here we will use idTAXA as part of the DECIPHER package to see how results compare to the base DADA2 workflow.
Download the DECIPHER training set
```bash
curl -o SILVA_SSU_r138_2019.RData http://www2.decipher.codes/Classification/TrainingSets/SILVA_SSU_r138_2019.RData
```

Now let's assign taxonomy
```{r}
dna <- DNAStringSet(getSequences(seqtab.nochim)) # Create a DNAStringSet from the ASVs
load("SILVA_SSU_r138_2019.RData") # CHANGE TO THE PATH OF YOUR TRAINING SET
ids <- IdTaxa(dna, trainingSet, strand="top", processors=NULL, verbose=FALSE) # use all processors
ranks <- c("domain", "phylum", "class", "order", "family", "genus", "species") # ranks of interest
# Convert the output object of class "Taxa" to a matrix analogous to the output from assignTaxonomy
taxid <- t(sapply(ids, function(x) {
        m <- match(ranks, x$rank)
        taxa <- x$taxon[m]
        taxa[startsWith(taxa, "unclassified_")] <- NA
        taxa
}))
colnames(taxid) <- ranks; rownames(taxid) <- getSequences(seqtab.nochim)
```

>**NOTE:** The `taxid` matrix from IdTaxa is synonymous with `taxa` from the base way DADA2 assigns taxonomy.

Quickly check to see how things look
```{r}
taxa.print <- taxid # Removing sequence rownames for display only
rownames(taxa.print) <- NULL
head(taxa.print)
```

From here we can export our data to be analyzed outside or R or continue to look at using packages such as `Phyloseq` and `vegan`.

  # giving our seq headers more manageable names (ASV_1, ASV_2...) -->
```{r}
asv_seqs <- colnames(seqtab.nochim)
asv_headers <- vector(dim(seqtab.nochim)[2], mode="character")

for (i in 1:dim(seqtab.nochim)[2]) {
asv_headers[i] <- paste(">ASV", i, sep="_")
}
#So this gives us the actually genetic sequence of our unique ASVs
# making and writing out a fasta of our final ASV seqs: -->
asv_fasta <- c(rbind(asv_headers, asv_seqs))
write(asv_fasta, "ASVs.fa")

# count table: -->
asv_tab <- t(seqtab.nochim)
row.names(asv_tab) <- sub(">", "", asv_headers)
write.table(asv_tab, "ASVs_counts.tsv", sep="\t", quote=F, col.names=NA)

# tax table: -->
asv_tax <- taxid
row.names(asv_tax) <- sub(">", "", asv_headers)
write.table(asv_tax, "ASVs_taxonomy.tsv", sep="\t", quote=F, col.names=NA)
  
```