---
title: "vegan-tutorial"
author: "Amy Solman"
date: "05/03/2021"
output: html_document
---

This is the follow along and some direct text of the vegan tutorial by Jari Oksanen (2013) Multivariate Analysis of Ecological Communities in R: vegan tutorial + my own notes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This tutorial demostrates the use of ordination methods in Rpack-agevegan.    The  tutorial  assumes  familiarity  both  withRandwith community ordination.  Packagevegansupports all basic or-dination methods, including non-metric multidimensional scaling.The constrained ordination methods include constrained analysis ofproximities, redundancy analysis and constrained correspondenceanalysis.  Packageveganalso has support functions for fitting en-vironmental variables and for ordination graphics

What is ordination? Google says: "a statistical technique in which data from a large number of sites or populations are represented as points in a multidimensional space."

Ordination: basic method
Non-metric Multidimensional scaling (NMDS)
From Rob K statistics (https://www.youtube.com/watch?v=YS89_2n9Tlw): Ordination techniques allow us to order samples of ecological communities and to make sense of complicated data sets
NMDS can show us how similar complicated samples are.
Using shape and colour we can indicate which group samples belong to (e.g. location) or something about their associated environmental variables (e.g. temp, pH)
NMDS is there to help us produce a low-dimensional representation of our high dimensional space. Plot similar samples close together. Preserve important patterns in the data.

I'm going to start by importing my phyloseq object to be used in this analysis
```{r}

rm(list=ls())
library(phyloseq)

physeq <- readRDS("my-phyloseq-object.rds")

```

```{r}
rm(list=ls())
```