---
title: "good-18S-6-community-profiles"
author: "Amy Solman"
date: "05/10/2021"
output: html_document
---

This script will visualize the taxa groups present in our samples, so that differences in total, abundant, intermediate and rare communities can be compared between glaciers, regions and poles.

This will be done by:
Abundance bar plots (bar plots showing the relative abundance of each phyla within each sample/glacier/region/pole

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear workspace and load packages
```{r}
rm(list=ls())
graphics.off()

# library(ggpubr) #for making my boxplots
# library(FSA) #for dunn's test
# library(vegan) #for diversity indices
library(ggplot2) #for making plots
library(dplyr) #for manipulating data where you see %>% and computing summary stats
#remotes::install_github("cpauvert/psadd")
#library(psadd)
```

Step Two: Load data 
```{r}
#TOTAL DATASET
ps <- readRDS("../../results/16S/phylo-objects/16S-phyloseq-object-rarefied-decontam.rds")

#Total Abundant
ps.abun <- readRDS("../../results/16S/phylo-objects/16S-total-abundant.rds")
#Total Intermediate
ps.int <- readRDS("../../results/16S/phylo-objects/16S-total-intermediate.rds")
#Total Rare
ps.rare <- readRDS("../../results/16S/phylo-objects/16S-total-rare.rds")
```

1 TOTAL TAXA
1.1 Abundance Bar Plots
```{r}
#Agglomerate ASVs by phyla into dataframes ready for plotting
ps.glom <- ps %>%
  tax_glom(taxrank = "Phylum", NArm=FALSE) %>%
  psmelt()

#Glacier
p1 <- ggplot(ps.glom, aes(x = Glacier, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Glacier")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("16S: Phyla Barplot by Glacier")
print(p1)
ggsave("../../results/16S/graphs/community-plots/phyla-glacier-barplot.pdf")

#Region
ggplot(ps.glom, aes(x = Region, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Region")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("16S: Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/phyla-region-barplot.pdf")

#Pole
ggplot(ps.glom, aes(x = Pole, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Pole")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("16S: Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/phyla-pole-barplot.pdf")

```
Stats
```{r}
#How many/which phyla total dataset? Relative abundance of each phyla.
#agglomerate data
my.glom = tax_glom(ps, "Phylum")
my.glom 
my.glom.phlum.table <- data.frame(otu_table(my.glom), row.names=unique(data.frame(tax_table(my.glom)))$Phylum)
total_phyla_table <- data.frame(Phylum=row.names(my.glom.phlum.table), Abundance=rowSums(my.glom.phlum.table), Relative_Abundance=rowSums(my.glom.phlum.table)/sum(rowSums(my.glom.phlum.table)))
rownames(total_phyla_table) <- c()

total_phyla_table <- total_phyla_table %>% arrange(desc(Relative_Abundance))
total_phyla_table$Relative_Abundance <- round(total_phyla_table$Relative_Abundance, digits=7)*100
colnames(total_phyla_table) <- c("Phylum", "Abundance", "Relative Abundance %")

write.csv(total_phyla_table, "../../results/16S/tables/phyla/total-phyla-data.csv")
```
1.2 Another bar plot
```{r}
#convert to relative abundance
ps_rel_abund = phyloseq::transform_sample_counts(ps, function(x){x / sum(x)})

#plot
#Glacier
phyloseq::plot_bar(ps_rel_abund, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Glacier, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("16S: Phyla Barplot by Glacier")
ggsave("../../results/16S/graphs/community-plots/phyla-glacier-barplot-2.pdf")

#Region
phyloseq::plot_bar(ps_rel_abund, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Region, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("16S: Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/phyla-region-barplot-2.pdf")

#Pole
phyloseq::plot_bar(ps_rel_abund, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Pole, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ggtitle("16S: Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/phyla-pole-barplot-2.pdf")

```
1.3 Comparing phyla for both poles
```{r}
#Agglomerate taxa to phylum level
ps_phylum <- phyloseq::tax_glom(ps, "Phylum")
phyloseq::taxa_names(ps_phylum) <- phyloseq::tax_table(ps_phylum)[, "Phylum"]

#plot
phyloseq::psmelt(ps_phylum) %>%
ggplot(data = ., aes(x = Pole, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = Phylum), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ Phylum, scales = "free")+
  ggtitle("16S: Phyla by pole")
ggsave("../../results/16S/graphs/community-plots/phyla-pole-scatterplot.pdf")

```
1.4 Pie chart
```{r}
counts <- data.frame(t(otu_table(ps_phylum)))
count_sums <- colSums(counts)
pct <- count_sums/sum(count_sums)*100
lbls <- names(counts) # add percents to labels
df <- data.frame(lbls, pct)
x <- subset(df, df$pct < 2)
sum_x <- sum(x$pct)
#remove rows with less than 1
y <- subset(df, df$pct > 2)
df2 <- data.frame("Other", sum_x)
names(df2) <- c("lbls", "pct")
newdf <- rbind(y, df2)
newdf$lbls <- paste(newdf$lbls, round(newdf$pct, digits=2)) # add percents to labels
newdf$lbls <- paste(newdf$lbls,"%",sep="") # ad % to labels 
#pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), radius = 0.9, cex = 0.4)
pdf("../../results/16S/graphs/community-plots/phyla-pie-chart.pdf")
pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), radius= 1, cex=0.4, main="Total Community Pie Chart")
dev.off()
```

2 ABUNDANT TAXA
2.1 Abundance Bar Plots
```{r}
#Agglomerate ASVs by phyla into dataframes ready for plotting
abun.glom <- ps.abun %>%
  tax_glom(taxrank = "Phylum", NArm=FALSE) %>%
  psmelt()

#Glacier
ggplot(abun.glom, aes(x = Glacier, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Glacier")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    ggtitle("16S: Abundant Phyla Barplot by Glacier")
ggsave("../../results/16S/graphs/community-plots/abundant-phyla-glacier-barplot.pdf")

#Region
ggplot(abun.glom, aes(x = Region, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Region")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("16S: Abundant Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/abundant-phyla-region-barplot.pdf")

#Pole
ggplot(abun.glom, aes(x = Pole, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Pole")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    ggtitle("16S: Abundant Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/abundant-phyla-pole-barplot.pdf")
```

Stats
```{r}
#How many/which phyla total dataset? Relative abundance of each phyla.
#agglomerate data
my.glom = tax_glom(ps.abun, "Phylum")
my.glom 
my.glom.phlum.table <- data.frame(otu_table(my.glom), row.names=unique(data.frame(tax_table(my.glom)))$Phylum)
total_phyla_table <- data.frame(Phylum=row.names(my.glom.phlum.table), Abundance=rowSums(my.glom.phlum.table), Relative_Abundance=rowSums(my.glom.phlum.table)/sum(rowSums(my.glom.phlum.table)))
rownames(total_phyla_table) <- c()

abun_phyla_table <- total_phyla_table %>% arrange(desc(Relative_Abundance))
abun_phyla_table$Relative_Abundance <- round(abun_phyla_table$Relative_Abundance, digits=7)*100
colnames(abun_phyla_table) <- c("Phylum", "Abundance", "Relative Abundance %")

write.csv(abun_phyla_table, "../../results/16S/tables/phyla/abundant-phyla-data.csv")
```

2.2 Another bar plot
```{r}
#convert to relative abundance
ps_rel_abund.ab = phyloseq::transform_sample_counts(ps.abun, function(x){x / sum(x)})

#plot
#Glacier
phyloseq::plot_bar(ps_rel_abund.ab, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Glacier, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
    ggtitle("16S: Abundant Phyla Barplot by Glacier")
ggsave("../../results/16S/graphs/community-plots/abundant-phyla-glacier-barplot-2.pdf")

#Region
phyloseq::plot_bar(ps_rel_abund.ab, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Region, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
    ggtitle("16S: Abundant Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/abundant-phyla-region-barplot-2.pdf")

#Pole
phyloseq::plot_bar(ps_rel_abund.ab, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Pole, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
    ggtitle("16S: Abundant Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/abundant-phyla-pole-barplot-2.pdf")

```

2.3 Comparing phyla for both poles
```{r}
#Agglomerate taxa to phylum level
ps_phylum.ab <- phyloseq::tax_glom(ps.abun, "Phylum")
phyloseq::taxa_names(ps_phylum.ab) <- phyloseq::tax_table(ps_phylum.ab)[, "Phylum"]

#plot
phyloseq::psmelt(ps_phylum.ab) %>%
ggplot(data = ., aes(x = Pole, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = Phylum), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ Phylum, scales = "free")+
    ggtitle("16S: Abundant Phyla by pole")
ggsave("../../results/16S/graphs/community-plots/abundant-phyla-pole-scatterplot.pdf")

```

2.4 Pie chart
```{r}
counts <- data.frame(t(otu_table(ps_phylum.ab)))
count_sums <- colSums(counts)
#pct <- round(count_sums/sum(count_sums)*100)
pct <- count_sums/sum(count_sums)*100
lbls <- names(counts) # add percents to labels
df <- data.frame(lbls, pct)
x <- subset(df, df$pct < 2)
sum_x <- sum(x$pct)
#remove rows with less than 1
y <- subset(df, df$pct > 2)
df2 <- data.frame("Other", sum_x)
names(df2) <- c("lbls", "pct")
newdf <- rbind(y, df2)
newdf$lbls <- paste(newdf$lbls, round(newdf$pct, digits=2)) # add percents to labels
newdf$lbls <- paste(newdf$lbls,"%",sep="") # ad % to labels 
#pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), radius = 1, cex = 0.3)
pdf("../../results/16S/graphs/community-plots/abundant-phyla-pie-chart.pdf")
pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), main="Abundant Community Pie Chart")
dev.off()
```

3 INTERMEDIATE TAXA
3.1 Abundance Bar Plots
```{r}
#Agglomerate ASVs by phyla into dataframes ready for plotting
ps.int.glom <- ps.int %>%
  tax_glom(taxrank = "Phylum", NArm=FALSE) %>%
  psmelt()

#Glacier
ggplot(ps.int.glom, aes(x = Glacier, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Glacier")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
      ggtitle("16S: Intermediate Phyla Barplot by Glacier")
ggsave("../../results/16S/graphs/community-plots/intermediate-phyla-glacier-barplot.pdf")

#Region
ggplot(ps.int.glom, aes(x = Region, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Region")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    ggtitle("16S: Intermediate Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/intermediate-phyla-region-barplot.pdf")

#Pole
ggplot(ps.int.glom, aes(x = Pole, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Pole")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
      ggtitle("16S: Intermediate Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/intermediate-phyla-pole-barplot.pdf")

```

Stats
```{r}
#How many/which phyla total dataset? Relative abundance of each phyla.
#agglomerate data
my.glom = tax_glom(ps.int, "Phylum")
my.glom 
my.glom.phlum.table <- data.frame(otu_table(my.glom), row.names=unique(data.frame(tax_table(my.glom)))$Phylum)
total_phyla_table <- data.frame(Phylum=row.names(my.glom.phlum.table), Abundance=rowSums(my.glom.phlum.table), Relative_Abundance=rowSums(my.glom.phlum.table)/sum(rowSums(my.glom.phlum.table)))
rownames(total_phyla_table) <- c()

int_phyla_table <- total_phyla_table %>% arrange(desc(Relative_Abundance))
int_phyla_table$Relative_Abundance <- round(int_phyla_table$Relative_Abundance, digits=7)*100
colnames(int_phyla_table) <- c("Phylum", "Abundance", "Relative Abundance %")

write.csv(int_phyla_table, "../../results/16S/tables/phyla/intermediate-phyla-data.csv")
```

3.2 Another bar plot
```{r}
#convert to relative abundance
ps_rel_abund.i = phyloseq::transform_sample_counts(ps.int, function(x){x / sum(x)})

#plot
#Glacier
phyloseq::plot_bar(ps_rel_abund.i, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Glacier, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggtitle("16S: Intermediate Phyla Barplot by Glacier")
ggsave("../../results/16S/graphs/community-plots/intermediate-phyla-glacier-barplot-2.pdf")

#Region
phyloseq::plot_bar(ps_rel_abund.i, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Region, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggtitle("16S: Intermediate Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/intermediate-phyla-region-barplot-2.pdf")

#Pole
phyloseq::plot_bar(ps_rel_abund.i, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Pole, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggtitle("16S: Intermediate Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/intermediate-phyla-pole-barplot-2.pdf")

```

3.3 Comparing phyla for both poles
```{r}
#Agglomerate taxa to phylum level
ps_phylum.i <- phyloseq::tax_glom(ps.int, "Phylum")
phyloseq::taxa_names(ps_phylum.i) <- phyloseq::tax_table(ps_phylum.i)[, "Phylum"]

#plot
phyloseq::psmelt(ps_phylum.i) %>%
ggplot(data = ., aes(x = Pole, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = Phylum), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ Phylum, scales = "free")+
      ggtitle("16S: Intermediate Phyla by pole")
ggsave("../../results/16S/graphs/community-plots/intermediate-phyla-pole-scatterplot.pdf")

```

3.4 Pie chart
```{r}
counts <- data.frame(t(otu_table(ps_phylum.i)))
count_sums <- colSums(counts)
pct <- count_sums/sum(count_sums)*100
lbls <- names(counts) # add percents to labels
df <- data.frame(lbls, pct)
x <- subset(df, df$pct < 2)
sum_x <- sum(x$pct)
#remove rows with less than 1
y <- subset(df, df$pct > 2)
df2 <- data.frame("Other", sum_x)
names(df2) <- c("lbls", "pct")
newdf <- rbind(y, df2)
newdf$lbls <- paste(newdf$lbls, round(newdf$pct, digits=2)) # add percents to labels
newdf$lbls <- paste(newdf$lbls,"%",sep="") # ad % to labels 
#pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), radius = 0.9, cex = 0.4)
pdf("../../results/16S/graphs/community-plots/intermediate-phyla-pie-chart.pdf")
pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), cex=0.6, main="Intermediate Community Pie Chart")
dev.off()
```

4 RARE TAXA
4.1 Abundance Bar Plots
```{r}
#Agglomerate ASVs by phyla into dataframes ready for plotting
rare.glom <- ps.rare %>%
  tax_glom(taxrank = "Phylum", NArm=FALSE) %>%
  psmelt()

#Glacier
ggplot(rare.glom, aes(x = Glacier, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Glacier")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
        ggtitle("16S: Rare Phyla Barplot by Glacier")
ggsave("../../results/16S/graphs/community-plots/rare-phyla-glacier-barplot.pdf")

#Region
ggplot(rare.glom, aes(x = Region, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Region")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
      ggtitle("16S: Rare Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/rare-phyla-region-barplot.pdf")

#Pole
ggplot(rare.glom, aes(x = Pole, y = Abundance, fill = Phylum)) +
  geom_bar(position="fill", stat="identity") +
  xlab("Pole")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
        ggtitle("16S: Rare Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/rare-phyla-pole-barplot.pdf")
```
Stats
```{r}
#How many/which phyla total dataset? Relative abundance of each phyla.
#agglomerate data
my.glom = tax_glom(ps.rare, "Phylum")
my.glom 
my.glom.phlum.table <- data.frame(otu_table(my.glom), row.names=unique(data.frame(tax_table(my.glom)))$Phylum)
total_phyla_table <- data.frame(Phylum=row.names(my.glom.phlum.table), Abundance=rowSums(my.glom.phlum.table), Relative_Abundance=rowSums(my.glom.phlum.table)/sum(rowSums(my.glom.phlum.table)))
rownames(total_phyla_table) <- c()

rare_phyla_table <- total_phyla_table %>% arrange(desc(Relative_Abundance))
rare_phyla_table$Relative_Abundance <- round(rare_phyla_table$Relative_Abundance, digits=7)*100
colnames(rare_phyla_table) <- c("Phylum", "Abundance", "Relative Abundance %")

write.csv(rare_phyla_table, "../../results/16S/tables/phyla/rare-phyla-data.csv")
```

4.2 Another bar plot
```{r}
#convert to relative abundance
ps_rel_abund.ra = phyloseq::transform_sample_counts(ps.rare, function(x){x / sum(x)})

#plot
#Glacier
phyloseq::plot_bar(ps_rel_abund.ra, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Glacier, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggtitle("16S: Rare Phyla Barplot by Glacier")
ggsave("../../results/16S/graphs/community-plots/rare-phyla-glacier-barplot-2.pdf")

#Region
phyloseq::plot_bar(ps_rel_abund.ra, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Region, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggtitle("16S: Rare Phyla Barplot by Region")
ggsave("../../results/16S/graphs/community-plots/rare-phyla-region-barplot-2.pdf")

#Pole
phyloseq::plot_bar(ps_rel_abund.ra, fill = "Phylum") +
  geom_bar(aes(color = Phylum, fill = Phylum), stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  facet_wrap(~ Pole, scales = "free") +
  theme(panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
        ggtitle("16S: Rare Phyla Barplot by Pole")
ggsave("../../results/16S/graphs/community-plots/rare-phyla-pole-barplot-2.pdf")

```
4.3 Comparing phyla for both poles
```{r}
#Agglomerate taxa to phylum level
ps_phylum.ra <- phyloseq::tax_glom(ps.rare, "Phylum")
phyloseq::taxa_names(ps_phylum.ra) <- phyloseq::tax_table(ps_phylum.ra)[, "Phylum"]

#plot
phyloseq::psmelt(ps_phylum.ra) %>%
ggplot(data = ., aes(x = Pole, y = Abundance)) +
  geom_boxplot(outlier.shape  = NA) +
  geom_jitter(aes(color = Phylum), height = 0, width = .2) +
  labs(x = "", y = "Abundance\n") +
  facet_wrap(~ Phylum, scales = "free")+
        ggtitle("16S: Rare Phyla by pole")
ggsave("../../results/16S/graphs/community-plots/rare-phyla-pole-scatterplot.pdf")

```
4.4 Pie chart
```{r}
counts <- data.frame(t(otu_table(ps_phylum.ra)))
count_sums <- colSums(counts)
#pct <- round(count_sums/sum(count_sums)*100)
pct <- count_sums/sum(count_sums)*100
lbls <- names(counts) # add percents to labels
df <- data.frame(lbls, pct)
x <- subset(df, df$pct < 2)
sum_x <- sum(x$pct)
#remove rows with less than 1
y <- subset(df, df$pct > 2)
df2 <- data.frame("Other", sum_x)
names(df2) <- c("lbls", "pct")
newdf <- rbind(y, df2)
newdf$lbls <- paste(newdf$lbls, round(newdf$pct, digits=2)) # add percents to labels
newdf$lbls <- paste(newdf$lbls,"%",sep="") # ad % to labels 
#pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), radius = 1, cex = 0.3)
pdf("../../results/16S/graphs/community-plots/rare-phyla-pie-chart.pdf")
pie(newdf$pct, labels = newdf$lbls, col=rainbow(length(newdf$pct)), radius=1, cex=0.5)
dev.off()
```