---
title: "18S-7-tree-plots"
author: "Amy Solman"
date: "06/10/2021"
output: html_document
---

This script will explore the taxa groups present in our samples, so that differences in total, abundant, intermediate and rare communities can be compared between glaciers, regions and poles.

This will be done using:
Annotated phylogenetic tree plots for each glacier, region, pole

THIS IS NOT ACTUALLY APPROPRIATE FOR THIS STAGE IN MY ANALYSIS SO I'M GOING TO MOVE ON

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear workspace and load packages
```{r}
rm(list=ls())
graphics.off()

# library(ggpubr) #for making my boxplots
# library(FSA) #for dunn's test
# library(vegan) #for diversity indices
# library(ggplot2) #for making plots
# library(dplyr) #for manipulating data where you see %>% and computing summary stats

#install ggtree
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ggtree")
```


Load data 
```{r}
#Load data files
load("../data-output/18S-my-sequence-data.RData") #load all our pre-processed data files

#load phyloseq object
##################################################BOTH POLES#################################################
ps <- readRDS("../data-output/18S-my-phyloseq-object.rds") #load unrarefied phyloseq object
ps.rar <- readRDS("../data-output/18S-my-phyloseq-object-rarefied.rds") #load rarefied phyloseq objectcol

#Global abundances i.e. globally abundant, globally rare, globally intermediate
ps.global.RA <- readRDS("../data-output/18S-all-rarefied-abundant.rds") #load rarefied globally abundant phyloseq object
ps.global.RR <- readRDS("../data-output/18S-all-rarefied-rare.rds") #load rarefied globally rare phyloseq object
ps.global.I <- readRDS("../data-output/18S-all-rarefied-intermediate.rds") #load rarefied globally abundant phyloseq object

#Both poles merged regional abundances
ps.both.merged.RA <- readRDS("../data-output/18S-all-merged-rarefied-abundant.rds") #merged poles abundant phyloseq object
ps.both.merged.RR <- readRDS("../data-output/18S-all-merged-rarefied-rare.rds") #merged poles rare phyloseq object
ps.both.merged.I <- readRDS("../data-output/18S-all-merged-rarefied-intermediate.rds") #merged poles intermediate phyloseq object

#All glaciers merged abundances
ps.glaciers.merged.RA <- readRDS("../data-output/18S-all-glaciers-merged-rarefied-abundant.rds") #merged glaciers abundant phyloseq object
ps.glaciers.merged.RR <- readRDS("../data-output/18S-all-glaciers-merged-rarefied-rare.rds") #merged glaciers rare phyloseq object
ps.glaciers.merged.I <- readRDS("../data-output/18S-all-glaciers-merged-rarefied-intermediate.rds") #merged glaciers intermediate phyloseq object

##################################################ARCTIC#################################################
ps.arc <- readRDS("../data-output/18S-arctic-rarefied.rds")
ps.arc.RA <- readRDS("../data-output/18S-arctic-rarefied-abundant.rds") #load rarefied arctic abundant phyloseq object
ps.arc.RR <- readRDS("../data-output/18S-arctic-rarefied-rare.rds") #load rarefied arctic rare phyloseq object
ps.arc.I <- readRDS("../data-output/18S-arctic-rarefied-intermediate.rds") #load rarefied arctic intermediate phyloseq object

##################################################ANTARCTIC#################################################
ps.ant <- readRDS("../data-output/18S-antarctic-rarefied.rds")
ps.ant.RA <- readRDS("../data-output/18S-antarctic-rarefied-abundant.rds") #load rarefied antarctic abundant phyloseq object
ps.ant.RR <- readRDS("../data-output/18S-antarctic-rarefied-rare.rds") #load rarefied antarctic rare phyloseq object
ps.ant.I <- readRDS("../data-output/18S-antarctic-rarefied-intermediate.rds") #load rarefied antarctic intermediate phyloseq object
```

Subset data (plotting too many taxa obscures meaning)
```{r}
plot_tree(ps.rar, color = "Region", shape = "Pole", label.tips = "Phylum", 
    size = "abundance", plot.margin = 0.5, ladderize = TRUE)

#Archaeplastida
sub_arch <- subset_taxa(ps.rar, Phylum == "Archaeplastida")

#map
plot_tree(sub_arch, color = "Region", shape = "Pole", label.tips = "Phylum", 
    size = "abundance", plot.margin = 0.5, ladderize = TRUE)

#subset by region
sub_green <- subset_samples(ps.rar, Region == "Greenland")

#subset by region
sub_swede <- subset_samples(ps.rar, Region == "Sweden")

#subset by glacier
sub_green <- subset_samples(ps.rar, Region == "Greenland")

#map
plot_tree(sub_swede, color = "Glacier", label.tips = "Phylum", 
    size = "abundance", plot.margin = 0.5, ladderize = TRUE)
```