---
title: "analysis"
author: "Amy Solman"
date: "07/04/2021"
output: html_document
---

#This document will follow the method set out by the Microbial Bioinformatics Course (https://mibwurrepo.github.io/Microbial-bioinformatics-introductory-course-Material-2018/alpha-diversities.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace 
```{r}
rm(list=ls())
graphics.off()
```

#Step Two: Load packages 
```{r}
library(phyloseq)
library(DT) #interactive tables in html and markdown


# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("microbiome")
library(microbiome) #useful phyloseq commands e.g. summarize_phyloseq()

library(RColorBrewer) #use for function brewer_pal() in core abundance heatmap
#install.packages("ggpubr")
library(ggpubr) #used for ggboxplot() function
library(data.table) #used for function data.table
# library(dplyr)

#Install microbiomeutilities
#install.packages("devtools")
#devtools::install_github("microsud/microbiomeutilities")
library(microbiomeutilities) #gives us the read_phyloseq function

library(picante) #for faith's phylogenetic diversity
library(vegan) #for our PERMANOVA (permutational multivariate analysis of variance)
library(viridis) #good colour palette, used for core microbiota heatmaps

#library(devtools)
#install_github("zdk123/SpiecEasi")


#install.packages("remotes")
remotes::install_github("zdk123/SpiecEasi")
library(SpiecEasi)

```

#Step Three: Load data and phyloseq objects  
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#my_physeq <- readRDS("results/my-phyloseq-object.rds")
```

#Step Four: Create new phyloseq object using count_table instead of hellinger transformed table: 
```{r}
# my.ps <- read_phyloseq(otu.file = "../Data-wrangling/work/export/table/feature-table.biom",
#                        taxonomy.file = taxonomy,
#                        metadata.file = metadata,
#                        type = "biom")
```

```{r}
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Five: Check for features of the phyloseq object 
```{r}
print(my.ps) #shows how many taxa and samples are in asv table, how many samples and variables are in metadata table, how many taxa and taxonomic ranks are in the taxonomy table and how many tips and nodes are in the phylogenetic tree
rank_names(my.ps) #shows the taxonomic rank names in the taxonomy table 
datatable(tax_table(my.ps)) #shows the ASV IDs and their taxonomic classifications
summarize_phyloseq(my.ps) #uses microbiome package
```

#Step Six: Check the spread of variation of the taxa.
#The coefficient of variation = sd(x)/mean(x) - widely used for quantifying heterogeneity in ASV abundance data
```{r}
p1 <- plot_taxa_cv(my.ps, plot.type = "scatter") #this function will convert the counts to relative abundances then calculate the coefficient of vairation
p1 + scale_x_log10()
```

#Step Seven: Check the distribution of number of sequences
```{r}
seqdepth.ps <- plot_read_distribution(my.ps, "Glacier", "density") #density here is the type of plot. Function from microbiomeutilities.
print(seqdepth.ps) 
```

#Step Eight: Plot a histogram of count abundance against frequency of ASVs
```{r}
# We make a data table with information on the OTUs
my.ps.table <- data.table(tax_table(my.ps),
                          ASVabundance = taxa_sums(my.ps),
                          ASV = taxa_names(my.ps))

my.ps.plot <- ggplot(my.ps.table, aes(ASVabundance)) +
  geom_histogram() + ggtitle("Histogram of ASVs (unique sequence) counts") +
  theme_bw() + scale_x_log10() + ylab("Frequency of ASVs") + xlab("Abundance (raw counts)")

print(my.ps.plot)

#The ASVs distribution is close to normal distribution and this has implications for downstream analysis (e.g. differential abundance testing)
```

#Step : Check for the prevalance abundance distribution of ASVs
```{r}
p <- plot_taxa_prevalence(my.ps, "Phylum")
p

#Here we see that taxa at low prevalence (y axis) are represented by high abundance in the sequencing data, suggesting that these are REAL observations and spurious ASVs.

#total number of reads in the dataset
tot_reads <- taxa_sums(my.ps)
print(sum(tot_reads))

#how many ASVs are less than 10 reads and how many reads do they contain?
print(length(tot_reads[tot_reads < 10]))
print(sum(tot_reads[tot_reads < 10]))

#how many singletons are there?
length(which(taxa_sums(my.ps) <= 1))

#how many doubletons are there?
length(which(taxa_sums(my.ps) == 2))

#how many singletons and doubletons are there?
length(which(taxa_sums(my.ps) <= 2))

round((127/2149)*100, digits = 2) #6% of ASVs are doubletons or singletons

sessionInfo()
```

#Step : Calculate alpha diversities 
```{r}
print(my.ps)
summary(sample_sums(my.ps)) #this shows us the number of reads where the minimum is 13401 and the max is 35129

#plot rarefaction curves for the observed ASVs so we can look at how richness is captured in the sequencing effort

asv_tab <- t(abundances(my.ps))
p <- vegan::rarecurve(asv_tab,
                      step=50, label = FALSE,
                      sample = min(rowSums(asv_tab),
                                   col = "blue", cex = 0.6))
```

#Step : Rarefy the samples to equal sampling depth
```{r}
set.seed(1961) #This will ensure the results of our normalisation step remain the same when repeated

ps.rar <- rarefy_even_depth(my.ps, sample.size = 13401)

print(ps.rar) #Let's so how our data have changed

barplot(sample_sums(ps.rar), las=2) #We can see all of our samples now have the same number of reads

#Check taxa prevelance in the rarefied data
p.ps.rar <- plot_taxa_prevalence(ps.rar, "Phylum")
p.ps.rar #We can see how the taxa prevalence plot had changed in comparison to before the reads were rarefied
```

#Step : Calculate non-phylogenetic diversities
```{r}
ps.div <- alpha(ps.rar, index="all")
datatable(ps.div) #So this gives us lots of diversity indices in one go!
write.csv(ps.div, "../Outputs/Sommers2020/diversities.csv")
```

#Step : Plot the diversity data 
```{r}
#get the metadata as a seperate object
ps.rar.meta <- meta(ps.rar)

#add row names as new column for easy integration later
ps.rar.meta$sam_name <- rownames(ps.rar.meta)

#add the row names to the diversity table
ps.div$sam_name <- rownames(ps.div)

#merge the dataframes
my.div.df <- merge(ps.div, ps.rar.meta, by = "sam_name")

colnames(my.div.df)

#now we can use the combined diversity and metadata to plot

p <- ggboxplot(my.div.df,
               x="Glacier",
               y="diversity_shannon",
               fill="Glacier",
               palette="jco") #this uses the ggpubr package

p

#Alternatively we could do this
#convert phyloseq object into long data format
my.div.df2 <- my.div.df[, c("Glacier", "diversity_inverse_simpson", "diversity_gini_simpson", "diversity_shannon", "diversity_fisher", "diversity_coverage")]

#Replace with better names
colnames(my.div.df2) <- c("Glacier", "Inverse Simpson", "Gini-Simpson", "Shannon", "Fisher", "Coverage")

#Check the col names
colnames(my.div.df2) 

melt.my.div.df <- reshape2::melt(my.div.df2)

head(melt.my.div.df)

#Now we'll plot using this dataframe

p <- ggboxplot(melt.my.div.df, x = "Glacier", y = "value", fill = "Glacier", palette = "jco", legend = "right", facet.by = "variable", scales = "free")
p <- p + rotate_x_text() 
p <- p + rremove("x.text")
p

#save 
ggsave("../Outputs/Sommers2020/graphs/Diversities.pdf", height=4, width =10)

lev <- levels(as.factor(melt.my.div.df$Glacier)) #get the glacier names

#make a pairwise list that we want to compare
lev.pairs <- combn(seq_along(lev), 2, simplify = FALSE, FUN = function(i) lev[i])

pval <- list(
  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
  symbols = c("****", "***", "**", "*", "n.s")
) #set our p-values

p2 <- p + stat_compare_means(
  comparisons = lev.pairs,
  label = "p.signif",
  symnum.args = list(
    cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
  symbols = c("****", "***", "**", "*", "n.s")
  )
)

print(p2)
```

#Step : Phylogenetic diversity
```{r}
ps.rar.asv.tab <- as.data.frame(ps.rar@otu_table)

ps.rar.tree <- ps.rar@phy_tree

#check if the tree is rooted

ps.rar@phy_tree #it is rooted

ps.df <- pd(t(ps.rar.asv.tab), ps.rar.tree, include.root = T) #transpose the asv table and find phylogetetic diversity

datatable(ps.df)

#now we plot the pd
#add the results of the pd to our metadata file

ps.rar.meta$Phylogenetic_Diversity <- ps.df$PD

#plot
pd.plot <- ggboxplot(ps.rar.meta,
                     x = "Glacier",
                     y = "Phylogenetic_Diversity",
                     fill = "Glacier",
                     palette = "jco",
                     ylab = "Phylogenetic Diversity",
                     xlab = "Glacier",
                     legend = "right")

pd.plot <- pd.plot + rotate_x_text()

pd.plot + stat_compare_means(
  comparisons = lev.pairs,
  label = "p.signif",
  symnum.args = list(
    cutpoints = c(0, 0.00001, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
    symbols = c("****", "***", "**", "*", "n.s")
  )
)

#save 
ggsave("../Outputs/Sommers2020/graphs/PD-Diversity.pdf", height=4, width =10)


```

#Step : Find diversities for non-rarefied data
```{r}
#Note from the bioinformatics guide: There are arguments both for and against the use of rarefying to equal library size.
#The application of normalization method will depend on the type of research question. 
#It is always good to check if there is a correlation between increasing library sizes and richness.
#Observed ASVs and Phylogenetic diversity can be affected by sibrary sizes. It is always good to check this before making a choice.

tot.div <- alpha(my.ps, index = "all")
tot.div2 <- richness(my.ps)

#add total number of reads/samples
tot.div$ReadsPerSample <- sample_sums(my.ps)
tot.div$Observed <- tot.div2$observed
colnames(tot.div)
```

#Step : Visualise shannon diversity, inverse simpson index and observed ASVs of non-rarefied data
```{r}
p1 <- ggscatter(tot.div, "diversity_shannon", "ReadsPerSample")+
  stat_cor(method="pearson")

p2 <- ggscatter(tot.div, "diversity_inverse_simpson", "ReadsPerSample", add="loess")+
  stat_cor(method="pearson")

p3 <- ggscatter(tot.div, "Observed", "ReadsPerSample", add = "loess")+
  stat_cor(
    method = "pearson",
    label.x = 100,
    label.y = 35000
  )

ggarrange(p1, p2, p3, ncol = 2, nrow = 2)
```

#Step : Composition Plots
```{r}
print(my.ps)

#Barplot counts
my.ps.com <- my.ps
taxa_names(my.ps.com) <- paste0("ASV_", rownames(tax_table(my.ps.com)))

#Set Palette
taxic <- as.data.frame(my.ps.com@tax_table) #help for setting colour options

taxic$ASV <- rownames(taxic)

#convert to matrix
taxmat <- as.matrix(taxic)
#convert to phyloseq compatible file
new.tax <- tax_table(taxmat)
#add to phyloseq object
tax_table(my.ps.com) <- new.tax

#Edit unclassified taxa
tax_table(my.ps.com)[tax_table(my.ps.com)[, "Family"] == "", "Family"] <- "Unclassified family"

#Put the taxa names in italics
guide_italics <- guides(fill = guide_legend(label.theme = element_text(
  size=15,
  face="italic", colour="Black", angle =0
)))

#plot counts at family level
#first remove the phylogenetic tree
my.ps.com@phy_tree <- NULL

#merge at the family level
my.ps.com.fam <- microbiome::aggregate_top_taxa(my.ps.com, "Family", top=10)

#merge at the class level
my.ps.com.class <- microbiome::aggregate_top_taxa(my.ps.com, "Class", top=10) #what does it mean if a function is deprecated

plot.composition.Count.Abun.Fam <- plot_composition(my.ps.com.fam) +
  theme(legend.position = "bottom")+
  scale_fill_brewer("Family", palette = "Paired")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("Count Abundance") +
  guide_italics + theme(legend.title = element_text(size=18))

plot.composition.Count.Abun.Fam

ggsave("../Outputs/Sommers2020/graphs/count-abundance-family-barplot.pdf", width = 10, height = 5)

plot.composition.Count.Abun.Class <- plot_composition(my.ps.com.class) +
  theme(legend.position = "bottom")+
  scale_fill_brewer("Class", palette = "Paired")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90))+
  ggtitle("Count Abundance") +
  guide_italics + theme(legend.title = element_text(size=18))

plot.composition.Count.Abun.Class

ggsave("../Outputs/Sommers2020/graphs/count-abundance-class-barplot.pdf", width = 10, height = 5)
```

#Step : Barplot of relative abundance
```{r}
#Let's convert our family merged count data to relative abundance data using the transform function of microbiome
ps1.com.fam.rel <- microbiome::transform(my.ps.com.fam, "compositional")

plot.composition.relAbun.Fam <- plot_composition(ps1.com.fam.rel, sample.sort = "Glacier", x.label = "Library.Name") 
plot.composition.relAbun.Fam <- plot.composition.relAbun.Fam + theme(legend.position = "bottom") 
plot.composition.relAbun.Fam <- plot.composition.relAbun.Fam + scale_fill_brewer("Family", palette = "Paired") + theme_bw() 
plot.composition.relAbun.Fam <- plot.composition.relAbun.Fam + theme(axis.text.x = element_text(angle = 90)) 
plot.composition.relAbun.Fam <- plot.composition.relAbun.Fam + ggtitle("Relative abundance") + guide_italics + theme(legend.title = element_text(size = 18))

plot.composition.relAbun.Fam
ggsave("../Outputs/Sommers2020/graphs/relative-abundance-plot-family.pdf", width=14, height=4)

ps1.com.class.rel <- microbiome::transform(my.ps.com.class, "compositional")

plot.composition.relAbun.Cla <- plot_composition(ps1.com.class.rel, sample.sort = "Glacier", x.label = "Glacier") 
plot.composition.relAbun.Cla <- plot.composition.relAbun.Cla + theme(legend.position = "bottom") 
plot.composition.relAbun.Cla <- plot.composition.relAbun.Cla + scale_fill_brewer("Class", palette = "Paired") + theme_bw() 
plot.composition.relAbun.Cla <- plot.composition.relAbun.Cla + theme(axis.text.x = element_text(angle = 90)) 
plot.composition.relAbun.Cla <- plot.composition.relAbun.Cla + ggtitle("Relative abundance") + guide_italics + theme(legend.title = element_text(size = 18))

plot.composition.relAbun.Cla
ggsave("../Outputs/Sommers2020/graphs/relative-abundance-plot-class.pdf", width=14, height=4)
```

#Step : Customize the barplot
```{r}
data.com <- plot.composition.relAbun.Cla$data
colnames(data.com)

p.com <- ggplot(data.com, aes(x=Sample, y=Abundance, fill=Tax))
p.com <- p.com + geom_bar(position="stack", stat = "identity")
p.com <- p.com + scale_x_discrete(labels = data.com$xlabels, breaks=data.com$Sample)
p.com <- p.com + scale_fill_brewer("Class", palette = "Paired") + theme_bw()
p.com <- p.com + facet_grid(~xlabel, scales = "free")
p.com <- p.com + rremove("x.text")

print(p.com)
ggsave("../Outputs/Sommers2020/graphs/relative-abundance-bar-plot-by-glacier.pdf", width=6, height=4)
```

#Step : Heat maps!
```{r}
#base plot
ps.heat <- ggplot(data.com, aes(x=Sample, y=Tax))+
  geom_tile(aes(fill=Abundance))

#change colour
ps.heat <- ps.heat + scale_fill_distiller("Abundace", palette = "RdYlBu")+
  theme_bw()

#Make bacterial names italics
ps.heat <- ps.heat + theme(axis.text.y = element_text(colour='black', size=10, face='italic'))

#Make separate samples based on main variable
ps.heat <- ps.heat + facet_grid(~xlabel, scales = "free")+ rremove("x.text")

ps.heat <- ps.heat + ylab("Class")

#Clean the axis
ps.heat <- ps.heat + theme(axis.title.x=element_blank(),
                           axis.text.x = element_blank(),
                           axis.ticks.x = element_blank())

#Clean the facet
ps.heat <- ps.heat + theme(legend.key = element_blank(),
                           strip.background = element_rect(colour="black", fill="white"))

print(ps.heat)

ggsave("../Outputs/Sommers2020/graphs/heatmap.pdf", height=4, width=6)
```

#Step : Let's try another plot
```{r}
ps.df <- microbiomeutilities::phy_to_ldf(my.ps.com.class,
                                         transform.counts = "compositional")

colnames(ps.df)

#this data frame can now be used to customize several plots
p.box <- ggstripchart(ps.df, "Glacier", "Abundance",
                      facet.by = "Class", color="Glacier", 
                      palette = "jco")
p.box + rremove("x.text")

ggsave("../Outputs/Sommers2020/graphs/strip-chart.pdf", height = 4, width = 6)
```

#Step : Beta diversity metrics - unweighted unifrac
```{r}
my.ps #non-rarefied data
ps.rar #rarefied data

#Unweighted unifrac is based on presence/absence not abundance - but it is sensitive to sequencing depth. So if a sample is of greater depth than other reads it may have more low abundance taxa and therefore effect the unifrac metric

#So we'll use our rarefied data for this.
#remove ASVs that are detected at least 10 times in 5% of samples
ps.rar.filter <- core(ps.rar, detection = 10, prevalence = 0.05)
summarize_phyloseq(ps.rar.filter) #look at the sparcity of the data, we want to reduce it. Sparcity = 0.5899
summarize_phyloseq(ps.rar) #sparcity = 0.8275

#do the unweighted unifrac with filtered phyloseq object
unw.uni.filter <- ordinate(ps.rar.filter, "PCoA", "unifrac", weighted=F)

#check for eigen values
barplot(unw.uni.filter$values$Eigenvalues[1:10])

unw.uni.filter.plot <- plot_ordination(ps.rar.filter, unw.uni.filter, color="Glacier")

unw.uni.filter.plot <- unw.uni.filter.plot + ggtitle("Unweighted UniFrac") + geom_point(size=2)
unw.uni.filter.plot <- unw.uni.filter.plot + theme_classic() + scale_color_brewer("Glacier", palette = "Set2")
print(unw.uni.filter.plot)
ggsave("../Outputs/Sommers2020/graphs/unweighted-unifrac-pcoa-filterd.pdf", width=6, height=4)

#do the unweighted unifrac with unfiltered phyloseq object
unw.uni <- ordinate(ps.rar, "PCoA", "unifrac", weighted=F)

#check for eigen values
barplot(unw.uni$values$Eigenvalues[1:10])

unw.uni.plot <- plot_ordination(ps.rar, unw.uni, color="Glacier")

unw.uni.plot <- unw.uni.plot + ggtitle("Unweighted UniFrac") + geom_point(size=2)
unw.uni.plot <- unw.uni.plot + theme_classic() + scale_color_brewer("Glacier", palette = "Set2")
print(unw.uni.plot)
ggsave("../Outputs/Sommers2020/graphs/unweighted-unifrac-pcoa-unfilterd.pdf", width=6, height=4)
```

#Step : Weighted Unifrac - considers abundances of taxa
```{r}
my.ps.rel <- microbiome::transform(my.ps, "compositional")

wt.uni <- ordinate(my.ps.rel, "PCoA", "unifrac", weighted=T)

#Check for eigen values
barplot(wt.uni$values$Eigenvalues[1:10])

wt.uni.plot <- plot_ordination(my.ps.rel, wt.uni, color="Glacier")
wt.uni.plot <- wt.uni.plot + ggtitle("Weighted UniFrac")+geom_point(size=2)
wt.uni.plot <- wt.uni.plot + theme_classic() + scale_color_brewer("Glacier", palette = "Set2")
print(wt.uni.plot)
print(wt.uni.plot + stat_ellipse())
ggsave("../Outputs/Sommers2020/graphs/weighted-unifrac-unfiltered-ellipse.pdf", width=8, height=4)
```

#Step : Population-level Density landscapes
```{r}
p <- plot_landscape(my.ps.rel,
                    "NMDS",
                    "bray",
                    col = "Glacier") +
  labs(title=paste("NMDS / Bray-Curtis"))

p <- p + scale_color_brewer(palette="Dark2") + scale_fill_gradient(low="#e0ecf4", high = "#6e016b")

p
ggsave("../Outputs/Sommers2020/graphs/NMDS-Bray-Curtis.pdf", width=8, height=4)
```

#Step : PERMANOVA
```{r}
meta.df <- data.frame(sample_data(my.ps.rel))

unifrac.dist <- UniFrac(my.ps.rel,
                        weighted=TRUE,
                        normalized=TRUE,
                        parallel=FALSE,
                        fast=TRUE)

permanova <- adonis(unifrac.dist ~ Glacier, data = meta.df)
permanova

```

#Step : Checking the homogeneity condition
```{r}

ps.disper <- betadisper(unifrac.dist, meta.df$Glacier)
permutest(ps.disper, pairwise = TRUE)
```

#Step : Core microbiota
```{r}
#non-rarefied data
my.ps

#subset data to keep only samples from Canada glacier
ps.Canada <- subset_samples(my.ps, Glacier == "Canada")

#convert to relative abundance
ps.Canada.rel <- microbiome::transform(ps.Canada, "compositional")

print(ps.Canada.rel)

ps.Canada.rel2 <- prune_taxa(taxa_sums(ps.Canada.rel) > 0, ps.Canada.rel)
print(ps.Canada.rel2) #keep only taxa with more than 0 occurrences

#Check for core ASVs
core.taxa.Canada <- core_members(ps.Canada.rel2, detection = 0.001, prevalence = 90/100) #changed this from 50/100 to 90/100

print(core.taxa.Canada) #This gives us the core 64 taxa IDs

#Let's get their classification
#Extract the taxonomy table
taxa.tab <- as.data.frame(tax_table(ps.Canada.rel2))
#Subset this taxonomy table to include only core ASVs
core_taxa_id <- subset(taxa.tab, rownames(taxa.tab) %in% core.taxa.Canada)
DT::datatable(core_taxa_id) #let's look at our core taxa classifications in an interactive table
```

#Step : Let's find the core abundance and diversity
```{r}
core.abundance <- sample_sums(core(ps.Canada.rel2, detection=0.001, prevalence = 90/100)) #changed this from 50/100 to 90/100
DT::datatable(as.data.frame(core.abundance)) #how abundant are the core taxa in each sample
```

#Step : Core visualization
```{r}
#Core heatmaps
#Core with compositionals
prevelances <- seq(.05, 1, .05)
detections <- 10^seq(log10(1e-3), log10(.2), length=10)

#Also define gray color palette
gray <- gray(seq(0,1,length=5))
p.core <- plot_core(ps.Canada.rel2,
                    plot.type = "heatmap",
                    colours=gray,
                    prevalences = prevalences,
                    detections = detections,
                    min.prevalence = .5) + 
  xlab("detection Threshold (Relative Abundance (%)")
print(p.core)

```

#Step : Let's use some better colour palettes
```{r}
print(p.core + scale_fill_viridis())

prevalences <- seq(.05, 1, .05)
detections <- 10^seq(log10(1e-03), log10(.2), length=10)

p.core <- plot_core(ps.Canada.rel2, 
                    plot.type = "heatmap", 
                    colours = rev(brewer.pal(5, "Spectral")),
                    prevalences = prevalences, 
                    detections = detections, 
                    min.prevalence = .5) + 
  xlab("Detection Threshold (Relative Abundance (%))")

print(p.core)

#Let's add our classificaition to our plot
ps.Canada.rel2.f <- microbiomeutilities::format_to_besthit(ps.Canada.rel2) #basically our ASV IDs are so long they make this plot look mad

p.core <- plot_core(ps.Canada.rel2.f, 
                    plot.type = "heatmap", 
                    colours = rev(brewer.pal(5, "Spectral")),
                    prevalences = prevalences, 
                    detections = detections, 
                    min.prevalence = .5) + 
  xlab("Detection Threshold (Relative Abundance (%))")

p.core + theme(axis.text.y = element_text(face="italic"))

# ggsave("../Outputs/Sommers2020/graphs/core-heatmap-Canada.pdf", width = 20, height = 5)

```

#Step : Inference of Microbial Ecological Networks
#Here I will be used SPIEC-EASI (SParse InversE Covariance Estimation for Ecological Association Inference), a statistical method for the inference of microbial ecological networks
```{r}
#The input for SPIEC-EASI is a counts table. The normalization and transformation is done by the function.

#let's get the Taylor Glacier data 
ps.Taylor <- subset_samples(my.ps, Glacier == "Taylor")

#for testing let's reduce the number of ASVs
ps.Taylor.small <- prune_taxa(taxa_sums(ps.Taylor) > 100, ps.Taylor)

#Add taxonomic classifications to ASV ID
ps.Taylor.small.tax <- microbiomeutilities::format_to_besthit(ps.Taylor.small)

head(tax_table(ps.Taylor.small.tax))
head(tax_table(ps.Taylor.small))
```

#Step : Prep data from SpiecEasi
```{r}
asv.c <- t(otu_table(ps.Taylor.small.tax)@.Data) #extract the asv table from the phyloseq object

tax.c <- as.data.frame(tax_table(ps.Taylor.small.tax)@.Data) #extract the taxonomy info

head(tax.c)
```

#Step : SPIEC-EASI network reconstruction
```{r}
#The input for SPIEC-EASI is a counts table. The normalization and transformation is done by the function. This process is very slow.

set.seed(1961)

net.c <- spiec.easi(asv.c, method='md', icov.select.params=list(rep.num=10)) #need more reps than this but using 10 as a starting point. Would usually use 99 or 999 reps for this.

#saveRDS(net.c, "../Outputs/Sommers2020/net.c.rds")

class(net.c)

#load custom function to get weights
source("scripts/symBeta.R")

n.c <- symBeta(getOptBeta(net.c))

#Add names to IDs
colnames(n.c) <- row.names(n.c) <- colnames(asv.c)


#add abundance values to vertex (nodes)
vsize <- log2(apply(asv.c, 2, mean)) #add log abundance as properties of vertex/nodes

```

#Step : Prepare data for plotting
```{r}

```

#Step : 
```{r}

```

#Step :
```{r}

```

#Step : 
```{r}

```

#Step :
```{r}

```

#Step : 
```{r}

```

#Step :
```{r}

```

#Step : 
```{r}

```

#Step :
```{r}

```

#Step : 
```{r}

```

#Step :
```{r}

```