---
title: "ordination"
author: "Amy Solman"
date: "08/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This script will perform ordination of the data I prepared in step two.
#Step One: Clear my working environment, load my packages and my data.

```{r}
rm(list=ls())

graphics.off()
library(tidyverse)
library(phyloseq)
library(ape)

#Load my data object
load("results/cleaned_files/cleanedfiles.RData")

#Load my phyloseq object
my_physeq <- readRDS("results/my-phyloseq-object.rds")

#phy_tree(my_physeq)
```

#Step Two: Check the distribution of phyla in my phyloseq object

```{r}
# Aglomerate the ASVs at the phylum level
my_physeq_glom = tax_glom(my_physeq, "Phylum")
# Plot
plot_bar(my_physeq_glom, x = "Sample", fill = "Phylum") #why does this look different from when I did it in the other ordination script (ordination-1.Rmd)?

#plot_bar(my_physeq_glom, x = "Sample") 


```
#Step Three: Calculate Bray-Curtis dissimilarity matrix

```{r}
bray_distance <- distance(my_physeq, method = "bray")
bray_distance
```

#Step Four: Carry out bray-curtis and unweighted unifrac PCoA analysis and plot, scaled by the eigenvalues
```{r}
#Bray-Curtis
my_pcoa_bray <- ordinate(my_physeq, method = "PCoA", distance = "bray")
my_eigenvalues_bray <- my_pcoa_bray$values$Eigenvalues
pcoa_plot_bray = plot_ordination(my_physeq, my_pcoa_bray, color = "Location") +
  geom_point(size = 3)+
  coord_fixed(sqrt(my_eigenvalues_bray[2]/my_eigenvalues_bray[1]))
pcoa_plot_bray
```

#Step Five: Carry out unweighted unifrac PCoA analysis and plot, scaled by the eigenvalues
```{r}
#Unweighted UniFrac
my_pcoa_wuni <- ordinate(my_physeq, method = "PCoA", distance = "wunifrac")
my_eigenvalues_wuni <- my_pcoa_wuni$values$Eigenvalues
pcoa_plot_wuni = plot_ordination(my_physeq, my_pcoa_wuni, color = "Location") +
  geom_point(size = 3)+
  coord_fixed(sqrt(my_eigenvalues_wuni[2]/my_eigenvalues_wuni[1]))
pcoa_plot_wuni
```

#Step Six: Create an NMDS plot 
```{r}
my_nmds <- ordinate(my_physeq, method = "NMDS", distance = "bray")
nmds_plot = plot_ordination(my_physeq, my_nmds, color = "Location") +
  geom_point(size = 3)
nmds_plot
```


