---
title: "data-prep"
author: "Amy Solman"
date: "05/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script will prepare our QIIME2 output for use in downstream analysis.

#Step One: Clear work space and load packages

```{r}
rm(list=ls())
graphics.off()

library(tidyverse)
library(phyloseq)
library(readr)
library(seqinr)
library(decontam)
library(ape)
library(vegan)
library(magrittr)
library(data.table)
```

#Step Two: Load the data from our QIIME2 step +metadata

```{r}
count_table <- read_tsv(file="export/table/table.tsv", skip=1)

taxonomy <- read_tsv(file="export/taxonomy/taxonomy.tsv")

tree <- read_tree("export/exported-tree/tree.nwk")

fasta <- read.fasta(file="export/rep-seqs.fasta/dna-sequences.fasta")

sample_info_tab <- read_csv(file="export/SraRunTable-wrangle.tsv")
```
#Step Three: Specify the first column in our count table is the row names
```{r}
count_table <- column_to_rownames(count_table, var = colnames(count_table)[1])
#rownames(count_table) <- paste0("ASV", 1:nrow(count_table)) <- removed
```

#Step Four: Load metadata table and remove unnecessary rows/columns/data

```{r}
#delete unneccessary row and column
sample_info_tab <- sample_info_tab[-c(1),]
sample_info_tab <- sample_info_tab[,-c(1)]
sample_info_tab <- sample_info_tab[!grepl("18S", sample_info_tab$Library.Name),] #only include 16S samples
sample_info_tab <- sample_info_tab[-c(42:80),] #remove reverse reads so we only have one entry per sample
#rename columns that didn't get the right name imported from the metadata
colnames(sample_info_tab)[10] <- "Chlorophyll"
colnames(sample_info_tab)[11] <- "Collection_date"
colnames(sample_info_tab)[16] <- "Depth"
colnames(sample_info_tab)[36] <- "pH"
colnames(sample_info_tab)[39] <- "Replicate"
colnames(sample_info_tab)[47] <- "Temp"

#Im also going to remove some of the text within these data so we can just work with the numbers
#we use the parse_number() function from the readr package
sample_info_tab$Chlorophyll <- parse_number(sample_info_tab$Chlorophyll)
sample_info_tab$org_matter <- parse_number(sample_info_tab$org_matter)
sample_info_tab$water_content <- parse_number(sample_info_tab$water_content)

#colnames(count_table) <- unique(sample_info_tab$Sample.Name) <- removed

#add a column with glacier name
Glacier <- NULL
for (i in 1:nrow(sample_info_tab)){
  if (startsWith(sample_info_tab$Sample.Name[[i]], "C")){
    Glacier <- c(Glacier, "Canada")
  } else {
    Glacier <- c(Glacier, "Taylor")
  }
}

sample_info_tab <- cbind(sample_info_tab, Glacier)

#Keep only the information from the metadata table that will be useful in later analysis and make into a phyloseq object
sampledata = sample_data(data.frame(
  Sample_name = sample_info_tab$Run, #changed this to Run from Sample. <--
  Location = sample_info_tab$Glacier,
  Lat_Long = sample_info_tab$lat_lon,
  Elevation = sample_info_tab$elev,
  Depth = sample_info_tab$Depth,
    Diameter1 = sample_info_tab$diameter1,
  Diameter2 = sample_info_tab$diameter2,
  Organic_matter = sample_info_tab$org_matter,
  Water_content = sample_info_tab$water_content,
  Chlorophyll = sample_info_tab$Chlorophyll,
  pH = sample_info_tab$pH,
  Temp = sample_info_tab$Temp,
  DNA_per_g = sample_info_tab$DNA_per_g,
  row.names = sample_info_tab$Run, #changed this to Run from Sample.Name <--
  stringsAsFactors = FALSE
))
```
#Step Five: Modifty taxonomy table
```{r}
taxonomy <- taxonomy %>%
  mutate(taxonomy=str_replace_all(string=Taxon, pattern="D_\\d*\\__", replacement="")) %>%
  mutate(taxonomy=str_replace_all(string=taxonomy, pattern=";$", replacement="")) %>%
separate(taxonomy, into=c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep=";") %>%
  select(-Taxon, -Confidence) %>%
  column_to_rownames(var='Feature ID')

row.names(taxonomy) <- row.names(count_table) #give taxonomy same ASV names as count_table
```

#Step Six: Identify contaminents in the count data
```{r}
# A logical vector showing which entries are 'TRUE' for controls and 'FALSE' for genuine samples
#decontam <- startsWith(sampledata$Sample_name, "NT")
decontam <- startsWith(sample_info_tab$Sample.Name, "NT")

# Run isContaminant (remember to transform the dataframe i.e. t())
contam <- isContaminant(t(count_table), neg=decontam)

# Look at the results
table(contam$contaminant) # how many ASVs were identified as contaminants? None. Why?

# Make a vector holding the contaminant IDs
contam_asvs <- row.names(contam[contam$contaminant == TRUE, ])

# What is the taxonomy of our contaminant ASVs?
taxonomy[taxonomy$`Feature ID` %in% contam_asvs] #Why don't I have any contaminants?!?

```

#Step Seven: Remove contaminents from count table,taxonomy table, sequence data, metadata and tree
```{r}

###COUNT TABLE
x <- rownames(count_table) %in% contam_asvs #which rows are contaminants?
count_table_clean <- count_table[!x,] #remove those rows

# and remove blank samples from count table
#what is the index of "TRUE" values in decontam?
true_index <- NULL
for (i in 1:length(decontam)){
  if (decontam[[i]] == "TRUE"){
    true_index <- c(true_index, i)
  }
}
count_table_clean <- count_table_clean[,-c(true_index)]  #remove those samples

###TAXONOMY TABLE
z <- taxonomy$`Feature ID` %in% contam_asvs
taxonomy_clean <- taxonomy[!row.names(taxonomy) %in% z,]#remove contaminating ASVs

###FASTA FILE
#Remove contaminants from dna sequences file
fasta_clean <- fasta
fasta_clean[contam_asvs] <- NULL

###METADATA TABLE
#y <- startsWith(sampledata$Sample_name, "NT")
y <- startsWith(sample_info_tab$Sample.Name, "NT")
true_index <- NULL
for (i in 1:length(y)){
  if (y[[i]] == "TRUE"){
    true_index <- c(true_index, i)
  }
}
sampledata_clean <- sampledata[-c(true_index),]  #remove those samples

###PHYLOGENETIC TREE
# This is a phyloseq object so we can use a phyloseq command called prune_taxa
tree_clean <- prune_taxa(count_table_clean$`#OTU ID`, tree)
```

#Step Eight: Plot rarefaction curves to see how deeply our samples have been sequenced
```{r}
#Plot the rarefaction curves, color-coding by sample and tranform using t()
#Col= option could also be used to look at different sample types/locations?
#pdf("results/rplot.pdf") 
rarecurve(t(count_table_clean), step=100, cex=0.5, col=palette(rainbow(nrow(sampledata_clean))), lwd=2, ylab="ASVs", label=T)
# Add line to indicate the fewest # of sequences from any sample
abline(v=(min(rowSums(t(count_table_clean)))))
#dev.off()
```

#Step Nine: Convert data to phyloseq objects and remove singletons

```{r}

###COUNT TABLE
# Convert count table to phyloseq package
count_table_phyloseq <- otu_table(as.matrix(count_table_clean), taxa_are_rows = TRUE)
ASV <- prune_taxa(taxa_sums(count_table_phyloseq) > 1, count_table_phyloseq) # Filter using prune_taxa

###TAXONOMY TABLE
TAX = tax_table(as.matrix(taxonomy_clean))
TAX <- prune_taxa(taxa_sums(count_table_phyloseq) > 1, TAX)

###METADATA TABLE 
META = sample_data(sampledata_clean)

###PHYLOGENETIC TREE
# This is a phyloseq object so we can use a phyloseq command called prune_taxa
#taxa_names(tree_clean) = taxa_names(count_table_phyloseq) #Change tree names to match ASVs
TREE <- prune_taxa(taxa_sums(count_table_phyloseq) > 1, tree_clean)

#Make into one phyloseq object
my_physeq = phyloseq(ASV, TAX, META, TREE)
```

#Step Ten: remove unassigned taxa and eukaryotes (becuase we're looking at 16S here)
If we check the data in our phyloseq object we see that there are some entries assigned to eukaryotes or not assigned at all. We want to remove these things and then save the object 
```{r}
#look at the data
unique(tax_table(my_physeq)[, "Kingdom"])
table(tax_table(my_physeq)[, "Kingdom"], exclude = NULL)

#remove what we don't want
my_physeq <- subset_taxa(my_physeq, !is.na(Kingdom) & !Kingdom %in% c("Unassigned", "Eukaryota"))

#check it again
table(tax_table(my_physeq)[, "Kingdom"], exclude = NULL)
```
Let's look at our phyla names
```{r}
table(tax_table(my_physeq)[, "Phylum"], exclude = NULL)
```
#Step Eleven: Remove sequences for which we don't know the phyla and those with very low abundances:
```{r}
#get names of phyla with abundances of 2 or less
x <- table(tax_table(my_physeq)[, "Phylum"], exclude = NULL)[table(tax_table(my_physeq)[, "Phylum"], exclude = NULL) < 3]
#remove data points without a phylum classification and with names in our x variable
my_physeq <- subset_taxa(my_physeq, !is.na(Phylum) & !Phylum %in% rownames(x))

#now we're just left with sequences with identified phyla and abundance of 3 or more
table(tax_table(my_physeq)[, "Phylum"], exclude = NULL)
```
#Step Twelve: Re-root the tree in case we lost the root while pruning and then save!
```{r}
# Define the function given here > (https://john-quensen.com/r/unifrac-and-tree-roots/) which picks the longest branch to root from 
pick_new_outgroup <- function(tree.unrooted){
  #   require("magrittr")
  # require("data.table")
  # require("ape") #appe::Ntip
  #tablif parts of the tree that we need
  treeDT <- 
    cbind(
      data.table(tree.unrooted$edge),
      data.table(length = tree.unrooted$edge.length)
    )[1:Ntip(tree.unrooted)] %>%
    cbind(data.table(id = tree.unrooted$tip.label))
  #Take out the longest terminal branch as outgroup
  new.outgroup <- treeDT[which.max(length)]$id
  return(new.outgroup)
}

#pull out the tree from my phyloseq object
my_tree <- phy_tree(my_physeq) 

#Run the function on the tree
new_outgroup <- pick_new_outgroup(my_tree)

# Use the outgroup to re-root the tree
new_tree <- ape::root(my_tree, outgroup=new_outgroup, resolve.root=TRUE)

# And convert to dichotomy tree - what does this mean?
new_tree1 <- ape::multi2di(new_tree)

#add the new tree to my phyloseq object
phy_tree(my_physeq) <- new_tree1

#Now I'm going to export my phyloseq object
saveRDS(my_physeq, "results/my-phyloseq-object.rds")
```

###Step Thirteen: Perform Hellinger transformation

```{r}
count_table_hellinger <- decostand(t(count_table_clean), "hellinger")
# Convert back to a tibble and switch rows for columns again 
count_table_hellinger <- as_tibble(t(count_table_hellinger), rownames = NA)
```

###Step Fourteen: Export data files 
```{r}
#Note singletons have not been removed from these files, only the phyloseq object
save(count_table_clean, count_table_hellinger, taxonomy_clean, fasta_clean, tree_clean, sampledata_clean, file = "results/cleaned_files/cleanedfiles_1.RData")
```