---
title: "Model Simplification"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/18-ModelSimp.ipynb
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#This script will look at MODEL SIMPLIFICATION
#Basically, we want to find our which terms or variables in our model are best at describing our data in the least amount of pvariables.
#So we start with a MAXIMAL model (one that contains everything that might be important) and simplifying it toward the NULL model (the one that says none of the variables are important). There should be a point in between the MAXIMAL MODEL and the NULL MODEL where we can't make the model simpler without making it worse = MINIMUM ADEQUATE MODEL

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library("picante")
#library(gplots)
#library(lattice)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Six: Prepare our data
```{r}
#get logs of the continuous variables we are interested in
my.data.sub$logChlo <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA <- log(my.data.sub$DNA_per_g)
my.data.sub$logOrg <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logArea <- log(my.data.sub$area)
my.data.sub$logRich <- log(my.data.sub$Richness)
my.data.sub$logWat <- log(my.data.sub$water_content)
```

#Step Seven: Maximal Model
```{r}
model <- lm(formula = logRich ~ logChlo * logDepth * logArea * logpH * logOrg * logDNA * logWat * Glacier, data = my.data.sub)

model <- lm(formula = logRich ~ logArea * logpH * logOrg * logWat * Glacier, data = my.data.sub)

summary(model)

anova(model)

#As we can see here, starting with all possible variables in the model doesn't work very well. It's difficult to interpret these results. Instead we can consider two way interactions - how do pairs of variables affect each other?

#Let's try this again!

#model <- lm(logRich ~ (logChlo + logDepth + logArea + logpH + logOrg + logDNA + as.factor(Glacier))^2, data = my.data.sub)
#model <- lm(logRich ~ (logChlo + logArea + logpH + logOrg + Glacier)^2, data = my.data.sub)
model <- lm(logRich ~ (logArea + logpH + logOrg + logWat + Glacier)^2, data = my.data.sub)

anova(model)
#The first lines here are our main effects, followed by the paired interactions
#We can see that area, organic matter and glacier are significant as main effects (pH might also be somewhat significant), whilst the slope of log Richness with organic matter differs with area (maybe, this significance is weak)
```

#Step Eight: Simplify the model!
#We can use F-tests to compare a model with a simplified model - significance of the F test is BAD, the model has got SIGNIFICANTLY worse
#So which terms can we remove from the model?
```{r}
drop.scope(model) #tells you what you can drop from a model

#Instead of re-typing the model we can use the update function to change it, for example...

# a simple model
f <- y ~ a + b + c + b:c
# remove b:c from the current model
update(f, . ~ . - b:c)
# model g as a response using the same explanatory variables.
update(f, g ~ .)

#from the ANOVA output I can see that the term logArea:Glacier is not significant (the drop output showed me that this was a valid term in the model)
model2 <- update(model, . ~ . - logpH:logWat)

#we can use ANOVA to compare the two models
anova(model, model2) #the p value = 0.58 so we know model 2 is not significantly worse that model

#let's get an anova table for model 2 to pick the next variable to remove!
anova(model2)

model3 <- update(model2, . ~ . - logWat:Glacier)
anova(model2, model3) #p = 0.79

anova(model3)
summary(model3)

model4 <- update(model3, . ~ . - logArea:logpH)
anova(model3, model4) #p = 0.94

anova(model4)
summary(model4)
model5 <- update(model4, . ~ . - logOrg:Glacier)
anova(model4, model5) #p = 0.19

anova(model5)
summary(model5)
model6 <- update(model5, . ~ . - logOrg:logWat)
anova(model5, model6) #p = 0.17

anova(model6)
summary(model6)
model7 <- update(model6, . ~ . - logpH:Glacier)
anova(model6, model7) #p = 0.21

anova(model7)
summary(model7)
model8 <- update(model7, . ~ . - logpH:logOrg)
anova(model7, model8) #p = 0.13

anova(model8)
summary(model8)
model9 <- update(model8, . ~ . - logArea:logOrg)
anova(model8, model9) #p = 0.09

anova(model9)
summary(model9)
model10 <- update(model9, . ~ . - logArea:logWat)
anova(model9, model10) #p = 0.14

anova(model10)
summary(model10)
model11 <- update(model10, . ~ . - logArea:Glacier)
anova(model10, model11) #p = 0.24

anova(model11)
summary(model11)
model12 <- update(model11, . ~ . - logWat)
anova(model11, model12) #p = 0.25

anova(model12)
summary(model12)
model13 <- update(model12, . ~ . - logpH)
anova(model12, model13) #p = 0.06

anova(model13)
summary(model13)
model14 <- update(model13, . ~ . - logOrg)
anova(model13, model14) #p = 0.83


###I CAN'T REMOVE ANYTHING ELSE HERE WITHOUT MAKING THE MODEL WORSE
anova(model14) #< FINAL MODEL = MINIMUM ADEQUATE MODEL
summary(model14)
# model15 <- update(model14, . ~ . - logArea)
# anova(model14, model15) #p = 0.83

```
