---
title: 'Linear Models: Looking at Multiple Variables with Interactions'
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/17-MulExplInter.ipynb) 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

######I WILL BE USING FAKE DATA FOR THIS ANALYSIS FOR DEMONSTRATION PURPOSES#####

#Here we're going to look at MULTI EXPLANATORY VARIABLE LINEAR MODELS with INTERACTIONS. We will also learn how to use ANCOVA.

#Two models we will build:
#Model 1 - Is species richness predicted by interactions between glacier and substrate?
#ANCOVA - Is species richness predicted by interactions between Glacier and hole area?

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library("picante")
library(gplots)
library(lattice)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Six: Add another fake Glacier of data so we can compare three glaciers - SKIP THIS STEP NO NECCESSARY
```{r}
# Chlorophyll <- sample(as.numeric(my.data.sub$Chlorophyll), size = 12)
# Depth <- sample(as.numeric(my.data.sub$Depth), size = 12)
# DNA_per_g <- sample(as.numeric(my.data.sub$DNA_per_g), size = 12)
# org_matter <- sample(as.numeric(my.data.sub$org_matter), size = 12)
# pH <- sample(as.numeric(my.data.sub$pH), size = 12)
# water_content <- sample(as.numeric(my.data.sub$water_content), size = 12)
# area <- sample(as.numeric(my.data.sub$area), size = 12)
# Faith <- sample(as.numeric(my.data.sub$Faith), size = 12)
# Richness <- sample(as.numeric(my.data.sub$Richness), size = 12)
# Shannon <- sample(as.numeric(my.data.sub$Shannon), size = 12)
# Simpson <- sample(as.numeric(my.data.sub$Simpson), size = 12)
# Inv.Simpson <- sample(as.numeric(my.data.sub$Inv.Simpson), size = 12)
# Pilou <- sample(as.numeric(my.data.sub$Pilou), size = 12)
# Glacier <- rep("Solman", 12)
# 
# df <- cbind(Chlorophyll, Depth, DNA_per_g, org_matter, pH, water_content, area, Faith, Richness, Shannon, Simpson, Inv.Simpson, Pilou, Glacier)
# 
# my.data.sub.X <- rbind(my.data.sub, df)
# my.data.sub.X.1 <- my.data.sub.X[,-c(14)] #remove glacier names
# 
# my.data.sub <- as.data.frame(lapply(my.data.sub.X.1, function(x) as.numeric(as.character(x))))
# #re-add glacier names 
# my.data.sub$Glacier <- as.factor(my.data.sub.X$Glacier)
```

#Step Seven: Add an addition factor for us to play with
```{r}
Substrate <- rep(c("Ice", "Snow", "Cryoconite"), 9) 
Substrate <- c(Substrate, rep("Cryoconite", 1))
my.data.sub$Substrate <- as.factor(Substrate)
```

#Step :
```{r}

```

#Step :
```{r}

```

#Step :
```{r}

```

#Step :
```{r}

```

#Step :
```{r}

```

#Step :
```{r}

```

#Step :
```{r}

```