---
title: "Model Comparison"
author: "Amy Solman"
date: "24/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/20-ModelFitting-NLLS.ipynb)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#THIS SCRIPT WILL DEFINE A SERIES OF MODELS, FIT THEM TO MY DATA AND ASSESS WHICH IS THE BEST FIT

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library(picante)
#library(gplots)
#library(lattice)
library(ggplot2)
library(minpack.lm)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Four: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Five: Prepare our data
```{r}
#get logs of the continuous variables we are interested in
my.data.sub$logChlo <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA <- log(my.data.sub$DNA_per_g)
my.data.sub$logOrg <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logArea <- log(my.data.sub$area)
my.data.sub$logRich <- log(my.data.sub$Richness)
my.data.sub$logWat <- log(my.data.sub$water_content)
```

#Step Six: Define our models as functions we can call when NLLS fitting
```{r}
#Straight Line Model
lineMod <- function(x,a,b){
  return(a + b*x)
}

#Power Law Model
powMod <- function(x, a, b) {
 return(a * x^b)
}

#Quadratic curve
quaMod <- function(x, a, b, c){
  return(a + b*x + c*x^2)
}

#Von Bertalanffy
vonbert.w <- function(t, Winf, c, K){
 Winf * (1 - exp(-K * t) + c * exp(-K * t))^3
}

#Logistic
logistic1 <- function(t, r, K, N0){
 N0 * K * exp(r * t)/(K+N0 * (exp(r * t)-1))
}
```

#Step Seven: Fit the five models to the data
```{r}
#subset our canada data 
Data2Fit <- subset(my.data.sub, Glacier == "Canada")

#We will scale the data before fitting to improve the stability of the estimates
#set a scale
scale <- 4000

#Straight Line Model
#LinFit <- nlsLM(Richness ~ lineMod(area, a, b), data = Data2Fit, start=list(a=.1, b=.1))
#we can achieve the same thing with a standard lm() fitting
LinFit2 <- lm(Richness/scale ~ area, data = Data2Fit)

#Power Model
PowFit <- nlsLM(Richness/scale ~ powMod(area, a, b), data = Data2Fit, start=list(a=.1, b=.1))

#Quadratic Curve
# QuaFit <- nlsLM(Richness ~ quaMod(area, a, b, c), data = Data2Fit, start=list(a=.1, b=.1, c=.1))
#we can achieve the same thing with the poly command 
QuaFit <- lm(Richness/scale ~ poly(area,2), data = Data2Fit)

#Von Bertalanffy Model - this model isn't working so we'll leave it for now
VonFit <- nlsLM(Richness/scale ~ vonbert.w(area, Winf, c, K), start=list(Winf=0.75, c=0.01, K=0.01), data=Data2Fit)

#Logistic Model
LogFit <- nlsLM(Richness ~logistic1(area, r, K, N0), start=list(K=600, r=0.1, N0=300), data=Data2Fit)

```

#Step Eight: Plot the models with our data
```{r}
#get areas for plotting
Areas <- seq(min(Data2Fit$area), max(Data2Fit$area), len=200)

#get points for plotting
#Straight Line
Predic2PlotLin <- predict.lm(LinFit2, data.frame(area = Areas)) * scale

#Power Model
#Predic2PlotPow <- predict.lm(PowFit, data.frame(area = Areas))
Predic2PlotPow <- powMod(Areas, coef(PowFit)["a"],coef(PowFit)["b"]) *scale

#Quadratic Model
Predic2PlotQua <- predict.lm(QuaFit, data.frame(area = Areas)) * scale

#Von Bertalanffy Model
Predic2PlotVon <- predict(VonFit, newdata = list(area=Areas)) * scale

#Logistic Model
# Predic2PlotLog <- predict.lm(LogFit, data.frame(area = Areas))
Predic2PlotLog <- predict(LogFit, newdata = list(area=Areas)) #*scale

plot(Data2Fit$area, Data2Fit$Richness)
lines(Areas, Predic2PlotLin, col="blue", lwd=2.5)
lines(Areas, Predic2PlotPow, col="red", lwd=2.5)
lines(Areas, Predic2PlotQua, col="pink", lwd=2.5)
lines(Areas, Predic2PlotVon, col="orange", lwd=2.5)
lines(Areas, Predic2PlotLog, col="green", lwd=2.5)
legend("topleft", legend = c("linear", "power", "quadratic", "Von Bert", "logistic"), lwd=2, lty=1, col=c("blue", "red", "pink", "orange", "green"))
```

#Step Nine: Get the R-Squared Values of the fitted models
```{r}
RSS_Lin <- sum(residuals(LinFit2)^2) # Residual sum of squares
TSS_Lin <- sum((Data2Fit$Richness - mean(Data2Fit$Richness))^2) # Total sum of squares
RSq_Lin <- 1 - (RSS_Lin/TSS_Lin) # R-squared value

RSS_Pow <- sum(residuals(PowFit)^2) # Residual sum of squares
TSS_Pow <- sum((Data2Fit$Richness - mean(Data2Fit$Richness))^2) # Total sum of squares
RSq_Pow <- 1 - (RSS_Pow/TSS_Pow) # R-squared value

RSS_Qua <- sum(residuals(QuaFit)^2) # Residual sum of squares
TSS_Qua <- sum((Data2Fit$Richness - mean(Data2Fit$Richness))^2) # Total sum of squares
RSq_Qua <- 1 - (RSS_Qua/TSS_Qua) # R-squared value

RSS_Von <- sum(residuals(VonFit)^2) # Residual sum of squares
TSS_Von <- sum((Data2Fit$Richness - mean(Data2Fit$Richness))^2) # Total sum of squares
RSq_Von <- 1 - (RSS_Von/TSS_Von) # R-squared value

RSS_Log <- sum(residuals(LogFit)^2) # Residual sum of squares
TSS_Log <- sum((Data2Fit$Richness - mean(Data2Fit$Richness))^2) # Total sum of squares
RSq_Log <- 1 - (RSS_Log/TSS_Log) # R-squared value

RSq_Lin
RSq_Pow 
RSq_Qua
RSq_Von
RSq_Log
```

#Step Ten: Use AIC to compare models
```{r}
#A difference of more than 2 tells us one more is significantly better than the first 
#value < -2 = the first model is better, value > 2 = the second model is better
AIC(LinFit2) - AIC(PowFit) #power fit is better
AIC(LinFit2) - AIC(QuaFit) #no significant difference
AIC(LinFit2) - AIC(VonFit) #the von model is better
AIC(LinFit2) - AIC(LogFit) #the line model is better
AIC(PowFit) - AIC(QuaFit) #the power model is better
AIC(PowFit) - AIC(VonFit) #the von model is better
AIC(PowFit) - AIC(LogFit) #the power model is better
AIC(QuaFit) - AIC(VonFit) #the von model is better
AIC(QuaFit) - AIC(LogFit) #the quadratic model is better
AIC(VonFit) - AIC(LogFit)  #the von model is better

#The von model seems most succcessful as describing our data

```

#Step Eleven: Let's compare the residuals of the models
```{r}
par(mfrow=c(4,1))
plot(Data2Fit$area, resid(LinFit2), main="LM resids", xlim=c(0,7000))
plot(Data2Fit$area, resid(PowFit), main="Power resids", xlim=c(0,7000))
plot(Data2Fit$area, resid(QuaFit), main="Quadratic resids", xlim=c(0,7000))
plot(Data2Fit$area, resid(VonFit), main="VB resids", xlim=c(0,7000))
#plot(Data2Fit$area, resid(LogFit), main="Logistic resids", xlim=c(0,7000)) #ouldn't fit all five models on here
```

#Step Twelve: Calculate the adjusted sum of squares for each mode
```{r}
n <- length(Data2Fit$Richness)
list(lin=signif(sum(resid(LinFit2)^2)/(n-2 * 2), 3), 
 pow= signif(sum(resid(PowFit)^2)/(n-2 * 3), 3), 
 qua= signif(sum(resid(QuaFit)^2)/(n-2 * 3), 3),
 vb= signif(sum(resid(VonFit)^2)/(n-2 * 3), 3),
 log= signif(sum(resid(LogFit)^2)/(n-2 * 3), 3)
 )

#The model with the lowest SSE is the best by this measure = VB model!
```

#Step Thirteen:
```{r}

```

#Step Fourteen:
```{r}

```

#Step Fifteen:
```{r}

```