---
title: "Model Fitting with Maximum Likelihood"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---

#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/21-ModelFitting-MLE.ipynb)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#MODEL FITTING USING MAXIMUM LIKELIHOOD
#When we have previously used ordinary least squares (OLS) or non-linear least squares (NLLS) methods for model fitting, we are essentially choosing model parameters to minimize the sum of the squares of the residuals/errors around the fitted model.

#An alternative to this is to find parameters to the function such that the likelihood of the parameters, given the data and the model, is maximized.

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library(picante)
#library(gplots)
#library(lattice)
library(ggplot2)
library(minpack.lm)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```


#Step Three: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Four: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Five: Prepare our data
```{r}
#get logs of the continuous variables we are interested in
my.data.sub$logChlo <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA <- log(my.data.sub$DNA_per_g)
my.data.sub$logOrg <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logArea <- log(my.data.sub$area)
my.data.sub$logRich <- log(my.data.sub$Richness)
my.data.sub$logWat <- log(my.data.sub$water_content)
```

#Step Six: Build an R function that returns the (negative log) likelihood for simple linear regression (it is negative log because the log of likelihood is itself negative)
#Standard linear regression assumes every observation in the dataset was generated by the model. That is, it is a model for the CONDITIONAL DISTRIBUTION of Y given X.
#Interpretting this function as a function of the parameters gives us the likelihood of the ith datapoint. We can use this to estimate the parameters of the model.
```{r}
nll.slr <- function(par, dat, ...){ #... is so we can use the optim() function later
 args <- list(...)
 
 b0 <- par[1]
 b1 <- par[2]
 X <- dat$X
 Y <- dat$Y
 if(!is.na(args$sigma)){
  sigma <- args$sigma
 } else 
  sigma <- par[3]

 mu <- b0+b1 * X
 
 return(-sum(dnorm(Y, mean=mu, sd=sigma, log=TRUE))) #dnorm function calculates the logged probability of observing Y given mu, sigma and that X
 #The negative sign on sum() is because the optim() function in R will minimize the negative log-likelihoog, which is a sum. 
}
```

#Step Seve: Generate simulated data
```{r}
set.seed(123)
n <- 30
b0 <- 10
b1 <- 3
sigma <- 2
X <- rnorm(n, mean=3, sd=7)
Y <- b0 + b1 * X + rnorm(n, mean=0, sd=sigma)
dat <- data.frame(X=X, Y=Y) # convert to a data frame

plot(X, Y)
```

#Step Eight: Build a likelihood profile
#We calculate the negative log-likeihood for fixed b1, across a range of b0.
#b1 and b0 = coefficients of the predictor variable X (the intercept and slope)
```{r}
N <- 50
b0s <- seq(5, 15, length=N)
mynll <- rep(NA, length=50)
for(i in 1:N){
 mynll[i] <- nll.slr(par=c(b0s[i],b1), dat=dat, sigma=sigma)
}

#plot the profile
plot(b0s, mynll, type="l")
abline(v=b0, col=2)
abline(v=b0s[which.min(mynll)], col=3)

#The true value for b0 is the red line (10). The value that minimizes the log-likelihood is the green line. They are not the same because maximum likelihood is estimating the true value given the measurement errors.
```

#Step Nine: Build a two-dimensional maximum likelihood surface
#If we want to estimate two parameters (b0 and b1) we need a two-dimensional maximum likelihood surface!
#We can do this by doing a grid search to fine the likelihood surface
```{r}
N0 <- 100
N1 <- 101
b0s <- seq(7,12, length=N0)
b1s <- seq(1,5, length=N1)

mynll <- matrix(NA, nrow=N0, ncol=N1)
for(i in 1:N0){
 for(j in 1:N1) mynll[i,j] <- nll.slr(par=c(b0s[i],b1s[j]), dat=dat, sigma=sigma)
}

ww <- which(mynll==min(mynll), arr.ind=TRUE)

b0.est <- b0s[ww[1]]
b1.est <- b1s[ww[2]]
rbind(c(b0, b1), c(b0.est, b1.est))

filled.contour(x = b0s, y = b1s, z= mynll, col=heat.colors(21), 
    plot.axes = {axis(1); axis(2); points(b0,b1, pch=21); 
       points(b0.est, b1.est, pch=8, cex=1.5); xlab="b0"; ylab="b1"})

#Asterisk = true parameter combination
#Circle = parameter combination that maximizes the negative log-likelihood
```

#Step Ten: Look at the conditional surface
#We look at the slice around whatever the best estimate is for the other parameter
```{r}
par(mfrow=c(1,2), bty="n")
plot(b0s, mynll[,ww[2]], type="l", xlab="b0", ylab="NLL")
plot(b1s, mynll[ww[1],], type="l", xlab="b1", ylab="NLL")
```

#Step Eleven: Do the model fitting
#Use R's optim() function to find the appropriate parameter values that achieve the maximum of the likelihood surface. 
#The first argument for optim() is the function that you want to minimize, and the second is a vector of starting values for your parameters. 
#After the main arguments, add what you need to evaluate your functions (e.g. sigma)
```{r}
fit <- optim(nll.slr, par=c(2, 1), method="L-BFGS-B", ## this is a n-D method
    lower=-Inf, upper=Inf, dat=dat, sigma=sigma)

fit

#We can fit sigma at the same time if we want
#we assign the starting values (b0 = 2, b1= 1, sigma =5) as we would with NLLS
#We have also bounded the arguments lower=, upper=
fit <- optim(nll.slr, par=c(2, 1, 5), method="L-BFGS-B", ## this is a n-D method
    lower=c(-Inf, -Inf, 0.1), upper=Inf, dat=dat, sigma=NA)
fit$par
```

#Step Twelve: Visualize the plot
```{r}
plot(X, Y)
abline(a=fit$par[1], b=fit$par[2], col=2, lwd=2)
```

#Step Thirteen: Estimate the confidence intervals
```{r}
fit <- optim(nll.slr, par=c(2, 1), method="L-BFGS-B", hessian=TRUE, lower=-Inf, upper=Inf, dat=dat, sigma=sigma)

fisher_info <- solve(fit$hessian)
est_sigma <- sqrt(diag(fisher_info))
upper <- fit$par+1.96 * est_sigma
lower <- fit$par-1.96 * est_sigma
interval <- data.frame(value=fit$par, upper=upper, lower=lower)
interval
```

#Step Fourteen: Compare with ordinary least squares

```{r}
lmfit <- lm(Y~X)

summary(lmfit)$coeff #the coefficients from OLS are almost the same as MLE
```


