---
title: "Model Fitting with Maximum Likelihood"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---

#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/21-ModelFitting-MLE.ipynb)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#MODEL FITTING USING MAXIMUM LIKELIHOOD
#When we have previously used ordinary least squares (OLS) or non-linear least squares (NLLS) methods for model fitting, we are essentially choosing model parameters to minimize the sum of the squares of the residuals/errors around the fitted model.

#An alternative to this is to find parameters to the function such that the likelihood of the parameters, given the data and the model, is maximized.

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library(picante)
#library(gplots)
#library(lattice)
library(ggplot2)
library(minpack.lm)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```


#Step Three: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Four: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Five: Prepare our data
```{r}
#get logs of the continuous variables we are interested in
my.data.sub$logChlo <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA <- log(my.data.sub$DNA_per_g)
my.data.sub$logOrg <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logArea <- log(my.data.sub$area)
my.data.sub$logRich <- log(my.data.sub$Richness)
my.data.sub$logWat <- log(my.data.sub$water_content)
```