---
title: "Direct or Constrained Ordination"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---
#This code will follow the tutorial and notes from Applied Quantitative Methods
(http://dmcglinn.github.io/quant_methods/lessons/multivariate_models.html)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#MULTIVARIATE ORDINATION ANALYSIS
#Two types of ordination:
#Indirect or unconstrained ordination in which only a single matrix is analyzed
#Direct on constrained ordination in which one matrix is used to explain the variance of another matrix

#Ordination is frequently used when exploring patterns in datasets graphically. It can also be used in hypothesis testing.
#The term ordination derives from the idea to ordinate or put things into order. With ordination approaches we are attempting to take a high-dimensional data matrix and explain its patterns with a small number of axis.

#Ordination is close to multiple regression. The key difference is that the response variable is multivariate rather than univariate.

#Clear workspace and load packages
```{r}
rm(list=ls())
graphics.off()

library(vegan)
```

#Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")

#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Transpose count table to that ASVs are columns and samples are rows
```{r}
comm <- t(count_table_clean)

```

#Visually explore our data
```{r}
uni_sp = unique(colnames(comm)) #get our unique species IDs
sp_sum <- apply(comm, 2, sum) #get the sum of each column (asv - how many individuals of the ASV are found across all samples)
site_sum <- apply(comm, 1, sum) #get the sum of each row (sample - how many individuals are there across all ASVs in the sample)
par(mfrow=c(2,2))
hist(sp_sum) #histogram of total ASV abundances
col=colorRamp(c('red', 'orange', 'blue'))
pc_cols = col(length(uni_sp))
plot(sp_sum[order(sp_sum, decreasing=T)], type='o', col='red', lwd=2, xlab='Sp Rank', ylab='Sum Cover') #plot the total ASV abundances by order
hist(site_sum) #histogram of total abundances across all samples
plot(site_sum[order(site_sum, decreasing=T)], type='o', col='red', lwd=2, xlab='Site Rank', ylab='Sum Cover') #plot total abundances per sample by order

#We can see from the upper right hand graph that most species have low sum cover (total covers), so most species are rare

#These plots are not terribly helpful at understanding the variability in the community matrix so let’s get our enviornmental variables ready to use and start to conduct some actual analyses.
```

#Create an explanatory matrix
#We know that each sample has an accompanying set of environmental measurements!
#Let's pull out the environmental data that we need
```{r}
cols_to_keep = c('Chlorophyll', 'Depth', 'DNA_per_g', 'org_matter', 'pH', 'water_content', 'area', 'Glacier')
env = aggregate(metadata[ , cols_to_keep], by = list(metadata$Run), function(x) x[1])
# aggregate does have a side effect of adding a column to the output which 
# is the unique plotID in this case. We just need to move that so it is a 
# row name instead. 
row.names(env) = env[ , 1]
env = env[ , -1]
head(env)

#Let's check to make sure our explanatory matrix has the same sample names as our count table

all.equal(rownames(comm), rownames(env)) #TRUE!

#Now we can rename the samples names to something simpler 
rownames(comm) = 1:nrow(comm)
rownames(env) = 1:nrow(env)
```

#DIRECT OR CONSTRAINED ORDINATION
#REDUNDANCY ANALYSIS
#Here we're going to carry out an RDA which expects a linear response of each species to the environmental variables. RDA is the most direct analog of OLS refression to the multivariate response variable.
```{r}
rda_cryo = rda(comm, env) #this doesn't work because I have NAs in my env data so let's remove them

c1 <- complete.cases(comm)
c2 <- complete.cases(env)
c12 <- c1 & c2

comm_sub <- comm[c12, ]
env_sub <- env[c12, ]

rda_cryo = rda(comm_sub, env_sub)

```


```{r}


```


```{r}


```


```{r}


```
