---
title: "28-02-21-metadata-analysis"
author: "Amy Solman"
date: "28/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script will prepare our QIIME2 output for use in downstream analysis.

#Step One: Clear work space and load packages
```{r}
rm(list=ls())
graphics.off()

library(tidyverse)
library(phyloseq)
library(readr)
library(seqinr)
library(decontam)
library(vegan)
library(magrittr)
library(data.table)
library(ape) #appe::Ntip
```

#Step Two: Load the data from our QIIME2 step +metadata
```{r}
count_table <- read_tsv(file="export/table/table.tsv", skip=1)

taxonomy <- read_tsv(file="export/taxonomy/taxonomy.tsv")

tree <- read_tree("export/exported-tree/tree.nwk")

fasta <- read.fasta(file="export/rep-seqs.fasta/dna-sequences.fasta")

metadata <- read_csv(file="export/SraRunTable-wrangle.tsv")

#Specify the first column in our count table is the row names
count_table <- column_to_rownames(count_table, var = colnames(count_table)[1])
```

#Step Three: Remove unnecessary rows/columns from our metadata
```{r}
#delete unneccessary row and column
metadata <- metadata[-c(1),]
metadata <- metadata[,-c(1)]
metadata <- metadata[!grepl("18S", metadata$Library.Name),] #only include 16S samples
metadata <- metadata[-c(42:80),] #remove reverse reads so we only have one entry per sample

#rename columns that didn't get the right name imported from the metadata
colnames(metadata)[10] <- "Chlorophyll"
colnames(metadata)[11] <- "Collection_date"
colnames(metadata)[16] <- "Depth"
colnames(metadata)[36] <- "pH"
colnames(metadata)[39] <- "Replicate"
colnames(metadata)[47] <- "Temp"

#Im also going to remove some of the text within these data so we can just work with the numbers
#we use the parse_number() function from the readr package
metadata$Chlorophyll <- parse_number(metadata$Chlorophyll)
metadata$org_matter <- parse_number(metadata$org_matter)
metadata$water_content <- parse_number(metadata$water_content)

#add a column with glacier name
Glacier <- NULL
for (i in 1:nrow(metadata)){
  if (startsWith(metadata$Sample.Name[[i]], "C")){
    Glacier <- c(Glacier, "Canada")
  } else {
    Glacier <- c(Glacier, "Taylor")
  }
}

metadata <- cbind(metadata, Glacier)
```

#Step Four: Identify contaminents in the count data
```{r}
# A logical vector showing which entries are 'TRUE' for controls and 'FALSE' for genuine samples
decontam <- startsWith(metadata$Sample.Name, "NT")

# convert count table to dataframe
count_table_df <- as.data.frame(count_table)

# Run isContaminant, while transforming the matrix with t()
contam_df <- isContaminant(t(count_table_df), neg=decontam)

# What do our results look like?
table(contam_df$contaminant) 

# Make a vector holding contaminant IDs
contam_asvs <- row.names(contam_df[contam_df$contaminant == TRUE, ])

# Check the taxonomy of these contaminants
taxonomy[taxonomy$`Feature ID` %in% contam_asvs, ]
```

#Step Five: Remove contaminents from count table,taxonomy table, sequence data, metadata and tree
```{r}
###COUNT TABLE
x <- rownames(count_table) %in% contam_asvs #which rows are contaminants?
count_table_clean <- count_table[!x,] #remove those rows

# and remove blank samples from count table
#what is the index of "TRUE" values in decontam?
true_index <- NULL
for (i in 1:length(decontam)){
  if (decontam[[i]] == "TRUE"){
    true_index <- c(true_index, i)
  }
}
count_table_clean <- count_table_clean[,-c(true_index)]  #remove those samples

###TAXONOMY TABLE
z <- taxonomy$`Feature ID` %in% contam_asvs
taxonomy_clean <- taxonomy[!row.names(taxonomy) %in% z,]#remove contaminating ASVs

###FASTA FILE
#Remove contaminants from dna sequences file
fasta_clean <- fasta
fasta_clean[contam_asvs] <- NULL

###METADATA TABLE
y <- startsWith(metadata$Sample.Name, "NT")
true_index <- NULL
for (i in 1:length(y)){
  if (y[[i]] == "TRUE"){
    true_index <- c(true_index, i)
  }
}
metadata <- metadata[-c(true_index),]  #remove those samples

###PHYLOGENETIC TREE
# This is a phyloseq object so we can use a phyloseq command called prune_taxa
tree_clean <- prune_taxa(count_table_clean$`#OTU ID`, tree)
```

#Step Six: Plot rarefaction curves to see how deeply our samples have been sequenced
```{r}
#Plot the rarefaction curves, color-coding by sample and tranform using t()
#Col= option could also be used to look at different sample types/locations?
pdf("results/rarefaction-plot.pdf") 
rarecurve(t(count_table_clean), step=100, cex=0.5, col=palette(rainbow(nrow(metadata))), lwd=2, ylab="ASVs", label=T)
# Add line to indicate the fewest # of sequences
abline(v=(min(rowSums(t(count_table_clean)))))
dev.off()

```

#Step Seven: Remove singletons
```{r}
# First convert to phyloseq package
count_table_no_contam_phyloseq <- otu_table(count_table_clean, taxa_are_rows = TRUE)

# Then filter using filter taxa
count_table_no_contam_no_singletons_phyloseq <- prune_taxa(taxa_sums(count_table_no_contam_phyloseq) > 1, count_table_no_contam_phyloseq)
```

#Step Eight: Modify taxonomy table 
```{r}
count_table_no_contam_no_singletons <- filter(count_table_clean, rowSums(count_table_clean)>1)
taxonomy_no_contam_no_singletons <- subset(taxonomy_clean, `Feature ID` %in% row.names(count_table_no_contam_no_singletons))

#Now to modify the taxonomic table for my data
taxonomy <- taxonomy_no_contam_no_singletons %>%
  mutate(taxonomy=str_replace_all(string=Taxon, pattern="D_\\d*\\__", replacement="")) %>%
  mutate(taxonomy=str_replace_all(string=taxonomy, pattern=";$", replacement="")) %>%
separate(taxonomy, into=c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep=";") %>%
  select(-Taxon, -Confidence) %>%
  column_to_rownames(var='Feature ID')

rownames(taxonomy) <- rownames(count_table_no_contam_no_singletons)
```

###Step Nine: Perform Hellinger transformation
```{r}
count_table_hellinger <- decostand(t(count_table_no_contam_no_singletons), "hellinger")
#Convert to tibble
count_table_hellinger <- as_tibble(t(count_table_hellinger), rownames = NA)

# Compare data
head(count_table_no_contam_no_singletons) #untransformed
head(count_table_hellinger)
```

#Step Ten: Convert count_table, taxonomy table and metadata table to phyloseq objects
```{r}
ASV = otu_table(data.frame(count_table_hellinger), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
```

#Step Eleven: Make into one phyloseq object
```{r}
my_phyloseq <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Twelve: remove abiguous annotations and low abundance phyla
```{r}
my_phyloseq <- subset_taxa(my_phyloseq, !is.na(Kingdom) & !Kingdom %in% c("Unassigned", "Eukaryota"))

#get names of phyla with abundances of 2 or less
x <- table(tax_table(my_phyloseq)[, "Phylum"], exclude = NULL)[table(tax_table(my_phyloseq)[, "Phylum"], exclude = NULL) < 3]

#remove phyla
my_phyloseq <- subset_taxa(my_phyloseq, !is.na(Phylum) & !Phylum %in% rownames(x))

```

#Step Thirteen: Re-root tree because the root may have been removed through pruning. Use this function > (https://john-quensen.com/r/unifrac-and-tree-roots/) < picks longest branch
```{r}
# First define the function from link above to find furthest outgroup
pick_new_outgroup <- function(tree.unrooted){
  #tablif parts of the tree that we need
  treeDT <- 
    cbind(
      data.table(tree.unrooted$edge),
      data.table(length = tree.unrooted$edge.length)
    )[1:Ntip(tree.unrooted)] %>%
    cbind(data.table(id = tree.unrooted$tip.label))
  #Take out the longest terminal branch as outgroup
  new.outgroup <- treeDT[which.max(length)]$id
  return(new.outgroup)
}

# Run on my_phyloseq tree
my_tree <- phy_tree(tree_clean)
new_outgroup <- pick_new_outgroup(my_tree)

# Re-root tree
new_tree <- ape::root(my_tree, outgroup=new_outgroup, resolve.root=TRUE)

# Convert to dichotomy tree
new_tree2 <- ape::multi2di(new_tree)
phy_tree(my_phyloseq) <- new_tree2
phy_tree(my_phyloseq)
```

#Save r data object and phyloseq object
```{r}
#r data object
save(metadata, fasta_clean, count_table_hellinger, count_table_clean, taxonomy, tree_clean, file = "results/cleaned_files/cleanedfiles_3.RData")

#Now I'm going to export my phyloseq object
saveRDS(my_phyloseq, "results/my-phyloseq-object.rds")
```
