---
title: "MVA Linear Models"
author: "Amy Solman"
date: "21/04/2021"
output: html_document
---

#This script will use linear models to explore the drivers of community assembly
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
#install.packages("repr")
library(repr) #for plotting our pairs of data
library("picante")
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Which of the environmental variables: hole area, hole depth, pH, DNA per gram, chlorophyll, organic matter content and water content, explains variation in community composition?

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(S) #Here we use the species count we got earlier

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
```

#Step Six: Plot richness with environmental data 
```{r}
#let's start with basic plots of richness with our evironmental variables
rich.chlo <- plot(my.data.sub$Chlorophyll, my.data.sub$Richness)
rich.depth <- plot(my.data.sub$Depth, my.data.sub$Richness)
rich.dna <- plot(my.data.sub$DNA_per_g, my.data.sub$Richness)
rich.org <- plot(my.data.sub$org_matter, my.data.sub$Richness)
rich.ph <- plot(my.data.sub$pH, my.data.sub$Richness)
rich.water <- plot(my.data.sub$water_content, my.data.sub$Richness)
rich.area <- plot(my.data.sub$area, my.data.sub$Richness)

#indices of other alpha metrics
alpha.met <- c(8, 10, 11, 12, 13)
pairs(my.data.sub[, -c(alpha.met)]) #so we can do a pairs plot with just richness and environmental data

```

#Step Seven: Look for correlations between richness and environmental data
```{r}
#cor looks for correlations between pairs of variables using Pearson's correlation coefficient
cor(my.data.sub[, -c(alpha.met)], use="pairwise")

#or we can use cor.test to compair a single pair of variables - this also carries out a t.test to see if the correlation is significant
cor.test(my.data.sub$Richness, my.data.sub$Chlorophyll, use="pairwise") #p >0.05
cor.test(my.data.sub$Richness, my.data.sub$Depth, use="pairwise") #p < 0.05
cor.test(my.data.sub$Richness, my.data.sub$DNA_per_g, use="pairwise") #p > 0.05
cor.test(my.data.sub$Richness, my.data.sub$org_matter, use="pairwise") #p < 0.05
cor.test(my.data.sub$Richness, my.data.sub$pH, use="pairwise") #p > 0.05
cor.test(my.data.sub$Richness, my.data.sub$water_content, use="pairwise") #p > 0.05
cor.test(my.data.sub$Richness, my.data.sub$area, use="pairwise") #p < 0.05

#only cryoconite hole depth, area and organic matter content are significant correlated with ASV richness

#WARNING The above correlation coefficient calculation assumes a straight line relationship
```

#Step Eight: Re-run above correlation analysis with log transformed data
```{r}
my.data.sub$logRichness <- log(my.data.sub$Richness)
my.data.sub$logChlorophyll <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA_per_g <- log(my.data.sub$DNA_per_g)
my.data.sub$logorg_matter <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logwater_content <- log(my.data.sub$water_content)
my.data.sub$logarea <- log(my.data.sub$area)

my.log.alpha <- c(14, 15, 16, 17, 18, 19, 20, 21)
pairs(my.data.sub[, (my.log.alpha)])

cor(my.data.sub[, (my.log.alpha)], use="pairwise")

```