---
title: "Network-Analysis"
author: "Amy Solman"
date: "21/04/2021"
output: html_document
---

#This code is developed from the previously published paper by Ju F, Xia Y, Guo F, Wang ZP, Zhang T (2014) 
#Taxonomic relatedness shapes bacterial assembly in activated sludge of globally distributed wastewater treatment plants.
#Environmental Microbiology. 16(8):2421-2432

#This script performs network analysis on my count table data and produces positive, negative and total network graphs in gml format, global and node tolopogical feature tables, as well as new count tables comprising only ASVs from positive or negative networks.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step 1: Prep your environment and load data
```{r}
#Clear Workspace
rm(list=ls())

#Load packages
library(vegan)
library(igraph)
library(Hmisc)

#Load Data
load("results/cleaned_files/cleanedfiles_3.RData")
```

#Step 2: Get your count table ready for network analysis
```{r}
#Turn your count table into a matrix
my.matrix <- as.matrix(count_table_clean)

#Turn your matrix count table into a presence absence table
my.working.matrix<-my.matrix 
my.working.matrix[my.working.matrix>0]<-1 #for every value in the table that is more than 0, make it a 1

#We want to only keep taxa in at least 20% of samples (I don't know why but this was done in the paper)
general.cut.off <- round(ncol(count_table_clean)/100*20)

#Keep only the ASVs that are in X number of samples 
taxa.generalist<-my.matrix[which(rowSums(my.working.matrix)>=general.cut.off),]
my.matrix<-taxa.generalist
```

#Step 3: Next we perform correlation analysis based on spearman's co-efficient
```{r}
#compute a matrix of spearman's rho rank correlation coefficients for all possible pairs of columns (ASVs)
matrix.dist<-rcorr(t(my.matrix),type="spearman")#this can also be done with pearsons - which is best?

matrix.cor<-matrix.dist$r #from the matrix extract r = the correlation coefficients
matrix.cor.p<-matrix.dist$P #from the matrix extract p = the p-values corresponding to the significance levels of correlations

#here we adjust the P-values for multiple comparisons using Benjamini-Hochberg standard false discovery rate correction ("FDR-BH")
matrix.cor.p <- p.adjust(matrix.cor.p, method="BH")
```

#Step 4: Define coefficient (cor.cutoff) and p-value cutoffs
```{r}
#define your coefficient cut offs and your p-value cut offs
cor.cutoff = 0.6 #only keep robust correlations =>0.6
p.cutoff = 0.01 #only keep significant p-values =>0.01
```

#Step 5: Now we look at our positive co-occurences at given coefficient (cor.cutoff) and p-value cutoffs
```{r}
matrix.cor1<-matrix.cor #give the correlation coefficients from our matrix to a new variable
matrix.cor1.p<-matrix.cor.p #give the o-values of our correlations to a new variable

matrix.cor1[which(matrix.cor1 <= cor.cutoff)]=0 #for those correlation coefficients below or equal to our cut off, give them the value of 0
matrix.cor1[which(matrix.cor1.p>p.cutoff)]=0 #for those correlation coefficients with corresponding p-values over the cut off, give them the value o

# delete those rows and columns with sum = 0
matrix.cor1<-matrix.cor1[which(rowSums(matrix.cor1)!=0),] #keep only rows with values that are not 0
#in the original code the above line had !=1 instead of !=0 and I thought this was a typo so I changed it
matrix.cor1<-matrix.cor1[,which(colSums(matrix.cor1)!=0)] #keep only columns with values that are not 0
```

#Step 6: Now we look at our negative co-occurences at given coefficient (cor.cutoff) and p-value cutoffs
```{r}
matrix.cor2<-matrix.cor #give the correlation coefficients from our matrix to a new variable
matrix.cor2.p<-matrix.cor.p #give the o-values of our correlations to a new variable

matrix.cor2[which(matrix.cor2 > (-cor.cutoff))]=0 #for those correlation coefficients above the negative value of our cut off, give them the value of 0
matrix.cor2[which(matrix.cor2.p>p.cutoff)]=0 #for those correlation coefficients with corresponding p-values over the cut off, give them the value 0

#delete those rows and columns with sum = 0
matrix.cor2<-matrix.cor2[which(rowSums(matrix.cor2)!=0),] #keep only rows with values that are not 0
#in the original code the above line had !=1 instead of !=0 and I thought this was a typo so I changed it
matrix.cor2<-matrix.cor2[,which(colSums(matrix.cor2)!=0)] #keep only columns with values that are not 0
```

#Step 7: Now we look at our positive and negative co-occurences at given coefficient (cor.cutoff) and p-value cutoffs
```{r}
matrix.cor3<-matrix.cor #give the correlation coefficients from our matrix to a new variable
matrix.cor3.p<-matrix.cor.p  #give the o-values of our correlations to a new variable

matrix.cor3[which(matrix.cor3>=(-cor.cutoff) & matrix.cor3 <= cor.cutoff)]=0 #for those correlation coefficients above and equal to the negative value of our cut off AND below or equal to the value of out cut off, give them the value 0
matrix.cor3[which(matrix.cor3.p>p.cutoff)]=0 #for those correlation coefficients with corresponding p-values over the cut off, give them the value 0
  
# delete those rows and columns with sum = 0
matrix.cor3<-matrix.cor3[which(rowSums(matrix.cor3)!=0),] #keep only rows with values that are not 0
#in the original code the above line had !=1 instead of !=0 and I thought this was a typo so I changed it
matrix.cor3<-matrix.cor3[,which(colSums(matrix.cor3)!=0)] #keep only columns with values that are not 0
```

#Step 8: Create graphs for positive, negative and total co-occurrences 
```{r}
#For positive co-occurrences within out cut off thresholds create graphs from adjacency matrices using the igraph package
g1<-graph.adjacency(matrix.cor1,weight=T,mode="undirected")
g1<-simplify(g1) #use the simplify function to removes the loop and/or multiple edges from a graph
V(g1)$label <- V(g1)$name #Create a vertex sequence (vs) containing all vertices of a graph (V) and get a list of ASV IDs and give those values to the graph vertices variable 'label' 
V(g1)$degree <- degree(g1) #Get the degree of a vertex (its most basic structural property, the number of its adjacent edges) and give those values to the graph vertices variable 'degree'
  
#For negative co-occurrences within out cut off thresholds create graphs from adjacency matrices using the igraph package
g2<-graph.adjacency(matrix.cor2,weight=T,mode="undirected")
g2<-simplify(g2)
V(g2)$label <- V(g2)$name
V(g2)$degree <- degree(g2)
  
#For positive and negative co-occurrences within out cut off thresholds create graphs from adjacency matrices using the igraph package
g3<-graph.adjacency(matrix.cor3,weight=T,mode="undirected")
g3<-simplify(g3)
V(g3)$label <- V(g3)$name
V(g3)$degree <- degree(g3)
```

#Step 9: Put the output into a results 
```{r}
result<-list() #create an empty list
result$matrix.cor<-matrix.cor #put the total correlation coefficients from our matrix to the list (the ones that haven't been reduced by our cutoffs)
result$matrix.cor.p<-matrix.cor.p #put the total p-values of our correlation coefficients from our matrix to the list (the ones that haven't been reduced by our cutoffs)
  
#Add the cut-off reduced positive occurrence correlation coefficients
result$matrix.cor1<-matrix.cor1
#Add the cut-off reduced positive occurrence correlation coefficients graph
result$graph1<-g1
  
#Add the cut-off reduced negative occurrence correlation coefficients
result$matrix.cor2<-matrix.cor2
#Add the cut-off reduced negative occurrence correlation coefficients graph
result$graph2<-g2
  
#Add the cut-off reduced positive and negative occurrence correlation coefficients
result$matrix.cor3<-matrix.cor3
#Add the cut-off reduced positive and negative occurrence correlation coefficients graph
result$graph3<-g3
```

#Step 10: Create gml network files for positive, negative and all associations
```{r}
write.graph(result$graph1,'../Outputs/Sommers2020/network-analysis/positive-network.gml',format='gml')
write.graph(result$graph2,'../Outputs/Sommers2020/network-analysis/negative-network.gml',format='gml')
write.graph(result$graph3,'../Outputs/Sommers2020/network-analysis/positive-negative-network.gml',format='gml') 
```

#Step 11: Calculate the global network topological properties for positive network
```{r}
g<-result$graph1   ###positive network

#This function tries to find densely connected subgraphs, also called communities in a graph via random walks. The idea is that short random walks tend to stay in the same community.
c <- cluster_walktrap(g) 

#MODULARITY
#Find the modularity of the subgraphs (or communities) in the positive network graph
modularity(c) #0.3581867

#Find the modularity of the positive network, with communities
md <- modularity(g, membership(c), weights = NULL) 

#CLUSTERING COEFFICIENT
#Transitivity measures the probability that the adjacent vertices of a vertex are connected. This is sometimes also called the clustering coefficient.
cc <- transitivity(g, vids = NULL,
                   weights = NULL)

#AVERAGE SHORTEST PATHS LENGTH
spl <- average.path.length(g, directed=FALSE, unconnected=TRUE)

#GRAPH DENSITY 
#The density of a graph is the ratio of the number of edges and the number of possible edges.
gd  <- graph.density(g, loops=FALSE)

#GRAPH DIAMETER
#The diameter of a graph is the length of the longest geodesic.
nd  <- diameter(g, directed = FALSE, unconnected = TRUE, weights = NA)

#NODE DEGREE
node.degree <- degree(g, v = V(g), mode="all")

#MEAN NODE DEGREE
ad  <- mean(node.degree)

#NUMBER OF EDGES
e <- ecount(g)

#NUMBER OF NODES
v <- vcount(g)

#PUT ALL TOPOLOGICAL FEATURES TOGETHER INTO A DATAFRAME
global.topology <- data.frame(e,v,cc,spl,md,gd,nd,ad)

#SAVE DATA FRAME
write.csv(global.topology, file="../Outputs/Sommers2020/network-analysis/positive-network-global-topology.csv")

```

#Step 12: Find the node topological features for the positive network
```{r}
#BETWEENNESS CENTRALITY
#the number of times a node is crossed by each of the shortest routes between nodes (between any two nodes there is a shortest distance to travel. How many shortest distances is the node of interest within?)
betweenness.centrality <- betweenness(g, v=V(g), 
                                      directed = FALSE, weights = NA,
                                      nobigint = TRUE, normalized = FALSE)

#CLOSENESS CENTRALITY 
#average length of the shortest path between the node and all other nodes in the graph (how long does it take me on average to reach another node)
closeness.centrality <- closeness(g, vids = V(g),
                                  weights = NA, normalized = FALSE)

#NODE TRANSITIVITY
#Transitivity measures the probability that the adjacent nodes of a node are connected. This is sometimes also called the clustering coefficient
node.transitivity <- transitivity(g, type = c("local"), vids = NULL,
                                  weights = NA)

#PUT ALL RESULTS TOGETHER INTO A DATAFRAME
node.topology <- data.frame(node.degree, betweenness.centrality, closeness.centrality, node.transitivity)

#SAVE DATAFRAME
write.csv(node.topology, file="../Outputs/Sommers2020/network-analysis/positive-network-node-topology.csv")
```

#Step 13: Calculate the global network topological properties for negative network
```{r}
g<-result$graph2   ###negative network

#This function tries to find densely connected subgraphs, also called communities in a graph via random walks. The idea is that short random walks tend to stay in the same community.
c <- cluster_walktrap(g) 

#MODULARITY
#Find the modularity of the subgraphs (or communities) in the negative network graph
modularity(c)

#Find the modularity of the negative network, with communities
md <- modularity(g, membership(c), weights = NULL) 

#CLUSTERING COEFFICIENT
#Transitivity measures the probability that the adjacent vertices of a vertex are connected. This is sometimes also called the clustering coefficient.
cc <- transitivity(g, vids = NULL,
                   weights = NULL)

#AVERAGE SHORTEST PATHS LENGTH
spl <- average.path.length(g, directed=FALSE, unconnected=TRUE)

#GRAPH DENSITY 
#The density of a graph is the ratio of the number of edges and the number of possible edges.
gd  <- graph.density(g, loops=FALSE)

#GRAPH DIAMETER
#The diameter of a graph is the length of the longest geodesic.
nd  <- diameter(g, directed = FALSE, unconnected = TRUE, weights = NA)

#NODE DEGREE
node.degree <- degree(g, v = V(g), mode="all")

#MEAN NODE DEGREE
ad  <- mean(node.degree)

#NUMBER OF EDGES
e <- ecount(g)

#NUMBER OF NODES
v <- vcount(g)

#PUT ALL TOPOLOGICAL FEATURES TOGETHER INTO A DATAFRAME
global.topology <- data.frame(e,v,cc,spl,md,gd,nd,ad)

#SAVE DATA FRAME
write.csv(global.topology, file="../Outputs/Sommers2020/network-analysis/negative-network-global-topology.csv")

```

#Step 14: Find the node topological features for the negative network
```{r}
#BETWEENNESS CENTRALITY
#the number of times a node is crossed by each of the shortest routes between nodes (between any two nodes there is a shortest distance to travel. How many shortest distances is the node of interest within?)
betweenness.centrality <- betweenness(g, v=V(g), 
                                      directed = FALSE, weights = NA,
                                      nobigint = TRUE, normalized = FALSE)

#CLOSENESS CENTRALITY 
#average length of the shortest path between the node and all other nodes in the graph (how long does it take me on average to reach another node)
closeness.centrality <- closeness(g, vids = V(g),
                                  weights = NA, normalized = FALSE)

#NODE TRANSITIVITY
#Transitivity measures the probability that the adjacent nodes of a node are connected. This is sometimes also called the clustering coefficient
node.transitivity <- transitivity(g, type = c("local"), vids = NULL,
                                  weights = NA)

#PUT ALL RESULTS TOGETHER INTO A DATAFRAME
node.topology <- data.frame(node.degree, betweenness.centrality, closeness.centrality, node.transitivity)

#SAVE DATAFRAME
write.csv(node.topology, file="../Outputs/Sommers2020/network-analysis/negative-network-node-topology.csv")
```


#Step 15: Create abundance (count) tables with only the ASVs found in our positive and negative networks
```{r}
my.list1 <- row.names(result$matrix.cor1) #add ASV IDs from positive network to list
my.list2 <- row.names(result$matrix.cor2) #add ASV IDs from negative network to list

logical1 <- row.names(my.matrix)  %in% my.list1 #which row names in my count table (my.matrix) are also found in the cut-off reduced positive network
logical2 <- row.names(my.matrix)  %in% my.list2 #which row names in my count table (my.matrix) are also found in the cut-off reduced negative network

tab.subset1 <- subset(my.matrix,logical1) #subset my count table by the TRUE/FALSE list - we we get a count table with only the ASVs found in the positive network
tab.subset2 <- subset(my.matrix,logical2) #subset my count table by the TRUE/FALSE list - we we get a count table with only the ASVs found in the positive network

#Save the count table with ASVs found in the positive network
write.table(tab.subset1,'../Outputs/Sommers2020/network-analysis/count-table-positive-network.txt',sep="\t")

#Save the count tanle with ASVs found in the negative network
write.table(tab.subset2,'../Outputs/Sommers2020/network-analysis/count-table-negative-network.txt',sep="\t")
```
