---
title: "MVA Analysis - Alpha"
author: "Amy Solman"
date: "21/04/2021"
output: html_document
---

#Script developed from that of BArnett et al., (2020) code available at https://github.com/seb369/landuse_comm_assembly/blob/master/Diversity_measures.md 
#This script will calculate alpha diversities 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

library(dplyr)
library(tibble)
library(vegan)
library(ggpubr)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity of each sample by OTU richness
```{r}
#get the metadata table
my.metadata <- data.frame(sample_data(my.ps))

#get the rarefied ASV table
ASV.table.bin <- otu_table(my.ps.rar)

#for any ASV count that does NOT equal 0, make it 1 (so make it into a presence/absence table)
ASV.table.bin[ASV.table.bin != 0] = 1

#make a new dataframe called ASV count with a column called "count"
#which is the column sums from the presence/absence table
#Give the columns the sample names
#join the metadata by sample name
#arrange by sample name
ASV.count = data.frame("num" = colSums(ASV.table.bin)) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps)), by="Run") %>%
  arrange(Run) %>%
  mutate(sample_ID=factor(Run, levels = Run))
```

#Step Five: #Compare ASV richness (alpha diversity) between glaciers
```{r}
#First get our summary statistics
summary.stats <- group_by(ASV.count, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(num, na.rm = TRUE),
    sd = sd(num, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot <- ggboxplot(ASV.count, x = "Glacier", y = "num",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="ASV Num", xlab = "Glacier")

my.plot

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(ASV.count, shapiro.test(num[Glacier == "Canada"])) #p=0.5263 (p >0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.
with(ASV.count, shapiro.test(num[Glacier == "Taylor"])) #p=0.9692 (p >0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES?
#Use F-test to test for homogeneity in variances
f.test <- var.test(num ~ Glacier, data = ASV.count)
f.test #p value = 0.7446 which is greater than 0.05 so there is no significant difference in the variances of the two datasets

#ALL THREE ASSUMPTIONS ARE MET FOR THE PARAMTERIC TWO INDEPENDENT SAMPLES TEST

#COMPARE THE MEANS OF TWO INDEPENDENT GROUPS
#UNPAIRED TWO SAMPLES T-TEST (PARAMETERIC)
my.t.test <- t.test(num ~ Glacier, data = ASV.count, var.equal = TRUE)
my.t.test #p = 0.0518 so p > 0.05, there is no significant difference between the ASV richnesses of the two glaciers

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
my.wilcoxon.test <- wilcox.test(num ~ Glacier, data = ASV.count, exact = FALSE)
my.wilcoxon.test # p = 0.04479 so p < 0.05, there is a significant difference between the ASV richness of the two glaciers.

#Even though the non-parametric test gave me a significant result, where the parameteric test did not, I should use the parametric test because it is more efficient.

####USE THIS WHEN COMPARING THREE OR MORE INDEPENDENT GROUPS WITH NON-PARAMETRIC DATA###
# writeLines("Kruskal-Wallis test:")
# k.test = kruskal.test(count ~ Glacier, data = ASV.count)
# print(k.test)

#The below Dunn's test can be used when comparing three or more groups
#That is not useful now.

# if (k.test$p.value <= 0.05){
#   writeLines(c("", "There is a significant effect of sample type (Glacier) on ASV number. Now running Dunn test for pairwise comparisons:"))
#   d.test = dunnTest(count ~ Glacier, data = ASV.count, method="bh")
#   print(d.test)
# } else {
#   writeLines(c("", "We did not find a significant effect of sample type (Glacier) on ASV number."))
# }
```
#Step Six: Plot the data
```{r}
ggplot(data=ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="Number of unique ASVs", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))
```

#Step Seven: Calculate alpha diversity of each sample by Shannon Diversity index
```{r}
#make a new dataframe with shannon as the shannon diversity index of each sample
ASV.table.bin = t(otu_table(my.ps.rar))
shannon.df = data.frame(shannon=diversity(ASV.table.bin, index="shannon")) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps)), by="Run") %>%
  arrange(Glacier)
```

#Step Eight: #Compare Shannon index (alpha diversity) between glaciers
```{r}
#First get our summary statistics
summary.stats.shan <- group_by(shannon.df, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(shannon, na.rm = TRUE),
    sd = sd(shannon, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot.shan <- ggboxplot(shannon.df, x = "Glacier", y = "shannon",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Shannon Index", xlab = "Glacier")

my.plot.shan

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(shannon.df, shapiro.test(shannon[Glacier == "Canada"])) #p=0.03166 (p < 0.05) so the distribution of the data are significantly different from normal distribution so must use a NON-PARAMETRIC test.
with(shannon.df, shapiro.test(shannon[Glacier == "Taylor"])) #p=0.6421 (p < 0.05) so the distribution of the data are not significantly different from normal distribution, but because the canada data is NOT normally distributed we'll use a non-parametric test

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES? <- Ignore because Canada data is non-parametric

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
my.wilcoxon.test.shan <- wilcox.test(shannon ~ Glacier, data = shannon.df, exact = FALSE)
my.wilcoxon.test.shan # p = 0.5808 so p > 0.05, there is no significant difference between the shannon diversity indexes of the two glaciers.
```
#Step Nine: Plot the data
```{r}
ggplot(data=shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))
```

#Step 10: Plot ASV richness and shannon diversity together
```{r}
#Glacier shapes
Glacier.shapes = c("Canada" = 15, "Taylor" = 16)

rich.plot = ggplot(data=ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(color=Glacier, shape= Glacier), width=0.2, height=0)+
  scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="Num unique ASVs", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

shan.plot = ggplot(data=shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour=Glacier, shape= Glacier), width=0.2, height=0)+
   scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

fonttheme = theme(legend.text = element_text(size=12),
                  legend.title = element_text(size=12),
                  axis.text = element_text(size=12),
                  axis.title = element_text(size=14))

#Merge the plots together

alpha.plot = cowplot::plot_grid(rich.plot + fonttheme, shan.plot + fonttheme, ncol=2, rel_widths = c(1, 1), labels = c("A", "B"))
alpha.plot
ggsave("../Outputs/Sommers2020/graphs/alpha-diversity.pdf", width=8, height=4)
```

