---
title: "MVA Analysis - Alpha"
author: "Amy Solman"
date: "21/04/2021"
output: html_document
---

#Script developed from that of BArnett et al., (2020) code available at https://github.com/seb369/landuse_comm_assembly/blob/master/Diversity_measures.md 
#This script will calculate alpha diversities 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

library(vegan)
library(FSA)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Calculate alpha diversity of each sample by OTU richness
```{r}
my.metadata <- data.frame(sample_data())
```

#Calculate alpha diversity of each sample by Shannon Diversity index

#Compare alpha diversity metrics between glaciers ysing Kruskal Wallis

#If there is significant vairation between habitats, run post-hoc pairwise comparisons using the Dunn test (FSA package)