---
title: "ordination-1"
author: "Amy Solman"
date: "01/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
#clean up the workspace
rm(list=ls())
graphics.off()
library(tidyverse)
library(phyloseq)
library(ape)
```

#Step Two: Load R data object and phyloseq object
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")

my_physeq <- readRDS("results/my-phyloseq-object.rds")
```

#Step Three: Make a bar chart of the agglomerated phyla
```{r}
glom = tax_glom(my_physeq, "Phylum")
# and plot
pdf("results/glom.pdf") 
plot_bar(glom, x = "Sample", fill = "Phylum")
dev.off()
```
#Step Four: Calculate Bray-Curtis dissimilarity matrix

```{r}
bray_distance <- distance(my_physeq, method = "bray")
bray_distance
```

#Step Five: Carry out bray-curtis PCoA analysis and plot, scaled by the eigenvalues
```{r}
#Bray-Curtis
my_pcoa_bray <- ordinate(my_physeq, method = "PCoA", distance = "bray")

my_eigenvalues_bray <- my_pcoa_bray$values$Eigenvalues

pdf("results/PCoA-bray.pdf") 
plot_ordination(my_physeq, my_pcoa_bray, color = "Glacier") +
  geom_point(size = 3)+
  coord_fixed(sqrt(my_eigenvalues_bray[2]/my_eigenvalues_bray[1]))
dev.off()

```

#Step Six: Carry out unweighted and weighted unifrac PCoA analysis and plot, scaled by the eigenvalues
```{r}
#Unweighted UniFrac
my_pcoa_uni <- ordinate(my_physeq, method = "PCoA", distance = "unifrac")
my_eigenvalues_uni <- my_pcoa_uni$values$Eigenvalues

pdf("results/PCoA-unifrac.pdf") 
plot_ordination(my_physeq, my_pcoa_uni, color = "Glacier") +
  geom_point(size = 3)+
  coord_fixed(sqrt(my_eigenvalues_uni[2]/my_eigenvalues_uni[1]))
dev.off()

#Weighted Unifrac
my_pcoa_wuni <- ordinate(my_physeq, method = "PCoA", distance = "wunifrac")
my_eigenvalues_wuni <- my_pcoa_wuni$values$Eigenvalues

pdf("results/PCoA-wunifrac.pdf") 
plot_ordination(my_physeq, my_pcoa_wuni, color = "Glacier") +
  geom_point(size = 3)+
  coord_fixed(sqrt(my_eigenvalues_wuni[2]/my_eigenvalues_wuni[1]))
dev.off()
```

#Step Seven: Create an NMDS plot 
```{r}
my_nmds <- ordinate(my_physeq, method = "NMDS", distance = "bray")

pdf("results/NMDS-bray.pdf")
plot_ordination(my_physeq, my_nmds, color = "Glacier") +
  geom_point(size = 3)
dev.off()
```