---
title: "3b-ordination-"
author: "Amy Solman"
date: "01/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
#clean up the workspace
rm(list=ls())
graphics.off()
library(compositions)
library(tidyverse)
library(vegan)
library(ape)
library(cowplot)
library(plotly)
library(RColorBrewer)
```

#Step Two: Load data.
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
```

#Step Three: Look for evidence of compositionality.
#By multiplying ASV table by itself we can look for evidence of covariance.
#We then find the determinant of the covariance matrix and if the answer is 0 we have evidence of compositionality in the data
```{r}
covariance_matrix <- as.matrix(count_table_clean) %*% t(count_table_clean)

cov_determinant <- det(covariance_matrix)
cov_determinant
```

#Because out data are compositional we then need to transform them for downstream analysis
#Step Four: Perform a Log-Ratio trasnformation
#This will normalise the ASV data within each sample
```{r}
#irl - isometric log ratio
# can also use clr for centred log ratio - centered around geometric average 
count_table_LR <- data.frame(compositions::ilr(t(count_table_clean)))

#compare the spread of 
hist(count_table_clean$SRR12805327) #here we see that a small number of ASVs had high frequency, while a large number had very low frequency
hist(count_table_LR$V1) #this normalises the data
```

#Step Five: Look at the difference the log-ratio transformation has made to the covariance of the data
```{r}
transformed_det <- det(as.matrix(count_table_LR) %*% t(count_table_LR))
cov_determinant #determinant from untransformed count data = 0 
transformed_det #determinant from transformed count data = !0
# Because none of the axis are equal to 0 the determinant is !0. The axes for PCoA explain variance.
```

#Step Six: Perform Principal Component Analysis on Log-Ratio Transformed Data
```{r}
my_pca <- prcomp(count_table_LR)
```

#Step Seven: Create visual representation of PCA
```{r}
my_pca_variances <- as.data.frame(my_pca$sdev^2/sum(my_pca$sdev^2)) %>% #Extract axes
  # Format to plot
  select(PercVar = 'my_pca$sdev^2/sum(my_pca$sdev^2)') %>% 
  rownames_to_column(var = "PCaxis") %>% 
  data.frame
```

#Step Eight: Plot the visual representation
```{r}
pdf("results/pca-screenplot-one.pdf")
ggplot(my_pca_variances, aes(x = as.numeric(PCaxis), y = PercVar)) + 
  geom_bar(stat = "identity", fill = "grey", color = "black") +
  theme_minimal() +
  theme(axis.title = element_text(color = "black", face = "bold", size = 10),
        axis.text.y = element_text(color = "black", face = "bold"),
        axis.text.x = element_blank()) +
  labs(x = "PC axis", y = "% Variance", title = "Log-Ratio PCA Screeplot")
dev.off()
```

#Step Nine: Visualise the PCA
```{r}
my_pca_frame <- data.frame(my_pca$x) %>% 
  rownames_to_column(var = "Glacier") %>% 
  separate(Glacier, into = c("Glacier", "excess"), sep = "_") %>% 
  select(-excess) %>% 
  data.frame


my_pca_frame$Glacier <- metadata$Glacier #here we'll change the sample names to the glacier names
head(my_pca_frame)

```

#Step Ten: Plot the PCA visualisation
```{r}
pdf("results/PCA-Log-Ratio.pdf")
ggplot(my_pca_frame) +
  geom_point(aes(x = PC1, y = PC2, fill = Glacier), size = 4, shape = 21, color = "black") +
  ylab(paste0('PC2 ', round(my_pca_variances[2,2]*100,2),'%')) + #Extract y axis value from variance
  xlab(paste0('PC1 ', round(my_pca_variances[1,2]*100,2),'%')) + #Extract x axis value from variance
  scale_fill_brewer(palette = 'Dark2', name = 'Glacier') +
  ggtitle('Log-Ratio PCA Ordination') +
  coord_fixed(ratio = 1) +
  theme_bw()
dev.off()
```

#Step Eleven: Compute Jaccard distance metric
```{r}
jaccard <- vegdist(t(count_table_clean), method="jaccard")
```

#Step Twelve: Perform PCoA with jaccard distance metric
```{r}
my_jaccard_pcoa <- pcoa(jaccard)
```

#Step Thirteen: Extract the variances from the jaccard PCoA
```{r}
my_jaccard_variances <- data.frame(my_jaccard_pcoa$values$Relative_eig) %>%
  select(PercVar='my_jaccard_pcoa.values.Relative_eig') %>%
  rownames_to_column(var="PCaxis") %>%
  data.frame
```

#Step Fourteen: Use jaccard variances to plot screenplot 
```{r}
pdf("results/pcoa-jaccard.pdf")
ggplot(my_jaccard_variances, aes(x=as.numeric(PCaxis), y=PercVar))+
  geom_bar(stat="identity", fill="grey", color="black")+
  theme_minimal()+
  theme(axis.title=element_text(color="black", face="bold", size=10),
        axis.text.y=element_text(color="black", face="bold"),
        axis.text.x=element_blank())+
  labs(x="PC axis", y= "% Variance", title = "Log-Ratio PCoA Screenplot")
dev.off()
```

#Step Fifteen: Plot this jaccard PCoA in 3D with plotly, firstly by putting jaccard pcoa vectors into a dataframe
```{r}
my_jaccard_pcoa_df <- data.frame(my_jaccard_pcoa$vectors) %>%
 rownames_to_column(var = "Glacier") %>% 
  separate(Glacier, into = c("Glacier", "excess"), sep = "_") %>% 
  select(-excess) %>% 
  data.frame

my_jaccard_pcoa_df$Glacier <- metadata$Glacier #here we'll change the sample names to the glacier names
head(my_jaccard_pcoa_df)
```

#Step Sixteen: Extract eigenvalues from jaccard pcoa and plot
```{r}
jaccard_eigenvalues <- round(my_jaccard_variances[,2], digits=4)*100

#Plotly - 3-D
plot_ly(my_jaccard_pcoa_df, type='scatter3d', mode='markers',
        x=~Axis.2, y=~Axis.3,z=~Axis.1,colors=~brewer.pal(6,'Set1'), color=~Glacier) %>%
  layout(font=list(size=18),
         title='PCoA Jaccard Distance',
         scene=list(xaxis=list(title=paste0('Co 2 ', jaccard_eigenvalues[2], '%'),
                               showticklabels=FALSE, zerolinecolor='black'),
                               yaxis=list(title=paste0('Co 3 ', jaccard_eigenvalues[3], '%'),
                                          showticklabels=FALSE,zerolinecolor='black'),
                    zaxis=list(title=paste0('Co 1 ', jaccard_eigenvalues[1], '%'),
                               showticklabels=FALSE,zerolinecolor='black')))
```

#Step Seventeen: Estimate Euclidean distances using irl transformed count data
```{r}
my_euclidean_dist <- dist(count_table_LR, method = "euclidean")
```

#Step Eighteen: Perform PCoA with euclidean distances
```{r}
my_euclidean_pcoa <- ape::pcoa(my_euclidean_dist)
```

#Step Nineteen: Extract the variances from the euclidean PCoA
```{r}
my_euclidean_variances <- data.frame(my_euclidean_pcoa$values$Relative_eig) %>%
  select(PercVar = 'my_euclidean_pcoa.values.Relative_eig') %>%
  rownames_to_column(var="PCaxis") %>%
  data.frame
```

#Step Twenty: Use euclidean variances to plot screenplot 
```{r}
pdf("results/pcoa-euclidean.pdf")
ggplot(my_euclidean_variances, aes(x=as.numeric(PCaxis), y=PercVar))+
  geom_bar(stat="identity", fill="grey", color="black")+
  theme_minimal()+
  theme(axis.title=element_text(color="black", face="bold", size=10),
        axis.text.y=element_text(color="black", face="bold"),
        axis.text.x=element_blank())+
  labs(z="PC axis", y="% Variance", title="Euclidean\nLog-Ratio PCoA Screenplot")
dev.off()
```

#Step Twenty One: Plot this euclidean PCoA in 3D with plotly, firstly by putting euclidean pcoa vectors into a dataframe
```{r}
my_euclidean_pcoa_df <- data.frame(my_euclidean_pcoa$vectors) %>%
  rownames_to_column(var="Glacier") %>%
  separate(Glacier, into =c("Glacier", "excess"), sep="_") %>%
  select(-excess) %>%
  data.frame

my_euclidean_pcoa_df$Glacier <- metadata$Glacier #here we'll change the sample names to the glacier names
head(my_euclidean_pcoa_df)
```

#Step Twenty Two: Extract eigenvalues from euclidean pcoa and plot
```{r}
euclidean_eigenvalues <- round(my_euclidean_variances[,2],digits=4)*100

#Plotly 3-d
plot_ly(my_euclidean_pcoa_df, type='scatter3d', mode='markers',
        x=~Axis.2,y=~Axis.3,z=~Axis.1,colors=~brewer.pal(6,'Set1'),color=~Glacier)%>%
  layout(font=list(size=18),
         title='PCoA Jaccard Distance',
         scene=list(xaxis=list(title=paste0('Co 2 ', euclidean_eigenvalues[2], '%'),
        showticklabels=FALSE, zerolinecolor='black'), yaxis=list(title=paste0('Co 3 ', euclidean_eigenvalues[3], '%'),
        showticklabels=FALSE, zerolinecolors='black'), zaxis=list(title=paste0('Co 1 ', euclidean_eigenvalues[1], '%'),
        showticklabels=FALSE,zerolinecolor='black')))
```

#Step Twenty Three: Set random seed to make sure NMDS plots come out the same each time (because they are iterative)
```{r}
set.seed(19614491)
```

#Step Twenty Four: Perform NMDS on eucidean distance matrix and jacard distance matrix
```{r}
NMDS_euclidean <-metaMDS(my_euclidean_dist,k=2,autotransform=FALSE)

NMDS_jaccard <-metaMDS(jaccard,k=2,autotransform=FALSE)
```

#Step Twenty Five: Look at the stress of each of the NMDS 
```{r}
NMDS_euclidean$stress 
NMDS_jaccard$stress
```

#Step Twenty Six: Plot the euclidean NMDS
```{r}
NMDS_euc_df <- data.frame(NMDS_euclidean$points) %>%
  rownames_to_column(var="Glacier") %>%
  separate(Glacier, into=c("Glacier", "excess"), sep="_") %>%
  select(-excess) %>%
  data.frame 

NMDS_euc_df$Glacier <- metadata$Glacier #here we'll change the sample names to the glacier names
head(NMDS_euc_df)

pdf("results/NMDS-euclidean.pdf")
ggplot(NMDS_euc_df, aes(x=MDS1, y=MDS2, fill=Glacier)) +
  geom_point(size=4, shape=21, color="black", aes(fill=Glacier)) +
  scale_fill_brewer(palette="Dark2")+
  theme_bw()+
  labs(x="NMDS 1", y="NMDS 2", title="Euclidean Distance NMDS")
dev.off()
```

#Step Twenty Seven: Plot the jaccard NMDS
```{r}
NMDS_jac_df <- data.frame(NMDS_jaccard$points) %>%
  rownames_to_column(var="Glacier") %>%
  separate(Glacier, into=c("Glacier", "excess"), sep="_") %>%
  select(-excess) %>%
  data.frame 

NMDS_jac_df$Glacier <- metadata$Glacier #here we'll change the sample names to the glacier names
head(NMDS_jac_df)

pdf("results/NMDS-jaccard.pdf")
ggplot(NMDS_jac_df, aes(x=MDS1, y=MDS2, fill=Glacier)) +
  geom_point(size=4, shape=21, color="black", aes(fill=Glacier)) +
  scale_fill_brewer(palette="Dark2")+
  theme_bw()+
  labs(x="NMDS 1", y="NMDS 2", title="Jaccard Distance NMDS")
dev.off()
```