---
title: "4-general-statistics"
author: "Amy Solman"
date: "01/04/2021"
output: html_document
---
#The below code and notes is taken directly in parts from 'The New Statistics with R: An Introduction for Biologists' by Andy Hector. These are not entirely MY notes, they are predominately ANDY HECTORS and I do not take credit.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#This script contains non-ordination exploratory statistics, including:
#Faith's phylogenetic diversity metric (Darcy et al., 2018)
#Linear regression of log-10 transformed cryoconite hole area and log-10 transformed faith's phylogenetic diversity (Darcy et al., 2018)
#Unweighted unifrac distances (Darcy et al., 2018)
#Mantel tests (vegan package) to look at relationship between geographic distance and unweighted unifrac distances (Darcy et al., 2018)
#Fit logistic model to log-10 transformed geographic distances and unweighted unifrac distances using downhill simplex optimization as implemented in R’s optim function (Darcy et al., 2018)
#Fit linear model to log-10 transformed geographic distances and unweighted unifrac distances (Darcy et al., 2018)
#Use log10 power law model to estimate c and z for data (Sommers et al., 2020)
#Analysis of Covariance (ANCOVA) with ASVs as proxy for unique species (Sommers et al., 2020)
#Use R package 'car' to calculate Type III errors, in which the variation assigned to each factor is not conditional on the order of terms in the model. Then estimate the values and standard errors of model parameters from fitting linear models seperately to the data from each domain of life on each glacier (Sommers et al., 2020)
#Add chlorophyl concentration as a covariate to the linear regression between log10S and log10A after accounting for glacier. Use Type I errors to evaluate the factors in these models sequentially - so what is the additional explainatory power of chlorophyl after accounting for hole area and glacier (Sommers et al., 2020)
#Inverse Simpson index using ‘hillR’ package as a measure of community evenness. Use sequential partitioning of variance (Type I errors) to test whether adding the evenness metric to the linear regression of log10 S against log10A was significantly related to additional variation after accounting for glacier (Sommers et al., 2020)
#ANCOVA on evenness with log10A to determine whether it co-varied with area and if eveness differed between glaciers when accounting for area (Sommers et al., 2020).
#Test whether the mass of sample or number of sequences in sample explain ISAR. 
#Linear model to explore whether total sediment mass collected varies with area of hole
#ANCOVA to ask if mass collected varied between glaciers while accounting for hole area.
#ANCOVA on whether sediment mass contributes to variation in the relationship between richness and area, use Type I errors to evaluate significance.
#Check normality of residuals using Shapiro–Wilk test
#Check for variance using Levene’s test for uneven variances
#Linear model of log10 S explained by glacier and log10A
#Linear model of log10 S explained by glacier and log10 sediment mass
#Remove sediment mass as a confounding covariate in log10A explaining log10S
#Is there collinearly between sediment mass and area?
#Does sediment mass differ between glaciers (use ANOVA)?
#Include sediment mass as covariate in richness-area ANCOVA to test whteer it is significantly related to additional variation in that model.
#Is there collinearity between number of sample reads and area?
#Does the number of sample reads differ between glaciers (use ANOVA)?
#Include number of sample reads as covariate in richness-area ANCOVA to test whether it is significantly related to additional variation in that model.


#Step One: Clear workspace and load packages
```{r}
rm(list=ls())
graphics.off()

library("phyloseq") #for working with phyloseq object
library("ggplot2") #for making our plots
#install.packages("arm") #use this if you've never installed the package before
library("arm") #use this to extract the key features of my linear models
library("picante") #for calculating faith's phylogenetic diversity 
library("vegan") #for caluclating our other diversity indices
#install.packages("DT")
library("DT") # interactive tables in html and markdown, for making a datatable with faiths diversity results
#install.packages("microbiome") #doesn't work
#library(microbiome) #useful phyloseq commands e.g. summarize_phyloseq()
#install.packages("geosphere")
library(geosphere) #for working with our geographic distances

```

#Step Two: Load data files and phyloseq object (I'm not sure which I'm going to work with at this stage!)
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
my_physeq <- readRDS("results/my-phyloseq-object.rds") #This is for caluclating phylogenetic diversity metrics
```

#Step: Calculate Faith's phylogenetic diversity 
```{r}
#How many unique ASVs do we have in each sample
S <- apply(t(count_table_clean) > 0, 1, sum)

#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(S) #Here we use the species count we got earlier

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my_physeq@otu_table) #Hellinger transformed asv table

my_tree <- my_physeq@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my_physeq@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

datatable(df.pd) #use DT package for this

```

#Step Three: Let's pull out all the data we need to perform our analyses and put it into it's own dataframe.
```{r}
#I could start by making a dataframe with sample name, glacier and number of unique ASVs
sample_name <- colnames(count_table_clean) #sample name
glacier <- metadata$Glacier #glacier
chlorophyll <- metadata$Chlorophyll #chlorophyll 
hole_depth <- metadata$Depth
DNA_g <- metadata$DNA_per_g

#calculate hole area 
#Method as per Darcy et al (2018)
hole_area_cm2 <- (pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2))*10000

#number of unique ASVs per sample
num_ASVs <- NULL
for (i in 1:ncol(count_table_clean)){
  num_ASVs <- c(num_ASVs, colSums(count_table_clean[i] > 0))
}

#now bind these into a dataframe
DF <- data.frame(sample_name, glacier, num_ASVs, hole_depth, hole_area_cm2, chlorophyll, DNA_g)

#Add alpha diversity indexes to dataframe
DF <- cbind(DF, div_shan, div_simp, div_inv_simp, div_pilou, df.pd$PD)

#remove any samples that have NA results
DF <- na.omit(DF)

```

#Step Four: Get our really basic statistics (mean, median, SD, variance, sample size, SE) See 'The new statistics with R' by Andy Hector for more detail on these statistics
```{r}
#ASVs
ASV_stats <- aggregate(num_ASVs ~ glacier, DF, function(x) c(mean = mean(x), median = median(x), sd = sd(x), variance = var(x), sample_size = length(x), se = sd(x)/sqrt(length(x))))

#hole depth
depth_stats <- aggregate(as.numeric(hole_depth) ~ glacier, DF, function(x) c(mean = mean(x), median = median(x), sd = sd(x), variance = var(x), sample_size = length(x), se = sd(x)/sqrt(length(x))))

#hole area
area_stats <- aggregate(hole_area_cm2 ~ glacier, DF, function(x) c(mean = mean(x), median = median(x), sd = sd(x), variance = var(x), sample_size = length(x), se = sd(x)/sqrt(length(x))))

#chlorophyll
chlorophyll_stats <- aggregate(chlorophyll ~ glacier, DF, function(x) c(mean = mean(x), median = median(x), sd = sd(x), variance = var(x), sample_size = length(x), se = sd(x)/sqrt(length(x))))

#DNA_g
DNA_stats <- aggregate(as.numeric(DNA_g) ~ glacier, DF, function(x) c(mean = mean(x), median = median(x), sd = sd(x), variance = var(x), sample_size = length(x), se = sd(x)/sqrt(length(x))))
```

#Step Five: Let's plot our data to visually assess any differences between our glaciers 
```{r}
# Create the loop.vector (all the columns)
loop.vector <- 3:12

for (i in loop.vector) { # Loop over loop.vector

  # store data in column.i as x
  x <- as.numeric(DF[,i])
  
qplot(glacier, x, data = DF, shape = glacier, colour = glacier)+
  theme_bw()

ggsave(file=paste0("../Outputs/Sommers2020/graphs/Sommers2020-glacier-", colnames(DF)[i], "-dotplots.pdf"))
}
```

#Step Six:  Let's create box and whisker plots for our data. These plots give the medians, 25th and 75th percentiles and highlight outlying data points.
```{r}
# # Create the loop.vector (all the columns)
# loop.vector <- 3:7

for (i in loop.vector) { # Loop over loop.vector

  # store data in column.i as x
  x <- as.numeric(DF[,i])
  
    qplot(glacier, x, data = DF, geom = "boxplot",
      main = paste(colnames(DF)[i], " by Glacier: Box-&-Whisker Plot"),
      xlab = "Glacier",
      ylab = colnames(DF)[i])+
  theme_bw()
ggsave(file=paste0("../Outputs/Sommers2020/graphs/Sommers2020-glacier-", colnames(DF)[i], "-boxplot.pdf"))
}
```

#Step Seven: Create linear models of our explanatory variable (glacier) and dependent variables (asv number, depth, area, chlorophyll, dna)

#Analysis of Variance (ANOVA), a linear model analysis. We want to ask: is the ASV richness/chlorophyll/hole size/hole depth/DNA per gram between glaciers significantly different? Plot the data divided into glacier groups
#ANOVA quantifies the signal to noise ratio. How different are our groups in ratio to how spread out the data are? ANOVA will tell us how confident we can be that we are not being fooled by a type I error (a false positive difference between the groups)
#ANOVA will: 1) quantify the overall variability of the dataset. 2) Divide that variability into the variability between and within groups. More signal = more condifent. More noise = less confident.
```{r}
#Least squares model 0
ls0 <- lm(num_ASVs ~ 1, data = DF) # here I have used '1' because I don't want to look at num_ASVs as a result of the explanatory variable 'glacier', I just want to estimate the grand mean of num_ASVs
display(ls0) #this uses the 'arm' package
#This gives us estimates of the model coefficients and their standard errors
#n = the number of data points 
#k = the number of parameters
#residual sd is the square root of the variance so it should equal var(DF$num_ASVs) which it more or less does (this may not be TOTALLY accurate due to rounding errors)
#basically, this linear model gives us the mean number of ASVs for all samples, the standard error of asvs across all samples, the standard deviation of asvs across all samples, the sample size and the number of parameters - so we get a lot of the statistics that we did earlier but in one easy step!

#But let's use ANOVA to look at the response variable (ASV) as a result of categorical explanatory variable (glacier)
lm_asv <- lm(num_ASVs ~ glacier, data=DF)
#Remember the ANOVA analysis makes a few assumptions about the data. If these assumptions are badly contravened then our results may not be trustworthy! 
display(lm_asv) #this gives us the mean asvs of each glacier where intercept coefficient = Canada mean. And glacier taylor coefficient (taylor mean) = Canada coefficient - 82.49. it also gives us the standard error of the mean and the standard error of the difference between the means.
summary(lm_asv) #R= 0.17 to 17% of the variation in number of ASV is explained by glacier as quantified by the total sum of squares. Adjusted R = 0.13 and this takes into account the complexity of the model

lm_depth <- lm(hole_depth ~ glacier, data=DF)
display(lm_depth)
summary(lm_depth) #summary gives a little more detail

lm_area <- lm(hole_area_cm2 ~ glacier, data=DF)
display(lm_area)
summary(lm_area)

lm_chlo <- lm(chlorophyll ~ glacier, data=DF)
display(lm_chlo)
summary(lm_chlo)

lm_dna <- lm(DNA_g ~ glacier, data=DF)
display(lm_dna)
summary(lm_dna)

#If samples were paired we could do a paired t test like so:
# anova(lm(num_ASVs ~ pair + glacier, data = DF))

# A Linear Model (linear regression analysis) to test for relationship between the categorical predictor variable (glacier) and the simpsons index
lm_div_simp <- lm(div_simp ~ glacier, data=DF)
display(lm_div_simp)
summary(lm_div_simp)

# A Linear Model (linear regression analysis) to test for relationship between the categorical predictor variable (glacier) and the inverse simpsons index
lm_div_inv_simp <- lm(div_inv_simp ~ glacier, data=DF)
display(lm_div_inv_simp)
summary(lm_div_inv_simp)

# A Linear Model (linear regression analysis) to test for relationship between the categorical predictor variable (glacier) and the shannons index
lm_div_shan <- lm(div_shan ~ glacier, data=DF)
display(lm_div_shan)
summary(lm_div_shan)

# A Linear Model (linear regression analysis) to test for relationship between the categorical predictor variable (glacier) and the pilou evenness
lm_div_pilou <- lm(div_pilou ~ glacier, data=DF)
display(lm_div_pilou)
summary(lm_div_pilou)

# A Linear Model (linear regression analysis) to test for relationship between the categorical predictor variable (glacier) and faith's pd
#rename the PD column from the DF dataframe other wise this won't work
names(DF)[12] <- "PD"

lm_pd <- lm(PD ~ glacier, data=DF)
display(lm_pd)
summary(lm_pd)
```

#Step Eight: View ANOVA tables for each of the linear models. One-way ANOVA!
```{r}
anova(lm_asv) #significant relationship F = 5.68 so the signal is 5.68 times larger than the noise
anova(lm_depth) #significant relationship
anova(lm_area) #no significant relationship F = 0.06 so the signal is 0.06 times larger than the noise
anova(lm_chlo) #no significant relationship
anova(lm_dna) #significant relationship
anova(lm_div_shan) #no significant relationship
anova(lm_div_simp) #no significant relationship
anova(lm_div_inv_simp) #no significant relationship
anova(lm_pd) #significant relationship

#Here the first column gives the source of the variation (the glacier and the leftover variation = the residual)
#The F values gives us our signal-to-noise ratio. 

#Of the asv ~ glacier ANOVA we could say: The Canada glacier cryoconite holes had an average of 392 unique ASVs while Taylor glacier cryoconite samples had an average of 309 unique ASVs - a difference of 83 unique ASVs which was statistically significant (F 1,28 = 5.7, p = 0.02)
```

#Step Nine: Look at the confidence intervals of the data 
```{r}
confint(lm(num_ASVs ~ glacier, data = DF)) #Confidence intervals 
coefplot(lm(num_ASVs ~ glacier, data = DF)) #This plot shows us that the Taylor glacier has almost 80 less unique ASVs than Canada. This shows the difference in mean ASVs (solid dot) with approximate 68% (thicker shorter bars) and 95% (thinner longer bars) confidence intervals
```

#Step Ten: Look at the residuals of our linear models (one-way ANOVA)
```{r}
par(mfrow=c(2,2))
plot(lm_asv) #Note that on our Normal Q-Q plot our residual differences fall approximately on a straight line 
```

#Step Eleven: Linear Regression Analysis
#We use linear regression analysis if our explanatory variable is continuous 
#For example, what if we want to see if hole area explains ASV richness?
#Linear regression analysis is not for hypothesis testing (like ANOVA and t-tests) but for estimating a relationship that can be used for prediction
#Can we use hole area to predict ASV number?
```{r}
area_asv_lm <- lm(num_ASVs ~ hole_area_cm2, data = DF)

#Let's create a graph of the linear model between ASV richness and hole area
xla <- expression(paste("Cryoconite hole area (", cm^2,")"))
yla <- "ASV abundance"
(qplot(x=hole_area_cm2, y = num_ASVs, data = DF,
       geom=c("point", "smooth"), method="lm")+
    theme_bw()+
    xlab(xla)+ylab(yla))
ggsave(file=paste0("../Outputs/Sommers2020/graphs/Sommers2020-asv-area-linear-regression.pdf"))

#Extract the point estimates from the linear regression
coef(area_asv_lm) #These results tell us that is we had a hole of area 0cm2 we would have 315 asvs
#so of the power law model y = a + b*x a = intercept (the number of asvs if our hole area was 0cm2), b = slope (change in asv number with each unit change in hole area)

#we can look at the model matrix 
model.matrix(area_asv_lm) #This contains a column of ones that indicate the intercept and a column of area values. 

#we can use the linear regression equation to predict the value of ASVs with y = (ax1) + (b*x)
coef(area_asv_lm)[1]*1 + coef(area_asv_lm)[2]*1000

#The fitted function extracts the fitted values for each observed value of the explanatory variable, this is the predicted value of y for each observed value of x
fitted(area_asv_lm) #these are the points the line is fitted on

#The residuals are the difference between the observed and fitted values and can be extracted with a function of the same name
residuals(area_asv_lm)
```

#Step:plot linear regressions of all variables will all variables
```{r}
#plot linear regressions of all variables will all variables
for (i in loop.vector) {
for (j in loop.vector) { # Loop over loop.vector

  # store data in column.i as x
  a <- as.numeric(DF[,i])
  b <- as.numeric(DF[,j])
  
#Let's create a graph of the linear model between ASV richness and our other continuous variables
xla <- colnames(DF)[i]
yla <- colnames(DF)[j]
(qplot(x=a, y = b, data = DF,
       geom=c("point", "smooth"), method="lm")+
    theme_bw()+
    xlab(xla)+ylab(yla))
ggsave(file=paste0("../Outputs/Sommers2020/graphs/Sommers2020", colnames(DF)[i], "-", colnames(DF)[j], "-linear-regression.pdf"))
}
}
```

#Step: Confidence intervals and prediction intervals
```{r}
#95% confidence intervals reflect our confidence only in the average relationship - the regression line itself. 
#In this case we want to use this regression relationship to predict the asv abundance of new cryoconite holes based on their area. We want to accompany our prediction with an interval that converys our confidence in it and the 95% CI does not do that. Instead we need a prediction interval (PI). The 95% CI can be said to convey our inferential uncertainty, since it conveys our confidence in our restimate of the regression relationship.
#However there is scatter around this line and because the PI applies to a prediction around a new point the 95% PI needs to include the predictive uncertainty. 

#Let's get hole areas with which to predict ASV abundance
plot.holes <- seq(min(DF$hole_area_cm2), max(DF$hole_area_cm2), length = 100)

#We can that substitute this sequence for the observed areas in the predict function
(predictions <- predict(area_asv_lm, list(hole_area_cm2 = plot.holes), se=TRUE)) #This gives us the predicted ASV abundance for 100 regular hole areas together with the standard error at each point. We also get the total DF 30 - 2 = 28 and the residual SD from the display() and summary() output, which is the square root of the residual vairance from the ANOVA table

anova(area_asv_lm) #$residual.scale = 72.01614
sqrt(5186) #72.01389

#Now we can plot the predicted ASV values to produce our regression line and use the standard error to caluclate 95% CI. 
t.value <- qt(0.975, 28) #The Student t Distribution = 2.05 (close enough to 2)

up.ci <- predictions$fit + t.value*predictions$se.fit
lo.ci <- predictions$fit - t.value*predictions$se.fit #This is showing what is going on behind the scenes but predict() can be used do it for you

(predictions <- predict(area_asv_lm, list(hole_area_cm2=plot.holes),
                        interval="confidence", se=TRUE)) #"confidence" = CI, replace with "predictions" for PI bounds
plot(DF$hole_area_cm2, DF$num_ASVs, xlab="Hole Area cm2", ylab = "ASV Abundance")+
lines(predictions$fit[,1]~plot.holes)+  #regression line
lines(predictions$fit[,2]~plot.holes, lty = 2)+ #upper CI
lines(predictions$fit[,3]~plot.holes, lty = 2) #lower CI

#PI is wider than CI due to the inclusion of predictive uncertainty 


```

#Step: Checking assumptions
```{r}
#The lm() function performs a normal least squares analysis which assumes that the residuals are approximately normally distributed. 
#The variance or SD should also be approximately constant in different treatments or across the range of response variables.
#Model checing is usually done by eyeballing the data! 

par(mfrow=c(1,2)) #divide page into 1x2 frames
plot(area_asv_lm, which = c(1:2)) #here we get the residual vs fitted and the normal Q-Q plot < these are made automatically when plot is applied to a linear model object.
#The Residual vs Fitted plot shows that the variability is not constant but increases a little at lower fitted values before decreasing at higher fitted values
#The Normal Q-Q plot the residuals are plotted against the normal distribution and the points mostly follow the line. So the data points are mostly
#normally distributed with a couple of outlying values numbered on the graph.

#With this info we can A) accept the shortcomings of the analysis and the uncertainty this brings to the results or 
# B) transform the data so they more closely follow the assumption that the residuals follow a normal distribution.

#Transformations may include: square-root transformation and log transformation.
area.asv.sqrt <- lm(sqrt(num_ASVs) ~ hole_area_cm2, data=DF)
area.asv.log <- lm(log(num_ASVs) ~ hole_area_cm2, data=DF)

(area.asv.sqrt.plot <- qplot(hole_area_cm2, sqrt(num_ASVs), data = DF,
                             geom = c("point", "smooth"), method="lm", formula= y~x,
                             main="Linear regression, square-rooted data")+theme_bw())

(area.asv.sqrt.res <- qplot(fitted(area.asv.sqrt), resid(area.asv.sqrt),
                            ylab="Residuals", xlab="Predicted values",
                            main = "Variance decreases with mean")+theme_bw()+
    geom_hline(yintercept = 0, colour="grey50"))

par(mfrow=c(1,2)) #divide page into 1x2 frames
plot(area.asv.sqrt, which = c(1:2))

(area.asv.log.plot <- qplot(hole_area_cm2, log(num_ASVs), data = DF,
                             geom = c("point", "smooth"), method="lm", formula= y~x,
                             main="Linear regression, log data")+theme_bw())

(area.asv.log.res <- qplot(fitted(area.asv.log), resid(area.asv.log),
                            ylab="Residuals", xlab="Predicted values",
                            main = "Variance decreases with mean")+theme_bw()+
    geom_hline(yintercept = 0, colour="grey50"))

par(mfrow=c(1,2)) #divide page into 1x2 frames
plot(area.asv.log, which = c(1:2))

```

#Step: Faith's phylogenetic diversity metric (Darcy et al., 2018)
```{r}

#To plot Faith's phylogenetic diversity metric we need:
#1) The rarefied count table data from a phyloseq object

#STEP ONE: GET OUR PHYLOSEQ OBJECT
#I can't use the my_physeq object because that contains Hellinger transformed data and I just want the count data
#So we'll make the phyloseq object from our imported data objects
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#STEP TWO: RAREFY THE DATA TO THE SMALLEST NUMBER OF READS
#What is the smallest number of reads for a sample?
how.many.reads <- t(abundances(my.ps)) #need microbiome package for this
min.reads <- min(rowSums(how.many.reads)) #what is our least amount of reads in any of our samples? This is what we will rarefy to.

set.seed(1961) #This will ensure the results of our normalisation step remain the same when repeated
#because this is going to randomly choose min.reads number of reads per sample

ps.rar <- rarefy_even_depth(my.ps, sample.size = 13401)

#compare rarefied phyloseq object to non-rarefied phyloseq objecy
my.ps
ps.rar

#STEP THREE: CHECK THAT THE PHYLOGENETIC TREE IS STILL ROOTED AFTER RAREFYING
ps.rar@phy_tree #it is rooted

#STEP FOUR: COMPUTE FAITHS PHYLOGENETIC DIVERSITY 
ps.pd <- pd(t(ps.rar@otu_table), ps.rar@phy_tree, include.root = T) #transpose the asv table and find phylogetetic diversity

datatable(ps.pd) #we can look at the results here
```

#Step: Linear regression of log-10 transformed cryoconite hole area and log-10 transformed faith's phylogenetic diversity (Darcy et al., 2018)
```{r}

#STEP ONE: MAKE A DATAFRAME WITH SAMPLE, HOLE AREA AND FAITHS PHYLOGENETIC DIVERSITY
PD <- ps.pd$PD
Area <- hole_area_cm2
Samples <- rownames(ps.pd)
# df1 <- cbind(rownames(ps.pd), as.numeric(hole_area_cm2), as.numeric(ps.pd$PD))
# colnames(df1) <- c("Sample", "Area", "PD")
df1 <- data.frame(Samples, Area, PD)
area.pd.lm <- lm(log(PD) ~ log(Area), data = df1)
area.pd.lm
summary(area.pd.lm)

```

#Step: Unweighted unifrac distances (Darcy et al., 2018)
```{r}
#We can do this with out rarefied data
unw.uni.rar <- ordinate(ps.rar, "PCoA", "unifrac", weighted=F)

#And non-rarefied data
unw.uni.non.rar <- ordinate(my.ps, "PCoA", "unifrac", weighted=F)

```

#Step: Mantel tests (vegan package) to look at relationship between geographic distance and unweighted unifrac distances (Darcy et al., 2018)
```{r}
#Let's get our geographic distances from our phyloseq object
latitude <- as.numeric(substr(my.ps@sam_data$lat_lon, 1, 8))
longitude <- as.numeric(substr(my.ps@sam_data$lat_lon, 12, 21))

#Find index of NA result
my.na <- which(is.na(my.ps@sam_data$lat_lon))
#Which sample ID does that equate to 
my.ID <- my.ps@sam_data$Run[my.na]
#remove from phyloseq object
#my.ps.prune = subset_samples(my.ps, sampleID != my.ID)
physeq.p2 = subset_samples(my.ps, sample_names(my.ps) != my.ID)

#let's get out geographic distances again
latitude <- as.numeric(substr(physeq.p2@sam_data$lat_lon, 1, 8))
longitude <- as.numeric(substr(physeq.p2@sam_data$lat_lon, 12, 21))

#put into a dataframe
geo = data.frame(longitude, latitude)

#convert into a distance matrix
d.geo = distm(geo, fun = distHaversine) #This function uses the geosphere package
dist.geo = as.dist(d.geo)

#And re-do the unweighted unifrac distances 

#We can do this with out rarefied data
unw.uni.rar <- ordinate(physeq.p2, "PCoA", "unifrac", weighted=F)

#And non-rarefied data
unw.uni.non.rar <- ordinate(physeq.p2, "PCoA", "unifrac", weighted=F)

#Now we can run our Mantel test

#abundance vs geographic 
abund_geo  = mantel(unw.uni.non.rar, dist.geo, method = "spearman", permutations = 9999, na.rm = TRUE)
abund_geo

```

#Step: 
```{r}

```

#Step: 
```{r}

```

#Step: 
```{r}

```

#Step: 
```{r}

```

#Step: 
```{r}

```

#Step: 
```{r}

```

#Step: 
```{r}

```

#Step: 
```{r}

```

