---
title: "Basic Hypothesis Test with T and F Tests"
author: "Amy Solman"
date: "22/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook (https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/13-t_F_tests.ipynb)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Basic Hypothesis Testing
#We're going to use t tests to compare differences in means. We're going to use F tests to compare the variance of samples. We're also going to look at non-parametric tests.

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
# library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
# library("picante")
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```
#T-tests
#Use to compare the means of two samples, or one sample to another value.
#T-test is best used for use with a two level factor (like my two glaciers) as multiple levels multiplies the significance level (0.05), increasing the chances of making a Type I error (a false positive)
# t = the difference between the values/the standard error
#If there is no difference then t = 0, so the larger the t value the bigger the difference between factors

#Step Three: One-sample t-test
#Is the mean of a sample difference from a specific value
```{r}
#get the mean species richness
mean.rich <- mean(my.data.sub$Richness)

#get the variance 
var.rich <- var(my.data.sub$Richness)

#get the number of datapoints
n.rich <- length(my.data.sub$Richness)

#now we can ask if the mean species richness is different from 400 (for example, I've just picked 400 out of nowhere!)
diff.rich <- mean.rich - 400

#get the standard error
se.rich <- sqrt(var.rich/n.rich)

#Now we can calculate the t value
t.rich <- diff.rich/se.rich
print(t.rich) #-2.38 this is a reasonably high t value so we can say that the mean species richness of our glaciers is different from 400

#or we can do all this in one go
t.test(my.data.sub$Richness, mu = 400) #t = -2.38. p = 0.02 so there is a significant different


```

#Step Four: Two-sample t-tests
#The two-sample t-test is more commonly used that the one sample t test to compare the means of two samples.
#This test makes the assumption that are data are normally distributed and have similar vairances
```{r}
t.test(Richness ~ Glacier, data = my.data.sub) # t = 2.067, p = 0.07 there is no significant difference between the mean species richnesses
t.test(Faith ~ Glacier, data = my.data.sub) #p > 0.05
t.test(Shannon ~ Glacier, data = my.data.sub) #p > 0.05
t.test(Simpson ~ Glacier, data = my.data.sub) #p > 0.05
t.test(Inv.Simpson ~ Glacier, data = my.data.sub) #p > 0.05
t.test(Pilou ~ Glacier, data = my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE
```

#Step Five: F Test 
#The F test can be used to compare variances of two samples or populations. This can be used when checking to see if data meet the variance assumption (like in the two sample t-test)
#If F value is < 0.5 or > 2 they have may have unequal variances, check the p value (p < 0.05 means there is a significant difference in the variances)
```{r}
var.test(Richness ~ Glacier, data = my.data.sub) #p > 0.05
var.test(Faith ~ Glacier, data = my.data.sub) #p > 0.05
var.test(Shannon ~ Glacier, data = my.data.sub) #p > 0.05
var.test(Simpson ~ Glacier, data = my.data.sub) #p > 0.05
var.test(Inv.Simpson ~ Glacier, data = my.data.sub) #p > 0.05
var.test(Pilou ~ Glacier, data = my.data.sub) # p < 0.05
```

#Non-parametric tests
#If our data are not normally distributed and don't have equal variances (and log transformations don't make our data normal) then we can use a non-parametric alternative
```{r}
wilcox.test(Richness ~ Glacier, data = my.data.sub)
```