---
title: "data-prep"
author: "Amy Solman"
date: "05/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script will prepare our QIIME2 output for use in downstream analysis.

#Step One: Clear work space and load packages

```{r}
rm(list=ls())
graphics.off()

library(tidyverse)
library(phyloseq)
library(readr)
library(seqinr)
library(decontam)
library(ape)
library(vegan)
```

#Step Two: Load the data from our QIIME2 step +metadata

```{r}
count_table <- read_tsv(file="export/table/table.tsv", skip=1)

taxonomy <- read_tsv(file="export/taxonomy/taxonomy.tsv")

tree <- read_tree("export/exported-tree/tree.nwk")

fasta <- read.fasta(file="export/rep-seqs.fasta/dna-sequences.fasta")

sample_info_tab <- read_csv(file="export/SraRunTable-wrangle.tsv")
```
#Step Three: Specify the first column in our count table is the row names
```{r}
count_table <- column_to_rownames(count_table, var = colnames(count_table)[1])
rownames(count_table) <- paste0("ASV", 1:nrow(count_table))
```

#Step Four: Load metadata table and remove unnecessary rows/columns/data

```{r}
#delete unneccessary row and column
sample_info_tab <- sample_info_tab[-c(1),]
sample_info_tab <- sample_info_tab[,-c(1)]
sample_info_tab <- sample_info_tab[!grepl("18S", sample_info_tab$Library.Name),] #only include 16S samples
sample_info_tab <- sample_info_tab[-c(42:80),] #remove reverse reads so we only have one entry per sample

colnames(count_table) <- unique(sample_info_tab$Sample.Name)

#add a column with glacier name
Glacier <- NULL
for (i in 1:nrow(sample_info_tab)){
  if (startsWith(sample_info_tab$Sample.Name[[i]], "C")){
    Glacier <- c(Glacier, "Canada")
  } else {
    Glacier <- c(Glacier, "Taylor")
  }
}

sample_info_tab <- cbind(sample_info_tab, Glacier)

#Keep only the information from the metadata table that will be useful in later analysis
sampledata = sample_data(data.frame(
  Sample_name = sample_info_tab$Sample.Name,
  Location = sample_info_tab$Glacier,
  Lat_Long = sample_info_tab$lat_lon,
  Elevation = sample_info_tab$elev,
    Diameter1 = sample_info_tab$diameter1,
  Diameter2 = sample_info_tab$diameter2,
  Organic_matter = sample_info_tab$org_matter,
  Water_content = sample_info_tab$water_content,
  Variable1 = sample_info_tab$undefined,
  Variable2 = sample_info_tab$undefined.2,
  Variable3 = sample_info_tab$undefined.3,
  row.names = sample_info_tab$Sample.Name,
  stringsAsFactors = FALSE
))
```
#Step Five: Modifty taxonomy table
```{r}
taxonomy <- taxonomy %>%
  mutate(taxonomy=str_replace_all(string=Taxon, pattern="D_\\d*\\__", replacement="")) %>%
  mutate(taxonomy=str_replace_all(string=taxonomy, pattern=";$", replacement="")) %>%
separate(taxonomy, into=c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep=";") %>%
  select(-Taxon, -Confidence) %>%
  column_to_rownames(var='Feature ID')

row.names(taxonomy) <- row.names(count_table) #give taxonomy same ASV names as count_table
```

#Step Six: Identify contaminents in the count data
```{r}
# A logical vector showing which entries are 'TRUE' for controls and 'FALSE' for genuine samples
decontam <- startsWith(sampledata$Sample_name, "NT")

# Run isContaminant (remember to transform the dataframe i.e. t())
contam <- isContaminant(t(count_table), neg=decontam)

# Look at the results
table(contam$contaminant) # how many ASVs were identified as contaminants? None. Why?

# Make a vector holding the contaminant IDs
contam_asvs <- row.names(contam[contam$contaminant == TRUE, ])

# What is the taxonomy of our contaminant ASVs?
taxonomy[taxonomy$`Feature ID` %in% contam_asvs] #Why don't I have any contaminants?!?

```

#Step Seven: Remove contaminents from count table,taxonomy table, sequence data, metadata and tree
```{r}

###COUNT TABLE
x <- rownames(count_table) %in% contam_asvs #which rows are contaminants?
count_table_clean <- count_table[!x,] #remove those rows

# and remove blank samples from count table
#what is the index of "TRUE" values in decontam?
true_index <- NULL
for (i in 1:length(decontam)){
  if (decontam[[i]] == "TRUE"){
    true_index <- c(true_index, i)
  }
}
count_table_clean <- count_table_clean[,-c(true_index)]  #remove those samples

###TAXONOMY TABLE
z <- taxonomy$`Feature ID` %in% contam_asvs
taxonomy_clean <- taxonomy[!row.names(taxonomy) %in% z,]#remove contaminating ASVs

###FASTA FILE
#Remove contaminants from dna sequences file
fasta_clean <- fasta
fasta_clean[contam_asvs] <- NULL

###METADATA TABLE
y <- startsWith(sampledata$Sample_name, "NT")
true_index <- NULL
for (i in 1:length(y)){
  if (y[[i]] == "TRUE"){
    true_index <- c(true_index, i)
  }
}
sampledata_clean <- sampledata[-c(true_index),]  #remove those samples

###PHYLOGENETIC TREE
# This is a phyloseq object so we can use a phyloseq command called prune_taxa
tree_clean <- prune_taxa(count_table_clean$`#OTU ID`, tree)
```

#Step Eight: Plot rarefaction curves to see how deeply our samples have been sequenced
```{r}
#Plot the rarefaction curves, color-coding by sample and tranform using t()
#Col= option could also be used to look at different sample types/locations?
pdf("rplot.pdf") 
rarecurve(t(count_table_clean), step=100, cex=0.5, col=palette(rainbow(nrow(sampledata_clean))), lwd=2, ylab="ASVs", label=T)
# Add line to indicate the fewest # of sequences from any sample
abline(v=(min(rowSums(t(count_table_clean)))))
dev.off()
```

#Step Nine: Convert data to phyloseq objects and remove singletons

```{r}

###COUNT TABLE
# Convert count table to phyloseq package
count_table_phyloseq <- otu_table(as.matrix(count_table_clean), taxa_are_rows = TRUE)
ASV <- prune_taxa(taxa_sums(count_table_phyloseq) > 1, count_table_phyloseq) # Filter using prune_taxa

###TAXONOMY TABLE
TAX = tax_table(as.matrix(taxonomy_clean))
TAX <- prune_taxa(taxa_sums(count_table_phyloseq) > 1, TAX)

###METADATA TABLE 
META = sample_data(sampledata_clean)

###PHYLOGENETIC TREE
# This is a phyloseq object so we can use a phyloseq command called prune_taxa
taxa_names(TREE) = taxa_names(ASV) #Change tree names to match ASVs
TREE <- prune_taxa(taxa_sums(count_table_phyloseq) > 1, tree_clean)

#Make into one phyloseq object
my_physeq = phyloseq(ASV, TAX, META, TREE)

#I'm going to export my phyloseq object
saveRDS(my_physeq, "my-phyloseq-object.rds")
```

###Step Ten: Perform Hellinger transformation

```{r}
count_table_hellinger <- decostand(t(count_table_clean), "hellinger")
# Convert back to a tibble and switch rows for columns again 
count_table_hellinger <- as_tibble(t(count_table_hellinger), rownames = NA)
```

###Step Eleven: Export data files 
```{r}
#Note singletons have not been removed from these files, only the phyloseq object
save(count_table_clean, count_table_hellinger, taxonomy_clean, fasta_clean, tree_clean, sampledata_clean, file = "results/cleaned_files/cleanedfiles.RData")
```