---
title: "Neutral"
author: "Amy Solman"
date: "16/04/2021"
output: html_document
---
The following code is derived from Barnett et al., (2020) Soil characteristics and land-use drive bacterial community assembly patterns. Code available at: https://github.com/seb369/landuse_comm_assembly/blob/master/Neutral_model.md. The neutral model code of Barnett et al., (2020) was adapted from Burns, A., Stephens, W., Stagaman, K. et al. Contribution of neutral processes to the assembly of gut microbial communities in the zebrafish over host development. ISME J 10, 655–664 (2016) doi:10.1038/ismej.2015.142
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages needed for analysis
```{r}
library(dplyr)
library(tibble)
library(phyloseq)
library(ape)
library(vegan)
library(FSA)
library(eulerr)

# Packages needed for plotting
library(ggplot2)
library(grid)
library(gridExtra)
```

#Step Two: Import data and make phyloseq object
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)

```

#Step Three: Check to make sure each sample has at least 10000 reads 
```{r}
#check number of reads in each sample
read.num <- data.frame("count" = colSums(otu_table(my.ps))) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps)), by="Run") %>%
  arrange(-count) %>%
  mutate(X.Sample=factor(Run, levels=Run))

#Check to make sure all of our samples are at or above 10000 reads
ggplot(data=read.num, aes(x=Run, y=log10(count), fill = Glacier)) +
  geom_bar(stat="identity")+
  labs(x="Sample", y="Log10(Read count)")+
  geom_hline(yintercept=log10(10000))+
  theme(text=element_text(size=16),
       axis.text.x=element_blank())
```

#Step Four: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

```{r}
## Code adapted from Adam Burns (http://dx.doi.org/10.1038/ismej.2015.142)
require(minpack.lm)
require(Hmisc)
require(stats4)
# Extract the OTU table from pthe phyloseq object
OTU.table = t(otu_table(bulk.physeq.rare))

# Calculate the number of individuals in the meta community (Average read depth)
N <- mean(apply(OTU.table, 1, sum))

# Calculate the average relative abundance of each taxa across communities
p.m <- apply(OTU.table, 2, mean)
p.m <- p.m[p.m != 0]
p <- p.m/N
p.df = data.frame(p) %>%
  rownames_to_column(var="OTU")

# Calculate the occurrence frequency of each taxa
OTU.table.bi <- 1*(OTU.table>0)
freq.table <- apply(OTU.table.bi, 2, mean)
freq.table <- freq.table[freq.table != 0]
freq.df = data.frame(OTU=names(freq.table), freq=freq.table)

#Combine
C <- inner_join(p.df,freq.df, by="OTU") %>%
  arrange(p)
# Remove rows with any zero (absent in either source pool or local communities). You already did this, but just to make sure we will do it again.
C.no0 <- C %>%
  filter(freq != 0, p != 0)

#Calculate the limit of detection
d <- 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
p.list <- C.no0$p
freq.list <- C.no0$freq
m.fit <- nlsLM(freq.list ~ pbeta(d, N*m*p.list, N*m*(1-p.list), lower.tail=FALSE), start=list(m=0.1))
m.ci <- confint(m.fit, 'm', level=0.95)
m.sum <- summary(m.fit)
m.coef = coef(m.fit)

freq.pred <- pbeta(d, N*coef(m.fit)*p.list, N*coef(m.fit)*(1-p.list), lower.tail=FALSE)
Rsqr <- 1 - (sum((freq.list - freq.pred)^2))/(sum((freq.list - mean(freq.list))^2))

# Get table of model fit stats
fitstats <- data.frame(m=m.coef, m.low.ci=m.ci[1], m.up.ci=m.ci[2], 
                       Rsqr=Rsqr, p.value=m.sum$parameters[4], N=N, 
                       Samples=nrow(OTU.table), Richness=length(p.list), 
                       Detect=d)

# Get confidence interval for predictions
freq.pred.ci <- binconf(freq.pred*nrow(OTU.table), nrow(OTU.table), alpha=0.05, method="wilson", return.df=TRUE)

# Get table of predictions
pred.df <- data.frame(metacomm_RA=p.list, frequency=freq.pred, 
                      frequency_lowerCI=freq.pred.ci[,2], 
                      frequency_upperCI=freq.pred.ci[,3]) %>%
  unique()

# Get table of observed occupancy and abundance
obs.df = C.no0 %>%
  rename(metacomm_RA = p, frequency=freq)

```

```{r}

fulldata.model.plot = ggplot(data=obs.df) +
    geom_point(data=obs.df, aes(x=log10(metacomm_RA), y=frequency), 
               alpha=.2, size=2, color="grey30") +
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency), color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_lowerCI), linetype=2, color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_upperCI), linetype=2, color="black") + 
    geom_text(data=fitstats, aes(label = paste("R^2 == ", round(Rsqr, 3))), 
              x=-4.9, y=0.95, size=4, parse=TRUE) +
    geom_text(data=fitstats, aes(label = paste("italic(m) ==", round(m, 3))), 
              x=-4.9, y=0.85, size=4, parse=TRUE) + 
    labs(x="Log10 abundance in\nmetacommunity", y="Frequency detected") +
    theme_bw() +
    theme(axis.line = element_line(color="black"),
          legend.position = "none",
          axis.title = element_text(size=14),
          axis.text = element_text(size=12))

fulldata.model.plot

```