---
title: 'Linear Models: Looking at Multiple Variables with Interactions'
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/17-MulExplInter.ipynb) 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

######I WILL BE USING FAKE DATA FOR THIS ANALYSIS FOR DEMONSTRATION PURPOSES#####

#Here we're going to look at MULTI EXPLANATORY VARIABLE LINEAR MODELS with INTERACTIONS. We will also learn how to use ANCOVA.

#Two models we will build:
#Model 1 - Is species richness predicted by interactions between glacier and substrate?
#ANCOVA - Is species richness predicted by interactions between Glacier and hole area?

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library("picante")
library(gplots)
library(lattice)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
# load("../Analyses/results/cleaned_files/cleanedfiles_3.RData")
# #hole area - calculate and add to dataframe
# #Method as per Darcy et al (2018)
# metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
# metadata$area <- metadata$area*10000
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
# #Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
# div_shan <- diversity(t(count_table_clean), index = "shannon")
# 
# #Use vegan package to get simpson diversity index of each sample
# div_simp <- diversity(t(count_table_clean), index = "simpson")
# 
# #Use vegan package to get inverted simpson diversity index of each sample
# div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")
# 
# #Faith's phylogenetic diversity 
# asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table
# 
# my_tree <- my.ps@phy_tree #get my phylogenetic tree
# 
# #Check if the tree is rooted
# my.ps@phy_tree #Rooted, includes branch lengths
# 
# #It is a rooted tree so...
# df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.
# 
# #Calculate Pilou evenness
# div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier
# 
# #bind alpha diversity indices into a dataframe
# diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
# colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
# my.data <- cbind(metadata, diversity)
# 
# #remove columns with data we're not interested in
# my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
# rownames(my.data.sub) <- my.data.sub$Run
# my.data.sub <- subset(my.data.sub, select = -c(1))
# my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
# my.data.sub$Glacier <- metadata$Glacier
# my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Six: Add another fake Glacier of data so we can compare three glaciers - SKIP THIS STEP NO NECCESSARY
```{r}
# Chlorophyll <- sample(as.numeric(my.data.sub$Chlorophyll), size = 12)
# Depth <- sample(as.numeric(my.data.sub$Depth), size = 12)
# DNA_per_g <- sample(as.numeric(my.data.sub$DNA_per_g), size = 12)
# org_matter <- sample(as.numeric(my.data.sub$org_matter), size = 12)
# pH <- sample(as.numeric(my.data.sub$pH), size = 12)
# water_content <- sample(as.numeric(my.data.sub$water_content), size = 12)
# area <- sample(as.numeric(my.data.sub$area), size = 12)
# Faith <- sample(as.numeric(my.data.sub$Faith), size = 12)
# Richness <- sample(as.numeric(my.data.sub$Richness), size = 12)
# Shannon <- sample(as.numeric(my.data.sub$Shannon), size = 12)
# Simpson <- sample(as.numeric(my.data.sub$Simpson), size = 12)
# Inv.Simpson <- sample(as.numeric(my.data.sub$Inv.Simpson), size = 12)
# Pilou <- sample(as.numeric(my.data.sub$Pilou), size = 12)
# Glacier <- rep("Solman", 12)
# 
# df <- cbind(Chlorophyll, Depth, DNA_per_g, org_matter, pH, water_content, area, Faith, Richness, Shannon, Simpson, Inv.Simpson, Pilou, Glacier)
# 
# my.data.sub.X <- rbind(my.data.sub, df)
# my.data.sub.X.1 <- my.data.sub.X[,-c(14)] #remove glacier names
# 
# my.data.sub <- as.data.frame(lapply(my.data.sub.X.1, function(x) as.numeric(as.character(x))))
# #re-add glacier names 
# my.data.sub$Glacier <- as.factor(my.data.sub.X$Glacier)
```

#Step Seven: Add an addition factor for us to play with
```{r}
# Substrate <- rep(c("Ice", "Snow", "Cryoconite"), 9) 
# Substrate <- c(Substrate, rep("Cryoconite", 1))
# my.data.sub$Substrate <- as.factor(Substrate)
```

#Step Eight: We're going to re-fit the model from the previous step but changing the '+' symbol to '*' to show that we want to look at INTERACTIONS betweent he predictor variables as well as their MAIN EFFECTS
```{r}
# model <- lm(Richness ~ Glacier * Substrate, data = my.data.sub)
# 
# #diagnostic plots
# par(mfrow=c(2,2), mar=c(3,3,1,1), mgp=c(2,0.8,0))
# plot(model)
# 
# #ANOVA
# anova(model) #glacier as a main effect seems like the only singificant predictor - glacier:Substrate explains about half as much variation as glacier alone)
# 
# #SUMMARY
# summary(model)
# #Intercept = the reference level for cryoconite on Canada glacier
```

#Step Nine: ANCOVA - a linear model with one categorical and one or more continuous predictor variables
```{r}
# my.data.sub$logRich <- log(my.data.sub$Richness)
# my.data.sub$logArea <- log(my.data.sub$area)

# #fit a model of log Richness as a function of both hole area and glacier
# model_1 <- lm(logRich ~ logArea * Glacier, data = my.data.sub)
# 
# anova(model_1) # There is a significant main effect of area and glacier but there is no strong interaction between area and glacier.
# 
# summary(model_1) #r2 value is high - the model explains three quaters of the variation in Richness
# #The intercept is for the first level of Glacier (Canada)
# #the next line for logArea is the slope for Canada
# #Then we have the difference in the intercept for Taylor
# #The last line is the interaction between Glacier and hole area, this shows how the slope for Canada differes from the slope for Taylor

###TOTAL TAXA 
#fit a model of log Richness as a function of both hole area and glacier
tot.model <- lm(logRichness ~ logArea * Glacier, data = tot.diversity.data)

anova(tot.model) # There is a significant main effect of area and glacier but there is no strong interaction between area and glacier.

summary(tot.model)

###ABUNDANT TAXA 
#fit a model of log Richness as a function of both hole area and glacier
abun.model <- lm(logRichness ~ logArea * Glacier, data = abun.diversity.data)

anova(abun.model) # There is a significant main effect of area and glacier but there is no strong interaction between area and glacier.

summary(abun.model)

###RARE TAXA 
#fit a model of log Richness as a function of both hole area and glacier
rare.model <- lm(logRichness ~ logArea * Glacier, data = rare.diversity.data)

anova(rare.model) # There is a significant main effect of area and glacier but there is no strong interaction between area and glacier.

summary(rare.model)
```

#Step 10: Get predicted lines from the model!
```{r}
###TOTAL TAXA 
#get the range of the data
rng <- range(tot.diversity.data$logArea)
#get a sequence from the min to the max with 100 equally spaced values:
LogAreaForFitting <- seq(rng[1], rng[2], length = 100)
print(LogAreaForFitting)

#we're going to use these with the model to predict values of richness at each of those hole areas for each of our glaciers
#get a data frame of new data for the order
CanVals <- data.frame(logArea = LogAreaForFitting, Glacier = "Canada")
#get the predictions and standard error
CanPred <- predict(tot.model, newdata = CanVals, se.fit = TRUE)

#get a data frame of new data for the order
TayVals <- data.frame(logArea = LogAreaForFitting, Glacier = "Taylor")
#get the predictions and standard error
TayPred <- predict(tot.model, newdata = TayVals, se.fit = TRUE)

# plot the scatterplot of the data
pdf("../Outputs/Sommers2020/graphs/linear-model-interaction-logarea--glacier-logrichness-total.pdf", width=6, height=4) 
plot(logRichness ~ logArea, data = tot.diversity.data, col=as.factor(Glacier))
# add the predicted lines
lines(CanPred$fit ~ LogAreaForFitting, col = "black")
lines(CanPred$fit + CanPred$se.fit ~ LogAreaForFitting, col = "black", lty = 2)
lines(CanPred$fit - CanPred$se.fit ~ LogAreaForFitting, col = "black", lty = 2)
lines(TayPred$fit ~ LogAreaForFitting, col = "red")
lines(TayPred$fit + TayPred$se.fit ~ LogAreaForFitting, col = "red", lty = 2)
lines(TayPred$fit - TayPred$se.fit ~ LogAreaForFitting, col = "red", lty = 2)
dev.off()

###ABUNDANT TAXA 
#get the range of the data
rng <- range(abun.diversity.data$logArea)
#get a sequence from the min to the max with 100 equally spaced values:
LogAreaForFitting <- seq(rng[1], rng[2], length = 100)
print(LogAreaForFitting)

#we're going to use these with the model to predict values of richness at each of those hole areas for each of our glaciers
#get a data frame of new data for the order
CanVals <- data.frame(logArea = LogAreaForFitting, Glacier = "Canada")
#get the predictions and standard error
CanPred <- predict(abun.model, newdata = CanVals, se.fit = TRUE)

#get a data frame of new data for the order
TayVals <- data.frame(logArea = LogAreaForFitting, Glacier = "Taylor")
#get the predictions and standard error
TayPred <- predict(abun.model, newdata = TayVals, se.fit = TRUE)

# plot the scatterplot of the data
pdf("../Outputs/Sommers2020/graphs/linear-model-interaction-logarea--glacier-logrichness-abundant.pdf", width=6, height=4) 
plot(logRichness ~ logArea, data = abun.diversity.data, col=as.factor(Glacier))
# add the predicted lines
lines(CanPred$fit ~ LogAreaForFitting, col = "black")
lines(CanPred$fit + CanPred$se.fit ~ LogAreaForFitting, col = "black", lty = 2)
lines(CanPred$fit - CanPred$se.fit ~ LogAreaForFitting, col = "black", lty = 2)
lines(TayPred$fit ~ LogAreaForFitting, col = "red")
lines(TayPred$fit + TayPred$se.fit ~ LogAreaForFitting, col = "red", lty = 2)
lines(TayPred$fit - TayPred$se.fit ~ LogAreaForFitting, col = "red", lty = 2)
dev.off()

###RARE TAXA 
#get the range of the data
rng <- range(rare.diversity.data$logArea)
#get a sequence from the min to the max with 100 equally spaced values:
LogAreaForFitting <- seq(rng[1], rng[2], length = 100)
print(LogAreaForFitting)

#we're going to use these with the model to predict values of richness at each of those hole areas for each of our glaciers
#get a data frame of new data for the order
CanVals <- data.frame(logArea = LogAreaForFitting, Glacier = "Canada")
#get the predictions and standard error
CanPred <- predict(rare.model, newdata = CanVals, se.fit = TRUE)

#get a data frame of new data for the order
TayVals <- data.frame(logArea = LogAreaForFitting, Glacier = "Taylor")
#get the predictions and standard error
TayPred <- predict(rare.model, newdata = TayVals, se.fit = TRUE)

# plot the scatterplot of the data
pdf("../Outputs/Sommers2020/graphs/linear-model-interaction-logarea--glacier-logrichness-rare.pdf", width=6, height=4) 
plot(logRichness ~ logArea, data = rare.diversity.data, col=as.factor(Glacier))
# add the predicted lines
lines(CanPred$fit ~ LogAreaForFitting, col = "black")
lines(CanPred$fit + CanPred$se.fit ~ LogAreaForFitting, col = "black", lty = 2)
lines(CanPred$fit - CanPred$se.fit ~ LogAreaForFitting, col = "black", lty = 2)
lines(TayPred$fit ~ LogAreaForFitting, col = "red")
lines(TayPred$fit + TayPred$se.fit ~ LogAreaForFitting, col = "red", lty = 2)
lines(TayPred$fit - TayPred$se.fit ~ LogAreaForFitting, col = "red", lty = 2)
dev.off()

```
