---
title: "Generalised Linear Models"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---

#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/19-glm.ipynb) 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

######I WILL BE USING FAKE DATA FOR THIS ANALYSIS FOR DEMONSTRATION PURPOSES#####

#This script will look at GENERALIZED LINEAR MODELS
#So far we have assumed that the response variable follows a normal (Gaussian) distribution. We sometimes transformed the data by taking a log so that it would approximate a normal distribution. However, sometimes this is not appropriate.

#Generalized linear model analysis accomodates response variables that have non-normal error distributions. 

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library(picante)
#library(gplots)
library(lattice)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Six: Prepare our data
```{r}
#get logs of the continuous variables we are interested in
my.data.sub$logChlo <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA <- log(my.data.sub$DNA_per_g)
my.data.sub$logOrg <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logArea <- log(my.data.sub$area)
my.data.sub$logRich <- log(my.data.sub$Richness)
my.data.sub$logWat <- log(my.data.sub$water_content)
```

#Step Seven: Add an addition factor for us to play with
```{r}
Substrate <- rep(c("Ice", "Snow"), 6) 
Substrate <- c(Substrate, rep("Cryoconite", 16))
my.data.sub$Substrate <- as.factor(Substrate)
```

#Step Seven: Let's start by looking at statistical distributions - Poisson Distribution
#Statistical distributions are important becuase when analysing data using GLMs we need to decide which statistical distribution best represents our dataset. 
#continuous numeric data (like bodyweight) are usually normally distributed
#count data (like number of species on an island) may be distributed according to the Poisson distribution
#The statistical distributions that GLM in R supports include: Binomial, Gaussian, Gamma, Inverse Gaussian, Poisson, Quasi, Quasi-Binomial and Quasi-Poisson
```{r}
#The Poisson Distribution
#This is appropriate for discrete, integer data
#We can generate data that follow this distribution using the below code...
par(mfrow=c(1,2)) 
hist(rpois(1000, lambda = 1),xlab = "Value", main="")
hist(rpois(1000, lambda = 100),xlab = "Value", main="")
#So this is appropriate for response data that only uses integer values (e.g. counts)
#This distribution can be right-skewedor bell shaped
```

#Step Eight: Bionomial Distribution
```{r}
#Bionomial distribution is appropriate for continuous data from repeated bionary (yes/no) outcomes
#rbinom(n, size, prob)
par(mfrow=c(1,2)) 
hist(rbinom(1000, 1, 0.5),xlab = "Value", main="") #single try of a 50/50 outcome, regenerated 1000 times (1000 flips of a coin) - each flip is a Bernoulli trial/experiment
hist(rbinom(1000, 100, 0.5),xlab = "Value", main="") #plots the number of successes if you tried something 100 times with a probability of 50, so the values hover around 50
#size = number of trials
#prob = probability of an outcome in each trial


```

#Step Nine: The Gamma Distribution
#Suitable for continuous, positive data
#The ratio of the variance and mean is constant so appropriate to model data where an increased mean values also increases the variance - useful for response variables that are times
```{r}
#rgamma(n, shape, rate) rate controls the scale of the distribution, over which values the distribution spreads
#shape parameter controls whether it is right-skewed or bell-shaped
par(mfrow=c(1,2)) 
hist(rgamma(1000,1,1),xlab = "Value", main = "")
hist(rgamma(1000,100,1),xlab = "Value", main = "")
```

#Step Ten: Exploring the data
```{r}
str(my.data.sub)
my.data.sub$Glacier <- as.factor(my.data.sub$Glacier)

#LEt's look at our data 
boxplot(Richness ~ Glacier, data = my.data.sub)

# #We can look at the minimum richnesses of the glaciers
# tapply(my.data.sub$Richness, my.data.sub$Glacier, min, na.rm = TRUE)
# 
# #and the maximum richnesses
# tapply(my.data.sub$Richness, my.data.sub$Glacier, max, na.rm = TRUE)

bwplot(logRich ~ Substrate | Glacier, data = my.data.sub)
#It looks like the substrates are doing different things on the two glaciers (FAKE NEWS) so perhaps the substrate and glacier are interacting to affect ASV richness
#The variances also look ver different

#We can also use barplots to look at the different means


tab <- tapply(my.data.sub$Richness, list(my.data.sub$Glacier, my.data.sub$Substrate), mean, na.rm=TRUE)

print(tab)

barplot(tab, beside=TRUE, log= 'y' )
```

#Step Eleven: Fit an ordinary linear model
```{r}
modLM <- lm(logRich ~ Substrate * Glacier, data=my.data.sub)
par(mfrow=c(2,2), mar=c(3,3,3,1), mgp=c(2,0.8,0))
plot(modLM)
#The normal q-q plot doesn't look great (it's not actually that bad but just pretend it's terrible) so it doesn't look like a log transformation would help our data.
```

#Step Twelve: Fit a Generalized Linear Model
#Some differences between continuous data and count data
#Counts can bever be negative but can be zero
#Counts are always integers rather than being continuous
#The variance of count data is not constant
#So if our data does not show constant normal variance we can't use a linear model
#note: we don't need logged data here
```{r}
modPois <- glm(Richness ~ Substrate * Glacier, data=my.data.sub, family= 'poisson')

summary(modPois)
#what is this table showing us?
#On Canada glacier, ice has the highest number of ASVs - because it had a positive difference with cryoconite (the intercept)
#The overall effect of taylor glacier is to decrease species richness (it has a negative coefficient) but then the positive interactions show an increase in richness for Taylor glacier with specific substrates
#There are two types of deviance in the model output. Deviance is a measure of goodness of fit in the GLM. Null deviance shows how well the response variable is predicted by a model that includes only the intercept (overall mean). We're looking for a significant reduction in deviation from the null deviance to the residual deviance. We see a significant decrease here so the model is significant.

```

#Step Thirteen: Let's look at overdispersion
```{r}
par(mfrow = c(2, 2), mar = c(3, 3, 3, 1), mgp = c(2, 0.8, 0))
plot(modPois)

#The q-q plot here actually doesn't look bad but just pretend it does for the purpose of this demo
#If our q-q plot had points way from the line and the residuals were massive we would have a problem of overdispersion - The Poisson distribution predicts that the variance at a point in the model is equal to the prediction — the mean count at that point.

#We can check the dispersion of the data using the summary table - ratio of the residual deviance to the residual degrees of freedom should be approximately 1.
residual.deviance = 295.76
residual.df = 22
residual.deviance/residual.df #= 13.44 this is supposed to be approximately 1 so our data ARE overdispersed!
#We can deal with this by using a different model...
```

#Step Fourteen: Fitting a GLM using a quasi-poisson distribution
#The quasi-poisson distribution uses the data to estimate the dispersion of the model
```{r}
modQPois <- glm(Richness ~ Substrate * Glacier, data = my.data.sub, family = 'quasipoisson')
summary(modQPois)
#We can see in this output that the dispersion parameter line has changed and the p values are less significant - this is the effect of controlling for overdispersion

par(mfrow = c(2, 2), mar = c(3, 3, 3, 1), mgp = c(2, 0.8, 0))
plot(modQPois) #the plot is better but not perfect. We can't improve this anymore so we accept the imperfections of the model

#ANOVA (technically an analysis of deviance) and F-test
anova(modQPois, test ="F") #the interaction is the only term we can drop from this model and it doesn't look significant. Let's try it!

modQPois2 <- update(modQPois, . ~ . - Substrate:Glacier)
anova(modQPois, modQPois2, test = "F") #p > 0.05 so this model is not worse!

anova(modQPois2, test ="F")
```

#Step Fifteen: Model predictions
# We can get model predictions and SEs using the predict function. This is done through a statistical method that I don't understand do don't worry about that for now. 
#These values are the same as those calculated for the bar plot. 
```{r}
# use expand.grid to get all combinations of factors
df <- expand.grid(Substrate = levels(my.data.sub$Substrate), Glacier = levels(my.data.sub$Glacier))

predict(modQPois, newdata = df, type = "response")
```

#Step Sixteen: Tips on reporting the model
```{r}
#Make sure the reader understands what has been done well enough to repeat the analysis
#Summarize and explain the results without pages of R output
#Think about what the data shows? Present a graph or a table to show the data you are about to model. Always include a figure or table legend and always refer to that figure or legend from the text.
#Have you transformed the data or used a subset? If so, why?
#What kind of model or statistical test have you used?
#With linear models, what is the response variable and what are the explanatory variables.
#Have you simplified the model and, if so, what was the most complex model you tried?
#How did you check the suitability of the model? Are there any problems with the model and, if so, what might cause them?
#If you summarise stats in text, you must include all the information about the test.
#For F tests, this is F, the two degrees of freedom and the p value. For example: 'There is a significant interaction between treatment and strain (F4,284=32.7,p<0.0001)'.
#For t tests, this is the coefficient, the standard error, t, the degrees of freedom and p. For example, 'Across strains, the main effect of nitroguanisine is to reduce colony counts relative to the control (estimate=-2.17, s.e= 0.51, t=−4.26, df=284, p<0.0001)'.
#With more complex models, it is common to present either the anova table or the coefficients table as a summary of the model output. Just include the tables from R output, not the information around it.
#Never just include chunks of raw output from R.
#Most importantly, you need to convey what the interpretation of the model is. What is it telling you about the data?
```