---
title: "Non-linear Least-squares Models"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/20-ModelFitting-NLLS.ipynb)

#Additional info on model fitting available above but not included here. Some great tips, particularly on finding start values and fitting one model to multiple datasets.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#This script will look at NON-LINEAR LEAST SQUARES MODEL FITTING
#How do we fit non-linear models to data using NLLS?
#We will be visualizing the data and the mathematical model
#Fitting a non-linear model
#Assessing the quality of the fit and whether the model is appropriate for the data

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library(picante)
#library(gplots)
#library(lattice)
library(ggplot2)
library(minpack.lm)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Six: Prepare our data
```{r}
#get logs of the continuous variables we are interested in
my.data.sub$logChlo <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA <- log(my.data.sub$DNA_per_g)
my.data.sub$logOrg <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logArea <- log(my.data.sub$area)
my.data.sub$logRich <- log(my.data.sub$Richness)
my.data.sub$logWat <- log(my.data.sub$water_content)
```

#Step Seven: Fitting the model
```{r}
#get our canada data
Data2Fit <- subset(my.data.sub, Glacier == "Canada")

#plot
plot(Data2Fit$area, Data2Fit$Richness, xlab = "Hole Area (cm2)", ylab="ASV Richness")

#or using ggplot
ggplot(Data2Fit, aes(x=area, y = Richness))+
  geom_point(size=(3), colour="red")+
  theme_bw()+
  labs(y="ASV Richness", x="Hole Area (cm2)")

#We can see that ASV richness does not increase proportionally with hole area. It increases quickly at small areas before levelling off.

#So we can fit the model to the data using NLLS
PowFit <- nlsLM(Richness ~ a * area^b, data = Data2Fit, start = list(a=.1, b=.1))
```

#Step Eight: Plot the model with the data 
```{r}
#get areas for plotting
Areas <- seq(min(Data2Fit$area), max(Data2Fit$area), len=200)

#get out coefficients from the model fitting
coef(PowFit)["a"] #intercept
coef(PowFit)["b"] #slope 

#create a function of our model
powMod <- function(x, a, b) {
 return(a * x^b)
}

#After defining the model function (this is the power law model by the way) we can run the nlls fitting like this
# PowFit <- nlsLM(Richness ~ powMod(area, a, b), data = Data2Fit, start=list(a=.1, b=.1))

#get out predicted ASV values for plotting
Predict2PlotPow <- powMod(Areas, coef(PowFit)["a"],coef(PowFit)["b"])

#plot with our observed data
plot(Data2Fit$area, Data2Fit$Richness)
lines(Areas, Predict2PlotPow, col="blue", lwd=2.5)

#get a summary of our model fitting
summary(PowFit)

#Commonly with data like this it is not necessary to use Nlls fitting. a log can be taken each side of the equation and the standard lm() fucntion can be used (this is ordinary least squares fitting)
#log(Richness) = log(a) + b*log(area)

model_lm <- lm(log(Richness) ~ log(area), data = Data2Fit)

summary(model_lm) #the intercept here = log(a)
#the slope below is not logged but multiplied to log(area) so log(richness) = log(a) + b*log(area)

```
