---
title: "Non-linear Least-squares Models"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/20-ModelFitting-NLLS.ipynb)

#Additional info on model fitting available above but not included here. Some great tips, particularly on finding start values and fitting one model to multiple datasets.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#This script will look at NON-LINEAR LEAST SQUARES MODEL FITTING
#How do we fit non-linear models to data using NLLS?
#We will be visualizing the data and the mathematical model
#Fitting a non-linear model
#Assessing the quality of the fit and whether the model is appropriate for the data

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library(picante)
#library(gplots)
#library(lattice)
library(ggplot2)
library(minpack.lm)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
# load("results/cleaned_files/cleanedfiles_3.RData")
# #hole area - calculate and add to dataframe
# #Method as per Darcy et al (2018)
# metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
# metadata$area <- metadata$area*10000
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
# #Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
# div_shan <- diversity(t(count_table_clean), index = "shannon")
# 
# #Use vegan package to get simpson diversity index of each sample
# div_simp <- diversity(t(count_table_clean), index = "simpson")
# 
# #Use vegan package to get inverted simpson diversity index of each sample
# div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")
# 
# #Faith's phylogenetic diversity 
# asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table
# 
# my_tree <- my.ps@phy_tree #get my phylogenetic tree
# 
# #Check if the tree is rooted
# my.ps@phy_tree #Rooted, includes branch lengths
# 
# #It is a rooted tree so...
# df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.
# 
# #Calculate Pilou evenness
# div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier
# 
# #bind alpha diversity indices into a dataframe
# diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
# colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
# my.data <- cbind(metadata, diversity)
# 
# #remove columns with data we're not interested in
# my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
# rownames(my.data.sub) <- my.data.sub$Run
# my.data.sub <- subset(my.data.sub, select = -c(1))
# my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
# my.data.sub$Glacier <- metadata$Glacier
# my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Six: Prepare our data
```{r}
# #get logs of the continuous variables we are interested in
# my.data.sub$logChlo <- log(my.data.sub$Chlorophyll)
# my.data.sub$logDepth <- log(my.data.sub$Depth)
# my.data.sub$logDNA <- log(my.data.sub$DNA_per_g)
# my.data.sub$logOrg <- log(my.data.sub$org_matter)
# my.data.sub$logpH <- log(my.data.sub$pH)
# my.data.sub$logArea <- log(my.data.sub$area)
# my.data.sub$logRich <- log(my.data.sub$Richness)
# my.data.sub$logWat <- log(my.data.sub$water_content)
```

#Step Seven: Fitting the model
```{r}
#get our canada data
Data2Fit <- subset(tot.diversity.data, Glacier == "Canada")

#plot
plot(Data2Fit$Area, Data2Fit$Richness, xlab = "Hole Area (cm2)", ylab="ASV Richness")

#or using ggplot
ggplot(Data2Fit, aes(x=Area, y = Richness))+
  geom_point(size=(3), colour="red")+
  theme_bw()+
  labs(y="ASV Richness", x="Hole Area (cm2)")

#We can see that ASV richness does not increase proportionally with hole area. It increases quickly at small areas before levelling off.

###TOTAL TAXA
#So we can fit the model to the data using NLLS
tot.PowFit <- nlsLM(Richness ~ a * Area^b, data = tot.diversity.data, start = list(a=.1, b=.1))

###ABUNDANT TAXA
#So we can fit the model to the data using NLLS
abun.PowFit <- nlsLM(Richness ~ a * Area^b, data = abun.diversity.data, start = list(a=.1, b=.1))

###RARE TAXA
#So we can fit the model to the data using NLLS
rare.PowFit <- nlsLM(Richness ~ a * Area^b, data = rare.diversity.data, start = list(a=.1, b=.1))
```

#Step Eight: Plot the model with the data 
```{r}
# #get areas for plotting
# Areas <- seq(min(Data2Fit$Area), max(Data2Fit$Area), len=200)
# 
# #get out coefficients from the model fitting
# coef(PowFit)["a"] #intercept
# coef(PowFit)["b"] #slope 
# 
# #create a function of our model
# powMod <- function(x, a, b) {
#  return(a * x^b)
# }
# 
# #After defining the model function (this is the power law model by the way) we can run the nlls fitting like this
# # PowFit <- nlsLM(Richness ~ powMod(area, a, b), data = Data2Fit, start=list(a=.1, b=.1))
# 
# #get out predicted ASV values for plotting
# Predict2PlotPow <- powMod(Areas, coef(PowFit)["a"],coef(PowFit)["b"])
# 
# #plot with our observed data
# plot(Data2Fit$Area, Data2Fit$Richness)
# lines(Areas, Predict2PlotPow, col="blue", lwd=2.5)
# 
# #get a summary of our model fitting
# summary(PowFit)
# 
# #Commonly with data like this it is not necessary to use Nlls fitting. a log can be taken each side of the equation and the standard lm() fucntion can be used (this is ordinary least squares fitting)
# #log(Richness) = log(a) + b*log(area)
# 
# model_lm <- lm(log(Richness) ~ log(Area), data = Data2Fit)
# 
# summary(model_lm) #the intercept here = log(a)
# #the slope below is not logged but multiplied to log(area) so log(richness) = log(a) + b*log(area)

###TOTAL TAXA
#get areas for plotting
Areas <- seq(min(tot.diversity.data$Area), max(tot.diversity.data$Area), len=200)

#get out coefficients from the model fitting
coef(tot.PowFit)["a"] #intercept
coef(tot.PowFit)["b"] #slope 

#create a function of our model
powMod <- function(x, a, b) {
 return(a * x^b)
}

#get out predicted ASV values for plotting
Predict2PlotPow <- powMod(Areas, coef(tot.PowFit)["a"],coef(tot.PowFit)["b"])

#plot with our observed data
pdf("../Outputs/Sommers2020/graphs/nlls-area-richness-total.pdf", width=6, height=4) 
plot(tot.diversity.data$Area, tot.diversity.data$Richness)
lines(Areas, Predict2PlotPow, col="blue", lwd=2.5)
dev.off()

#get a summary of our model fitting
summary(tot.PowFit)

###ABUNDANT TAXA
#get areas for plotting
Areas <- seq(min(abun.diversity.data$Area), max(abun.diversity.data$Area), len=200)

#get out coefficients from the model fitting
coef(abun.PowFit)["a"] #intercept
coef(abun.PowFit)["b"] #slope 

#get out predicted ASV values for plotting
Predict2PlotPow <- powMod(Areas, coef(abun.PowFit)["a"],coef(abun.PowFit)["b"])

#plot with our observed data
pdf("../Outputs/Sommers2020/graphs/nlls-area-richness-abundant.pdf", width=6, height=4) 
plot(abun.diversity.data$Area, abun.diversity.data$Richness)
lines(Areas, Predict2PlotPow, col="blue", lwd=2.5)
dev.off()

#get a summary of our model fitting
summary(abun.PowFit)

###RARE TAXA
#get areas for plotting
Areas <- seq(min(rare.diversity.data$Area), max(rare.diversity.data$Area), len=200)

#get out coefficients from the model fitting
coef(rare.PowFit)["a"] #intercept
coef(rare.PowFit)["b"] #slope 

#get out predicted ASV values for plotting
Predict2PlotPow <- powMod(Areas, coef(rare.PowFit)["a"],coef(rare.PowFit)["b"])

#plot with our observed data
pdf("../Outputs/Sommers2020/graphs/nlls-area-richness-rare.pdf", width=6, height=4) 
plot(rare.diversity.data$Area, rare.diversity.data$Richness)
lines(Areas, Predict2PlotPow, col="blue", lwd=2.5)
dev.off()

#get a summary of our model fitting
summary(rare.PowFit)


```
