---
title: "MVA Beta Analysis"
author: "Amy Solman"
date: "21/04/2021"
output: html_document
---
#Script developed from that of BArnett et al., (2020) code available at https://github.com/seb369/landuse_comm_assembly/blob/master/Diversity_measures.md 
#This script will calculate beta diversities and carry out ordination on the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

library(dplyr)
library(tibble)
library(vegan)
library(ggpubr)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
# load("results/cleaned_files/cleanedfiles_3.RData")
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```
#Step Four: Weighted UniFrac Ordination
```{r}
###TOTAL TAXA 
meta.data = data.frame(sample_data(my.ps.rar.rel))
set.seed(1961)
wuf.ord = ordinate(my.ps.rar.rel, "PCoA", distance="unifrac", weighted=TRUE)
Xaxis = paste("PCo1 (", round(wuf.ord$values[1,2]*100, digits=2), "%)", sep="")
Yaxis = paste("PCo2 (", round(wuf.ord$values[2,2]*100, digits=2), "%)", sep="")

wuf.ord.df = data.frame(wuf.ord$vectors) %>%
  rownames_to_column(var="Run") %>%
  select(Run, Axis.1, Axis.2)

wuf.ord.df = full_join(wuf.ord.df, meta.data, by ="Run")

ord.plot = ggplot(data=wuf.ord.df, aes(x=Axis.1, y=Axis.2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=2, alpha=.5)+
  labs(x=Xaxis, y=Yaxis, colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background=NULL)

#### Function for pulling out legends
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
###

ord.leg = g_legend(ord.plot)
ord.plot = ord.plot + theme(legend.position = "none")

ord.leg.plot = ord.plot + annotation_custom(ord.leg, xmin=0.015, ymin=0.01)

ord.leg.plot

###TOTAL TAXA - PERMANOVA analysis of weighted unifrac results to see if glacier significantly explains variation in community composition.
wuf.dist = phyloseq::distance(my.ps.rar.rel, "wunifrac")
set.seed(1961)
wuf.adon = adonis(wuf.dist ~ Glacier, data=meta.data)
wuf.adon

ord.leg.plot + annotate("text", label=paste("R2 = ", round(wuf.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(wuf.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.057, y=0.05, size=3)
ggsave("../Outputs/Sommers2020/graphs/weighted-unifrac-pcoa-total.pdf", width=8, height=4)

```


```{r}
###ABUNDANT TAXA 
#get the relative abundance of our abundant taxa 
my.ps.rar.rel.abun <- transform_sample_counts(my.ps.rar.abun, function(x) x/sum(x))

meta.data = data.frame(sample_data(my.ps.rar.rel.abun))
set.seed(1961)
wuf.ord = ordinate(my.ps.rar.rel.abun, "PCoA", distance="unifrac", weighted=TRUE)
Xaxis = paste("PCo1 (", round(wuf.ord$values[1,2]*100, digits=2), "%)", sep="")
Yaxis = paste("PCo2 (", round(wuf.ord$values[2,2]*100, digits=2), "%)", sep="")

wuf.ord.df = data.frame(wuf.ord$vectors) %>%
  rownames_to_column(var="Run") %>%
  select(Run, Axis.1, Axis.2)

wuf.ord.df = full_join(wuf.ord.df, meta.data, by ="Run")

ord.plot = ggplot(data=wuf.ord.df, aes(x=Axis.1, y=Axis.2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=2, alpha=.5)+
  labs(x=Xaxis, y=Yaxis, colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background=NULL)

#### Function for pulling out legends
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
###

ord.leg = g_legend(ord.plot)
ord.plot = ord.plot + theme(legend.position = "none")

ord.leg.plot.abun = ord.plot + annotation_custom(ord.leg, xmin=0.15, ymin=0.04)

ord.leg.plot.abun


###TOTAL TAXA - PERMANOVA analysis of weighted unifrac results to see if glacier significantly explains variation in community composition.
wuf.dist = phyloseq::distance(my.ps.rar.rel.abun, "wunifrac")
set.seed(1961)
wuf.adon = adonis(wuf.dist ~ Glacier, data=meta.data)
wuf.adon

ord.leg.plot.abun + annotate("text", label=paste("R2 = ", round(wuf.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(wuf.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.057, y=0.05, size=3)
ggsave("../Outputs/Sommers2020/graphs/weighted-unifrac-pcoa-abundant.pdf", width=8, height=4)

```

```{r}
###RARE TAXA 
#get the relative abundance of our abundant taxa 
my.ps.rar.rel.rare <- transform_sample_counts(my.ps.rar.rare, function(x) x/sum(x))

meta.data = data.frame(sample_data(my.ps.rar.rel.rare))
set.seed(1961)
wuf.ord = ordinate(my.ps.rar.rel.rare, "PCoA", distance="unifrac", weighted=TRUE)
Xaxis = paste("PCo1 (", round(wuf.ord$values[1,2]*100, digits=2), "%)", sep="")
Yaxis = paste("PCo2 (", round(wuf.ord$values[2,2]*100, digits=2), "%)", sep="")

wuf.ord.df = data.frame(wuf.ord$vectors) %>%
  rownames_to_column(var="Run") %>%
  select(Run, Axis.1, Axis.2)

wuf.ord.df = full_join(wuf.ord.df, meta.data, by ="Run")

ord.plot = ggplot(data=wuf.ord.df, aes(x=Axis.1, y=Axis.2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=2, alpha=.5)+
  labs(x=Xaxis, y=Yaxis, colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background=NULL)

#### Function for pulling out legends
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
###

ord.leg = g_legend(ord.plot)
ord.plot = ord.plot + theme(legend.position = "none")

ord.leg.plot.rare = ord.plot + annotation_custom(ord.leg, xmin=0.2, ymin=0.1)

ord.leg.plot.rare

###TOTAL RARE - PERMANOVA analysis of weighted unifrac results to see if glacier significantly explains variation in community composition.
wuf.dist = phyloseq::distance(my.ps.rar.rel.rare, "wunifrac")
set.seed(1961)
wuf.adon = adonis(wuf.dist ~ Glacier, data=meta.data)
wuf.adon

ord.leg.plot.rare + annotate("text", label=paste("R2 = ", round(wuf.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(wuf.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.057, y=0.05, size=3)
ggsave("../Outputs/Sommers2020/graphs/weighted-unifrac-pcoa-rare.pdf", width=8, height=4)

```

#Step Five: PERMANOVA analysis of weighted unifrac results to see if glacier significantly explains variation in community composition.
```{r}

# ###TOTAL TAXA - PERMANOVA analysis of weighted unifrac results to see if glacier significantly explains variation in community composition.
# wuf.dist = phyloseq::distance(my.ps.rar.rel, "wunifrac")
# set.seed(1961)
# wuf.adon = adonis(wuf.dist ~ Glacier, data=meta.data)
# wuf.adon
# 
# ord.leg.plot + annotate("text", label=paste("R2 = ", round(wuf.adon$aov.tab$R2[1],3),
#                                             "\npvalue = ", round(wuf.adon$aov.tab$`Pr(>F)`[1],3)), 
#                         x=-0.057, y=0.05, size=3)
```

#Step Six: Unweighted UniFrac Ordination for Total TAXA
```{r}
# set.seed(1961)
# uf.ord = ordinate(my.ps.rar.rel, "PCoA", distance="unifrac", weighted=FALSE)
# Xaxis = paste("PCo1 (", round(uf.ord$values[1,2]*100, digits=2), "%)", sep="")
# Yaxis = paste("PCo2 (", round(uf.ord$values[2,2]*100, digits=2), "%)", sep="")
# 
# uf.ord.df = data.frame(uf.ord$vectors) %>%
#   rownames_to_column(var="Run") %>%
#   select(Run, Axis.1, Axis.2)
# 
# uf.ord.df = full_join(uf.ord.df, meta.data, by ="Run")
# 
# ord.plot = ggplot(data=uf.ord.df, aes(x=Axis.1, y=Axis.2, colour=Glacier))+
#   geom_point(size=3)+
#   stat_ellipse(linetype=2, alpha=.5)+
#   labs(x=Xaxis, y=Yaxis, colour="Glacier")+
#   theme(text=element_text(size=16),
#         legend.background=NULL)
# 
# ord.leg = g_legend(ord.plot)
# ord.plot = ord.plot + theme(legend.position = "none")
# 
# ord.leg.plot = ord.plot + annotation_custom(ord.leg, xmin=0.2, ymin=0.1)
# 
# ord.leg.plot
# 
# ggsave("../Outputs/Sommers2020/graphs/unweighted-unifrac-pcoa.pdf", width=8, height=4)

###TOTAL TAXA 
meta.data = data.frame(sample_data(my.ps.rar.rel))
set.seed(1961)
uf.ord = ordinate(my.ps.rar.rel, "PCoA", distance="unifrac", weighted=FALSE)
Xaxis = paste("PCo1 (", round(uf.ord$values[1,2]*100, digits=2), "%)", sep="")
Yaxis = paste("PCo2 (", round(uf.ord$values[2,2]*100, digits=2), "%)", sep="")

uf.ord.df = data.frame(uf.ord$vectors) %>%
  rownames_to_column(var="Run") %>%
  select(Run, Axis.1, Axis.2)

uf.ord.df = full_join(uf.ord.df, meta.data, by ="Run")

ord.plot = ggplot(data=uf.ord.df, aes(x=Axis.1, y=Axis.2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=2, alpha=.5)+
  labs(x=Xaxis, y=Yaxis, colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background=NULL)

#### Function for pulling out legends
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
###

ord.leg = g_legend(ord.plot)
ord.plot = ord.plot + theme(legend.position = "none")

ord.leg.plot = ord.plot + annotation_custom(ord.leg, xmin=0.015, ymin=0.01)

ord.leg.plot

###TOTAL TAXA - PERMANOVA analysis of weighted unifrac results to see if glacier significantly explains variation in community composition.
uf.dist = phyloseq::distance(my.ps.rar.rel, "unifrac")
set.seed(1961)
uf.adon = adonis(uf.dist ~ Glacier, data=meta.data)
uf.adon

ord.leg.plot + annotate("text", label=paste("R2 = ", round(uf.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(uf.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.057, y=0.05, size=3)
ggsave("../Outputs/Sommers2020/graphs/unweighted-unifrac-pcoa-total.pdf", width=8, height=4)
```

#Step Six: Unweighted UniFrac Ordination for ABUNDANT TATA
```{r}
###ABUNDANT TAXA 
meta.data = data.frame(sample_data(my.ps.rar.rel.abun))
set.seed(1961)
uf.ord = ordinate(my.ps.rar.rel.abun, "PCoA", distance="unifrac", weighted=FALSE)
Xaxis = paste("PCo1 (", round(uf.ord$values[1,2]*100, digits=2), "%)", sep="")
Yaxis = paste("PCo2 (", round(uf.ord$values[2,2]*100, digits=2), "%)", sep="")

uf.ord.df = data.frame(uf.ord$vectors) %>%
  rownames_to_column(var="Run") %>%
  select(Run, Axis.1, Axis.2)

uf.ord.df = full_join(uf.ord.df, meta.data, by ="Run")

ord.plot = ggplot(data=uf.ord.df, aes(x=Axis.1, y=Axis.2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=2, alpha=.5)+
  labs(x=Xaxis, y=Yaxis, colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background=NULL)

#### Function for pulling out legends
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
###

ord.leg = g_legend(ord.plot)
ord.plot = ord.plot + theme(legend.position = "none")

ord.leg.plot.abun = ord.plot + annotation_custom(ord.leg, xmin=0.015, ymin=0.01)

ord.leg.plot.abun

###TOTAL TAXA - PERMANOVA analysis of unweighted unifrac results to see if glacier significantly explains variation in community composition.
uf.dist = phyloseq::distance(my.ps.rar.rel.abun, "unifrac")
set.seed(1961)
uf.adon = adonis(uf.dist ~ Glacier, data=meta.data)
uf.adon

ord.leg.plot.abun + annotate("text", label=paste("R2 = ", round(uf.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(uf.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.057, y=0.05, size=3)
ggsave("../Outputs/Sommers2020/graphs/unweighted-unifrac-pcoa-abundant.pdf", width=8, height=4)
```

#Step Six: Unweighted UniFrac Ordination for RARE TATA
```{r}
###RARE TAXA 
meta.data = data.frame(sample_data(my.ps.rar.rel.rare))
set.seed(1961)
uf.ord = ordinate(my.ps.rar.rel.rare, "PCoA", distance="unifrac", weighted=FALSE)
Xaxis = paste("PCo1 (", round(uf.ord$values[1,2]*100, digits=2), "%)", sep="")
Yaxis = paste("PCo2 (", round(uf.ord$values[2,2]*100, digits=2), "%)", sep="")

uf.ord.df = data.frame(uf.ord$vectors) %>%
  rownames_to_column(var="Run") %>%
  select(Run, Axis.1, Axis.2)

uf.ord.df = full_join(uf.ord.df, meta.data, by ="Run")

ord.plot = ggplot(data=uf.ord.df, aes(x=Axis.1, y=Axis.2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=2, alpha=.5)+
  labs(x=Xaxis, y=Yaxis, colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background=NULL)

#### Function for pulling out legends
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
###

ord.leg = g_legend(ord.plot)
ord.plot = ord.plot + theme(legend.position = "none")

ord.leg.plot.rare = ord.plot + annotation_custom(ord.leg, xmin=0.015, ymin=0.01)

ord.leg.plot.rare

###TOTAL TAXA - PERMANOVA analysis of unweighted unifrac results to see if glacier significantly explains variation in community composition.
uf.dist = phyloseq::distance(my.ps.rar.rel.rare, "unifrac")
set.seed(1961)
uf.adon = adonis(uf.dist ~ Glacier, data=meta.data)
uf.adon

ord.leg.plot.rare + annotate("text", label=paste("R2 = ", round(uf.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(uf.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.057, y=0.05, size=3)
ggsave("../Outputs/Sommers2020/graphs/unweighted-unifrac-pcoa-rare.pdf", width=8, height=4)
```

#Step Seven: PERMANOVA analysis of unweighted unifrac results to see if glacier significantly explains variation in community composition.
```{r}
# uf.dist = phyloseq::distance(my.ps.rar.rel, "unifrac")
# set.seed(1961)
# uf.adon = adonis(uf.dist ~ Glacier, data=meta.data)
# uf.adon
# 
# ord.leg.plot + annotate("text", label=paste("R2 = ", round(uf.adon$aov.tab$R2[1],3),
#                                             "\npvalue = ", round(uf.adon$aov.tab$`Pr(>F)`[1],3)), 
#                         x=-0.057, y=0.05, size=3)
```

#Step Eight:Bray-Curtis Ordination TOTAL TAXA
```{r}
#Extract ASV table from phyloseq object
asv.table <- t(otu_table(my.ps.rar.rel))

#make the distance matrix
bc.dist <- vegdist(asv.table, method="bray", binary=FALSE)

#Scale with NMDS and plot
set.seed(1961)
bc.ord = metaMDS(bc.dist)
bc.ord.df = data.frame(bc.ord$points) %>%
  rownames_to_column(var="Run") %>%
  full_join(meta.data)

ord.plot <- ggplot(data=bc.ord.df, aes(x=MDS1, y=MDS2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=1, alpha=.5)+
  labs(x="NMDS1", y="NMDS2", colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background = NULL)

ord.leg <- g_legend(ord.plot)
ord.plot <- ord.plot + theme(legend.position = "none")

ord.leg.plot <- ord.plot + annotation_custom(ord.leg, xmin=0.45, ymin=-0.8)
ord.leg.plot
ggsave("../Outputs/Sommers2020/graphs/bray-curtis-nmds-plot-total.pdf", width=8, height=4)

#Step Nine: PERMANOVA analysis of bray-curtis results to see if glacier significantly explains variation in community composition.
set.seed(1961)
bc.adon = adonis(bc.dist ~ Glacier, meta.data)
bc.adon
ord.leg.plot + annotate("text", label=paste("R2 = ", round(bc.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(bc.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.4, y=-0.3, size=5)
```

#Step Eight:Bray-Curtis Ordination ABUNDANT TAXA
```{r}
#Extract ASV table from phyloseq object
asv.table <- t(otu_table(my.ps.rar.rel.abun))

#make the distance matrix
bc.dist <- vegdist(asv.table, method="bray", binary=FALSE)

#Scale with NMDS and plot
set.seed(1961)
bc.ord = metaMDS(bc.dist)
bc.ord.df = data.frame(bc.ord$points) %>%
  rownames_to_column(var="Run") %>%
  full_join(meta.data)

ord.plot <- ggplot(data=bc.ord.df, aes(x=MDS1, y=MDS2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=1, alpha=.5)+
  labs(x="NMDS1", y="NMDS2", colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background = NULL)

ord.leg <- g_legend(ord.plot)
ord.plot <- ord.plot + theme(legend.position = "none")

ord.leg.plot.abun <- ord.plot + annotation_custom(ord.leg, xmin=0.45, ymin=-0.8)
ord.leg.plot.abun
ggsave("../Outputs/Sommers2020/graphs/bray-curtis-nmds-plot-abundant.pdf", width=8, height=4)

#Step Nine: PERMANOVA analysis of bray-curtis results to see if glacier significantly explains variation in community composition.
set.seed(1961)
bc.adon = adonis(bc.dist ~ Glacier, meta.data)
bc.adon
ord.leg.plot.abun + annotate("text", label=paste("R2 = ", round(bc.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(bc.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.4, y=-0.3, size=5)
```

#Step Eight:Bray-Curtis Ordination RARE TAXA
```{r}
#Extract ASV table from phyloseq object
asv.table <- t(otu_table(my.ps.rar.rel.rare))

#make the distance matrix
bc.dist <- vegdist(asv.table, method="bray", binary=FALSE)

#Scale with NMDS and plot
set.seed(1961)
bc.ord = metaMDS(bc.dist)
bc.ord.df = data.frame(bc.ord$points) %>%
  rownames_to_column(var="Run") %>%
  full_join(meta.data)

ord.plot <- ggplot(data=bc.ord.df, aes(x=MDS1, y=MDS2, colour=Glacier))+
  geom_point(size=3)+
  stat_ellipse(linetype=1, alpha=.5)+
  labs(x="NMDS1", y="NMDS2", colour="Glacier")+
  theme(text=element_text(size=16),
        legend.background = NULL)

ord.leg <- g_legend(ord.plot)
ord.plot <- ord.plot + theme(legend.position = "none")

ord.leg.plot.rare <- ord.plot + annotation_custom(ord.leg, xmin=0.45, ymin=-0.8)
ord.leg.plot.rare
ggsave("../Outputs/Sommers2020/graphs/bray-curtis-nmds-plot-rare.pdf", width=8, height=4)

#Step Nine: PERMANOVA analysis of bray-curtis results to see if glacier significantly explains variation in community composition.
set.seed(1961)
bc.adon = adonis(bc.dist ~ Glacier, meta.data)
bc.adon
ord.leg.plot.rare + annotate("text", label=paste("R2 = ", round(bc.adon$aov.tab$R2[1],3),
                                            "\npvalue = ", round(bc.adon$aov.tab$`Pr(>F)`[1],3)), 
                        x=-0.4, y=-0.3, size=5)
```

#Save our two new phyloseq objects
```{r}
saveRDS(my.ps.rar.rel.abun, "../Outputs/Sommers2020/my-ps-rarefied-relative-abundant.rds")
saveRDS(my.ps.rar.rel.rare, "../Outputs/Sommers2020/my-ps-rarefied-relative-rare.rds")
```
