---
title: "MVA Beta Ordination Enviro Variables"
author: "Amy Solman"
date: "21/04/2021"
output: html_document
---
#This script will look at the impact of environmental variables on community composition 
#Script developed from that of Barnett et al., (2020) code available at https://github.com/seb369/landuse_comm_assembly/
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

library(dplyr)
library(tibble)
library(ggpubr)
library(phyloseq)
library(ape)
library(vegan)
library(ggrepel)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Get your data together for analysis
```{r}
#Glacier shapes
Glacier.shapes = c("Canada" = 15, "Taylor" = 16)

#which enviro variables am I going to look at:
#hole area
#pH
#Chlorophyll
#Organic matter content
#Water content
#DNA per gram

#Get the glacier metadata
meta.data = data.frame(sample_data(my.ps.rar.rel), stringsAsFactors = F)
#meta.data$area <- hole_area_cm2
meta.data$pH <- as.numeric(meta.data$pH)
meta.data$DNA_per_g <- as.numeric(meta.data$DNA_per_g)

#%>%
  # rename(soil_temp = in_situ_soil_temp,
  #        percent_organic_matter = organic_content_perc,
  #        sand = sand__perc, silt = silt__perc, clay = clay__perc) %>%
  # mutate(C.N_ratio = percent_C/percent_N,
  #        location = gsub("_", " ", location))

# ecosystem.conv = data.frame(ecosystem = c("agriculture", "meadow", "forest"), ecosystem2 = c("Cropland", "Old-field", "Forest"))
# metadata = left_join(metadata, ecosystem.conv)
# metadata$ecosystem2 = factor(metadata$ecosystem2, levels = c("Cropland", "Old-field", "Forest"))
# 
# metadata$ecosystem = factor(metadata$ecosystem, levels = c("agriculture", "meadow", "forest"))
```

#Step Five: Compare cryoconite hole area between the glaciers
```{r}
#First get our summary statistics
summary.stats.area <- group_by(meta.data, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(area, na.rm = TRUE),
    sd = sd(area, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot.area <- ggboxplot(meta.data, x = "Glacier", y = "area",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Cryoconite Hole Area (cm2)", xlab = "Glacier")

my.plot.area

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(meta.data, shapiro.test(area[Glacier == "Canada"])) #p=0.003 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NON-PARAMETRIC tests.
with(meta.data, shapiro.test(area[Glacier == "Taylor"])) #p=0.0006 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NONPARAMETRIC tests.

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
my.wilcoxon.test <- wilcox.test(area ~ Glacier, data = meta.data, exact = FALSE)
my.wilcoxon.test # p = 0.7966 so p > 0.05, there is no significant difference between cryoconite hole areas of the two glaciers.
```

#Step Six: Plot area data
```{r}
area.plot = ggplot(data=meta.data, aes(x=Glacier, y=area))+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(shape = Glacier, colour=Glacier), width = 0.2, alpha=0.3, size=2) +
  scale_shape_manual(values = Glacier.shapes) +
  labs(x="Glacier", y="Cryoconite Hole Area (cm2)", title="Area") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none")

area.plot
ggsave("../Outputs/Sommers2020/graphs/cryoconite-area-glacier.pdf", width=8, height=4)
```

#Step Seven: pH between the glaciers
```{r}
#First get our summary statistics
summary.stats.pH <- group_by(meta.data, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(pH, na.rm = TRUE),
    sd = sd(pH, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot.pH <- ggboxplot(meta.data, x = "Glacier", y = "pH",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="pH", xlab = "Glacier")

my.plot.pH

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(meta.data, shapiro.test(pH[Glacier == "Canada"])) #p=0.378 (p > 0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.
with(meta.data, shapiro.test(pH[Glacier == "Taylor"])) #p=0.7411 (p > 0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES?
#Use F-test to test for homogeneity in variances
f.test <- var.test(pH ~ Glacier, data = meta.data)
f.test #p value = 0.03839 which is less than 0.05 so there is a significant difference in the variances of the two datasets. Because the variances of the two groups are unequal we must use Welch's t-test (just omit the var.equal = TRUE bit)

#WELCH'S T-TEST
my.t.test <- t.test(pH ~ Glacier, data = meta.data)
my.t.test #p = 0.0785 P< 0.05 so there is a significant difference in the pH of the two glaciers
```

#Step Eight: Plot pH data
```{r}
pH.plot = ggplot(data=meta.data, aes(x=Glacier, y=pH))+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(shape = Glacier, colour=Glacier), width = 0.2, alpha=0.3, size=2) +
  scale_shape_manual(values = Glacier.shapes) +
  labs(x="Glacier", y="pH", title="pH") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none")

pH.plot
ggsave("../Outputs/Sommers2020/graphs/pH-glacier.pdf", width=8, height=4)
```

#Step Nine: Compare chlorophyll between the glaciers
```{r}
#First get our summary statistics
summary.stats.cho <- group_by(meta.data, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(Chlorophyll, na.rm = TRUE),
    sd = sd(Chlorophyll, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot.cho <- ggboxplot(meta.data, x = "Glacier", y = "Chlorophyll",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Chlorophyll", xlab = "Glacier")

my.plot.cho

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(meta.data, shapiro.test(Chlorophyll[Glacier == "Canada"])) #p=0.0006 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NON-PARAMETRIC tests.
with(meta.data, shapiro.test(Chlorophyll[Glacier == "Taylor"])) #p=0.004 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NONPARAMETRIC tests.

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
my.wilcoxon.test <- wilcox.test(Chlorophyll ~ Glacier, data = meta.data, exact = FALSE)
my.wilcoxon.test # p = 0.141 so p > 0.05, there is no significant difference between cryoconite hole areas of the two glaciers.
```

#Step Ten: Plot chlorophyll data
```{r}
cho.plot = ggplot(data=meta.data, aes(x=Glacier, y=Chlorophyll))+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(shape = Glacier, colour=Glacier), width = 0.2, alpha=0.3, size=2) +
  scale_shape_manual(values = Glacier.shapes) +
  labs(x="Glacier", y="Chlorophyll", title="Chlorophyll") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none")

cho.plot
ggsave("../Outputs/Sommers2020/graphs/chlorophyll-glacier.pdf", width=8, height=4)
```

#Step Eleven: Compare organic matter content between the glaciers
```{r}
#First get our summary statistics
summary.stats.org <- group_by(meta.data, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(org_matter, na.rm = TRUE),
    sd = sd(org_matter, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot.org <- ggboxplot(meta.data, x = "Glacier", y = "org_matter",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Organic Matter Content", xlab = "Glacier")

my.plot.org

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(meta.data, shapiro.test(org_matter[Glacier == "Canada"])) #p=0.03 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NON-PARAMETRIC tests.
with(meta.data, shapiro.test(org_matter[Glacier == "Taylor"])) #p=0.7224 (p > 0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
my.wilcoxon.test <- wilcox.test(org_matter ~ Glacier, data = meta.data, exact = FALSE)
my.wilcoxon.test # p = 0.000000838 so p < 0.05, there is a significant difference between organic matter content of the two glaciers.
```

#Step Twelve: Plot organic matter content data
```{r}
org.plot = ggplot(data=meta.data, aes(x=Glacier, y=org_matter))+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(shape = Glacier, colour=Glacier), width = 0.2, alpha=0.3, size=2) +
  scale_shape_manual(values = Glacier.shapes) +
  labs(x="Glacier", y="Organic Matter Content", title="Org Matter") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none")

org.plot
ggsave("../Outputs/Sommers2020/graphs/organic-matter-glacier.pdf", width=8, height=4)
```

#Step Thirteen: Compare water content between the glaciers
```{r}
#First get our summary statistics
summary.stats.water <- group_by(meta.data, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(water_content, na.rm = TRUE),
    sd = sd(water_content, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot.water <- ggboxplot(meta.data, x = "Glacier", y = "water_content",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Water Content", xlab = "Glacier")

my.plot.water

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(meta.data, shapiro.test(water_content[Glacier == "Canada"])) #p=0.957 (p > 0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.
with(meta.data, shapiro.test(water_content[Glacier == "Taylor"])) #p=0.00005 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NONPARAMETRIC tests.

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
my.wilcoxon.test <- wilcox.test(water_content ~ Glacier, data = meta.data, exact = FALSE)
my.wilcoxon.test # p = 0.000004 so p < 0.05, there is a significant difference between water content of the two glaciers.
```

#Step Fourteen: Plot water content data
```{r}
water.plot = ggplot(data=meta.data, aes(x=Glacier, y=water_content))+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(shape = Glacier, colour=Glacier), width = 0.2, alpha=0.3, size=2) +
  scale_shape_manual(values = Glacier.shapes) +
  labs(x="Glacier", y="Water Content", title="Water") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none")

water.plot
ggsave("../Outputs/Sommers2020/graphs/water-content-glacier.pdf", width=8, height=4)
```

#Step Fifteen: Compare DNA per gram cryoconite between the glaciers
```{r}
#First get our summary statistics
summary.stats.dna <- group_by(meta.data, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(DNA_per_g, na.rm = TRUE),
    sd = sd(DNA_per_g, na.rm = TRUE)
  )

#Then plot our glaciers
my.plot.dna <- ggboxplot(meta.data, x = "Glacier", y = "DNA_per_g",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="DNA per gram", xlab = "Glacier")

my.plot.dna

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(meta.data, shapiro.test(DNA_per_g[Glacier == "Canada"])) #p=0.002 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NONPARAMETRIC tests.
with(meta.data, shapiro.test(DNA_per_g[Glacier == "Taylor"])) #p=0.00002 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NONPARAMETRIC tests.

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
my.wilcoxon.test <- wilcox.test(DNA_per_g ~ Glacier, data = meta.data, exact = FALSE)
my.wilcoxon.test # p = 0.7966 so p > 0.05, there is no significant difference between cryoconite hole areas of the two glaciers.
```

#Step Sixteen: Plot dna data
```{r}
dna.plot = ggplot(data=meta.data, aes(x=Glacier, y=DNA_per_g))+
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(shape = Glacier, colour=Glacier), width = 0.2, alpha=0.3, size=2) +
  scale_shape_manual(values = Glacier.shapes) +
  labs(x="Glacier", y="DNA per gram cryoconite", title="DNA") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.text = element_text(size=12),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none")

dna.plot
ggsave("../Outputs/Sommers2020/graphs/dna-glacier.pdf", width=8, height=4)
```

#Step Seventeen: Plot all together and save
```{r}
# Now plot all together
glacier.properties.plot = cowplot::plot_grid(area.plot, cho.plot, dna.plot, org.plot, pH.plot, water.plot, ncol = 6, rel_widths = c(1,1,1,1,1,1))
glacier.properties.plot
ggsave("../Outputs/Sommers2020/graphs/all-enviro-var-plots.pdf", width=8, height=4)

```

#Step Eighteen: Use canonical analysis of principal coordinates (CAP) to explore how differences in community composition are explained by environmental variables - for Canada Glacier ONLY
```{r}
#Run CAP analysis for Canada Glacier
my.ps.can = subset_samples(my.ps.rar.rel, Glacier == "Canada")
  
  meta = data.frame(sample_data(my.ps.can), stringsAsFactors = F) 
  #meta$area <- meta.data$area[meta.data$Glacier == "Canada"]
  
  rownames(meta) = meta$Run
  sample_data(my.ps.can) = sample_data(meta)
  
  # Extract which habitat you are looking at
  glacier = unique(as.character(meta$Glacier))[1]
  
  writeLines(paste("Running", glacier, "\n"))
  
    #remove samples with missing data
 my.ps.can = subset_samples(my.ps.can, org_matter != "NA")
 my.ps.can = subset_samples(my.ps.can, water_content != "NA")
  my.ps.can <- subset_samples(my.ps.can, Chlorophyll != "NA")
  
  # Get weighted unifrac distance matrix
  wuf.dist = phyloseq::distance(my.ps.can, "wunifrac")
  

  # Ordinate
  set.seed(1961)
  cap_ord.full <- ordinate(physeq = my.ps.can, method = "CAP", distance = wuf.dist, 
                      formula = ~ DNA_per_g + pH + area + Chlorophyll + org_matter + water_content) 
  
  set.seed(1961)
  cap_ord.null <- ordinate(physeq = my.ps.can, method = "CAP", distance = wuf.dist, formula = ~ 1)
  
  # Model selection to get just significant variables
  set.seed(1961)
  ordistep.res = ordistep(cap_ord.null, scope = formula(cap_ord.full), perm.max = 1000, trace=F)
  
  goodform = ordistep.res$call$formula
  
  set.seed(1961)
  cap_ord <- ordinate(physeq = my.ps.can, method = "CAP", distance = wuf.dist, formula = goodform)
  
  # CAP plot
  cap_plot <- plot_ordination(physeq = my.ps.can, ordination = cap_ord, 
                              color = "Glacier", axes = c(1,2))
  
  # Now add the environmental variables as arrows
  arrowmat <- vegan::scores(cap_ord, display = "bp")
  
  # Add labels, make a data.frame
  arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat) %>%
    mutate(labels = gsub("\\.", ":", labels))
  colnames(arrowdf) = c("labels", "xend", "yend")
  
  # Define the arrow aesthetic mapping
  arrow_map <- aes(xend = xend, yend = yend, x = 0, y = 0, 
                   color = NULL)
  
  label_map <- aes(x = 1.1 * xend, y = 1.1 * yend, 
                   color = NULL, label = labels)
  
  arrowhead = arrow(length = unit(0.02, "npc"))
  
  # Make a new graphic
  cap_plot_1 = cap_plot + 
    geom_segment(mapping = arrow_map, size = .5, data = arrowdf, color = "red", arrow = arrowhead) + 
    geom_text_repel(mapping = label_map, size = 4, data = arrowdf, color = "red", show.legend = FALSE) +
    ggtitle("Canada")
  print(cap_plot_1)
  ggsave("../Outputs/Sommers2020/graphs/CAP-canada.pdf", width=8, height=4)
  
  print(anova(cap_ord))
  
  writeLines(paste("\nTested environmental variabibles explain", round(100*RsquareAdj(cap_ord)$r.squared, 3), "% of the variation"))
  writeLines("\n--------------------\n")

```

#Step Nineteen: Use canonical analysis of principal coordinates (CAP) to explore how differences in community composition are explained by environmental variables - for Taylor Glacier ONLY
```{r}
#Run CAP analysis for Canada Glacier
my.ps.can = subset_samples(my.ps.rar.rel, Glacier == "Taylor")
  
  meta = data.frame(sample_data(my.ps.can), stringsAsFactors = F) 
  #meta$area <- meta.data$area[meta.data$Glacier == "Canada"]
  
  rownames(meta) = meta$Run
  sample_data(my.ps.can) = sample_data(meta)
  
  # Extract which habitat you are looking at
  glacier = unique(as.character(meta$Glacier))[1]
  
  writeLines(paste("Running", glacier, "\n"))
  
    #remove samples with missing data
 my.ps.can = subset_samples(my.ps.can, org_matter != "NA")
 my.ps.can = subset_samples(my.ps.can, water_content != "NA")
  my.ps.can <- subset_samples(my.ps.can, Chlorophyll != "NA")
  
  # Get weighted unifrac distance matrix
  wuf.dist = phyloseq::distance(my.ps.can, "wunifrac")
  

  # Ordinate
  set.seed(1961)
  cap_ord.full <- ordinate(physeq = my.ps.can, method = "CAP", distance = wuf.dist, 
                      formula = ~ DNA_per_g + pH + area + Chlorophyll + org_matter + water_content) 
  
  set.seed(1961)
  cap_ord.null <- ordinate(physeq = my.ps.can, method = "CAP", distance = wuf.dist, formula = ~ 1)
  
  # Model selection to get just significant variables
  set.seed(1961)
  ordistep.res = ordistep(cap_ord.null, scope = formula(cap_ord.full), perm.max = 1000, trace=F)
  
  goodform = ordistep.res$call$formula
  
  set.seed(1961)
  cap_ord <- ordinate(physeq = my.ps.can, method = "CAP", distance = wuf.dist, formula = goodform)
  
  # CAP plot
  cap_plot <- plot_ordination(physeq = my.ps.can, ordination = cap_ord, 
                              color = "Glacier", axes = c(1,2))
  
  # Now add the environmental variables as arrows
  arrowmat <- vegan::scores(cap_ord, display = "bp")
  
  # Add labels, make a data.frame
  arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat) %>%
    mutate(labels = gsub("\\.", ":", labels))
  colnames(arrowdf) = c("labels", "xend", "yend")
  
  # Define the arrow aesthetic mapping
  arrow_map <- aes(xend = xend, yend = yend, x = 0, y = 0, 
                   color = NULL)
  
  label_map <- aes(x = 1.1 * xend, y = 1.1 * yend, 
                   color = NULL, label = labels)
  
  arrowhead = arrow(length = unit(0.02, "npc"))
  
  # Make a new graphic
  cap_plot_2 = cap_plot + 
    geom_segment(mapping = arrow_map, size = .5, data = arrowdf, color = "red", arrow = arrowhead) + 
    geom_text_repel(mapping = label_map, size = 4, data = arrowdf, color = "red", show.legend = FALSE) +
    ggtitle("Taylor")
  print(cap_plot_2)
  ggsave("../Outputs/Sommers2020/graphs/CAP-talor.pdf", width=8, height=4)
  
  print(anova(cap_ord))
  
  writeLines(paste("\nTested environmental variabibles explain", round(100*RsquareAdj(cap_ord)$r.squared, 3), "% of the variation"))
  writeLines("\n--------------------\n")

```


#Step Twenty: Put our CAP plots together
```{r}
ag.cap.plot.nl = cap_plot_1 + theme(legend.position="none")
f.cap.plot.nl = cap_plot_2 + theme(legend.position="none")
cowplot::plot_grid(ag.cap.plot.nl, f.cap.plot.nl)
ggsave("../Outputs/Sommers2020/graphs/CAP-all.pdf", width=8, height=4)
```