---
title: "DDR"
author: "Amy Solman"
date: "03/03/2021"
output: html_document
---
THIS SCRIPT BUILDS ON THE CODE FROM LOCEY ET AL., (2020) DORMANCY DAMPENS THE DISTANCE DECAY RELATIONSHIP (https://github.com/LennonLab/DormancyDecay)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First clean the working environment, load packages and data
```{r}
rm(list=ls())
graphics.off()

#install.packages('simba', dependencies = TRUE)
require("simba")
#install.packages("fossil")
require("fossil")
require("dplyr")
#install.packages("picante")
require("picante")

load("results/cleaned_files/cleanedfiles.RData")
sample_info_tab <- read_tsv("results/cleaned_files/SraRunTable-mod.tsv")
```

What I need first of all is a matrix with OTUs as columns and samples as rows with absolute abundance data

```{r}
#generate generic ASV names
name_my_rows <- NULL
for (i in 1:nrow(count_table_clean)){
  name_my_rows <- c(name_my_rows, paste0("ASV_", i))
}

#Apply to count_table_no_contam row names
row.names(count_table_clean) <- name_my_rows

#Change sample names to something more useful
colnames(count_table_clean) <- sample_info_tab$Sample.Name

#Transpose the data 
cryo = t(count_table_clean)
```

Remove any samples that don't have corresponding coordinates for this analysis
```{r}
#Remove all rows of data with NA in them
env_info <- sample_info_tab[complete.cases(sample_info_tab$lat_lon),]

#Remove samples that don't have corresponding environmental data so we know all the samples we're using for our analysis, we have corresponding geographic distances
cryo <- cryo[rownames(cryo)  %in% env_info$Sample.Name, ]

#Remove any taxa that don't occur in any remaining samples
cryo <- cryo[ , colSums(cryo) > 0]

#Turn the absolute abundance data into relative abundance data
cryo <- sweep(cryo, 1, rowSums(cryo), '/')

#Check that this has worked
sum(cryo[1,]) #1 > gives the sum of the first row which should = 1
```
Now let's get our geographic distances
###THIS ANALYSIS WON'T WORK FOR THIS DATASET BECAUSE THERE IS NO APPRECIATEBALE DISTANCE BETWEEN CRYOCONITE HOLES
I will finish writing the script anyway for use with the next dataset/total datasets
```{r}

#First pull the latitude and longitude coordinates from the strong data in env_info
#I put a minus at the front of the latitude because it's in the southern hemisphere
latitude <- -(as.numeric(substr(env_info$lat_lon, 1, 8)))
longitude <- as.numeric(substr(env_info$lat_lon, 12, 21))
long.lat <- as.matrix(cbind(longitude, latitude))

#Let's alter the long.lat  distances for the purposes of this analysis 
# num <- seq(from=1, to=33)/100
# new_latitude <- -(as.numeric(substr(env_info$lat_lon, 1, 8)))+num
# new_longitude <- as.numeric(substr(env_info$lat_lon, 12, 21))+num
# long.lat <- as.matrix(cbind(new_longitude, new_latitude))

#Create a distance matrix between a list of points, how far is each point from every other point
geo.dist <- earth.dist(long.lat, dist = TRUE) 

 #Distances relatavized to values between 0 and 1 by dividing each pairwise distance by the maximum distance
geo.dist <- geo.dist/max(geo.dist)
geo.dist[which(!is.finite(geo.dist))] = NA #what does this do?
```

Now we can workout the community dissimilarities for comparison with our geographic distances!
```{r}
#workout the taxonomic distances between the cryoconite holes using Bray-Curtis similarity, Sørensen metric and Canberra similarity metric
# Taxonomic Distances Among Ponds
cryo.bray.dist <- 1 - vegdist(cryo, method="bray", binary=F) #Bray-Curtis
cryo.sore.dist <- 1 - vegdist(cryo, method="bray", binary=T) #Sørensen
cryo.canb.dist <- 1 - vegdist(cryo, method="canberra", binary=F) #Canberra
```

The following is a null model function taken from Locey et al., (2020). We'll use this for plotting our DDR graph
```{r}
m_and_b <- function(com, dist, metric, bin1) {
  
  m <- 0
  b <- 0
  n <- 40
  for (i in 1:n){
    
    com <- randomizeMatrix(com, null.model = "richness", iterations = 1000)
    com.dist <- 1 - vegdist(com, method=metric, binary=bin1)
    #dist <- sample(dist)
    
    lmResult <- lm(com.dist ~ dist)
    b1 <- as.numeric(coef(lmResult)["(Intercept)"])
    b1 <- ifelse(is.na(b1), 0, b1)
    b <- b + b1
    
    m1 <- as.numeric(coef(lmResult)["dist"])
    print(m1)
    m1 <- ifelse(is.na(m1), 0, m1)
    m <- m + m1
    
  }
  
  b <- b/n
  m <- m/n
  return(c(m,b))
}
```

Now we get regression stats for our community dissimilarities + geographic distances
```{r}
# Regression for bray community similarity and geographic distance
DD.cryo.geo.bray <- lm(cryo.bray.dist ~ geo.dist)
summary(DD.cryo.geo.bray, correlation = TRUE)

# Regression for Sørensen community similarity and geographic distance
DD.cryo.geo.sore <- lm(cryo.sore.dist ~ geo.dist)
summary(DD.cryo.geo.sore, correlation = TRUE)

# Regression for Canberra community similarity and geographic distance
DD.cryo.geo.canb <- lm(cryo.canb.dist ~ geo.dist)
summary(DD.cryo.geo.canb, correlation = TRUE)
```

Next we can plot our results

```{r}
#Some colours
clr <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan",
              "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

# Let's plot bray community similarity and geographic distance
slope1 <- round(coefficients(DD.cryo.geo.bray)[2], 3)
int1 <- round(coefficients(DD.cryo.geo.bray)[1], 3)
p <- round(summary(DD.cryo.geo.bray)$coefficients[8],3)
colors <- densCols(geo.dist, cryo.bray.dist, colramp=clr)
res <- m_and_b(cryo, geo.dist, "bray", FALSE)
slope <- res[1]
intercept <- res[2]
plot(geo.dist, cryo.bray.dist, ylab="Bray-Curtis similarity", pch=20,
  xlab="Geographic distance", col = colors, cex=1, cex.lab=1.2,
  ylim=c(0, max(sapply(cryo.bray.dist, max))))+
title(paste("Slope = ", slope1), line = 1, cex=1)+
abline(DD.cryo.geo.bray, col = "black", lwd=3)+
abline(a=intercept, b=slope, col = "red", lwd=2)

# Repeat for Sørensen community similarity and geographic distance
slope2 <- round(coefficients(DD.cryo.geo.sore)[2], 3)
int2 <- round(coefficients(DD.cryo.geo.sore)[1], 3)
p <- round(summary(DD.cryo.geo.sore)$coefficients[8],3)
colors <- densCols(geo.dist, cryo.sore.dist, colramp=clr)
res <- m_and_b(cryo, geo.dist, "bray", TRUE)
slope <- res[1]
intercept <- res[2]
p2 <- plot(geo.dist, cryo.sore.dist, ylab="Sørensen similarity", pch=20,
  xlab="Geographic distance", col = colors, cex=1, cex.lab=1.2,
  ylim=c(0, max(sapply(cryo.sore.dist, max))))+
title(paste("Slope = ", slope2), line = 1, cex=1)+
abline(DD.cryo.geo.sore, col = "black", lwd=3)+
abline(a=intercept, b=slope, col = "red", lwd=2)

# Repeat for Canberra community similarity and geographic distance
slope3 <- round(coefficients(DD.cryo.geo.canb)[2], 3)
int3 <- round(coefficients(DD.cryo.geo.canb)[1], 3)
p <- round(summary(DD.cryo.geo.canb)$coefficients[8],3)
colors <- densCols(geo.dist, cryo.canb.dist, colramp=clr)
res <- m_and_b(cryo, geo.dist, "canberra", FALSE)
slope <- res[1]
intercept <- res[2]
p3 <- plot(geo.dist, cryo.canb.dist, ylab="Canberra similarity", pch=20,
  xlab="Geographic distance", col = colors, cex=1, cex.lab=1.2,
  ylim=c(0, max(sapply(cryo.canb.dist, max))))+
title(paste("Slope = ", slope3), line = 1, cex=1)+
abline(DD.cryo.geo.canb, col = "black", lwd=3)+
abline(a=intercept, b=slope, col = "red", lwd=2)
```


