---
title: "ISAR"
author: "Amy Solman"
date: "02/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's clean up the work environment and install packages
```{r}
rm(list=ls())
graphics.off()

#install nlls fitting package
#install.packages("minpack.lm") #<- if you've never installed minpack before
library("minpack.lm") #for nlls fitting
library("ggplot2") #for plots 
library(readr)
```

I'm going to read in the Sommers et al., (2020) data. I need the absolute ASV numbers for each hole, the size of each hole and the glacier

```{r}
# load("results/cleaned_files/cleanedfiles.RData")
# sample_info_tab <- read_tsv("results/cleaned_files/SraRunTable-mod.tsv")

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")

#Phyloseq objects
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.rel.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-abundant.rds")#rarefied relative abundance abundant taxa ONLY
my.ps.rar.rel.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-rare.rds") #rarefied relative abundance rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")
```

#I'm going to get my data into an appropriate format - TOTAL TAXA
```{r}
# sample_names <- sample_info_tab$Sample.Name
# #ASV_num <- apply(count_table_no_contam,2,function(x) sum(x > 0))
# ASV_num <- apply(count_table_clean,2,function(x) sum(x > 0))
# glacier <- sample_info_tab$Glacier
# 
# #calculate hole area 
# #Method as per Darcy et al (2018)
# hole_area_m2 <- pi*(sample_info_tab$diameter1/2)*(sample_info_tab$diameter2/2)
#   
# hole_area_cm2 <- hole_area_m2*10000
# data <- as.data.frame(cbind(sample_names, ASV_num, hole_area_cm2, glacier))
# names(data) <- c("Sample", "ASVs", "Area_cm2", "Glacier")
# #make area and ASV data numeric for plotting
# data$Area_cm2=as.numeric(data$Area_cm2)
# data$ASVs=as.numeric(data$ASVs)

Area <- as.numeric(tot.diversity.data$Area)
ASVs <- as.numeric(tot.diversity.data$Richness)
Glacier <- tot.diversity.data$Glacier
data <- as.data.frame(cbind(Area, ASVs, Glacier))
data$Area <- as.numeric(data$Area)
data$ASVs <- as.numeric(data$ASVs)

```

Let's plot them to take a look at the data 
```{r}
ggplot(data, aes(x=Area, y=ASVs, color=Glacier))+
  geom_point()+
  xlab("Area (cm2)")+
  ylab("ASVs")
```
One way to fit the Power Law model to the data is using non-linear least squares fitting
Let's do this for Canada and Taylor Glaciers separately
```{r}
#First define the power law function
powMod <- function(x, c, z) {
 return(c * x^z)
}

#Now use the function to perform non-linear least squares fitting
PowFit_Canada <- nlsLM(ASVs ~ powMod(Area,c,z), data = data[data$Glacier == "Canada", ], start = list(c = .1, z = .1))
PowFit_Taylor <- nlsLM(ASVs ~ powMod(Area,c,z), data = data[data$Glacier == "Taylor", ], start = list(c = .1, z = .1))

#We can look at the coefficients of the fit
coef(PowFit_Canada)["c"]
coef(PowFit_Canada)["z"]
coef(PowFit_Taylor)["c"]
coef(PowFit_Taylor)["z"]

#we can look at the model summary 
summary(PowFit_Canada)
summary(PowFit_Taylor)

#get R2 of the model - BE WARNED THIS IS UNRELIABLE WITH NLLS FITTING
RSS_Pow <- sum(residuals(PowFit_Canada)^2) # Residual sum of squares
TSS_Pow <- sum((data$ASVs - mean(data$ASVs))^2) # Total sum of squares
RSq_Pow <- 1 - (RSS_Pow/TSS_Pow) # R-squared value
```
Now let's get ready to plot our power law model fit!
```{r}

#Let's get ready to plot! Generate areas to use
Areas <- seq(min(data$Area),max(data$Area),len=200)

#Now we use the coefficients from our fit to generate predictive points to plot with our data
Predic2PlotPow_Canada <- powMod(Areas,coef(PowFit_Canada)["c"],coef(PowFit_Canada)["z"])
Predic2PlotPow_Taylor <- powMod(Areas,coef(PowFit_Taylor)["c"],coef(PowFit_Taylor)["z"])

df1 <- data.frame(Areas, Predic2PlotPow_Canada, "Canada")
names(df1) <- c("Area", "ASVs", "Glacier")
df2 <- data.frame(Areas, Predic2PlotPow_Taylor, "Taylor")
names(df2) <- c("Area", "ASVs", "Glacier")

ggplot(data, aes(x=Area, y=ASVs, color=Glacier))+
  geom_point()+
  geom_line(data = df1, aes(x = Area, y = ASVs, color=Glacier), size = 1) +
  geom_line(data = df2, aes(x = Area, y = ASVs, color=Glacier), size = 1) +
  xlab(expression(~Area(cm^2)))+
  ylab("ASVs")
ggsave("../Outputs/Sommers2020/graphs/Area-ASV-total.pdf", width=5, height=4)
```

REPEAT THE POWER LAW MODEL FIT USING LOGARITHM OF THE DATA
Get the logarithm of ASV and Area
```{r}
data$logArea <- log10(data$Area)
data$logAsv <- log10(data$ASVs)
```

This is a different way of fitting the power law model
Here we start by fitting a linear model to the log10 transformed area and ASV data, then use the coefficients from the linear model to generate plotting points and using the power law equation log(S) = log(c) + z * log(A) as described by Sommers et al., (2020)
```{r}
#perform linear model on log asv and area data of each glacier  
Canada_model_log <- lm(logAsv ~ logArea, data=data[data$Glacier == "Canada", ])
summary(Canada_model_log)

Taylor_model_log <- lm(logAsv ~ logArea, data=data[data$Glacier == "Taylor", ])
summary(Taylor_model_log)

#divide up the dataset for each glacier
Canada_data <- data[data$Glacier == "Canada", ]
Taylor_data <- data[data$Glacier == "Taylor", ]

#calculate the plotting points of the log transformed power law mode
Canada_x <- Canada_model_log$coefficients[1] + Canada_model_log$coefficients[2] * Canada_data$logArea
Taylor_x <- Taylor_model_log$coefficients[1] + Taylor_model_log$coefficients[2] * Taylor_data$logArea

#put the points into dataframes
df3 <- data.frame(Canada_data$logArea, Canada_x, "Canada")
names(df3) <- c("Area_log", "ASVs_log", "Glacier")
df4 <- data.frame(Taylor_data$logArea, Taylor_x, "Taylor")
names(df4) <- c("Area_log", "ASVs_log", "Glacier")


ggplot(data, aes(x=logArea, y=logAsv, color=Glacier))+
  geom_point()+
  geom_line(data = df3, aes(x = Area_log, y = ASVs_log, color=Glacier), size = 1)+
  geom_line(data = df4, aes(x = Area_log, y = ASVs_log, color=Glacier), size = 1)+
    xlab("Log10 Area (cm2)")+
  xlab(expression(~log[10]~(Area(cm^2))))+
  ylab(expression(~log[10]~(ASVs)))
ggsave("../Outputs/Sommers2020/graphs/logArea-logASVs-total.pdf", width=5, height=4)

summary(Canada_model_log)
summary(Taylor_model_log)
```

#I'm going to get my data into an appropriate format - ABUNDANT TAXA
```{r}
Area <- as.numeric(abun.diversity.data$Area)
ASVs <- as.numeric(abun.diversity.data$Richness)
Glacier <- abun.diversity.data$Glacier
data <- as.data.frame(cbind(Area, ASVs, Glacier))
data$Area <- as.numeric(data$Area)
data$ASVs <- as.numeric(data$ASVs)

```

Let's plot them to take a look at the data 
```{r}
ggplot(data, aes(x=Area, y=ASVs, color=Glacier))+
  geom_point()+
  xlab("Area (cm2)")+
  ylab("ASVs")
```
One way to fit the Power Law model to the data is using non-linear least squares fitting
Let's do this for Canada and Taylor Glaciers separately
```{r}
#First define the power law function
powMod <- function(x, c, z) {
 return(c * x^z)
}

#Now use the function to perform non-linear least squares fitting
PowFit_Canada <- nlsLM(ASVs ~ powMod(Area,c,z), data = data[data$Glacier == "Canada", ], start = list(c = .1, z = .1))
PowFit_Taylor <- nlsLM(ASVs ~ powMod(Area,c,z), data = data[data$Glacier == "Taylor", ], start = list(c = .1, z = .1))

#We can look at the coefficients of the fit
coef(PowFit_Canada)["c"]
coef(PowFit_Canada)["z"]
coef(PowFit_Taylor)["c"]
coef(PowFit_Taylor)["z"]

#we can look at the model summary 
summary(PowFit_Canada)
summary(PowFit_Taylor)

#get R2 of the model - BE WARNED THIS IS UNRELIABLE WITH NLLS FITTING
RSS_Pow <- sum(residuals(PowFit_Canada)^2) # Residual sum of squares
TSS_Pow <- sum((data$ASVs - mean(data$ASVs))^2) # Total sum of squares
RSq_Pow <- 1 - (RSS_Pow/TSS_Pow) # R-squared value
```
Now let's get ready to plot our power law model fit!
```{r}

#Let's get ready to plot! Generate areas to use
Areas <- seq(min(data$Area),max(data$Area),len=200)

#Now we use the coefficients from our fit to generate predictive points to plot with our data
Predic2PlotPow_Canada <- powMod(Areas,coef(PowFit_Canada)["c"],coef(PowFit_Canada)["z"])
Predic2PlotPow_Taylor <- powMod(Areas,coef(PowFit_Taylor)["c"],coef(PowFit_Taylor)["z"])

df1 <- data.frame(Areas, Predic2PlotPow_Canada, "Canada")
names(df1) <- c("Area", "ASVs", "Glacier")
df2 <- data.frame(Areas, Predic2PlotPow_Taylor, "Taylor")
names(df2) <- c("Area", "ASVs", "Glacier")

ggplot(data, aes(x=Area, y=ASVs, color=Glacier))+
  geom_point()+
  geom_line(data = df1, aes(x = Area, y = ASVs, color=Glacier), size = 1) +
  geom_line(data = df2, aes(x = Area, y = ASVs, color=Glacier), size = 1) +
  xlab(expression(~Area(cm^2)))+
  ylab("ASVs")
ggsave("../Outputs/Sommers2020/graphs/Area-ASV-abundant.pdf", width=7, height=4)
```

REPEAT THE POWER LAW MODEL FIT USING LOGARITHM OF THE DATA
Get the logarithm of ASV and Area
```{r}
data$logArea <- log10(data$Area)
data$logAsv <- log10(data$ASVs)
```

This is a different way of fitting the power law model
Here we start by fitting a linear model to the log10 transformed area and ASV data, then use the coefficients from the linear model to generate plotting points and using the power law equation log(S) = log(c) + z * log(A) as described by Sommers et al., (2020)
```{r}
#perform linear model on log asv and area data of each glacier  
Canada_model_log <- lm(logAsv ~ logArea, data=data[data$Glacier == "Canada", ])
summary(Canada_model_log)

Taylor_model_log <- lm(logAsv ~ logArea, data=data[data$Glacier == "Taylor", ])
summary(Taylor_model_log)

#divide up the dataset for each glacier
Canada_data <- data[data$Glacier == "Canada", ]
Taylor_data <- data[data$Glacier == "Taylor", ]

#calculate the plotting points of the log transformed power law mode
Canada_x <- Canada_model_log$coefficients[1] + Canada_model_log$coefficients[2] * Canada_data$logArea
Taylor_x <- Taylor_model_log$coefficients[1] + Taylor_model_log$coefficients[2] * Taylor_data$logArea

#put the points into dataframes
df3 <- data.frame(Canada_data$logArea, Canada_x, "Canada")
names(df3) <- c("Area_log", "ASVs_log", "Glacier")
df4 <- data.frame(Taylor_data$logArea, Taylor_x, "Taylor")
names(df4) <- c("Area_log", "ASVs_log", "Glacier")


ggplot(data, aes(x=logArea, y=logAsv, color=Glacier))+
  geom_point()+
  geom_line(data = df3, aes(x = Area_log, y = ASVs_log, color=Glacier), size = 1)+
  geom_line(data = df4, aes(x = Area_log, y = ASVs_log, color=Glacier), size = 1)+
    xlab("Log10 Area (cm2)")+
  xlab(expression(~log[10]~(Area(cm^2))))+
  ylab(expression(~log[10]~(ASVs)))
ggsave("../Outputs/Sommers2020/graphs/logArea-logASVs-abundant.pdf", width=7, height=4)

summary(Canada_model_log)
summary(Taylor_model_log)
```


#I'm going to get my data into an appropriate format - RARE TAXA
```{r}
Area <- as.numeric(rare.diversity.data$Area)
ASVs <- as.numeric(rare.diversity.data$Richness)
Glacier <- rare.diversity.data$Glacier
data <- as.data.frame(cbind(Area, ASVs, Glacier))
data$Area <- as.numeric(data$Area)
data$ASVs <- as.numeric(data$ASVs)

```

Let's plot them to take a look at the data 
```{r}
ggplot(data, aes(x=Area, y=ASVs, color=Glacier))+
  geom_point()+
  xlab("Area (cm2)")+
  ylab("ASVs")
```
One way to fit the Power Law model to the data is using non-linear least squares fitting
Let's do this for Canada and Taylor Glaciers separately
```{r}
#First define the power law function
powMod <- function(x, c, z) {
 return(c * x^z)
}

#Now use the function to perform non-linear least squares fitting
PowFit_Canada <- nlsLM(ASVs ~ powMod(Area,c,z), data = data[data$Glacier == "Canada", ], start = list(c = .1, z = .1))
PowFit_Taylor <- nlsLM(ASVs ~ powMod(Area,c,z), data = data[data$Glacier == "Taylor", ], start = list(c = .1, z = .1))

#We can look at the coefficients of the fit
coef(PowFit_Canada)["c"]
coef(PowFit_Canada)["z"]
coef(PowFit_Taylor)["c"]
coef(PowFit_Taylor)["z"]

#we can look at the model summary 
summary(PowFit_Canada)
summary(PowFit_Taylor)

#get R2 of the model - BE WARNED THIS IS UNRELIABLE WITH NLLS FITTING
RSS_Pow <- sum(residuals(PowFit_Canada)^2) # Residual sum of squares
TSS_Pow <- sum((data$ASVs - mean(data$ASVs))^2) # Total sum of squares
RSq_Pow <- 1 - (RSS_Pow/TSS_Pow) # R-squared value
```
Now let's get ready to plot our power law model fit!
```{r}

#Let's get ready to plot! Generate areas to use
Areas <- seq(min(data$Area),max(data$Area),len=200)

#Now we use the coefficients from our fit to generate predictive points to plot with our data
Predic2PlotPow_Canada <- powMod(Areas,coef(PowFit_Canada)["c"],coef(PowFit_Canada)["z"])
Predic2PlotPow_Taylor <- powMod(Areas,coef(PowFit_Taylor)["c"],coef(PowFit_Taylor)["z"])

df1 <- data.frame(Areas, Predic2PlotPow_Canada, "Canada")
names(df1) <- c("Area", "ASVs", "Glacier")
df2 <- data.frame(Areas, Predic2PlotPow_Taylor, "Taylor")
names(df2) <- c("Area", "ASVs", "Glacier")

ggplot(data, aes(x=Area, y=ASVs, color=Glacier))+
  geom_point()+
  geom_line(data = df1, aes(x = Area, y = ASVs, color=Glacier), size = 1) +
  geom_line(data = df2, aes(x = Area, y = ASVs, color=Glacier), size = 1) +
  xlab(expression(~Area(cm^2)))+
  ylab("ASVs")
ggsave("../Outputs/Sommers2020/graphs/Area-ASV-rare.pdf", width=7, height=4)
```

REPEAT THE POWER LAW MODEL FIT USING LOGARITHM OF THE DATA
Get the logarithm of ASV and Area
```{r}
data$logArea <- log10(data$Area)
data$logAsv <- log10(data$ASVs)
```

This is a different way of fitting the power law model
Here we start by fitting a linear model to the log10 transformed area and ASV data, then use the coefficients from the linear model to generate plotting points and using the power law equation log(S) = log(c) + z * log(A) as described by Sommers et al., (2020)
```{r}
#perform linear model on log asv and area data of each glacier  
Canada_model_log <- lm(logAsv ~ logArea, data=data[data$Glacier == "Canada", ])
summary(Canada_model_log)

Taylor_model_log <- lm(logAsv ~ logArea, data=data[data$Glacier == "Taylor", ])
summary(Taylor_model_log)

#divide up the dataset for each glacier
Canada_data <- data[data$Glacier == "Canada", ]
Taylor_data <- data[data$Glacier == "Taylor", ]

#calculate the plotting points of the log transformed power law mode
Canada_x <- Canada_model_log$coefficients[1] + Canada_model_log$coefficients[2] * Canada_data$logArea
Taylor_x <- Taylor_model_log$coefficients[1] + Taylor_model_log$coefficients[2] * Taylor_data$logArea

#put the points into dataframes
df3 <- data.frame(Canada_data$logArea, Canada_x, "Canada")
names(df3) <- c("Area_log", "ASVs_log", "Glacier")
df4 <- data.frame(Taylor_data$logArea, Taylor_x, "Taylor")
names(df4) <- c("Area_log", "ASVs_log", "Glacier")


ggplot(data, aes(x=logArea, y=logAsv, color=Glacier))+
  geom_point()+
  geom_line(data = df3, aes(x = Area_log, y = ASVs_log, color=Glacier), size = 1)+
  geom_line(data = df4, aes(x = Area_log, y = ASVs_log, color=Glacier), size = 1)+
    xlab("Log10 Area (cm2)")+
  xlab(expression(~log[10]~(Area(cm^2))))+
  ylab(expression(~log[10]~(ASVs)))
ggsave("../Outputs/Sommers2020/graphs/logArea-logASVs-rare.pdf", width=7, height=4)

summary(Canada_model_log)
anova(Canada_model_log)
summary(Taylor_model_log)
```






