---
title: "Neutral"
author: "Amy Solman"
date: "16/04/2021"
output: html_document
---
The following code is derived from Barnett et al., (2020) Soil characteristics and land-use drive bacterial community assembly patterns. Code available at: https://github.com/seb369/landuse_comm_assembly/blob/master/Neutral_model.md. The neutral model code of Barnett et al., (2020) was adapted from Burns, A., Stephens, W., Stagaman, K. et al. Contribution of neutral processes to the assembly of gut microbial communities in the zebrafish over host development. ISME J 10, 655–664 (2016) doi:10.1038/ismej.2015.142
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages needed for analysis
```{r}

rm(list=ls())

library(dplyr)
library(tibble)
library(phyloseq)
library(ape)
library(vegan)
#install.packages("FSA")
library(FSA)
#install.packages("eulerr")
library(eulerr)

# Packages needed for plotting
library(ggplot2)
library(grid)
library(gridExtra)
```

#Step Two: Import data and make phyloseq object
```{r}
# load("results/cleaned_files/cleanedfiles_3.RData")
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")

#Phyloseq objects
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.rel.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-abundant.rds")#rarefied relative abundance abundant taxa ONLY
my.ps.rar.rel.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-rare.rds") #rarefied relative abundance rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")

```

#Step Three: Check to make sure each sample has at least 10000 reads 
```{r}
# #check number of reads in each sample
# read.num <- data.frame("count" = colSums(otu_table(my.ps.rar))) %>%
#   rownames_to_column(var="Run") %>%
#   inner_join(data.frame(sample_data(my.ps.rar)), by="Run") %>%
#   arrange(-count) %>%
#   mutate(X.Sample=factor(Run, levels=Run))
# 
# #Check to make sure all of our samples are at or above 10000 reads
# ggplot(data=read.num, aes(x=Run, y=log10(count), fill = Glacier)) +
#   geom_bar(stat="identity")+
#   labs(x="Sample", y="Log10(Read count)")+
#   geom_hline(yintercept=log10(10000))+
#   theme(text=element_text(size=16),
#        axis.text.x=element_blank())
```

#Step Four: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

###TOTAL TAXA
```{r}
## Code adapted from Adam Burns (http://dx.doi.org/10.1038/ismej.2015.142)
require(minpack.lm)
require(Hmisc)
require(stats4)
# Extract the OTU table from the phyloseq object
OTU.table = t(otu_table(my.ps.rar))

# Calculate the number of individuals in the meta community (Average read depth)
N <- mean(apply(OTU.table, 1, sum)) #find the mean (mean) of the sum (sum) of all (apply) the rows (1) in the asv table (OTU.table)

# Calculate the average relative abundance of each taxa across communities
p.m <- apply(OTU.table, 2, mean) #find the mean of each column (ASV reads)
p.m <- p.m[p.m != 0] #Remove any zeros
p <- p.m/N #divide the mean number of reads per ASV by the size of the metacommunity 
p.df = data.frame(p) %>%
  rownames_to_column(var="OTU") #put the average relative abundance of each taxa into a dataframe

# Calculate the occurrence frequency of each taxa
OTU.table.bi <- 1*(OTU.table>0) #get the OTU table but make each occurrence of an ASV have the value of 1
freq.table <- apply(OTU.table.bi, 2, mean) #find the mean of frequence of each column (ASV)
freq.table <- freq.table[freq.table != 0] #only keep data that isn't zero
freq.df = data.frame(OTU=names(freq.table), freq=freq.table) #Make into dataframe with OTU name and frequence of occurence (percentage of samples the ASV is present in )

#Combine
C <- inner_join(p.df,freq.df, by="OTU") %>%
  arrange(p)
# Remove rows with any zero (absent in either source pool or local communities). You already did this, but just to make sure we will do it again.
C.no0 <- C %>%
  filter(freq != 0, p != 0)

#Calculate the limit of detection
d <- 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
p.list <- C.no0$p #this creates a vector of the mean relative abundances of each ASV
freq.list <- C.no0$freq #this creates a list of the mean frequency of each ASV
m.fit <- nlsLM(freq.list ~ pbeta(d, N*m*p.list, N*m*(1-p.list), lower.tail=FALSE), start=list(m=0.1)) #using the mean relative abundances of each ASV, the mean frequency of each ASV, the size of the metacommunity and the detection limit, estimate the dispersal rate of the study area using non-linear least squares fitting. 
m.ci <- confint(m.fit, 'm', level=0.95) #get the confidence intervals for m 
m.sum <- summary(m.fit) #get a summary of the model fitting 
m.coef = coef(m.fit) #get the estimated coefficient of our model fitting

freq.pred <- pbeta(d, N*coef(m.fit)*p.list, N*coef(m.fit)*(1-p.list), lower.tail=FALSE) #get the predicted ASV frequencies under stochastic processes
Rsqr <- 1 - (sum((freq.list - freq.pred)^2))/(sum((freq.list - mean(freq.list))^2)) #get the R2 result of observed frequencies against predicted frequencies

# Get table of model fit stats
fitstats <- data.frame(m=m.coef, m.low.ci=m.ci[1], m.up.ci=m.ci[2], 
                       Rsqr=Rsqr, p.value=m.sum$parameters[4], N=N, 
                       Samples=nrow(OTU.table), Richness=length(p.list), 
                       Detect=d)

# Get confidence interval for predictions
freq.pred.ci <- binconf(freq.pred*nrow(OTU.table), nrow(OTU.table), alpha=0.05, method="wilson", return.df=TRUE)

# Get table of predictions
pred.df <- data.frame(metacomm_RA=p.list, frequency=freq.pred, 
                      frequency_lowerCI=freq.pred.ci[,2], 
                      frequency_upperCI=freq.pred.ci[,3]) %>%
  unique()

# Get table of observed occupancy and abundance
obs.df = C.no0 %>%
  rename(metacomm_RA = p, frequency=freq)

```

```{r}

fulldata.model.plot = ggplot(data=obs.df) +
    geom_point(data=obs.df, aes(x=log10(metacomm_RA), y=frequency), 
               alpha=.2, size=2, color="grey30") +
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency), color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_lowerCI), linetype=2, color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_upperCI), linetype=2, color="black") + 
    geom_text(data=fitstats, aes(label = paste("R^2 == ", round(Rsqr, 3))), 
              x=-4.9, y=0.95, size=4, parse=TRUE) +
    geom_text(data=fitstats, aes(label = paste("italic(m) ==", round(m, 3))), 
              x=-4.9, y=0.85, size=4, parse=TRUE) + 
    labs(x="Log10 abundance in\nmetacommunity", y="Frequency detected") +
    theme_bw() +
    theme(axis.line = element_line(color="black"),
          legend.position = "none",
          axis.title = element_text(size=14),
          axis.text = element_text(size=12))

fulldata.model.plot

ggsave(file= "../Outputs/Sommers2020/graphs/neutral-model-total.pdf", width=6, height=4)

```

###ABUNDANT TAXA
```{r}
## Code adapted from Adam Burns (http://dx.doi.org/10.1038/ismej.2015.142)
#require(minpack.lm)
#require(Hmisc)
#require(stats4)
# Extract the OTU table from the phyloseq object
OTU.table = t(otu_table(my.ps.rar.abun))

# Calculate the number of individuals in the meta community (Average read depth)
N <- mean(apply(OTU.table, 1, sum)) #find the mean (mean) of the sum (sum) of all (apply) the rows (1) in the asv table (OTU.table)

# Calculate the average relative abundance of each taxa across communities
p.m <- apply(OTU.table, 2, mean) #find the mean of each column (ASV reads)
p.m <- p.m[p.m != 0] #Remove any zeros
p <- p.m/N #divide the mean number of reads per ASV by the size of the metacommunity 
p.df = data.frame(p) %>%
  rownames_to_column(var="OTU") #put the average relative abundance of each taxa into a dataframe

# Calculate the occurrence frequency of each taxa
OTU.table.bi <- 1*(OTU.table>0) #get the OTU table but make each occurrence of an ASV have the value of 1
freq.table <- apply(OTU.table.bi, 2, mean) #find the mean of frequence of each column (ASV)
freq.table <- freq.table[freq.table != 0] #only keep data that isn't zero
freq.df = data.frame(OTU=names(freq.table), freq=freq.table) #Make into dataframe with OTU name and frequence of occurence (percentage of samples the ASV is present in )

#Combine
C <- inner_join(p.df,freq.df, by="OTU") %>%
  arrange(p)
# Remove rows with any zero (absent in either source pool or local communities). You already did this, but just to make sure we will do it again.
C.no0 <- C %>%
  filter(freq != 0, p != 0)

#Calculate the limit of detection
d <- 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
p.list <- C.no0$p #this creates a vector of the mean relative abundances of each ASV
freq.list <- C.no0$freq #this creates a list of the mean frequency of each ASV
m.fit <- nlsLM(freq.list ~ pbeta(d, N*m*p.list, N*m*(1-p.list), lower.tail=FALSE), start=list(m=0.1)) #using the mean relative abundances of each ASV, the mean frequency of each ASV, the size of the metacommunity and the detection limit, estimate the dispersal rate of the study area using non-linear least squares fitting. 
m.ci <- confint(m.fit, 'm', level=0.95) #get the confidence intervals for m 
m.sum <- summary(m.fit) #get a summary of the model fitting 
m.coef = coef(m.fit) #get the estimated coefficient of our model fitting

freq.pred <- pbeta(d, N*coef(m.fit)*p.list, N*coef(m.fit)*(1-p.list), lower.tail=FALSE) #get the predicted ASV frequencies under stochastic processes
Rsqr <- 1 - (sum((freq.list - freq.pred)^2))/(sum((freq.list - mean(freq.list))^2)) #get the R2 result of observed frequencies against predicted frequencies

# Get table of model fit stats
fitstats <- data.frame(m=m.coef, m.low.ci=m.ci[1], m.up.ci=m.ci[2], 
                       Rsqr=Rsqr, p.value=m.sum$parameters[4], N=N, 
                       Samples=nrow(OTU.table), Richness=length(p.list), 
                       Detect=d)

# Get confidence interval for predictions
freq.pred.ci <- binconf(freq.pred*nrow(OTU.table), nrow(OTU.table), alpha=0.05, method="wilson", return.df=TRUE)

# Get table of predictions
pred.df <- data.frame(metacomm_RA=p.list, frequency=freq.pred, 
                      frequency_lowerCI=freq.pred.ci[,2], 
                      frequency_upperCI=freq.pred.ci[,3]) %>%
  unique()

# Get table of observed occupancy and abundance
obs.df = C.no0 %>%
  rename(metacomm_RA = p, frequency=freq)

```

```{r}

fulldata.model.plot = ggplot(data=obs.df) +
    geom_point(data=obs.df, aes(x=log10(metacomm_RA), y=frequency), 
               alpha=.2, size=2, color="grey30") +
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency), color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_lowerCI), linetype=2, color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_upperCI), linetype=2, color="black") + 
    geom_text(data=fitstats, aes(label = paste("R^2 == ", round(Rsqr, 3))), 
              x=-1.5, y=0.5, size=4, parse=TRUE) +
    geom_text(data=fitstats, aes(label = paste("italic(m) ==", round(m, 3))), 
              x=-1.5, y=0.3, size=4, parse=TRUE) + 
    labs(x="Log10 abundance in\nmetacommunity", y="Frequency detected") +
    theme_bw() +
    theme(axis.line = element_line(color="black"),
          legend.position = "none",
          axis.title = element_text(size=14),
          axis.text = element_text(size=12))

fulldata.model.plot

ggsave(file= "../Outputs/Sommers2020/graphs/neutral-model-abundant.pdf", width=6, height=4)

```

###RARE TAXA
```{r}
## Code adapted from Adam Burns (http://dx.doi.org/10.1038/ismej.2015.142)
#require(minpack.lm)
#require(Hmisc)
#require(stats4)
# Extract the OTU table from the phyloseq object
OTU.table = t(otu_table(my.ps.rar.rare))

# Calculate the number of individuals in the meta community (Average read depth)
N <- mean(apply(OTU.table, 1, sum)) #find the mean (mean) of the sum (sum) of all (apply) the rows (1) in the asv table (OTU.table)

# Calculate the average relative abundance of each taxa across communities
p.m <- apply(OTU.table, 2, mean) #find the mean of each column (ASV reads)
p.m <- p.m[p.m != 0] #Remove any zeros
p <- p.m/N #divide the mean number of reads per ASV by the size of the metacommunity 
p.df = data.frame(p) %>%
  rownames_to_column(var="OTU") #put the average relative abundance of each taxa into a dataframe

# Calculate the occurrence frequency of each taxa
OTU.table.bi <- 1*(OTU.table>0) #get the OTU table but make each occurrence of an ASV have the value of 1
freq.table <- apply(OTU.table.bi, 2, mean) #find the mean of frequence of each column (ASV)
freq.table <- freq.table[freq.table != 0] #only keep data that isn't zero
freq.df = data.frame(OTU=names(freq.table), freq=freq.table) #Make into dataframe with OTU name and frequence of occurence (percentage of samples the ASV is present in )

#Combine
C <- inner_join(p.df,freq.df, by="OTU") %>%
  arrange(p)
# Remove rows with any zero (absent in either source pool or local communities). You already did this, but just to make sure we will do it again.
C.no0 <- C %>%
  filter(freq != 0, p != 0)

#Calculate the limit of detection
d <- 1/N

##Fit model parameter m (or Nm) using Non-linear least squares (NLS)
p.list <- C.no0$p #this creates a vector of the mean relative abundances of each ASV
freq.list <- C.no0$freq #this creates a list of the mean frequency of each ASV
m.fit <- nlsLM(freq.list ~ pbeta(d, N*m*p.list, N*m*(1-p.list), lower.tail=FALSE), start=list(m=0.1)) #using the mean relative abundances of each ASV, the mean frequency of each ASV, the size of the metacommunity and the detection limit, estimate the dispersal rate of the study area using non-linear least squares fitting. 
m.ci <- confint(m.fit, 'm', level=0.95) #get the confidence intervals for m 
m.sum <- summary(m.fit) #get a summary of the model fitting 
m.coef = coef(m.fit) #get the estimated coefficient of our model fitting

freq.pred <- pbeta(d, N*coef(m.fit)*p.list, N*coef(m.fit)*(1-p.list), lower.tail=FALSE) #get the predicted ASV frequencies under stochastic processes
Rsqr <- 1 - (sum((freq.list - freq.pred)^2))/(sum((freq.list - mean(freq.list))^2)) #get the R2 result of observed frequencies against predicted frequencies

# Get table of model fit stats
fitstats <- data.frame(m=m.coef, m.low.ci=m.ci[1], m.up.ci=m.ci[2], 
                       Rsqr=Rsqr, p.value=m.sum$parameters[4], N=N, 
                       Samples=nrow(OTU.table), Richness=length(p.list), 
                       Detect=d)

# Get confidence interval for predictions
freq.pred.ci <- binconf(freq.pred*nrow(OTU.table), nrow(OTU.table), alpha=0.05, method="wilson", return.df=TRUE)

# Get table of predictions
pred.df <- data.frame(metacomm_RA=p.list, frequency=freq.pred, 
                      frequency_lowerCI=freq.pred.ci[,2], 
                      frequency_upperCI=freq.pred.ci[,3]) %>%
  unique()

# Get table of observed occupancy and abundance
obs.df = C.no0 %>%
  rename(metacomm_RA = p, frequency=freq)

```

```{r}

fulldata.model.plot = ggplot(data=obs.df) +
    geom_point(data=obs.df, aes(x=log10(metacomm_RA), y=frequency), 
               alpha=.2, size=2, color="grey30") +
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency), color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_lowerCI), linetype=2, color="black") + 
    geom_line(data=pred.df, aes(x=log10(metacomm_RA), y=frequency_upperCI), linetype=2, color="black") + 
    geom_text(data=fitstats, aes(label = paste("R^2 == ", round(Rsqr, 3))), 
              x=-3, y=0.25, size=4, parse=TRUE) +
    geom_text(data=fitstats, aes(label = paste("italic(m) ==", round(m, 3))), 
              x=-3, y=0.2, size=4, parse=TRUE) + 
    labs(x="Log10 abundance in\nmetacommunity", y="Frequency detected") +
    theme_bw() +
    theme(axis.line = element_line(color="black"),
          legend.position = "none",
          axis.title = element_text(size=14),
          axis.text = element_text(size=12))

fulldata.model.plot

ggsave(file= "../Outputs/Sommers2020/graphs/neutral-model-rare.pdf", width=6, height=4)

```