---
title: "Phylogenetic Signal"
author: "Amy Solman"
date: "13/04/2021"
output: html_document
---
#Before we perform null model analysis we need to see if there is phylogenetic signal within our data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

#I had to remove these packages before loading dplyr so I could use the mutate function
# remove.packages("tidyr")
# remove.packages("microbiome")
# remove.packages("microbiomeutilities")
# remove.packages("ggpubr")

library(dplyr)
#install.packages("tidyr")
library(tidyr)
library(tibble)
library(phyloseq)
library(vegan)
library(ape)
library(ggplot2)
```

#Step Two: Import data and make phyloseq object
```{r}
# load("results/cleaned_files/cleanedfiles_3.RData")
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")

#Phyloseq objects
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.rel.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-abundant.rds")#rarefied relative abundance abundant taxa ONLY
my.ps.rar.rel.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-rare.rds") #rarefied relative abundance rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")

```

#Step Three: Check to make sure each sample has at least 10000 reads 
```{r}
# #check number of reads in each sample
# read.num <- data.frame("count" = colSums(otu_table(my.ps))) %>%
#   rownames_to_column(var="Run") %>%
#   inner_join(data.frame(sample_data(my.ps)), by="Run") %>%
#   arrange(-count) %>%
#   mutate(X.Sample=factor(Run, levels=Run))
# 
# #Check to make sure all of our samples are at or above 10000 reads
# ggplot(data=read.num, aes(x=Run, y=log10(count), fill = Glacier)) +
#   geom_bar(stat="identity")+
#   labs(x="Sample", y="Log10(Read count)")+
#   geom_hline(yintercept=log10(10000))+
#   theme(text=element_text(size=16),
#        axis.text.x=element_blank())
```

#Step Four: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Five: Remove samples that don't have the environmental data we need
```{r}
# NA.Chlor <- which(is.na(META$Chlorophyll), arr.ind=TRUE)
# NA.Depth <- which(is.na(META$Depth), arr.ind=TRUE)
# NA.pH <- which(is.na(META$pH), arr.ind=TRUE)
# NA.water <- which(is.na(META$water_content), arr.ind=TRUE)
# my.NA <- as.numeric(c(NA.Chlor, NA.Depth, NA.pH, NA.water))

NA.Chlor <- which(is.na(my.ps.rar@sam_data$Chlorophyll), arr.ind=TRUE)
NA.Depth <- which(is.na(my.ps.rar@sam_data$Depth), arr.ind=TRUE)
NA.pH <- which(is.na(my.ps.rar@sam_data$pH), arr.ind=TRUE)
NA.water <- which(is.na(my.ps.rar@sam_data$water_content), arr.ind=TRUE)
my.NA <- as.numeric(c(NA.Chlor, NA.Depth, NA.pH, NA.water))

#Get sample names as those indexes
samples.to.remove <- my.ps.rar@sam_data$Run[my.NA]

#Remove those samples
for (i in 1:length(samples.to.remove)){
  my.ps.rar <- subset_samples(my.ps.rar, Run != samples.to.remove[i])
}

```

#Step Six: Get phylogenetic distances
```{r}
# #Remove ASVs not found in at least 3 samples
# asv.tab <- otu_table(my.ps.rar.rel)
# asv.tab[asv.tab > 0] = 1
# asv.frq <- rowSums(asv.tab)
# asv.frq <- asv.frq[asv.frq > 2]
# my.ps.edit <- prune_taxa(names(asv.frq), my.ps.rar.rel)

# #Remove ASVs not found in at least 15 samples (this will reduce the size of the dataset)
set.seed(1961)
asv.tab <- otu_table(my.ps.rar)
asv.tab[asv.tab > 0] = 1
asv.frq <- rowSums(asv.tab)
asv.frq <- asv.frq[asv.frq > 20]
#asv.frq <- sample(asv.frq, 150)
#asv.frq <- floor(runif(150, min=1, max=nrow(my.ps.rar@otu_table)))
my.ps.edit <- prune_taxa(names(asv.frq), my.ps.rar)

# set.seed(1961)
# myTaxa = names(sort(taxa_sums(my.ps.rar), decreasing = TRUE))
# tax.to.keep <- sample(myTaxa, 150)
# my.ps.edit = prune_taxa(tax.to.keep, my.ps.rar)

# set.seed(1961)
# #myTaxa = names(sort(taxa_sums(my.ps.rar), decreasing = TRUE))
# #tax.to.keep <- sample(myTaxa, 150)
# tax.to.keep <- floor(runif(150, min=1, max=nrow(my.ps.rar@otu_table)))
# my.ps.edit = prune_taxa(names(tax.to.keep), my.ps.rar)

#check tree is rooted
is.rooted(phy_tree(my.ps.edit))  #TRUE

#Get phylogenetic distances
my.tree <- phy_tree(my.ps.edit)
my.distances <- cophenetic(my.tree)
my.asvs <- my.tree$tip.label
my.phylo.dists <- my.distances[my.asvs, my.asvs]
my.phylo.dists[upper.tri(my.phylo.dists, diag=TRUE)] = NA

```

#Step Seven: Get niche differences and run mantel correlograms
```{r}
#Make a dataframe of the variables we'll use as niche preferences
lovely.niches <- data.frame(sample_data(my.ps.edit)) %>%
  dplyr::select(Run, Chlorophyll, Depth, pH, water_content)

rownames(lovely.niches) <- lovely.niches$Run #make sample names the row names
lovely.niches$Run <- NULL #remove the sample column
lovely.niches <- as.matrix(lovely.niches) #turn the dataframe into a matrix
m <- apply(lovely.niches, 2, as.numeric) #make the columns in the matrix numeric not characters

asv.table <- t(otu_table(my.ps.edit))
asv.niche <- wascores(m, asv.table) #this will find the weighted mean environmental parameter for each asv
asv.niche.df <- data.frame(asv.niche)


best.chlorophyll <- asv.niche.df$Chlorophyll
names(best.chlorophyll) <- rownames(asv.niche.df)
dist.chlorophyll <- as.matrix(dist(best.chlorophyll), labels=TRUE)
x <- dist.chlorophyll[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]
# new.my.phylo.dists <- my.phylo.dists[,1:450]
# new.my.phylo.dists <- new.my.phylo.dists[1:450,]

chlor.corlg <- mantel.correlog(x, my.phylo.dists)
#chlor.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("chlorophyll-crlg.rds")
saveRDS(chlor.corlg, file=filename)

best.depth <- asv.niche.df$Depth
names(best.depth) <- rownames(asv.niche.df)
dist.depth <- as.matrix(dist(best.depth), labels=TRUE)
x <- dist.depth[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]

depth.corlg <- mantel.correlog(x, my.phylo.dists)
#depth.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("depth-crlg.rds")
saveRDS(depth.corlg, file=filename)

best.pH <- asv.niche.df$pH
names(best.pH) <- rownames(asv.niche.df)
dist.pH <- as.matrix(dist(best.pH), labels=TRUE)
x <- dist.pH[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]

pH.corlg <- mantel.correlog(x, my.phylo.dists)
#pH.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("pH-crlg.rds")
saveRDS(pH.corlg, file=filename)

best.water <- asv.niche.df$water_content
names(best.water) <- rownames(asv.niche.df)
dist.water <- as.matrix(dist(best.water), labels=TRUE)
x <- dist.water[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,] #this took 11 hours to run

water.corlg <- mantel.correlog(x, my.phylo.dists)
#water.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("water-crlg.rds")
saveRDS(water.corlg, file=filename)

```

#Step Eight: Import data from correlograms and plot
```{r}
# Import data
chlorophyll.crlg <- data.frame(readRDS(file="chlorophyll-crlg.rds")$mantel.res) %>%
  mutate(property = "chlorophyll")

depth.crlg <- data.frame(readRDS(file="depth-crlg.rds")$mantel.res) %>%
  mutate(property = "depth")

pH.crlg <- data.frame(readRDS(file="pH-crlg.rds")$mantel.res) %>%
  mutate(property = "pH")

water.crlg <- data.frame(readRDS(file="water-crlg.rds")$mantel.res) %>%
  mutate(property = "water")

crlg <- rbind(chlorophyll.crlg, depth.crlg, pH.crlg, water.crlg) %>%
  mutate(sig = ifelse(Pr.corrected. <= 0.05, "significant", "non-significant")) %>%
  filter(!(is.na(Pr.corrected.)))

ggplot(data=crlg, aes(x=class.index, y=Mantel.cor)) +
  geom_point(data=crlg[crlg$sig=="significant",], color = "black", size=2, shape=16) +
  geom_point(data=crlg[crlg$sig=="non-significant",], color = "black",size=2, shape=1) +
  geom_line(data=crlg, aes(color=property)) +
  geom_hline(yintercept = 0, linetype=2) +
  labs(x = "Phylogenetic distance class", y="Mantel correlation", color="Cryoconite property")+
 theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "bottom",
        axis.text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_text(size=14),
        strip.text.x = element_text(size = 14))
ggsave(file="../Outputs/Sommers2020/graphs/mantel-crlg-total.pdf", width=10, height=6)
```

#######ABUNDANT TAXA
#Step Five: Remove samples that don't have the environmental data we need
```{r}
# NA.Chlor <- which(is.na(META$Chlorophyll), arr.ind=TRUE)
# NA.Depth <- which(is.na(META$Depth), arr.ind=TRUE)
# NA.pH <- which(is.na(META$pH), arr.ind=TRUE)
# NA.water <- which(is.na(META$water_content), arr.ind=TRUE)
# my.NA <- as.numeric(c(NA.Chlor, NA.Depth, NA.pH, NA.water))

NA.Chlor <- which(is.na(my.ps.rar.abun@sam_data$Chlorophyll), arr.ind=TRUE)
NA.Depth <- which(is.na(my.ps.rar.abun@sam_data$Depth), arr.ind=TRUE)
NA.pH <- which(is.na(my.ps.rar.abun@sam_data$pH), arr.ind=TRUE)
NA.water <- which(is.na(my.ps.rar.abun@sam_data$water_content), arr.ind=TRUE)
my.NA <- as.numeric(c(NA.Chlor, NA.Depth, NA.pH, NA.water))

#Get sample names as those indexes
samples.to.remove <- my.ps.rar.abun@sam_data$Run[my.NA]

#Remove those samples
for (i in 1:length(samples.to.remove)){
  my.ps.rar.abun <- subset_samples(my.ps.rar.abun, Run != samples.to.remove[i])
}

```

#Step Six: Get phylogenetic distances
```{r}
# #Remove ASVs not found in at least 3 samples
# asv.tab <- otu_table(my.ps.rar.rel)
# asv.tab[asv.tab > 0] = 1
# asv.frq <- rowSums(asv.tab)
# asv.frq <- asv.frq[asv.frq > 2]
# my.ps.edit <- prune_taxa(names(asv.frq), my.ps.rar.rel)

# #Remove ASVs not found in at least 15 samples (this will reduce the size of the dataset)
# asv.tab <- otu_table(my.ps.rar)
# asv.tab[asv.tab > 0] = 1
# asv.frq <- rowSums(asv.tab)
# asv.frq <- asv.frq[asv.frq > 20]
# my.ps.edit <- prune_taxa(names(asv.frq), my.ps.rar)

#set.seed(1961)
# asv.tab <- otu_table(my.ps.rar)
# asv.tab[asv.tab > 0] = 1
# asv.frq <- rowSums(asv.tab)
# asv.frq <- asv.frq[asv.frq > 20]
#asv.frq <- sample(asv.frq, 150)
#asv.frq <- floor(runif(150, min=1, max=nrow(my.ps.rar@otu_table)))
my.ps.rar.abun <- prune_taxa(names(asv.frq), my.ps.rar.abun)

#check tree is rooted
is.rooted(phy_tree(my.ps.rar.abun))  #TRUE

#Get phylogenetic distances
my.tree <- phy_tree(my.ps.rar.abun)
my.distances <- cophenetic(my.tree)
my.asvs <- my.tree$tip.label
my.phylo.dists <- my.distances[my.asvs, my.asvs]
my.phylo.dists[upper.tri(my.phylo.dists, diag=TRUE)] = NA

```

#Step Seven: Get niche differences and run mantel correlograms
```{r}
#Make a dataframe of the variables we'll use as niche preferences
lovely.niches <- data.frame(sample_data(my.ps.rar.abun)) %>%
  dplyr::select(Run, Chlorophyll, Depth, pH, water_content)

rownames(lovely.niches) <- lovely.niches$Run #make sample names the row names
lovely.niches$Run <- NULL #remove the sample column
lovely.niches <- as.matrix(lovely.niches) #turn the dataframe into a matrix
m <- apply(lovely.niches, 2, as.numeric) #make the columns in the matrix numeric not characters

asv.table <- t(otu_table(my.ps.rar.abun))
asv.niche <- wascores(m, asv.table) #this will find the weighted mean environmental parameter for each asv
asv.niche.df <- data.frame(asv.niche)


best.chlorophyll <- asv.niche.df$Chlorophyll
names(best.chlorophyll) <- rownames(asv.niche.df)
dist.chlorophyll <- as.matrix(dist(best.chlorophyll), labels=TRUE)
x <- dist.chlorophyll[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]
# new.my.phylo.dists <- my.phylo.dists[,1:450]
# new.my.phylo.dists <- new.my.phylo.dists[1:450,]

chlor.corlg <- mantel.correlog(x, my.phylo.dists)
#chlor.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("chlorophyll-crlg-abundant.rds")
saveRDS(chlor.corlg, file=filename)

best.depth <- asv.niche.df$Depth
names(best.depth) <- rownames(asv.niche.df)
dist.depth <- as.matrix(dist(best.depth), labels=TRUE)
x <- dist.depth[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]

depth.corlg <- mantel.correlog(x, my.phylo.dists)
#depth.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("depth-crlg-abundant.rds")
saveRDS(depth.corlg, file=filename)

best.pH <- asv.niche.df$pH
names(best.pH) <- rownames(asv.niche.df)
dist.pH <- as.matrix(dist(best.pH), labels=TRUE)
x <- dist.pH[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]

pH.corlg <- mantel.correlog(x, my.phylo.dists)
#pH.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("pH-crlg-abundant.rds")
saveRDS(pH.corlg, file=filename)

best.water <- asv.niche.df$water_content
names(best.water) <- rownames(asv.niche.df)
dist.water <- as.matrix(dist(best.water), labels=TRUE)
x <- dist.water[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,] #this took 11 hours to run

water.corlg <- mantel.correlog(x, my.phylo.dists)
#water.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("water-crlg-abundant.rds")
saveRDS(water.corlg, file=filename)

```

#Step Eight: Import data from correlograms and plot
```{r}
# Import data
chlorophyll.crlg <- data.frame(readRDS(file="chlorophyll-crlg-abundant.rds")$mantel.res) %>%
  mutate(property = "chlorophyll")

depth.crlg <- data.frame(readRDS(file="depth-crlg-abundant.rds")$mantel.res) %>%
  mutate(property = "depth")

pH.crlg <- data.frame(readRDS(file="pH-crlg-abundant.rds")$mantel.res) %>%
  mutate(property = "pH")

water.crlg <- data.frame(readRDS(file="water-crlg-abundant.rds")$mantel.res) %>%
  mutate(property = "water")

crlg <- rbind(chlorophyll.crlg, depth.crlg, pH.crlg, water.crlg) %>%
  mutate(sig = ifelse(Pr.corrected. <= 0.05, "significant", "non-significant")) %>%
  filter(!(is.na(Pr.corrected.)))

ggplot(data=crlg, aes(x=class.index, y=Mantel.cor)) +
  geom_point(data=crlg[crlg$sig=="significant",], color = "black", size=2, shape=16) +
  geom_point(data=crlg[crlg$sig=="non-significant",], color = "black",size=2, shape=1) +
  geom_line(data=crlg, aes(color=property)) +
  geom_hline(yintercept = 0, linetype=2) +
  labs(x = "Phylogenetic distance class", y="Mantel correlation", color="Cryoconite property")+
 theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "bottom",
        axis.text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_text(size=14),
        strip.text.x = element_text(size = 14))
ggsave(file="../Outputs/Sommers2020/graphs/mantel-crlg-abundant.pdf", width=10, height=6)
```

#######RARE TAXA
#Step Five: Remove samples that don't have the environmental data we need
```{r}
# NA.Chlor <- which(is.na(META$Chlorophyll), arr.ind=TRUE)
# NA.Depth <- which(is.na(META$Depth), arr.ind=TRUE)
# NA.pH <- which(is.na(META$pH), arr.ind=TRUE)
# NA.water <- which(is.na(META$water_content), arr.ind=TRUE)
# my.NA <- as.numeric(c(NA.Chlor, NA.Depth, NA.pH, NA.water))

NA.Chlor <- which(is.na(my.ps.rar.rare@sam_data$Chlorophyll), arr.ind=TRUE)
NA.Depth <- which(is.na(my.ps.rar.rare@sam_data$Depth), arr.ind=TRUE)
NA.pH <- which(is.na(my.ps.rar.rare@sam_data$pH), arr.ind=TRUE)
NA.water <- which(is.na(my.ps.rar.rare@sam_data$water_content), arr.ind=TRUE)
my.NA <- as.numeric(c(NA.Chlor, NA.Depth, NA.pH, NA.water))

#Get sample names as those indexes
samples.to.remove <- my.ps.rar.rare@sam_data$Run[my.NA]

#Remove those samples
for (i in 1:length(samples.to.remove)){
  my.ps.rar.rare <- subset_samples(my.ps.rar.rare, Run != samples.to.remove[i])
}

```

#Step Six: Get phylogenetic distances
```{r}
# #Remove ASVs not found in at least 3 samples
# asv.tab <- otu_table(my.ps.rar.rel)
# asv.tab[asv.tab > 0] = 1
# asv.frq <- rowSums(asv.tab)
# asv.frq <- asv.frq[asv.frq > 2]
# my.ps.edit <- prune_taxa(names(asv.frq), my.ps.rar.rel)

# #Remove ASVs not found in at least 15 samples (this will reduce the size of the dataset)
# asv.tab <- otu_table(my.ps.rar)
# asv.tab[asv.tab > 0] = 1
# asv.frq <- rowSums(asv.tab)
# asv.frq <- asv.frq[asv.frq > 20]
# my.ps.edit <- prune_taxa(names(asv.frq), my.ps.rar)

#Remove ASVs not found in only 2 sample (this will reduce the size of the dataset)
asv.tab <- otu_table(my.ps.rar.rare)
asv.tab[asv.tab > 0] = 1
asv.frq <- rowSums(asv.tab)
asv.frq <- asv.frq[1:200]
my.ps.rar.rare <- prune_taxa(names(asv.frq), my.ps.rar.rare)


#check tree is rooted
is.rooted(phy_tree(my.ps.rar.rare))  #TRUE

#Get phylogenetic distances
my.tree <- phy_tree(my.ps.rar.rare)
my.distances <- cophenetic(my.tree)
my.asvs <- my.tree$tip.label
my.phylo.dists <- my.distances[my.asvs, my.asvs]
my.phylo.dists[upper.tri(my.phylo.dists, diag=TRUE)] = NA

```

#Step Seven: Get niche differences and run mantel correlograms
```{r}
#Make a dataframe of the variables we'll use as niche preferences
lovely.niches <- data.frame(sample_data(my.ps.rar.rare)) %>%
  dplyr::select(Run, Chlorophyll, Depth, pH, water_content)

rownames(lovely.niches) <- lovely.niches$Run #make sample names the row names
lovely.niches$Run <- NULL #remove the sample column
lovely.niches <- as.matrix(lovely.niches) #turn the dataframe into a matrix
m <- apply(lovely.niches, 2, as.numeric) #make the columns in the matrix numeric not characters

###For some reason my phyloseq object wouldn't convert to an ASV table so I've included a work around here
#asv.table <- t(otu_table(my.ps.rar.rare))
library(metagMisc)
asv.table <- phyloseq_to_df(my.ps.rar.rare)
row.names(asv.table) <- row.names(my.ps.rar.rare@otu_table)
asv.table <- asv.table[-c(1:8)]
asv.table <- t(asv.table)

asv.niche <- wascores(m, asv.table) #this will find the weighted mean environmental parameter for each asv
asv.niche.df <- data.frame(asv.niche)


best.chlorophyll <- asv.niche.df$Chlorophyll
names(best.chlorophyll) <- rownames(asv.niche.df)
dist.chlorophyll <- as.matrix(dist(best.chlorophyll), labels=TRUE)
x <- dist.chlorophyll[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]
# new.my.phylo.dists <- my.phylo.dists[,1:450]
# new.my.phylo.dists <- new.my.phylo.dists[1:450,]

chlor.corlg <- mantel.correlog(x, my.phylo.dists)
#chlor.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("chlorophyll-crlg-rare.rds")
saveRDS(chlor.corlg, file=filename)

best.depth <- asv.niche.df$Depth
names(best.depth) <- rownames(asv.niche.df)
dist.depth <- as.matrix(dist(best.depth), labels=TRUE)
x <- dist.depth[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]

depth.corlg <- mantel.correlog(x, my.phylo.dists)
#depth.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("depth-crlg-rare.rds")
saveRDS(depth.corlg, file=filename)

best.pH <- asv.niche.df$pH
names(best.pH) <- rownames(asv.niche.df)
dist.pH <- as.matrix(dist(best.pH), labels=TRUE)
x <- dist.pH[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,]

pH.corlg <- mantel.correlog(x, my.phylo.dists)
#pH.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("pH-crlg-rare.rds")
saveRDS(pH.corlg, file=filename)

best.water <- asv.niche.df$water_content
names(best.water) <- rownames(asv.niche.df)
dist.water <- as.matrix(dist(best.water), labels=TRUE)
x <- dist.water[my.asvs, my.asvs]
x[upper.tri(x, diag = TRUE)] = NA
# new.x <- x[,1:450]
# new.x <- new.x[1:450,] #this took 11 hours to run

water.corlg <- mantel.correlog(x, my.phylo.dists)
#water.corlg <- mantel.correlog(new.x, new.my.phylo.dists)
filename = ("water-crlg-rare.rds")
saveRDS(water.corlg, file=filename)

```

#Step Eight: Import data from correlograms and plot
```{r}
# Import data
chlorophyll.crlg <- data.frame(readRDS(file="chlorophyll-crlg-rare.rds")$mantel.res) %>%
  mutate(property = "chlorophyll")

depth.crlg <- data.frame(readRDS(file="depth-crlg-rare.rds")$mantel.res) %>%
  mutate(property = "depth")

pH.crlg <- data.frame(readRDS(file="pH-crlg-rare.rds")$mantel.res) %>%
  mutate(property = "pH")

water.crlg <- data.frame(readRDS(file="water-crlg-rare.rds")$mantel.res) %>%
  mutate(property = "water")

crlg <- rbind(chlorophyll.crlg, depth.crlg, pH.crlg, water.crlg) %>%
  mutate(sig = ifelse(Pr.corrected. <= 0.05, "significant", "non-significant")) %>%
  filter(!(is.na(Pr.corrected.)))

ggplot(data=crlg, aes(x=class.index, y=Mantel.cor)) +
  geom_point(data=crlg[crlg$sig=="significant",], color = "black", size=2, shape=16) +
  geom_point(data=crlg[crlg$sig=="non-significant",], color = "black",size=2, shape=1) +
  geom_line(data=crlg, aes(color=property)) +
  geom_hline(yintercept = 0, linetype=2) +
  labs(x = "Phylogenetic distance class", y="Mantel correlation", color="Cryoconite property")+
 theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "bottom",
        axis.text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_text(size=14),
        strip.text.x = element_text(size = 14))
ggsave(file="../Outputs/Sommers2020/graphs/mantel-crlg-rare.pdf", width=10, height=6)
```