---
title: "BNTI"
author: "Amy Solman"
date: "14/04/2021"
output: html_document
---

#The following script is developed from Barrnet et al., (2020) (https://github.com/seb369/landuse_comm_assembly/blob/master/bNTI_calculation.Rmd) who developed their script from from Stegen et al., (2013) (https://github.com/stegen/Stegen_etal_ISME_2013/blob/master/bNTI_Local_Machine.r)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# ## change to the directory on your computer that contains the OTU table and phylogeny
# ## note that the 'slash' needs to be changed to a forward slash like this /
# #setwd("C:/Users/steg815/Desktop/Stegen_PNNL/")
# 
# ## load this library
# ## if not already installed, use install.packages('picante')
# library(picante)
# 
# #Read in our data 
# 
# load("Analyses/results/cleaned_files/cleanedfiles_3.RData")
# 
# ## read in OTU table
# 
# #otu = read.csv("Analyses/results/cleaned_files/cleanedfiles_3.RData", header=T,row.names=1);
# otu <- count_table_clean
# #otu <- taxonomy
# dim(otu); # this gives the dimensions
# otu[1:5,1:5]; # this gives a look at the first 5 rows and columns
```


```{r}
# ## read in the phylogeny
# 
# #phylo = read.tree("bacteria-phylogeny.txt");
# phylo <- tree_clean
# phylo; # a summary of the phylogeny
# plot.phylo(phylo,typ="fan"); # a quick plot
# 
# ## make sure the names on the phylogeny are ordered the same as the names in otu table
# 
# match.phylo.otu = match.phylo.data(phylo, t(otu));
# str(match.phylo.otu);
```

```{r}
# ## calculate empirical betaMNTD
# 
# beta.mntd.weighted = as.matrix(comdistnt(t(match.phylo.otu$data),cophenetic(match.phylo.otu$phy),abundance.weighted=T));
# dim(beta.mntd.weighted);
# beta.mntd.weighted[1:5,1:5];
# write.csv(beta.mntd.weighted,'betaMNTD_weighted.csv',quote=F);
# 
# identical(colnames(match.phylo.otu$data),colnames(beta.mntd.weighted)); # just a check, should be TRUE
# identical(colnames(match.phylo.otu$data),rownames(beta.mntd.weighted)); # just a check, should be TRUE
# ```
# 
# ```{r}
# # calculate randomized betaMNTD
# 
# beta.reps = 999; # number of randomizations
# 
# rand.weighted.bMNTD.comp = array(c(-999),dim=c(ncol(match.phylo.otu$data),ncol(match.phylo.otu$data),beta.reps));
# dim(rand.weighted.bMNTD.comp);
# 
# for (rep in 1:beta.reps) {
#   
#   rand.weighted.bMNTD.comp[,,rep] = as.matrix(comdistnt(t(match.phylo.otu$data),taxaShuffle(cophenetic(match.phylo.otu$phy)),abundance.weighted=T,exclude.conspecifics = F));
#   
#   print(c(date(),rep));
#   
# }
# 
# weighted.bNTI = matrix(c(NA),nrow=ncol(match.phylo.otu$data),ncol=ncol(match.phylo.otu$data));
# dim(weighted.bNTI);
# 
# for (columns in 1:(ncol(match.phylo.otu$data)-1)) {
#   for (rows in (columns+1):ncol(match.phylo.otu$data)) {
#     
#     rand.vals = rand.weighted.bMNTD.comp[rows,columns,];
#     weighted.bNTI[rows,columns] = (beta.mntd.weighted[rows,columns] - mean(rand.vals)) / sd(rand.vals);
#     rm("rand.vals");
#     
#   };
# };
# 
# rownames(weighted.bNTI) = colnames(match.phylo.otu$data);
# colnames(weighted.bNTI) = colnames(match.phylo.otu$data);
# weighted.bNTI;
# write.csv(weighted.bNTI,"weighted_bNTI.csv",quote=F);
# 
# pdf("weighted_bNTI_Histogram.pdf")
#   hist(weighted.bNTI)
# dev.off()
```

## Introduction

In this notebook we are calculating the βNTI across samples to examine soil bacterial community assembly in a finger lakes region. The data generated here will be used in other notebooks for analysis.

This analysis uses the same (-2, 2) range for significance testing as in Stegen et al. 2012. This means that the following conclusison can be drawn from this data:
  βNTI > 2: Community assembly driven by variable selection
  βNTI < -2: Community assembly driven by homgenizing selection
  |βNTI| < 2: Community assembly is stochastic

Please note that some of these cell have eval=FALSE in the rmarkdown version so that they wont run when knitted. This was done because they take a really long time and are better run in terminal.

### Initiate libraries
```{r}
rm(list=ls())
# Packages needed for analysis
library(dplyr)
library(tidyr)
library(tibble)
library(phyloseq)
library(picante)
library(parallel)
# Packages needed for plotting
library(ggplot2)
```

### Import data

```{r}


# # Import bulk soil phyloseq data
# #bulk.physeq = readRDS("/home/sam/data/fullCyc2_data/bulk_soil_physeq.RDS")
# load("results/cleaned_files/cleanedfiles_3.RData")
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)
# 
# #check number of reads in each sample
# read.num <- data.frame("count" = colSums(otu_table(my.ps))) %>%
#   rownames_to_column(var="Run") %>%
#   inner_join(data.frame(sample_data(my.ps)), by="Run") %>%
#   arrange(-count) %>%
#   mutate(X.Sample=factor(Run, levels=Run))
# 
# #Check to make sure all of our samples are at or above 10000 reads
# ggplot(data=read.num, aes(x=Run, y=log10(count), fill = Glacier)) +
#   geom_bar(stat="identity")+
#   labs(x="Sample", y="Log10(Read count)")+
#   geom_hline(yintercept=log10(10000))+
#   theme(text=element_text(size=16),
#        axis.text.x=element_blank())

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")

#Phyloseq objects
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification,
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.rel.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-abundant.rds")#rarefied relative abundance abundant taxa ONLY
my.ps.rar.rel.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative-rare.rds") #rarefied relative abundance rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")



## Check how many reads you have in each of the samples. This will tell you if you need to re-do anything
# Get read counts and make a new dataframe with this data
# read_count = data.frame("count" = colSums(otu_table(bulk.physeq))) %>%
#   rownames_to_column(var="X.Sample") %>%
#   inner_join(data.frame(sample_data(bulk.physeq)), by="X.Sample") %>%
#   arrange(-count) %>%
#   mutate(X.Sample=factor(X.Sample, levels=X.Sample))
# # Now plot read count for each sample. The horizontal line represents a 2000 read threshold
# ggplot(data=read_count, aes(x=X.Sample, y=log10(count), fill=ecosystem)) +
#   geom_bar(stat="identity") +
#   labs(x="Sample", y="Log10(Read count)") +
#   geom_hline(yintercept=log10(10000)) +
#   theme(text = element_text(size=16),
#         axis.text.x = element_blank())
# # Everything seems to be at or above 10000 total reads
# bulk.physeq
```


#Now we need to rarefy the data to normalize the sequencing depth. We should also get a normalized dataset which gives relative abundance rather than readcounts.

```{r}
# # Rarefy to an even depth
# set.seed(72)  # setting seed for reproducibility
# bulk.physeq.rare = rarefy_even_depth(bulk.physeq)
# # Normalize read counts (this gives relative abundance)
# bulk.physeq.norm = transform_sample_counts(bulk.physeq.rare, function(x) x/sum(x))

# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))

#subset to reduce size of dataset
#my.ps.sub <- subset_samples(my.ps.rar, Glacier == "Taylor")
#Remove ASVs not found in at least 3 samples
# asv.tab <- otu_table(my.ps.rar)
# asv.tab[asv.tab > 0] = 1
# asv.frq <- rowSums(asv.tab)
# asv.frq <- asv.frq[asv.frq > 20]
# my.ps.edit <- prune_taxa(names(asv.frq), my.ps.rar)

# tax.ID <- row.names(my.ps.edit@otu_table)
# tax.to.keep <- sample(tax.ID, 50)
# my.ps.rar.edit <- prune_taxa(tax.to.keep, my.ps.rar)
set.seed(1961)
myTaxa = names(sort(taxa_sums(my.ps.rar), decreasing = TRUE))
tax.to.keep <- sample(myTaxa, 100)
my.ps.edit = prune_taxa(tax.to.keep, my.ps.rar)

#check tree is rooted
is.rooted(phy_tree(my.ps.edit))  #TRUE

# GPr  = transform_sample_counts(my.ps.rar, function(x) x / sum(x) )
# GPfr = filter_taxa(GPr, function(x) mean(x) > 1e-5, TRUE)
# 
# GPr  = transform_sample_counts(GlobalPatterns, function(x) sample(x, 100) )
# GPfr = filter_taxa(my.ps.rar, sample(my.ps.rar, 5))
# 
# sample(my.ps.rar, 5)
# 
# sampsums = sample_sums(my.ps.rar)
# keepsamples = sample_data(my.ps.rar)$SampleType %in% sampletypes
# template = prune_samples(keepsamples, GlobalPatterns)
# 
# GP.chl = subset_taxa(my.ps.rar, Phylum=="Chlamydiae")
# GP.chl = prune_samples(sample_sums(GP.chl)>=20, GP.chl)
```

## Calculating βNTI

### The function
This will the the function used to calculate the βNTI for samples from a phyloseq.

```{r}
# Function for calculating the βMNTD for each random null community
bMNTD_null_func <- function(i, OTU.table, tree){
  tree$tip.label = sample(tree$tip.label)
  bMNTD_s = comdistnt(OTU.table, cophenetic(tree), abundance.weighted = TRUE)
  A <- attr(bMNTD_s, "Size")
  B <- if (is.null(attr(bMNTD_s, "Labels"))) sequence(A) else attr(bMNTD_s, "Labels")
  if (isTRUE(attr(bMNTD_s, "Diag"))) attr(bMNTD_s, "Diag") <- FALSE
  if (isTRUE(attr(bMNTD_s, "Upper"))) attr(bMNTD_s, "Upper") <- FALSE
  bMNTD_s.df = data.frame(Sample_1 = B[unlist(lapply(sequence(A)[-1], function(x) x:A))],
                          Sample_2 = rep(B[-length(B)], (length(B)-1):1),
                          bMNTD = as.vector(bMNTD_s),
                          rep=i)
  return(bMNTD_s.df)
}
# The main function for calculating βNTI
Phylo_turnover <- function(physeq, reps, nproc){
  # Extract OTU table
  OTU.table = t(otu_table(physeq))
  # Extract phylogenetic tree
  tree = phy_tree(physeq)
  # Get βMNTD between all communities
  bMNTD_o = comdistnt(OTU.table, cophenetic(tree), abundance.weighted = TRUE)
  A <- attr(bMNTD_o, "Size")
  B <- if (is.null(attr(bMNTD_o, "Labels"))) sequence(A) else attr(bMNTD_o, "Labels")
  if (isTRUE(attr(bMNTD_o, "Diag"))) attr(bMNTD_o, "Diag") <- FALSE
  if (isTRUE(attr(bMNTD_o, "Upper"))) attr(bMNTD_o, "Upper") <- FALSE
  bMNTD_o.df = data.frame(Sample_1 = B[unlist(lapply(sequence(A)[-1], function(x) x:A))],
                          Sample_2 = rep(B[-length(B)], (length(B)-1):1),
                          bMNTD = as.vector(bMNTD_o))
  
  # Get βMNTD for randomized null communities
  rep.list = seq(1, reps)
  bMNTD_s.df.list = mclapply(rep.list, bMNTD_null_func, OTU.table=OTU.table, tree=tree, mc.cores=nproc)
  
  # Combine all data together and calculate βNTI for each sample pair
  bMNTD_s.df <- do.call("rbind", bMNTD_s.df.list)
  bMNTD_s.means.df = bMNTD_s.df %>%
    group_by(Sample_1, Sample_2) %>%
    dplyr::summarize(mean_bMNTD = mean(bMNTD),
                     sd_bMNTD = sd(bMNTD))
  
  bMNTD_o.df = inner_join(bMNTD_o.df, bMNTD_s.means.df, by=c("Sample_1", "Sample_2")) %>%
    mutate(bNTI = (bMNTD - mean_bMNTD)/sd_bMNTD)
  return(bMNTD_o.df)
}
```


### Running function to get βNTI

Only run this cell once (eval=FALSE) because it take a really long time. Also best to do this in terminal not in RStudio

```{r}
# Get permutation beta null deviations
full.bNTI.df = Phylo_turnover(my.ps.edit, 999, 10)
# Save bNTI table
write.table(full.bNTI.df, "../Outputs/Sommers2020/full_bNTI-total.txt")
```

#REDUCE DATASET - ABUNDANT TAXA

```{r}
set.seed(1961)

myTaxa = names(sort(taxa_sums(my.ps.rar.abun), decreasing = TRUE))
tax.to.keep <- sample(myTaxa, 100)
my.ps.edit.abun = prune_taxa(tax.to.keep, my.ps.rar.abun)

#check tree is rooted
is.rooted(phy_tree(my.ps.edit.abun))  #TRUE
```

### Running function to get βNTI

Only run this cell once (eval=FALSE) because it take a really long time. Also best to do this in terminal not in RStudio

```{r}
# Get permutation beta null deviations
full.bNTI.df = Phylo_turnover(my.ps.edit.abun, 999, 10)
# Save bNTI table
write.table(full.bNTI.df, "../Outputs/Sommers2020/full_bNTI-abundant.txt")
```

#REDUCE DATASET - RARE TAXA

```{r}
set.seed(1961)

myTaxa = names(sort(taxa_sums(my.ps.rar.rare), decreasing = TRUE))
tax.to.keep <- sample(myTaxa, 100)
my.ps.edit.rare = prune_taxa(tax.to.keep, my.ps.rar.rare)

#check tree is rooted
is.rooted(phy_tree(my.ps.edit.rare))  #TRUE
```

### Running function to get βNTI

Only run this cell once (eval=FALSE) because it take a really long time. Also best to do this in terminal not in RStudio

```{r}
# Get permutation beta null deviations
full.bNTI.df = Phylo_turnover(my.ps.edit.rare, 999, 10)
# Save bNTI table
write.table(full.bNTI.df, "../Outputs/Sommers2020/full_bNTI-rare.txt")
```