---
title: "Rare and Abundant"
author: "Amy Solman"
date: "27/04/2021"
output: html_document
---

#Script differentiating between rare and abundant ASVs in a data set
#Some code taken from https://github.com/shuojiao/Rare-and-abundant-bacteria-/blob/master/Alpha-diversity.R 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Clear workspace and load data
```{r}
rm(list=ls())

#install.packages("remotes")
#remotes::install_github("vmikk/metagMisc") #for transforming phyloseq object to data frame
library(metagMisc)
library(tibble) #for subsetting the data
#install.packages("compare")
#library(compare) #for comparing ASV IDs in my dataframes
library(dplyr)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles.RData")

#load phyloseq object
my.ps <- readRDS("../Outputs/Sommers2020/my-phyloseq-object.rds")

```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)

#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)

#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x)) #
```

#Define rare and abundant taxa
Abundant and rare taxa will be classified based on their relative abundance in local and metacommunities (M. Li et al., 2021). Locally abundant ASVs are those that comprise >1% of a sample, while abundant ASVs in the metacommunity are those that comprise >0.1% of average local abundance. Locally rare ASVs are those that comprise <0.01% of a sample, whilst rare ASVs in the metacommunity are those that comprise <0.001% of average local abundance

For bacterial and archaeal communities, abundant and rare ASVs were defined based on their local and regional relative abundance. Locally abundant and rare ASVs were defined as all the ASVs with abundance >1% and <0.01% within a sample, respectively (Galand et al., 2009; Liu et al., 2015b). Subsequently, the average relative abundance of locally abundant and rare ASVs was calculated. Regionally abundant and rare ASVs were defined as all the ASVs with an average abundance >0.1% and <0.001%, respectively (Logares et al., 2014).
```{r}
#convert rarefied phyloseq object with relative abundances to data frame
otu <- phyloseq_to_df(my.ps.rar.rel) 

#get our rarefied relative abundance data
a <- otu[-c(1:8)] 

#get the mean relative abundances of all ASVs across all samples
mean.abun <- data.frame(rowMeans(a, na.rm = TRUE))
row.names(mean.abun) <- otu$OTU#we now have a list of ASV IDs and their mean relative abundances across samples
names(mean.abun)[1] <- "mean_abun"

#Locally abundant ASVs >1% per sample
#Regionally abundance ASVs have mean relative abundance across all samples >0.1%
regional.abun <- mean.abun %>% filter(mean_abun > 0.001)

#Locally rare ASVs <0.01% per sample
#Regionally rare ASVs have mean relative abundance across all samples < 0.001
regional.rare <- mean.abun %>% filter(mean_abun < .00001)


#regionally intermediate ASVs with mean relative abundance across all samples > 0.001% & < 0.1%
regional.inter <- mean.abun %>% filter(mean_abun > 0.00001 & mean_abun < 0.001)

#Make a dataframe with every ASV as a rowname, then if they're either abundant, rare or intermediate in the metacommunity

regional.rare.df <- regional.rare
regional.rare.df$Meta <- "Rare"

regional.abun.df <- regional.abun
regional.abun.df$Meta <- "Abundant"

regional.inter.df <- regional.inter
regional.inter.df$Meta <- "Intermediate"

df <- rbind(regional.rare.df, regional.abun.df, regional.inter.df)
```


#Now let's make a data frame with our locally rare, abundant or intermediate taxa
```{r}

my.data <- data.frame(otu$OTU)
row.names(my.data) <- otu$OTU

for (i in 1:ncol(a)){
  
blah <- data.frame(a)[i]
names(blah) <- "Sample"
row.names(blah) <- otu$OTU

local.abs <- blah %>% filter(Sample == 0)

#Locally abundant = >1% relative abundance per sample
local.abun <- blah %>% filter(Sample > 0.01)

#Locally rare ASVs <0.01% per sample
local.rare <- blah %>% filter(Sample > 0 & Sample < 0.0001)

#locallt intermediate ASVs with relative abundance > 0.01% & < 1%
local.inter <- blah %>% filter(Sample > 0.0001 & Sample < 0.01)

local.abs.df <- local.abs
local.abs.df$Sample <- "Absent"

local.rare.df <- local.rare
local.rare.df$Sample <- "Rare"

local.abun.df <- local.abun
local.abun.df$Sample <- "Abundant"

local.inter.df <- local.inter
local.inter.df$Sample <- "Intermediate"

#df2 <- cbind.data.frame(local.abs.df, local.rare.df, local.abun.df, local.inter.df, row.names = TRUE)
df2 <- rbind(local.abs.df, local.rare.df, local.abun.df, local.inter.df)

names(df2) <- names((a)[i])

#my.data[ , ncol(my.data) + 1] <- df2
my.data <- merge(my.data, df2, by=0, all=TRUE)
row.names(my.data) <- my.data$Row.names
my.data <- my.data[-c(1)]

}

#merge our local abundances with the regional abundance
my.data <- merge(my.data, df, by=0, all=TRUE)
row.names(my.data) <- my.data$Row.names
my.data <- my.data[-c(1,2)]

#Add taxonomic classifications
my.data.tax <- merge(my.data, taxonomy, by=0, all=TRUE)
row.names(my.data.tax) <- my.data.tax$Row.names
my.data.tax <- my.data.tax[-c(1)]

#remove rows with NA in any particular sample (these are the ASVs that got removed from the count table in the rarification step. Remember that they were NOT removed from the taxonomy table)
my.data_complete <- my.data.tax[complete.cases(my.data.tax$SRR12805327), ]

```

#Make new phyloseq objects with total, abundant and rare microbial communities
```{r}
#Subset the count_table_clean, metadata, taxonomy and tree_clean dataframes by only ASVs that are abundant in the metacommunity
abun.IDs <- my.data_complete[my.data_complete$Meta== "Abundant",]
abun.IDs <- row.names(abun.IDs)
abun.count <- subset(count_table_clean.tax, rownames(count_table_clean.tax) %in% abun.IDs)
abun.tax <- subset(taxonomy, rownames(taxonomy) %in% abun.IDs)
abun.tree <- prune_taxa(abun.IDs, tree_clean)

#remove taxonomy from the count table
abun.count.no.tax <- abun.count[-c(35:41)]

#make a phyloseq object with just abundant taxa
abun.ASV = otu_table(data.frame(abun.count.no.tax), taxa_are_rows = TRUE)
abun.TAX = tax_table(as.matrix(abun.tax))
abun.META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
ps.abun <- phyloseq(abun.ASV, abun.TAX, abun.META, abun.tree)

#Subset the count_table_clean, metadata, taxonomy and tree_clean dataframes by only ASVs that are rare in the metacommunity
rare.IDs <- my.data_complete[my.data_complete$Meta== "Rare",]
rare.IDs <- row.names(rare.IDs)
rare.count <- subset(count_table_clean.tax, rownames(count_table_clean.tax) %in% abun.IDs)
rare.tax <- subset(taxonomy, rownames(taxonomy) %in% rare.IDs)
rare.tree <- prune_taxa(rare.IDs, tree_clean)

#remove taxonomy from the count table
rare.count.no.tax <- rare.count[-c(35:41)]

#make a phyloseq object with just abundant taxa
rare.ASV = otu_table(data.frame(rare.count.no.tax), taxa_are_rows = TRUE)
rare.TAX = tax_table(as.matrix(rare.tax))
rare.META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
ps.rare <- phyloseq(rare.ASV, rare.TAX, rare.META, rare.tree)
```

#Export data: I need the phyloseq object with original data, the rarefied phyloseq object, the rarefied relative abundance phyloseq object. I also want my metadata table, fasta file, count_table with taxonomy, phylogenetic table and abundance table with taxonomy.
```{r}

#r data object
save(metadata, fasta_clean, count_table_clean.tax, taxonomy, tree_clean, my.data_complete, file = "../Outputs/Sommers2020/cleanedfiles-abundance.RData")

#Now I'm going to export my phyloseq object
saveRDS(my.ps, "../Outputs/Sommers2020/my-ps-original.rds")
saveRDS(my.ps.rar, "../Outputs/Sommers2020/my-ps-rarefied-total.rds") #this is essentially the total community as the abundant and rare phyloseq object are also rarefied which is important for downstream analysis
saveRDS(my.ps.rar.rel, "../Outputs/Sommers2020/my-ps-rarefied-relative.rds")
saveRDS(ps.abun, "../Outputs/Sommers2020/my-ps-rarefied-abundant.rds")
saveRDS(ps.rare, "../Outputs/Sommers2020/my-ps-rarefied-rare.rds")


```