---
title: "MVA Analysis - Alpha"
author: "Amy Solman"
date: "21/04/2021"
output: html_document
---

#Script developed from that of BArnett et al., (2020) code available at https://github.com/seb369/landuse_comm_assembly/blob/master/Diversity_measures.md 
#This script will calculate alpha diversities 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

library(dplyr)
library(tibble)
library(vegan)
library(ggpubr)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
#load("results/cleaned_files/cleanedfiles_3.RData")

#Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")

#load phyloseq objects
# my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
# my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied.rds") #rarefied dataset, total taxa
# my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance dataset
# my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rda") #rarefied abundant taxa ONLY
# my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY

my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity of each sample by OTU richness
```{r}
####FOR TOTAL DATA 
#get the metadata table
tot.my.metadata <- data.frame(sample_data(my.ps.rar))

#get the rarefied ASV table
tot.ASV.table.bin <- otu_table(my.ps.rar)

#for any ASV count that does NOT equal 0, make it 1 (so make it into a presence/absence table)
tot.ASV.table.bin[tot.ASV.table.bin != 0] = 1

#make a new dataframe called ASV count with a column called "num"
#which is the column sums from the presence/absence table
#Give the columns the sample names
#join the metadata by sample name
#arrange by sample name
tot.ASV.count = data.frame("num" = colSums(tot.ASV.table.bin)) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps.rar)), by="Run") %>%
  arrange(Run) %>%
  mutate(sample_ID=factor(Run, levels = Run))


###FOR ABUNDANT DATA
#get the metadata table
abun.my.metadata <- data.frame(sample_data(my.ps.rar.abun))

#get the rarefied ASV table
abun.ASV.table.bin <- otu_table(my.ps.rar.abun)

#for any ASV count that does NOT equal 0, make it 1 (so make it into a presence/absence table)
abun.ASV.table.bin[abun.ASV.table.bin != 0] = 1

#make a new dataframe called ASV count with a column called "num"
#which is the column sums from the presence/absence table
#Give the columns the sample names
#join the metadata by sample name
#arrange by sample name
abun.ASV.count = data.frame("num" = colSums(abun.ASV.table.bin)) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps.rar.abun)), by="Run") %>%
  arrange(Run) %>%
  mutate(sample_ID=factor(Run, levels = Run))

###FOR RARE DATA 
#get the metadata table
rare.my.metadata <- data.frame(sample_data(my.ps.rar.rare))

#get the rarefied ASV table
rare.ASV.table.bin <- otu_table(my.ps.rar.rare)

#for any ASV count that does NOT equal 0, make it 1 (so make it into a presence/absence table)
rare.ASV.table.bin[rare.ASV.table.bin != 0] = 1

#make a new dataframe called ASV count with a column called "num"
#which is the column sums from the presence/absence table
#Give the columns the sample names
#join the metadata by sample name
#arrange by sample name
rare.ASV.count = data.frame("num" = colSums(rare.ASV.table.bin)) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps.rar.rare)), by="Run") %>%
  arrange(Run) %>%
  mutate(sample_ID=factor(Run, levels = Run))
```

#Step Five: #Compare ASV richness (alpha diversity) between glaciers
```{r}

####FOR TOTAL DATA 
#First get our summary statistics
tot.summary.stats <- group_by(tot.ASV.count, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(num, na.rm = TRUE),
    sd = sd(num, na.rm = TRUE)
  )

#Then plot our glaciers
tot.my.plot <- ggboxplot(tot.ASV.count, x = "Glacier", y = "num",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="ASV Num", xlab = "Glacier")

tot.my.plot
ggsave("../Outputs/Sommers2020/graphs/ASV-richness-boxplot.pdf", width=6, height = 4)

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(tot.ASV.count, shapiro.test(num[Glacier == "Canada"])) #p=0.56 (p >0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.
with(tot.ASV.count, shapiro.test(num[Glacier == "Taylor"])) #p=0.97 (p >0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES?
#Use F-test to test for homogeneity in variances
tot.f.test <- var.test(num ~ Glacier, data = tot.ASV.count)
tot.f.test #p value = 0.81 which is greater than 0.05 so there is no significant difference in the variances of the two datasets

#ALL THREE ASSUMPTIONS ARE MET FOR THE PARAMTERIC TWO INDEPENDENT SAMPLES TEST

#COMPARE THE MEANS OF TWO INDEPENDENT GROUPS
#UNPAIRED TWO SAMPLES T-TEST (PARAMETERIC)
tot.my.t.test <- t.test(num ~ Glacier, data = tot.ASV.count, var.equal = TRUE)
tot.my.t.test #p = 0.057 so p > 0.05, there is no significant difference between the ASV richnesses of the two glaciers

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
tot.my.wilcoxon.test <- wilcox.test(num ~ Glacier, data = tot.ASV.count, exact = FALSE)
tot.my.wilcoxon.test # p = 0.04479 so p < 0.05, there is a significant difference between the ASV richness of the two glaciers.

#Even though the non-parametric test gave me a significant result, where the parameteric test did not, I should use the parametric test because it is more efficient.

####USE THIS WHEN COMPARING THREE OR MORE INDEPENDENT GROUPS WITH NON-PARAMETRIC DATA###
# writeLines("Kruskal-Wallis test:")
# k.test = kruskal.test(count ~ Glacier, data = ASV.count)
# print(k.test)

#The below Dunn's test can be used when comparing three or more groups
#That is not useful now.

# if (k.test$p.value <= 0.05){
#   writeLines(c("", "There is a significant effect of sample type (Glacier) on ASV number. Now running Dunn test for pairwise comparisons:"))
#   d.test = dunnTest(count ~ Glacier, data = ASV.count, method="bh")
#   print(d.test)
# } else {
#   writeLines(c("", "We did not find a significant effect of sample type (Glacier) on ASV number."))
# }

####FOR ABUNDANT DATA 
#First get our summary statistics
abun.summary.stats <- group_by(abun.ASV.count, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(num, na.rm = TRUE),
    sd = sd(num, na.rm = TRUE)
  )

#Then plot our glaciers
abun.my.plot <- ggboxplot(abun.ASV.count, x = "Glacier", y = "num",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="ASV Num", xlab = "Glacier")

abun.my.plot

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(abun.ASV.count, shapiro.test(num[Glacier == "Canada"])) #p=0.001 (p < 0.05) so the distribution of the data are significantly different from normal distribution so we can use NON-PARAMETRIC tests.
with(abun.ASV.count, shapiro.test(num[Glacier == "Taylor"])) #p=0.85 (p >0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.

#Because one group was not normally distributed we'll use the non-parametric test

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES?
#Use F-test to test for homogeneity in variances
abun.f.test <- var.test(num ~ Glacier, data = abun.ASV.count)
abun.f.test #p value = 0.81 which is greater than 0.05 so there is no significant difference in the variances of the two datasets

#ALL THREE ASSUMPTIONS ARE NOT MET FOR THE PARAMTERIC TWO INDEPENDENT SAMPLES TEST

#COMPARE THE MEANS OF TWO INDEPENDENT GROUPS
#UNPAIRED TWO SAMPLES T-TEST (PARAMETERIC)
# abun.my.t.test <- t.test(num ~ Glacier, data = abun.ASV.count, var.equal = TRUE)
# abun.my.t.test #p = 0.0518 so p > 0.05, there is no significant difference between the ASV richnesses of the two glaciers

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
abun.my.wilcoxon.test <- wilcox.test(num ~ Glacier, data = abun.ASV.count, exact = FALSE)
abun.my.wilcoxon.test # p = 0.053 so p > 0.05, there is no significant difference between the ASV richness of the two glaciers.

####FOR RARE DATA 
#First get our summary statistics
rare.summary.stats <- group_by(rare.ASV.count, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(num, na.rm = TRUE),
    sd = sd(num, na.rm = TRUE)
  )

#Then plot our glaciers
rare.my.plot <- ggboxplot(rare.ASV.count, x = "Glacier", y = "num",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="ASV Num", xlab = "Glacier")

rare.my.plot

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(rare.ASV.count, shapiro.test(num[Glacier == "Canada"])) #p=0.61 (p >0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.
with(rare.ASV.count, shapiro.test(num[Glacier == "Taylor"])) #p=0.067 (p >0.05) so the distribution of the data are not significantly different from normal distribution so we can use PARAMETRIC tests.

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES?
#Use F-test to test for homogeneity in variances
rare.f.test <- var.test(num ~ Glacier, data = rare.ASV.count)
rare.f.test #p value = 0.73 which is greater than 0.05 so there is no significant difference in the variances of the two datasets

#ALL THREE ASSUMPTIONS ARE MET FOR THE PARAMTERIC TWO INDEPENDENT SAMPLES TEST

#COMPARE THE MEANS OF TWO INDEPENDENT GROUPS
#UNPAIRED TWO SAMPLES T-TEST (PARAMETERIC)
rare.my.t.test <- t.test(num ~ Glacier, data = rare.ASV.count, var.equal = TRUE)
rare.my.t.test #p = 0.31 so p > 0.05, there is no significant difference between the ASV richnesses of the two glaciers

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
# rare.my.wilcoxon.test <- wilcox.test(num ~ Glacier, data = rare.ASV.count, exact = FALSE)
# rare.my.wilcoxon.test # p = 0.04479 so p < 0.05, there is a significant difference between the ASV richness of the two glaciers.

# #let's look at the three plots together
# par(mfrow=(1,3))
# tot.my.plot
# abun.my.plot
# rare.my.plot
```

#Put all ASV richness stats together into a dataframe
```{r}
tot.summary.stats$Abundance <- "Total"
abun.summary.stats$Abundance <- "Abundant"
rare.summary.stats$Abundance <- "Rare"
asv.richness.stats.total <- rbind(tot.summary.stats, abun.summary.stats, rare.summary.stats)

```
#Step Six: Plot the data
```{r}
tot.plot <- ggplot(data=tot.ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="ASV Rich", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))

abun.plot <- ggplot(data=abun.ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="ASV Rich", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))

rare.plot <- ggplot(data=rare.ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="ASV Rich", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))

ggarrange(tot.plot, abun.plot, rare.plot + rremove("x.text"), 
          labels = c("Total", "Abundant", "Rare"),
          ncol = 2, nrow = 2)
ggsave("../Outputs/Sommers2020/graphs/ASV-Richness-total-abundant-rare.pdf")
```

#Step Seven: Calculate alpha diversity of each sample by Shannon Diversity index
```{r}
###TOTAL TAXA
#make a new dataframe with shannon as the shannon diversity index of each sample
tot.ASV.table.bin = t(otu_table(my.ps.rar))
tot.shannon.df = data.frame(shannon=diversity(tot.ASV.table.bin, index="shannon")) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps.rar)), by="Run") %>%
  arrange(Glacier)

###ABUNDANT TAXA
#make a new dataframe with shannon as the shannon diversity index of each sample
abun.ASV.table.bin = t(otu_table(my.ps.rar.abun))
abun.shannon.df = data.frame(shannon=diversity(abun.ASV.table.bin, index="shannon")) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps.rar.abun)), by="Run") %>%
  arrange(Glacier)

###RARE TAXA
#make a new dataframe with shannon as the shannon diversity index of each sample
rare.ASV.table.bin = t(otu_table(my.ps.rar.rare))
rare.shannon.df = data.frame(shannon=diversity(rare.ASV.table.bin, index="shannon")) %>%
  rownames_to_column(var="Run") %>%
  inner_join(data.frame(sample_data(my.ps.rar.rare)), by="Run") %>%
  arrange(Glacier)
```

#Step Eight: #Compare Shannon index (alpha diversity) between glaciers
```{r}
###TOTAL TAXA 
#First get our summary statistics
tot.summary.stats.shan <- group_by(tot.shannon.df, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(shannon, na.rm = TRUE),
    sd = sd(shannon, na.rm = TRUE)
  )

#Then plot our glaciers
tot.my.plot.shan <- ggboxplot(tot.shannon.df, x = "Glacier", y = "shannon",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Shannon Index", xlab = "Glacier")

tot.my.plot.shan

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(tot.shannon.df, shapiro.test(shannon[Glacier == "Canada"])) #p=0.043 (p < 0.05) so the distribution of the data are significantly different from normal distribution so must use a NON-PARAMETRIC test.
with(tot.shannon.df, shapiro.test(shannon[Glacier == "Taylor"])) #p=0.71 (p > 0.05) so the distribution of the data are not significantly different from normal distribution, but because the canada data is NOT normally distributed we'll use a non-parametric test

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES? <- Ignore because Canada data is non-parametric

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
tot.my.wilcoxon.test.shan <- wilcox.test(shannon ~ Glacier, data = tot.shannon.df, exact = FALSE)
tot.my.wilcoxon.test.shan # p = 0.5808 so p > 0.05, there is no significant difference between the shannon diversity indexes of the two glaciers.

###ABUNDANT TAXA
#First get our summary statistics
abun.summary.stats.shan <- group_by(abun.shannon.df, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(shannon, na.rm = TRUE),
    sd = sd(shannon, na.rm = TRUE)
  )

#Then plot our glaciers
abun.my.plot.shan <- ggboxplot(abun.shannon.df, x = "Glacier", y = "shannon",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Shannon Index", xlab = "Glacier")

abun.my.plot.shan

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(abun.shannon.df, shapiro.test(shannon[Glacier == "Canada"])) #p=0.008 (p < 0.05) so the distribution of the data are significantly different from normal distribution so must use a NON-PARAMETRIC test.
with(abun.shannon.df, shapiro.test(shannon[Glacier == "Taylor"])) #p=0.78 (p > 0.05) so the distribution of the data are not significantly different from normal distribution, but because the canada data is NOT normally distributed we'll use a non-parametric test

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES? <- Ignore because Canada data is non-parametric

#UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
#Let's look at the non-parametric result
abun.my.wilcoxon.test.shan <- wilcox.test(shannon ~ Glacier, data = abun.shannon.df, exact = FALSE)
abun.my.wilcoxon.test.shan # p = 0.5808 so p > 0.05, there is no significant difference between the shannon diversity indexes of the two glaciers.


###RARE TAXA
#First get our summary statistics
rare.summary.stats.shan <- group_by(rare.shannon.df, Glacier) %>%
  summarise(
    count = n(),
    mean = mean(shannon, na.rm = TRUE),
    sd = sd(shannon, na.rm = TRUE)
  )

#Then plot our glaciers
rare.my.plot.shan <- ggboxplot(rare.shannon.df, x = "Glacier", y = "shannon",
          color="Glacier", palette = c("#00AFBB", "#E7B800"),
          ylab="Shannon Index", xlab = "Glacier")

rare.my.plot.shan

#ASUMPTION 1: ARE THE GROUPS INDEPENDENT?
#Yes the glacier samples are independent.

#ASSUMPTION 2: ARE THE GROUPS NORMALLY DISTRIBUTED?
#Should I use parameteric or non-parametric tests?
#We're going to use a Shapiro-Wilk's Test to see if the data is normally distributed for each group

with(rare.shannon.df, shapiro.test(shannon[Glacier == "Canada"])) #p=0.075 (p > 0.05) so the distribution of the data are not significantly different from normal distribution so must use a PARAMETRIC test.
with(rare.shannon.df, shapiro.test(shannon[Glacier == "Taylor"])) #p=0.09 (p > 0.05) so the distribution of the data are not significantly different from normal distribution

#ASSUMPTION 3: DO THE TWO POPULATIONS HAVE THE SAME VARIANCES?
#Use F-test to test for homogeneity in variances
rare.f.test.shan <- var.test(shannon ~ Glacier, data = rare.shannon.df)
rare.f.test.shan #p value = 0.9 which is greater than 0.05 so there is no significant difference in the variances of the two datasets

# #UNPARIED TWO SAMPLES WILCOXON TEST (NON-PARAMTEROC)
# #Let's look at the non-parametric result
# rare.my.wilcoxon.test.shan <- wilcox.test(shannon ~ Glacier, data = rare.shannon.df, exact = FALSE)
# rare.my.wilcoxon.test.shan # p = 0.5808 so p > 0.05, there is no significant difference between the shannon diversity indexes of the two glaciers.

#COMPARE THE MEANS OF TWO INDEPENDENT GROUPS
#UNPAIRED TWO SAMPLES T-TEST (PARAMETERIC)
rare.my.t.test.shan <- t.test(shannon ~ Glacier, data = rare.shannon.df, var.equal = TRUE)
rare.my.t.test.shan #p = 0.31 so p > 0.05, there is no significant difference between the ASV richnesses of the two glaciers
```


#Put all ASV richness stats together into a dataframe
```{r}
tot.summary.stats.shan$Abundance <- "Total"
abun.summary.stats.shan$Abundance <- "Abundant"
rare.summary.stats.shan$Abundance <- "Rare"
shannon.stats.total <- rbind(tot.summary.stats.shan, abun.summary.stats.shan, rare.summary.stats.shan)

```

#Step Nine: Plot the data
```{r}
# ggplot(data=shannon.df, aes(x=Glacier, y=shannon))+
#   geom_boxplot(outlier.shape = 8)+
#   geom_jitter(aes(color=Glacier), width=0.2, height=0)+
#   labs(x="Glacier", y="Shannon Index", colour="Glacier")+
#   theme(axis.text.x = element_text(angle=90, vjust=0.5))

tot.plot.shan <- ggplot(data=tot.shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))

abun.plot.shan <- ggplot(data=abun.shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))

rare.plot.shan <- ggplot(data=rare.shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = 8)+
  geom_jitter(aes(color=Glacier), width=0.2, height=0)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme(axis.text.x = element_text(angle=90, vjust=0.5))

ggarrange(tot.plot.shan, abun.plot.shan, rare.plot.shan + rremove("x.text"), 
          labels = c("Total", "Abundant", "Rare"),
          ncol = 2, nrow = 2)
ggsave("../Outputs/Sommers2020/graphs/Shannon-total-abundant-rare.pdf", width=6, height=6)
```

#Step 10: Plot ASV richness and shannon diversity together
```{r}
###TOTAL TAXA

#Glacier shapes
Glacier.shapes = c("Canada" = 15, "Taylor" = 16)

tot.rich.plot = ggplot(data=tot.ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(color=Glacier, shape= Glacier), width=0.2, height=0)+
  scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="ASV Rich", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

tot.shan.plot = ggplot(data=tot.shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour=Glacier, shape= Glacier), width=0.2, height=0)+
   scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

fonttheme = theme(legend.text = element_text(size=12),
                  legend.title = element_text(size=12),
                  axis.text = element_text(size=12),
                  axis.title = element_text(size=14))

#Merge the plots together

tot.alpha.plot = cowplot::plot_grid(tot.rich.plot + fonttheme, tot.shan.plot + fonttheme, ncol=2, rel_widths = c(1, 1), labels = c("A", "B"))
tot.alpha.plot
ggsave("../Outputs/Sommers2020/graphs/total-alpha-diversity.pdf", width=8, height=4)

###ABUNDANT TAXA

#Glacier shapes
Glacier.shapes = c("Canada" = 15, "Taylor" = 16)

abun.rich.plot = ggplot(data=abun.ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(color=Glacier, shape= Glacier), width=0.2, height=0)+
  scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="ASV Rich", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

abun.shan.plot = ggplot(data=abun.shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour=Glacier, shape= Glacier), width=0.2, height=0)+
   scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

fonttheme = theme(legend.text = element_text(size=12),
                  legend.title = element_text(size=12),
                  axis.text = element_text(size=12),
                  axis.title = element_text(size=14))

#Merge the plots together

abun.alpha.plot = cowplot::plot_grid(abun.rich.plot + fonttheme, abun.shan.plot + fonttheme, ncol=2, rel_widths = c(1, 1), labels = c("A", "B"))
abun.alpha.plot
ggsave("../Outputs/Sommers2020/graphs/abundant-alpha-diversity.pdf", width=8, height=4)

###RARE TAXA

#Glacier shapes
Glacier.shapes = c("Canada" = 15, "Taylor" = 16)

rare.rich.plot = ggplot(data=rare.ASV.count, aes(x=Glacier, y=num))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(color=Glacier, shape= Glacier), width=0.2, height=0)+
  scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="ASV Rich", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

rare.shan.plot = ggplot(data=rare.shannon.df, aes(x=Glacier, y=shannon))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(colour=Glacier, shape= Glacier), width=0.2, height=0)+
   scale_shape_manual(values = Glacier.shapes)+
  labs(x="Glacier", y="Shannon Index", colour="Glacier")+
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        legend.position="none")

fonttheme = theme(legend.text = element_text(size=12),
                  legend.title = element_text(size=12),
                  axis.text = element_text(size=12),
                  axis.title = element_text(size=14))

#Merge the plots together

rare.alpha.plot = cowplot::plot_grid(rare.rich.plot + fonttheme, rare.shan.plot + fonttheme, ncol=2, rel_widths = c(1, 1), labels = c("A", "B"))
rare.alpha.plot
ggsave("../Outputs/Sommers2020/graphs/rare-alpha-diversity.pdf", width=8, height=4)
```

