---
title: "Linear Regression Models"
author: "Amy Solman"
date: "22/04/2021"
output: html_document
---

#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook (https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/14-regress.ipynb)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#We're going to use linear regression models 
#Linear regression models are useful, especially if both your response (dependent) and predictor (independent) variables are continuous.
#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
# library(vegan)
# library(ggrepel)
# install.packages("repr")
library(repr) #for plotting our pairs of data
# library("picante")
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```
#Step Six: Plot richness with environmental data 
```{r}
#let's start with basic plots of richness with our evironmental variables
rich.chlo <- plot(my.data.sub$Chlorophyll, my.data.sub$Richness)
rich.depth <- plot(my.data.sub$Depth, my.data.sub$Richness)
rich.dna <- plot(my.data.sub$DNA_per_g, my.data.sub$Richness)
rich.org <- plot(my.data.sub$org_matter, my.data.sub$Richness)
rich.ph <- plot(my.data.sub$pH, my.data.sub$Richness)
rich.water <- plot(my.data.sub$water_content, my.data.sub$Richness)
rich.area <- plot(my.data.sub$area, my.data.sub$Richness)

#indices of other alpha metrics
alpha.met <- c(8, 10, 11, 12, 13, 14)
pairs(my.data.sub[, -c(alpha.met)]) #so we can do a pairs plot with just richness and environmental data

```

#Step Seven: Look for correlations between richness and environmental data
```{r}
#cor looks for correlations between pairs of variables using Pearson's correlation coefficient
cor(my.data.sub[, -c(alpha.met)], use="pairwise")

#or we can use cor.test to compair a single pair of variables - this also carries out a t.test to see if the correlation is significant
cor.test(my.data.sub$Richness, my.data.sub$Chlorophyll, use="pairwise") #p >0.05
cor.test(my.data.sub$Richness, my.data.sub$Depth, use="pairwise") #p < 0.05
cor.test(my.data.sub$Richness, my.data.sub$DNA_per_g, use="pairwise") #p > 0.05
cor.test(my.data.sub$Richness, my.data.sub$org_matter, use="pairwise") #p < 0.05
cor.test(my.data.sub$Richness, my.data.sub$pH, use="pairwise") #p > 0.05
cor.test(my.data.sub$Richness, my.data.sub$water_content, use="pairwise") #p > 0.05
cor.test(my.data.sub$Richness, my.data.sub$area, use="pairwise") #p < 0.05

#only cryoconite hole depth, area and organic matter content are significant correlated with ASV richness

#WARNING The above correlation coefficient calculation assumes a straight line relationship
```

#Step Eight: Re-run above correlation analysis with log transformed data
```{r}
my.data.sub$logRichness <- log(my.data.sub$Richness)
my.data.sub$logChlorophyll <- log(my.data.sub$Chlorophyll)
my.data.sub$logDepth <- log(my.data.sub$Depth)
my.data.sub$logDNA_per_g <- log(my.data.sub$DNA_per_g)
my.data.sub$logorg_matter <- log(my.data.sub$org_matter)
my.data.sub$logpH <- log(my.data.sub$pH)
my.data.sub$logwater_content <- log(my.data.sub$water_content)
my.data.sub$logarea <- log(my.data.sub$area)

my.log.alpha <- c(15, 16, 17, 18, 19, 20, 21, 22)
pairs(my.data.sub[, (my.log.alpha)])

cor(my.data.sub[, (my.log.alpha)], use="pairwise")

```

#Step Nine: Plot the data 
```{r}
myColours <- c('red', 'blue')
mySymbols <- c(1,3)

my.data.sub$Glacier <- as.factor(my.data.sub$Glacier)

plot(logRichness ~ area, data = my.data.sub,
      col = myColours[Glacier], pch=mySymbols[Glacier],
      xlab = "Hole Area", ylab = "log(ASV Richness)")

legend("topleft", legend=levels(my.data.sub$Glacier),
       col=myColours, pch=mySymbols)

#Visual inspection of the plots shows the glaciers have similar but slightly different relationships between hoel area and log(ASV Richness)

```
#Step Nine: Create linear models of richness and area
```{r}
#This is a null model of the logRichness relationship on Canada Glacier
nullModelRichnessCan <- lm(logRichness ~ 1, data = my.data.sub, subset = Glacier == "Canada")

#This is the linear regression model of the logarea relationship on Canada Glacier
RichnessAreaModelCan <- lm(logRichness ~ logarea, data = my.data.sub, subset = Glacier == "Canada")

#let's look at the outputs of our model
summary(RichnessAreaModelCan)

#Info we have in this table:
#Call - the model description
#Residuals - a summary of the residuals
#Coefficients - a table of coefficients
#Info on residual error
#R-squared
#F-statistic
#P-value

#there are two rows in the coefficients table
#these are the two coefficients of our equation y = a + b * x where a is the intercept and b is the slope. We get the estimated value of the intercept and the slope, the standard error pf those estimates, the t test value (and associated p value) of whether the slope and intercept are significantly different from zero.

```
#Step Ten: Look at the terms of the model using the anova function
#Anova test how much variation in the response variable (richness) is explained by each explanatory variable (area)
```{r}
anova(RichnessAreaModelCan)
#So here the ANOVA table is comparing the vairation in log Richness explained by log area to the total variation in log Richness.
#We want to see how much smaller the residuals are for the area model than the null model in explaining Richness.
#In the ANOVA table we have:
#DF: The degrees of freedom, each parameter/coefficient takes up a degree of freedom. This model adds area (1 parameter) to the null model, so has 1 DF
#Sum Sq: The sums of squares. The top line shows the variation in richness explained by area. The bottom line shows the residual sum of squares for the model.
#Mean Sq: Sum of squares values divided by the degrees of freedom. So if we explain lots of variation in the model with one coefficient (null model) that's great! If we explain a little variation with a loss of degrees of freedom by adding more parameters that's not so good...
#F value: Ratio of the mean sq for the variable and residual mean sq. Can show us if the explained variation is large or small. 
#Pr(>F): The p value shows us how likely the model explained this much variance by chance.

#We can get the sums of squares of the residuals from the null and area model using the function resid, square them and add them
sum(resid(nullModelRichnessCan) ^ 2) #0.81
sum(resid(RichnessAreaModelCan) ^ 2) #0.28 - is one is the same as seen in our ANOVA table

#TAKE AWAY: P < 0.05 so area explains a large amount of variation in ASV richness
```

#Step Eleven: Repeat the above for Taylor Glacier
```{r}
#This is a null model of the logRichness relationship on Taylor Glacier
nullModelRichnessTay <- lm(logRichness ~ 1, data = my.data.sub, subset = Glacier == "Taylor")

#This is the linear regression model of the logarea relationship on Taylor Glacier
RichnessAreaModelTay <- lm(logRichness ~ logarea, data = my.data.sub, subset = Glacier == "Taylor")

#let's look at the outputs of our model
summary(RichnessAreaModelTay)

#get ANOVA table
anova(RichnessAreaModelTay)

#We can get the sums of squares of the residuals from the null and area model using the function resid, square them and add them
sum(resid(nullModelRichnessTay) ^ 2) #0.67
sum(resid(RichnessAreaModelTay) ^ 2) #0.10 - is one is the same as seen in our ANOVA table

#TAKE AWAY: P < 0.05 so area explains a large amount of variation in ASV richness
```

#Step Twelve: Plot the fitted regression model
```{r}
myCol <- c('red', 'blue')
plot(logRichness ~ logarea, data = my.data.sub, col=myCol[Glacier], xlab='log Area', ylab = 'log Richness')
abline(RichnessAreaModelCan, col='red')
abline(RichnessAreaModelTay, col='blue')

#Example model reports:
#log area explains significant variation in log Richness on Canada Glacier (F=35.1, df=1,19, p<0.01) and shows that species richness increases with hole area (intercept: 5.0, se=0.16; slope: 0.13, se=0.02).

#log area explains significant variation in log Richness on Taylor Glacier (F=29.8, df=1,5, p=0.003) and shows that species richness increases with hole area (intercept: 4.3, se=0.26; slope: 0.21, se=0.04).
```

#Step Thirteen: Model Diagnostics
#We need to check that our models are appropriate for our data
#We will use diagnostic plots to do this.
```{r}
par(mfrow=c(2,2), mar=c(5,5,1.5,1.5))
plot(RichnessAreaModelCan)

par(mfrow=c(2,2), mar=c(5,5,1.5,1.5))
plot(RichnessAreaModelTay)

#Residuals vs Fitted: This plot looks to see if the distribution of the residuals has a similar variance for different predited values. We want NO obvious pattern or big gaps. 
#Normal Q-Q: This shows if the residuals are normally distributed. We want the points as close to a straight line as possible.
#Scale-Location: x axis is the fitted values, y axis is the square root of the standardized residuals. Basically we want the red line to be relatively flat.
#Residuals vs Leverage: This shows the standardized residuals against leverage - leverage means how much eaxch data point influences the linear model's coefficient estimates. Numbered points are those to pay special attention to. They are potential outliers

```

#Step Fourteen: Repeat the above using Faith's diversity metric
```{r}
#First get logs of our diversity metrics
my.data.sub$logFaith <- log(my.data.sub$Faith)
my.data.sub$logShannon <- log(my.data.sub$Shannon)
my.data.sub$logSimpson <- log(my.data.sub$Simpson)
my.data.sub$logInvSimp <- log(my.data.sub$Inv.Simpson)
my.data.sub$logPilou <- log(my.data.sub$Pilou)

#Linear Regression
FaithAreaModelCan <- lm(logFaith ~ logarea, data = my.data.sub, subset = Glacier == "Canada")
#Summary
summary(FaithAreaModelCan)
#get ANOVA table
anova(FaithAreaModelCan)

#Linear Regression
FaithAreaModelTay <- lm(logFaith ~ logarea, data = my.data.sub, subset = Glacier == "Taylor")
#Summary
summary(FaithAreaModelTay)
#get ANOVA table
anova(FaithAreaModelTay)

myCol <- c('red', 'blue')
plot(logFaith ~ logarea, data = my.data.sub, col=myCol[Glacier], xlab='log Area', ylab = 'log Faith')
abline(FaithAreaModelCan, col='red')
abline(FaithAreaModelTay, col='blue')
```

#Step Fifteen: Repeat using Shannon
```{r}
#Linear Regression
ShanAreaModelCan <- lm(logShannon ~ logarea, data = my.data.sub, subset = Glacier == "Canada")
#Summary
summary(ShanAreaModelCan)
#get ANOVA table
anova(ShanAreaModelCan)

#Linear Regression
ShanAreaModelTay <- lm(logShannon ~ logarea, data = my.data.sub, subset = Glacier == "Taylor")
#Summary
summary(ShanAreaModelTay)
#get ANOVA table
anova(ShanAreaModelTay)

myCol <- c('red', 'blue')
plot(logShannon ~ logarea, data = my.data.sub, col=myCol[Glacier], xlab='log Area', ylab = 'log Shannon')
abline(ShanAreaModelCan, col='red')
abline(ShanAreaModelTay, col='blue')
```

#Step Fifteen: Repeat using Simpson
```{r}
#Linear Regression
SimpAreaModelCan <- lm(logSimpson ~ logarea, data = my.data.sub, subset = Glacier == "Canada")
#Summary
summary(SimpAreaModelCan)
#get ANOVA table
anova(SimpAreaModelCan)

#Linear Regression
SimpAreaModelTay <- lm(logSimpson ~ logarea, data = my.data.sub, subset = Glacier == "Taylor")
#Summary
summary(SimpAreaModelTay)
#get ANOVA table
anova(SimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logSimpson ~ logarea, data = my.data.sub, col=myCol[Glacier], xlab='log Area', ylab = 'log Simpson')
abline(SimpAreaModelCan, col='red')
abline(SimpAreaModelTay, col='blue')
```

#Step Fifteen: Repeat using Inv Simpson
```{r}
#Linear Regression
InvSimpAreaModelCan <- lm(logInvSimp ~ logarea, data = my.data.sub, subset = Glacier == "Canada")
#Summary
summary(InvSimpAreaModelCan)
#get ANOVA table
anova(InvSimpAreaModelCan)

#Linear Regression
InvSimpAreaModelTay <- lm(logInvSimp ~ logarea, data = my.data.sub, subset = Glacier == "Taylor")
#Summary
summary(InvSimpAreaModelTay)
#get ANOVA table
anova(InvSimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logInvSimp ~ logarea, data = my.data.sub, col=myCol[Glacier], xlab='log Area', ylab = 'log Inv Simpson')
abline(InvSimpAreaModelCan, col='red')
abline(InvSimpAreaModelTay, col='blue')
```

#Step Fifteen: Repeat using Pilou
```{r}
#Linear Regression
PilouAreaModelCan <- lm(logPilou ~ logarea, data = my.data.sub, subset = Glacier == "Canada")
#Summary
summary(PilouAreaModelCan)
#get ANOVA table
anova(PilouAreaModelCan)

#Linear Regression
PilouAreaModelTay <- lm(logPilou ~ logarea, data = my.data.sub, subset = Glacier == "Taylor")
#Summary
summary(PilouAreaModelTay)
#get ANOVA table
anova(PilouAreaModelTay)

myCol <- c('red', 'blue')
plot(logPilou ~ logarea, data = my.data.sub, col=myCol[Glacier], xlab='log Area', ylab = 'log Pilou')
abline(PilouAreaModelCan, col='red')
abline(PilouAreaModelTay, col='blue')
```