---
title: "Basic Hypothesis Test with T and F Tests"
author: "Amy Solman"
date: "22/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook (https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/13-t_F_tests.ipynb)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Basic Hypothesis Testing
#We're going to use t tests to compare differences in means. We're going to use F tests to compare the variance of samples. We're also going to look at non-parametric tests.

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
# library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library("picante")
```

#Step Two: Load Data and make into phyloseq object 
```{r}
# load("results/cleaned_files/cleanedfiles_3.RData")
# #hole area - calculate and add to dataframe
# #Method as per Darcy et al (2018)
# metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
# metadata$area <- metadata$area*10000
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
###TOTAL TAXA
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
tot.div_shan <- diversity(t(otu_table(my.ps.rar)), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
tot.div_simp <- diversity(t(otu_table(my.ps.rar)), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
tot.div_inv_simp <- diversity(t(otu_table(my.ps.rar)), index = "invsimpson")

#Faith's phylogenetic diversity 
tot.asv_tab <- as.data.frame(my.ps.rar@otu_table)

tot.my_tree <- my.ps.rar@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps.rar@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
tot.df.pd <- pd(t(tot.asv_tab), tot.my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
tot.div_pilou <- diversity(t(otu_table(my.ps.rar)), index = "simpson")/log(tot.df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
tot.diversity <- cbind(tot.df.pd, tot.div_shan, tot.div_simp, tot.div_inv_simp, tot.div_pilou)
colnames(tot.diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

###ABUNDANT TAXA
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
abun.div_shan <- diversity(t(otu_table(my.ps.rar.abun)), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
abun.div_simp <- diversity(t(otu_table(my.ps.rar.abun)), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
abun.div_inv_simp <- diversity(t(otu_table(my.ps.rar.abun)), index = "invsimpson")

#Faith's phylogenetic diversity 
abun.asv_tab <- as.data.frame(my.ps.rar.abun@otu_table)

abun.my_tree <- my.ps.rar.abun@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps.rar.abun@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
abun.df.pd <- pd(t(abun.asv_tab), abun.my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
abun.div_pilou <- diversity(t(otu_table(my.ps.rar.abun)), index = "simpson")/log(abun.df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
abun.diversity <- cbind(abun.df.pd, abun.div_shan, abun.div_simp, abun.div_inv_simp, abun.div_pilou)
colnames(abun.diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

###RARE TAXA
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
rare.div_shan <- diversity(t(otu_table(my.ps.rar.rare)), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
rare.div_simp <- diversity(t(otu_table(my.ps.rar.rare)), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
rare.div_inv_simp <- diversity(t(otu_table(my.ps.rar.rare)), index = "invsimpson")

#Faith's phylogenetic diversity 
rare.asv_tab <- as.data.frame(my.ps.rar.rare@otu_table)

rare.my_tree <- my.ps.rar.rare@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps.rar.rare@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
rare.df.pd <- pd(t(rare.asv_tab), rare.my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
rare.div_pilou <- diversity(t(otu_table(my.ps.rar.rare)), index = "simpson")/log(rare.df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
rare.diversity <- cbind(rare.df.pd, rare.div_shan, rare.div_simp, rare.div_inv_simp, rare.div_pilou)
colnames(rare.diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}

###TOTAL TAXA
tot.my.data <- cbind(metadata, tot.diversity)

#remove columns with data we're not interested in
tot.my.data.sub <- subset(tot.my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(tot.my.data.sub) <- tot.my.data.sub$Run
tot.my.data.sub <- subset(tot.my.data.sub, select = -c(1))
tot.my.data.sub <- as.data.frame(lapply(tot.my.data.sub, function(x) as.numeric(as.character(x))))
tot.my.data.sub$Glacier <- metadata$Glacier
tot.my.data.sub <- tot.my.data.sub[complete.cases(tot.my.data.sub), ]

###ABUNDANT TAXA
abun.my.data <- cbind(metadata, abun.diversity)

#remove columns with data we're not interested in
abun.my.data.sub <- subset(abun.my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(abun.my.data.sub) <- abun.my.data.sub$Run
abun.my.data.sub <- subset(abun.my.data.sub, select = -c(1))
abun.my.data.sub <- as.data.frame(lapply(abun.my.data.sub, function(x) as.numeric(as.character(x))))
abun.my.data.sub$Glacier <- metadata$Glacier
abun.my.data.sub <- abun.my.data.sub[complete.cases(abun.my.data.sub), ]

###RARE TAXA
rare.my.data <- cbind(metadata, rare.diversity)

#remove columns with data we're not interested in
rare.my.data.sub <- subset(rare.my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(rare.my.data.sub) <- rare.my.data.sub$Run
rare.my.data.sub <- subset(rare.my.data.sub, select = -c(1))
rare.my.data.sub <- as.data.frame(lapply(rare.my.data.sub, function(x) as.numeric(as.character(x))))
rare.my.data.sub$Glacier <- metadata$Glacier
rare.my.data.sub <- rare.my.data.sub[complete.cases(rare.my.data.sub), ]

```

#Export our diversity dataframes
```{r}
write.csv(tot.my.data.sub,"../Outputs/Sommers2020/total-diversity-metadata.csv", row.names = FALSE)
write.csv(abun.my.data.sub,"../Outputs/Sommers2020/abundant-diversity-metadata.csv", row.names = FALSE)
write.csv(rare.my.data.sub,"../Outputs/Sommers2020/rare-diversity-metadata.csv", row.names = FALSE)
```
#T-tests
#Use to compare the means of two samples, or one sample to another value.
#T-test is best used for use with a two level factor (like my two glaciers) as multiple levels multiplies the significance level (0.05), increasing the chances of making a Type I error (a false positive)
# t = the difference between the values/the standard error
#If there is no difference then t = 0, so the larger the t value the bigger the difference between factors

#Step Three: One-sample t-test
#Is the mean of a sample difference from a specific value
```{r}
#get the mean species richness
mean.rich <- mean(my.data.sub$Richness)

#get the variance 
var.rich <- var(my.data.sub$Richness)

#get the number of datapoints
n.rich <- length(my.data.sub$Richness)

#now we can ask if the mean species richness is different from 400 (for example, I've just picked 400 out of nowhere!)
diff.rich <- mean.rich - 400

#get the standard error
se.rich <- sqrt(var.rich/n.rich)

#Now we can calculate the t value
t.rich <- diff.rich/se.rich
print(t.rich) #-2.38 this is a reasonably high t value so we can say that the mean species richness of our glaciers is different from 400

#or we can do all this in one go
t.test(my.data.sub$Richness, mu = 400) #t = -2.38. p = 0.02 so there is a significant different


```

#Step Four: Two-sample t-tests
#The two-sample t-test is more commonly used that the one sample t test to compare the means of two samples.
#This test makes the assumption that are data are normally distributed and have similar vairances
```{r}
###TOTAL TAXA
t.test(Richness ~ Glacier, data = tot.my.data.sub) # #p > 0.05
t.test(Faith ~ Glacier, data = tot.my.data.sub) #p > 0.05
t.test(Shannon ~ Glacier, data = tot.my.data.sub) #p > 0.05
t.test(Simpson ~ Glacier, data = tot.my.data.sub) #p > 0.05
t.test(Inv.Simpson ~ Glacier, data = tot.my.data.sub) #p > 0.05
t.test(Pilou ~ Glacier, data = tot.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE

###ABUNDANT TAXA
t.test(Richness ~ Glacier, data = abun.my.data.sub) #p > 0.05
t.test(Faith ~ Glacier, data = abun.my.data.sub) #p > 0.05
t.test(Shannon ~ Glacier, data = abun.my.data.sub) #p > 0.05
t.test(Simpson ~ Glacier, data = abun.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE
t.test(Inv.Simpson ~ Glacier, data = abun.my.data.sub) #p > 0.05
t.test(Pilou ~ Glacier, data = abun.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE

###RARE TAXA
t.test(Richness ~ Glacier, data = rare.my.data.sub) #p > 0.05
t.test(Faith ~ Glacier, data = rare.my.data.sub) #p > 0.05
t.test(Shannon ~ Glacier, data = rare.my.data.sub) #p > 0.05
t.test(Simpson ~ Glacier, data = rare.my.data.sub) #p > 0.05
t.test(Inv.Simpson ~ Glacier, data = rare.my.data.sub) #p > 0.05
t.test(Pilou ~ Glacier, data = rare.my.data.sub) #p > 0.05
```

#Step Five: F Test 
#The F test can be used to compare variances of two samples or populations. This can be used when checking to see if data meet the variance assumption (like in the two sample t-test)
#If F value is < 0.5 or > 2 they have may have unequal variances, check the p value (p < 0.05 means there is a significant difference in the variances)
```{r}
# var.test(Richness ~ Glacier, data = my.data.sub) #p > 0.05
# var.test(Faith ~ Glacier, data = my.data.sub) #p > 0.05
# var.test(Shannon ~ Glacier, data = my.data.sub) #p > 0.05
# var.test(Simpson ~ Glacier, data = my.data.sub) #p > 0.05
# var.test(Inv.Simpson ~ Glacier, data = my.data.sub) #p > 0.05
# var.test(Pilou ~ Glacier, data = my.data.sub) # p < 0.05

###TOTAL TAXA
var.test(Richness ~ Glacier, data = tot.my.data.sub) # #p > 0.05
var.test(Faith ~ Glacier, data = tot.my.data.sub) #p > 0.05
var.test(Shannon ~ Glacier, data = tot.my.data.sub) #p > 0.05
var.test(Simpson ~ Glacier, data = tot.my.data.sub) #p > 0.05
var.test(Inv.Simpson ~ Glacier, data = tot.my.data.sub) #p > 0.05
var.test(Pilou ~ Glacier, data = tot.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE

###ABUNDANT TAXA
var.test(Richness ~ Glacier, data = abun.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE
var.test(Faith ~ Glacier, data = abun.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE
var.test(Shannon ~ Glacier, data = abun.my.data.sub) #p > 0.05
var.test(Simpson ~ Glacier, data = abun.my.data.sub) #p > 0.05
var.test(Inv.Simpson ~ Glacier, data = abun.my.data.sub) #p > 0.05
var.test(Pilou ~ Glacier, data = abun.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE

###RARE TAXA
var.test(Richness ~ Glacier, data = rare.my.data.sub) #p > 0.05
var.test(Faith ~ Glacier, data = rare.my.data.sub) #p > 0.05
var.test(Shannon ~ Glacier, data = rare.my.data.sub) #p > 0.05
var.test(Simpson ~ Glacier, data = rare.my.data.sub) #p > 0.05
var.test(Inv.Simpson ~ Glacier, data = rare.my.data.sub) #p > 0.05
var.test(Pilou ~ Glacier, data = rare.my.data.sub) #p > 0.05
```

#Non-parametric tests
#If our data are not normally distributed and don't have equal variances (and log transformations don't make our data normal) then we can use a non-parametric alternative
```{r}
# wilcox.test(Richness ~ Glacier, data = my.data.sub)
# wilcox.test(Faith ~ Glacier, data = my.data.sub) #p > 0.05
# wilcox.test(Shannon ~ Glacier, data = my.data.sub) #p > 0.05
# wilcox.test(Simpson ~ Glacier, data = my.data.sub) #p > 0.05
# wilcox.test(Inv.Simpson ~ Glacier, data = my.data.sub) #p > 0.05
# wilcox.test(Pilou ~ Glacier, data = my.data.sub) # p < 0.05

###TOTAL TAXA
wilcox.test(Richness ~ Glacier, data = tot.my.data.sub) # #p > 0.05
wilcox.test(Faith ~ Glacier, data = tot.my.data.sub) #p > 0.05
wilcox.test(Shannon ~ Glacier, data = tot.my.data.sub) #p > 0.05
wilcox.test(Simpson ~ Glacier, data = tot.my.data.sub) #p > 0.05
wilcox.test(Inv.Simpson ~ Glacier, data = tot.my.data.sub) #p > 0.05
wilcox.test(Pilou ~ Glacier, data = tot.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE

###ABUNDANT TAXA
wilcox.test(Richness ~ Glacier, data = abun.my.data.sub) #p > 0.05
wilcox.test(Faith ~ Glacier, data = abun.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE
wilcox.test(Shannon ~ Glacier, data = abun.my.data.sub) #p > 0.05
wilcox.test(Simpson ~ Glacier, data = abun.my.data.sub) #p > 0.05
wilcox.test(Inv.Simpson ~ Glacier, data = abun.my.data.sub) #p > 0.05
wilcox.test(Pilou ~ Glacier, data = abun.my.data.sub) #p < 0.05 SIGNIFICANT DIFFERENCE

###RARE TAXA
wilcox.test(Richness ~ Glacier, data = rare.my.data.sub) #p > 0.05
wilcox.test(Faith ~ Glacier, data = rare.my.data.sub) #p > 0.05
wilcox.test(Shannon ~ Glacier, data = rare.my.data.sub) #p > 0.05
wilcox.test(Simpson ~ Glacier, data = rare.my.data.sub) #p > 0.05
wilcox.test(Inv.Simpson ~ Glacier, data = rare.my.data.sub) #p > 0.05
wilcox.test(Pilou ~ Glacier, data = rare.my.data.sub) #p > 0.05
```
