---
title: "Linear Regression Models"
author: "Amy Solman"
date: "22/04/2021"
output: html_document
---

#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook (https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/14-regress.ipynb)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#We're going to use linear regression models 
#Linear regression models are useful, especially if both your response (dependent) and predictor (independent) variables are continuous.
#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
# library(vegan)
# library(ggrepel)
# install.packages("repr")
library(repr) #for plotting our pairs of data
# library("picante")
```

#Step Two: Load Data and make into phyloseq object 
```{r}
# load("results/cleaned_files/cleanedfiles_3.RData")
# #hole area - calculate and add to dataframe
# #Method as per Darcy et al (2018)
# metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
# metadata$area <- metadata$area*10000
# 
# #Let's make our data into a phyloseq object
# ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
# TAX = tax_table(as.matrix(taxonomy))
# META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
# my.ps <- phyloseq(ASV, TAX, META, tree_clean)

#Load data files
load("../Outputs/Sommers2020/cleanedfiles-abundance.RData")
my.ps <- readRDS("../Outputs/Sommers2020/my-ps-original.rds") #original full dataset, no rareification, 
my.ps.rar <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-total.rds") #rarefied dataset, total taxa
my.ps.rar.rel <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-relative.rds") #rarefied relative abundance, total taxa
my.ps.rar.abun <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-abundant.rds") #rarefied abundant taxa ONLY
my.ps.rar.rare <- readRDS("../Outputs/Sommers2020/my-ps-rarefied-rare.rds") #rarefied rare taxa ONLY

#Load dataframes
tot.diversity.data <- read.csv("../Outputs/Sommers2020/total-diversity-metadata.csv")
abun.diversity.data <- read.csv("../Outputs/Sommers2020/abundant-diversity-metadata.csv")
rare.diversity.data <- read.csv("../Outputs/Sommers2020/rare-diversity-metadata.csv")

#Combine our dataframes
# tot.diversity.data$Abundance <- "Total"
# abun.diversity.data$Abundance <- "Abundant"
# rare.diversity.data$Abundance <- "Rare"
# combined.diversity.data <- rbind(tot.diversity.data, abun.diversity.data, rare.diversity.data)
```

#Step Three: Rarefy data and get relative abundances
```{r}
# #set a seed so we get the same results each time we run this
# set.seed(1961)
# #rarefy the data to equal sampling depths
# my.ps.rar <- rarefy_even_depth(my.ps)
# #get relative abundances
# my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
# #Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
# div_shan <- diversity(t(count_table_clean), index = "shannon")
# 
# #Use vegan package to get simpson diversity index of each sample
# div_simp <- diversity(t(count_table_clean), index = "simpson")
# 
# #Use vegan package to get inverted simpson diversity index of each sample
# div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")
# 
# #Faith's phylogenetic diversity 
# asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table
# 
# my_tree <- my.ps@phy_tree #get my phylogenetic tree
# 
# #Check if the tree is rooted
# my.ps@phy_tree #Rooted, includes branch lengths
# 
# #It is a rooted tree so...
# df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.
# 
# #Calculate Pilou evenness
# div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier
# 
# #bind alpha diversity indices into a dataframe
# diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
# colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
# my.data <- cbind(metadata, diversity)
# 
# #remove columns with data we're not interested in
# my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
# rownames(my.data.sub) <- my.data.sub$Run
# my.data.sub <- subset(my.data.sub, select = -c(1))
# my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
# my.data.sub$Glacier <- metadata$Glacier
# my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```
#Step Six: Plot richness with environmental data 
```{r}
# #let's start with basic plots of richness with our evironmental variables
# rich.chlo <- plot(my.data.sub$Chlorophyll, my.data.sub$Richness)
# rich.depth <- plot(my.data.sub$Depth, my.data.sub$Richness)
# rich.dna <- plot(my.data.sub$DNA_per_g, my.data.sub$Richness)
# rich.org <- plot(my.data.sub$org_matter, my.data.sub$Richness)
# rich.ph <- plot(my.data.sub$pH, my.data.sub$Richness)
# rich.water <- plot(my.data.sub$water_content, my.data.sub$Richness)
# rich.area <- plot(my.data.sub$area, my.data.sub$Richness)
# 
# #indices of other alpha metrics
# alpha.met <- c(8, 10, 11, 12, 13, 14)
# pairs(my.data.sub[, -c(alpha.met)]) #so we can do a pairs plot with just richness and environmental data

#Through visual inspection can we see correlations between ASV richness and any environmental variables?
###TOTAL TAXA
tot.rich.chlo <- plot(tot.diversity.data$Chlorophyll, tot.diversity.data$Richness)
tot.rich.depth <- plot(tot.diversity.data$Depth, tot.diversity.data$Richness)
tot.rich.dna <- plot(tot.diversity.data$DNA_per_g, tot.diversity.data$Richness)
tot.rich.org <- plot(tot.diversity.data$org_matter, tot.diversity.data$Richness)
tot.rich.ph <- plot(tot.diversity.data$pH, tot.diversity.data$Richness)
tot.rich.water <- plot(tot.diversity.data$water_content, tot.diversity.data$Richness)
tot.rich.area <- plot(tot.diversity.data$Area, tot.diversity.data$Richness)

###ABUNDANT TAXA
abun.rich.chlo <- plot(abun.diversity.data$Chlorophyll, abun.diversity.data$Richness)
abun.rich.depth <- plot(abun.diversity.data$Depth, abun.diversity.data$Richness)
abun.rich.dna <- plot(abun.diversity.data$DNA_per_g, abun.diversity.data$Richness)
abun.rich.org <- plot(abun.diversity.data$org_matter, abun.diversity.data$Richness)
abun.rich.ph <- plot(abun.diversity.data$pH, abun.diversity.data$Richness)
abun.rich.water <- plot(abun.diversity.data$water_content, abun.diversity.data$Richness)
abun.rich.area <- plot(abun.diversity.data$Area, abun.diversity.data$Richness)

###RARE TAXA
rare.rich.chlo <- plot(rare.diversity.data$Chlorophyll, rare.diversity.data$Richness)
rare.rich.depth <- plot(rare.diversity.data$Depth, rare.diversity.data$Richness)
rare.rich.dna <- plot(rare.diversity.data$DNA_per_g, rare.diversity.data$Richness)
rare.rich.org <- plot(rare.diversity.data$org_matter, rare.diversity.data$Richness)
rare.rich.ph <- plot(rare.diversity.data$pH, rare.diversity.data$Richness)
rare.rich.water <- plot(rare.diversity.data$water_content, rare.diversity.data$Richness)
rare.rich.area <- plot(rare.diversity.data$Area, rare.diversity.data$Richness)

```

#Step Seven: Look for correlations between richness and environmental data
```{r}
# #cor looks for correlations between pairs of variables using Pearson's correlation coefficient
# cor(my.data.sub[, -c(alpha.met)], use="pairwise")
# 
# #or we can use cor.test to compair a single pair of variables - this also carries out a t.test to see if the correlation is significant
# cor.test(my.data.sub$Richness, my.data.sub$Chlorophyll, use="pairwise") #p >0.05
# cor.test(my.data.sub$Richness, my.data.sub$Depth, use="pairwise") #p < 0.05
# cor.test(my.data.sub$Richness, my.data.sub$DNA_per_g, use="pairwise") #p > 0.05
# cor.test(my.data.sub$Richness, my.data.sub$org_matter, use="pairwise") #p < 0.05
# cor.test(my.data.sub$Richness, my.data.sub$pH, use="pairwise") #p > 0.05
# cor.test(my.data.sub$Richness, my.data.sub$water_content, use="pairwise") #p > 0.05
# cor.test(my.data.sub$Richness, my.data.sub$area, use="pairwise") #p < 0.05
# 
# #only cryoconite hole depth, area and organic matter content are significant correlated with ASV richness
# 
# #WARNING The above correlation coefficient calculation assumes a straight line relationship
```

#Step Eight: Re-run above correlation analysis with log transformed data
```{r}
###TOTAL TAXA
tot.diversity.data$logRichness <- log(tot.diversity.data$Richness)
tot.diversity.data$logChlorophyll <- log(tot.diversity.data$Chlorophyll)
tot.diversity.data$logDepth <- log(tot.diversity.data$Depth)
tot.diversity.data$logDNA_per_g <- log(tot.diversity.data$DNA_per_g)
tot.diversity.data$logorg_matter <- log(tot.diversity.data$org_matter)
tot.diversity.data$logpH <- log(tot.diversity.data$pH)
tot.diversity.data$logwater_content <- log(tot.diversity.data$water_content)
tot.diversity.data$logArea <- log(tot.diversity.data$Area)

cor.test(tot.diversity.data$logRichness, tot.diversity.data$logChlorophyll, use="pairwise") #p >0.05
cor.test(tot.diversity.data$logRichness, tot.diversity.data$logDepth, use="pairwise") #p < 0.05
cor.test(tot.diversity.data$logRichness, tot.diversity.data$logDNA_per_g, use="pairwise") #p < 0.05
cor.test(tot.diversity.data$logRichness, tot.diversity.data$logorg_matter, use="pairwise") #p < 0.05
cor.test(tot.diversity.data$logRichness, tot.diversity.data$logpH, use="pairwise") #p > 0.05
cor.test(tot.diversity.data$logRichness, tot.diversity.data$logwater_content, use="pairwise") #p > 0.05
cor.test(tot.diversity.data$logRichness, tot.diversity.data$logArea, use="pairwise") #p < 0.05

###ABUNDANT TAXA
abun.diversity.data$logRichness <- log(abun.diversity.data$Richness)
abun.diversity.data$logChlorophyll <- log(abun.diversity.data$Chlorophyll)
abun.diversity.data$logDepth <- log(abun.diversity.data$Depth)
abun.diversity.data$logDNA_per_g <- log(abun.diversity.data$DNA_per_g)
abun.diversity.data$logorg_matter <- log(abun.diversity.data$org_matter)
abun.diversity.data$logpH <- log(abun.diversity.data$pH)
abun.diversity.data$logwater_content <- log(abun.diversity.data$water_content)
abun.diversity.data$logArea <- log(abun.diversity.data$Area)

cor.test(abun.diversity.data$logRichness, abun.diversity.data$logChlorophyll, use="pairwise") #p >0.05
cor.test(abun.diversity.data$logRichness, abun.diversity.data$logDepth, use="pairwise") #p < 0.05
cor.test(abun.diversity.data$logRichness, abun.diversity.data$logDNA_per_g, use="pairwise") #p < 0.05
cor.test(abun.diversity.data$logRichness, abun.diversity.data$logorg_matter, use="pairwise") #p < 0.05
cor.test(abun.diversity.data$logRichness, abun.diversity.data$logpH, use="pairwise") #p > 0.05
cor.test(abun.diversity.data$logRichness, abun.diversity.data$logwater_content, use="pairwise") #p > 0.05
cor.test(abun.diversity.data$logRichness, abun.diversity.data$logArea, use="pairwise") #p < 0.05

###RARE TAXA
rare.diversity.data$logRichness <- log(rare.diversity.data$Richness)
rare.diversity.data$logChlorophyll <- log(rare.diversity.data$Chlorophyll)
rare.diversity.data$logDepth <- log(rare.diversity.data$Depth)
rare.diversity.data$logDNA_per_g <- log(rare.diversity.data$DNA_per_g)
rare.diversity.data$logorg_matter <- log(rare.diversity.data$org_matter)
rare.diversity.data$logpH <- log(rare.diversity.data$pH)
rare.diversity.data$logwater_content <- log(rare.diversity.data$water_content)
rare.diversity.data$logArea <- log(rare.diversity.data$Area)

cor.test(rare.diversity.data$logRichness, rare.diversity.data$logChlorophyll, use="pairwise") #p >0.05
cor.test(rare.diversity.data$logRichness, rare.diversity.data$logDepth, use="pairwise") #p < 0.05
cor.test(rare.diversity.data$logRichness, rare.diversity.data$logDNA_per_g, use="pairwise") #p < 0.05
cor.test(rare.diversity.data$logRichness, rare.diversity.data$logorg_matter, use="pairwise") #p < 0.05
cor.test(rare.diversity.data$logRichness, rare.diversity.data$logpH, use="pairwise") #p > 0.05
cor.test(rare.diversity.data$logRichness, rare.diversity.data$logwater_content, use="pairwise") #p > 0.05
cor.test(rare.diversity.data$logRichness, rare.diversity.data$logArea, use="pairwise") #p < 0.05

# my.log.alpha <- c(15, 16, 17, 18, 19, 20, 21, 22)
# pairs(my.data.sub[, (my.log.alpha)])
# 
# cor(my.data.sub[, (my.log.alpha)], use="pairwise")

```

#Step Nine: Plot the data 
```{r}

###TOTAL TAXA
myColours <- c('red', 'blue')
mySymbols <- c(1,3)

tot.diversity.data$Glacier <- as.factor(tot.diversity.data$Glacier)

plot(logRichness ~ logArea, data = tot.diversity.data,
      col = myColours[Glacier], pch=mySymbols[Glacier],
      xlab = "log(Area)", ylab = "log(ASV Richness)")

legend("topleft", legend=levels(tot.diversity.data$Glacier),
       col=myColours, pch=mySymbols)

#Visual inspection of the plots shows the glaciers have similar but slightly different relationships between hoel area and log(ASV Richness)

###ABUNDANT TAXA
myColours <- c('red', 'blue')
mySymbols <- c(1,3)

#my.data.sub$Glacier <- as.factor(my.data.sub$Glacier)
abun.diversity.data$Glacier <- as.factor(abun.diversity.data$Glacier)

plot(logRichness ~ logArea, data = abun.diversity.data,
      col = myColours[Glacier], pch=mySymbols[Glacier],
      xlab = "log(Area)", ylab = "log(ASV Richness)")

legend("topleft", legend=levels(abun.diversity.data$Glacier),
       col=myColours, pch=mySymbols)

#Visual inspection of the plots shows the glaciers have similar but slightly different relationships between hoel area and log(ASV Richness)

###RARE TAXA
myColours <- c('red', 'blue')
mySymbols <- c(1,3)

#my.data.sub$Glacier <- as.factor(my.data.sub$Glacier)
rare.diversity.data$Glacier <- as.factor(rare.diversity.data$Glacier)

plot(logRichness ~ logArea, data = rare.diversity.data,
      col = myColours[Glacier], pch=mySymbols[Glacier],
      xlab = "log(Area)", ylab = "log(ASV Richness)")

legend("topleft", legend=levels(rare.diversity.data$Glacier),
       col=myColours, pch=mySymbols)

#Visual inspection of the plots shows the glaciers have similar but slightly different relationships between hoel area and log(ASV Richness)


```
#Step Nine: Create linear models of richness and area
```{r}
# #This is a null model of the logRichness relationship on Canada Glacier
# nullModelRichnessCan <- lm(logRichness ~ 1, data = my.data.sub, subset = Glacier == "Canada")
# 
# #This is the linear regression model of the logarea relationship on Canada Glacier
# RichnessAreaModelCan <- lm(logRichness ~ logarea, data = my.data.sub, subset = Glacier == "Canada")
# 
# #let's look at the outputs of our model
# summary(RichnessAreaModelCan)
# 
# #Info we have in this table:
# #Call - the model description
# #Residuals - a summary of the residuals
# #Coefficients - a table of coefficients
# #Info on residual error
# #R-squared
# #F-statistic
# #P-value
# 
# #there are two rows in the coefficients table
# #these are the two coefficients of our equation y = a + b * x where a is the intercept and b is the slope. We get the estimated value of the intercept and the slope, the standard error pf those estimates, the t test value (and associated p value) of whether the slope and intercept are significantly different from zero.

###TOTAL TAXA
#This is the linear regression model of the logarea relationship on Canada Glacier
tot.RichnessAreaModelCan <- lm(logRichness ~ logArea, data = tot.diversity.data, subset = Glacier == "Canada")

#let's look at the outputs of our model
summary(tot.RichnessAreaModelCan)
anova(tot.RichnessAreaModelCan)
#Info we have in this table:
#Call - the model description
#Residuals - a summary of the residuals
#Coefficients - a table of coefficients
#Info on residual error
#R-squared
#F-statistic
#P-value

#there are two rows in the coefficients table
#these are the two coefficients of our equation y = a + b * x where a is the intercept and b is the slope. We get the estimated value of the intercept and the slope, the standard error pf those estimates, the t test value (and associated p value) of whether the slope and intercept are significantly different from zero.

###ABUNDANT TAXA
#This is the linear regression model of the logarea relationship on Canada Glacier
abun.RichnessAreaModelCan <- lm(logRichness ~ logArea, data = abun.diversity.data, subset = Glacier == "Canada")

#let's look at the outputs of our model
summary(abun.RichnessAreaModelCan)
anova(abun.RichnessAreaModelCan)


###RARE TAXA
#This is the linear regression model of the logarea relationship on Canada Glacier
rare.RichnessAreaModelCan <- lm(logRichness ~ logArea, data = rare.diversity.data, subset = Glacier == "Canada")

#let's look at the outputs of our model
summary(rare.RichnessAreaModelCan)


```
#Step Ten: Look at the terms of the model using the anova function
#Anova test how much variation in the response variable (richness) is explained by each explanatory variable (area)
```{r}
anova(tot.RichnessAreaModelCan)
anova(abun.RichnessAreaModelCan)
anova(rare.RichnessAreaModelCan)
#So here the ANOVA table is comparing the vairation in log Richness explained by log area to the total variation in log Richness.
#We want to see how much smaller the residuals are for the area model than the null model in explaining Richness.
#In the ANOVA table we have:
#DF: The degrees of freedom, each parameter/coefficient takes up a degree of freedom. This model adds area (1 parameter) to the null model, so has 1 DF
#Sum Sq: The sums of squares. The top line shows the variation in richness explained by area. The bottom line shows the residual sum of squares for the model.
#Mean Sq: Sum of squares values divided by the degrees of freedom. So if we explain lots of variation in the model with one coefficient (null model) that's great! If we explain a little variation with a loss of degrees of freedom by adding more parameters that's not so good...
#F value: Ratio of the mean sq for the variable and residual mean sq. Can show us if the explained variation is large or small. 
#Pr(>F): The p value shows us how likely the model explained this much variance by chance.

#We can get the sums of squares of the residuals from the null and area model using the function resid, square them and add them
# sum(resid(nullModelRichnessCan) ^ 2) #0.81
# sum(resid(RichnessAreaModelCan) ^ 2) #0.28 - is one is the same as seen in our ANOVA table

#TAKE AWAY: P < 0.05 so area explains a large amount of variation in ASV richness
```

#Step Eleven: Repeat the above for Taylor Glacier
```{r}
# #This is a null model of the logRichness relationship on Taylor Glacier
# nullModelRichnessTay <- lm(logRichness ~ 1, data = my.data.sub, subset = Glacier == "Taylor")
# 
# #This is the linear regression model of the logarea relationship on Taylor Glacier
# RichnessAreaModelTay <- lm(logRichness ~ logarea, data = my.data.sub, subset = Glacier == "Taylor")
# 
# #let's look at the outputs of our model
# summary(RichnessAreaModelTay)
# 
# #get ANOVA table
# anova(RichnessAreaModelTay)
# 
# #We can get the sums of squares of the residuals from the null and area model using the function resid, square them and add them
# sum(resid(nullModelRichnessTay) ^ 2) #0.67
# sum(resid(RichnessAreaModelTay) ^ 2) #0.10 - is one is the same as seen in our ANOVA table
# 
# #TAKE AWAY: P < 0.05 so area explains a large amount of variation in ASV richness

###TOTAL TAXA
#This is the linear regression model of the logarea relationship on Canada Glacier
tot.RichnessAreaModelTay <- lm(logRichness ~ logArea, data = tot.diversity.data, subset = Glacier == "Taylor")

#let's look at the outputs of our model
summary(tot.RichnessAreaModelTay)
anova(tot.RichnessAreaModelTay)

###ABUNDANT TAXA
#This is the linear regression model of the logarea relationship on Canada Glacier
abun.RichnessAreaModelTay <- lm(logRichness ~ logArea, data = abun.diversity.data, subset = Glacier == "Taylor")

#let's look at the outputs of our model
summary(abun.RichnessAreaModelTay)
anova(abun.RichnessAreaModelTay)

###RARE TAXA
#This is the linear regression model of the logarea relationship on Canada Glacier
rare.RichnessAreaModelTay <- lm(logRichness ~ logArea, data = rare.diversity.data, subset = Glacier == "Taylor")

#let's look at the outputs of our model
summary(rare.RichnessAreaModelTay)
anova(rare.RichnessAreaModelTay)
```

#Step Twelve: Plot the fitted regression model
```{r}
myColours <- c('red', 'blue')
mySymbols <- c(1,3)

###TOTAL TAXA
tot.diversity.data$Glacier <- as.factor(tot.diversity.data$Glacier)

pdf("../Outputs/Sommers2020/graphs/linear-model-logarea-logrichness-total.pdf", width=6, height=4) 
plot(logRichness ~ logArea, data = tot.diversity.data,
      col = myColours[Glacier], pch=mySymbols[Glacier],
      xlab = "log(Area)", ylab = "log(ASV Richness)")
abline(tot.RichnessAreaModelCan, col='red')
abline(tot.RichnessAreaModelTay, col='blue')

legend("topleft", legend=levels(tot.diversity.data$Glacier),
       col=myColours, pch=mySymbols)
#ggsave("../Outputs/Sommers2020/graphs/linear-model-logarea-logrichness-total.pdf", width=6, height=4)
dev.off()

###ABUNDANT TAXA
abun.diversity.data$Glacier <- as.factor(abun.diversity.data$Glacier)

pdf("../Outputs/Sommers2020/graphs/linear-model-logarea-logrichness-abundant.pdf", width=6, height=4) 
plot(logRichness ~ logArea, data = abun.diversity.data,
      col = myColours[Glacier], pch=mySymbols[Glacier],
      xlab = "log(Area)", ylab = "log(ASV Richness)")
abline(abun.RichnessAreaModelCan, col='red')
abline(abun.RichnessAreaModelTay, col='blue')

legend("topleft", legend=levels(abun.diversity.data$Glacier),
       col=myColours, pch=mySymbols)
#ggsave("../Outputs/Sommers2020/graphs/linear-model-logarea-logrichness-abundant.pdf", width=6, height=4)
dev.off()

###RARE TAXA
rare.diversity.data$Glacier <- as.factor(rare.diversity.data$Glacier)

pdf("../Outputs/Sommers2020/graphs/linear-model-logarea-logrichness-rare.pdf", width=6, height=4) 
plot(logRichness ~ logArea, data = rare.diversity.data,
      col = myColours[Glacier], pch=mySymbols[Glacier],
      xlab = "log(Area)", ylab = "log(ASV Richness)")
abline(rare.RichnessAreaModelCan, col='red')
abline(rare.RichnessAreaModelTay, col='blue')

legend("topleft", legend=levels(rare.diversity.data$Glacier),
       col=myColours, pch=mySymbols)

#ggsave("../Outputs/Sommers2020/graphs/linear-model-logarea-logrichness-rare.pdf", width=6, height=4)
dev.off()
```

#Step Thirteen: Model Diagnostics
#We need to check that our models are appropriate for our data
#We will use diagnostic plots to do this.
```{r}
# par(mfrow=c(2,2), mar=c(5,5,1.5,1.5))
# plot(RichnessAreaModelCan)
# 
# par(mfrow=c(2,2), mar=c(5,5,1.5,1.5))
# plot(RichnessAreaModelTay)

#Residuals vs Fitted: This plot looks to see if the distribution of the residuals has a similar variance for different predited values. We want NO obvious pattern or big gaps. 
#Normal Q-Q: This shows if the residuals are normally distributed. We want the points as close to a straight line as possible.
#Scale-Location: x axis is the fitted values, y axis is the square root of the standardized residuals. Basically we want the red line to be relatively flat.
#Residuals vs Leverage: This shows the standardized residuals against leverage - leverage means how much eaxch data point influences the linear model's coefficient estimates. Numbered points are those to pay special attention to. They are potential outliers

```

#Step Fourteen: Repeat the above using Faith's diversity metric
```{r}

###TOTAL TAXA 
#First get logs of our diversity metrics
tot.diversity.data$logFaith <- log(tot.diversity.data$Faith)
tot.diversity.data$logShannon <- log(tot.diversity.data$Shannon)
tot.diversity.data$logSimpson <- log(tot.diversity.data$Simpson)
tot.diversity.data$logInvSimp <- log(tot.diversity.data$Inv.Simpson)
tot.diversity.data$logPilou <- log(tot.diversity.data$Pilou)

#Linear Regression
tot.FaithAreaModelCan <- lm(logFaith ~ logArea, data = tot.diversity.data, subset = Glacier == "Canada")
#Summary
summary(tot.FaithAreaModelCan)
#get ANOVA table
anova(tot.FaithAreaModelCan)

#Linear Regression
tot.FaithAreaModelTay <- lm(logFaith ~ logArea, data = tot.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(tot.FaithAreaModelTay)
#get ANOVA table
anova(tot.FaithAreaModelTay)

myCol <- c('red', 'blue')
plot(logFaith ~ logArea, data = tot.diversity.data, col=myCol[Glacier], xlab='log(Area)', ylab = 'log(Faith)')
abline(tot.FaithAreaModelCan, col='red')
abline(tot.FaithAreaModelTay, col='blue')

###ABUNDANT TAXA 
#First get logs of our diversity metrics
abun.diversity.data$logFaith <- log(abun.diversity.data$Faith)
abun.diversity.data$logShannon <- log(abun.diversity.data$Shannon)
abun.diversity.data$logSimpson <- log(abun.diversity.data$Simpson)
abun.diversity.data$logInvSimp <- log(abun.diversity.data$Inv.Simpson)
abun.diversity.data$logPilou <- log(abun.diversity.data$Pilou)

#Linear Regression
abun.FaithAreaModelCan <- lm(logFaith ~ logArea, data = abun.diversity.data, subset = Glacier == "Canada")
#Summary
summary(abun.FaithAreaModelCan)
#get ANOVA table
anova(abun.FaithAreaModelCan)

#Linear Regression
abun.FaithAreaModelTay <- lm(logFaith ~ logArea, data = abun.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(abun.FaithAreaModelTay)
#get ANOVA table
anova(abun.FaithAreaModelTay)

myCol <- c('red', 'blue')
plot(logFaith ~ logArea, data = abun.diversity.data, col=myCol[Glacier], xlab='log(Area)', ylab = 'log(Faith)')
abline(abun.FaithAreaModelCan, col='red')
abline(abun.FaithAreaModelTay, col='blue')

###RARE TAXA 
#First get logs of our diversity metrics
rare.diversity.data$logFaith <- log(rare.diversity.data$Faith)
rare.diversity.data$logShannon <- log(rare.diversity.data$Shannon)
rare.diversity.data$logSimpson <- log(rare.diversity.data$Simpson)
rare.diversity.data$logInvSimp <- log(rare.diversity.data$Inv.Simpson)
rare.diversity.data$logPilou <- log(rare.diversity.data$Pilou)

#Linear Regression
rare.FaithAreaModelCan <- lm(logFaith ~ logArea, data = rare.diversity.data, subset = Glacier == "Canada")
#Summary
summary(rare.FaithAreaModelCan)
#get ANOVA table
anova(rare.FaithAreaModelCan)

#Linear Regression
rare.FaithAreaModelTay <- lm(logFaith ~ logArea, data = rare.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(rare.FaithAreaModelTay)
#get ANOVA table
anova(rare.FaithAreaModelTay)

myCol <- c('red', 'blue')
plot(logFaith ~ logArea, data = rare.diversity.data, col=myCol[Glacier], xlab='log(Area)', ylab = 'log(Faith)')
abline(rare.FaithAreaModelCan, col='red')
abline(rare.FaithAreaModelTay, col='blue')
```

#Step Fifteen: Repeat using Shannon
```{r}
###TOTAL TAXA
#Linear Regression
tot.ShanAreaModelCan <- lm(logShannon ~ logArea, data = tot.diversity.data, subset = Glacier == "Canada")
#Summary
summary(tot.ShanAreaModelCan)
#get ANOVA table
anova(tot.ShanAreaModelCan)

#Linear Regression
tot.ShanAreaModelTay <- lm(logShannon ~ logArea, data = tot.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(tot.ShanAreaModelTay)
#get ANOVA table
anova(tot.ShanAreaModelTay)

myCol <- c('red', 'blue')
plot(logShannon ~ logArea, data = tot.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Shannon')
abline(tot.ShanAreaModelCan, col='red')
abline(tot.ShanAreaModelTay, col='blue')

###ABUNDANT TAXA
#Linear Regression
abun.ShanAreaModelCan <- lm(logShannon ~ logArea, data = abun.diversity.data, subset = Glacier == "Canada")
#Summary
summary(abun.ShanAreaModelCan)
#get ANOVA table
anova(abun.ShanAreaModelCan)

#Linear Regression
abun.ShanAreaModelTay <- lm(logShannon ~ logArea, data = abun.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(abun.ShanAreaModelTay)
#get ANOVA table
anova(abun.ShanAreaModelTay)

myCol <- c('red', 'blue')
plot(logShannon ~ logArea, data = abun.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Shannon')
abline(abun.ShanAreaModelCan, col='red')
abline(abun.ShanAreaModelTay, col='blue')

###RARE TAXA
#Linear Regression
rare.ShanAreaModelCan <- lm(logShannon ~ logArea, data = rare.diversity.data, subset = Glacier == "Canada")
#Summary
summary(rare.ShanAreaModelCan)
#get ANOVA table
anova(rare.ShanAreaModelCan)

#Linear Regression
rare.ShanAreaModelTay <- lm(logShannon ~ logArea, data = rare.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(rare.ShanAreaModelTay)
#get ANOVA table
anova(rare.ShanAreaModelTay)

myCol <- c('red', 'blue')
plot(logShannon ~ logArea, data = rare.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Shannon')
abline(rare.ShanAreaModelCan, col='red')
abline(rare.ShanAreaModelTay, col='blue')


```

#Step Fifteen: Repeat using Simpson
```{r}
###TOTAL TAXA
#Linear Regression
tot.SimpAreaModelCan <- lm(logSimpson ~ logArea, data = tot.diversity.data, subset = Glacier == "Canada")
#Summary
summary(tot.SimpAreaModelCan)
#get ANOVA table
anova(tot.SimpAreaModelCan)

#Linear Regression
tot.SimpAreaModelTay <- lm(logSimpson ~ logArea, data = tot.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(tot.SimpAreaModelTay)
#get ANOVA table
anova(tot.SimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logSimpson ~ logArea, data = tot.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Simpson')
abline(tot.SimpAreaModelCan, col='red')
abline(tot.SimpAreaModelTay, col='blue')

###TOTAL ABUNDANT
#Linear Regression
abun.SimpAreaModelCan <- lm(logSimpson ~ logArea, data = abun.diversity.data, subset = Glacier == "Canada")
#Summary
summary(abun.SimpAreaModelCan)
#get ANOVA table
anova(abun.SimpAreaModelCan)

#Linear Regression
abun.SimpAreaModelTay <- lm(logSimpson ~ logArea, data = abun.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(abun.SimpAreaModelTay)
#get ANOVA table
anova(abun.SimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logSimpson ~ logArea, data = abun.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Simpson')
abline(abun.SimpAreaModelCan, col='red')
abline(abun.SimpAreaModelTay, col='blue')

###TOTAL RARE
#Linear Regression
rare.SimpAreaModelCan <- lm(logSimpson ~ logArea, data = rare.diversity.data, subset = Glacier == "Canada")
#Summary
summary(rare.SimpAreaModelCan)
#get ANOVA table
anova(rare.SimpAreaModelCan)

#Linear Regression
rare.SimpAreaModelTay <- lm(logSimpson ~ logArea, data = rare.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(rare.SimpAreaModelTay)
#get ANOVA table
anova(rare.SimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logSimpson ~ logArea, data = rare.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Simpson')
abline(rare.SimpAreaModelCan, col='red')
abline(rare.SimpAreaModelTay, col='blue')
```

#Step Fifteen: Repeat using Inv Simpson
```{r}
###TOTAL TAXA
#Linear Regression
tot.InvSimpAreaModelCan <- lm(logInvSimp ~ logArea, data = tot.diversity.data, subset = Glacier == "Canada")
#Summary
summary(tot.InvSimpAreaModelCan)
#get ANOVA table
anova(tot.InvSimpAreaModelCan)

#Linear Regression
tot.InvSimpAreaModelTay <- lm(logInvSimp ~ logArea, data = tot.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(tot.InvSimpAreaModelTay)
#get ANOVA table
anova(tot.InvSimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logInvSimp ~ logArea, data = tot.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Inv Simpson')
abline(tot.InvSimpAreaModelCan, col='red')
abline(tot.InvSimpAreaModelTay, col='blue')

###ABUNDANT TAXA
#Linear Regression
abun.InvSimpAreaModelCan <- lm(logInvSimp ~ logArea, data = abun.diversity.data, subset = Glacier == "Canada")
#Summary
summary(abun.InvSimpAreaModelCan)
#get ANOVA table
anova(abun.InvSimpAreaModelCan)

#Linear Regression
abun.InvSimpAreaModelTay <- lm(logInvSimp ~ logArea, data = abun.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(abun.InvSimpAreaModelTay)
#get ANOVA table
anova(abun.InvSimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logInvSimp ~ logArea, data = abun.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Inv Simpson')
abline(abun.InvSimpAreaModelCan, col='red')
abline(abun.InvSimpAreaModelTay, col='blue')

###RARE TAXA
#Linear Regression
rare.InvSimpAreaModelCan <- lm(logInvSimp ~ logArea, data = rare.diversity.data, subset = Glacier == "Canada")
#Summary
summary(rare.InvSimpAreaModelCan)
#get ANOVA table
anova(rare.InvSimpAreaModelCan)

#Linear Regression
rare.InvSimpAreaModelTay <- lm(logInvSimp ~ logArea, data = rare.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(rare.InvSimpAreaModelTay)
#get ANOVA table
anova(rare.InvSimpAreaModelTay)

myCol <- c('red', 'blue')
plot(logInvSimp ~ logArea, data = rare.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Inv Simpson')
abline(rare.InvSimpAreaModelCan, col='red')
abline(rare.InvSimpAreaModelTay, col='blue')
```

#Step Fifteen: Repeat using Pilou
```{r}
###TOTAL TAXA
#Linear Regression
tot.PilouAreaModelCan <- lm(logPilou ~ logArea, data = tot.diversity.data, subset = Glacier == "Canada")
#Summary
summary(tot.PilouAreaModelCan)
#get ANOVA table
anova(tot.PilouAreaModelCan)

#Linear Regression
tot.PilouAreaModelTay <- lm(logPilou ~ logArea, data = tot.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(tot.PilouAreaModelTay)
#get ANOVA table
anova(tot.PilouAreaModelTay)

myCol <- c('red', 'blue')
plot(logPilou ~ logArea, data = tot.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Pilou')
abline(tot.PilouAreaModelCan, col='red')
abline(tot.PilouAreaModelTay, col='blue')

###ABUNDANT TAXA
#Linear Regression
abun.PilouAreaModelCan <- lm(logPilou ~ logArea, data = abun.diversity.data, subset = Glacier == "Canada")
#Summary
summary(abun.PilouAreaModelCan)
#get ANOVA table
anova(abun.PilouAreaModelCan)

#Linear Regression
abun.PilouAreaModelTay <- lm(logPilou ~ logArea, data = abun.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(abun.PilouAreaModelTay)
#get ANOVA table
anova(abun.PilouAreaModelTay)

myCol <- c('red', 'blue')
plot(logPilou ~ logArea, data = abun.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Pilou')
abline(abun.PilouAreaModelCan, col='red')
abline(abun.PilouAreaModelTay, col='blue')

###RARE TAXA
#Linear Regression
rare.PilouAreaModelCan <- lm(logPilou ~ logArea, data = rare.diversity.data, subset = Glacier == "Canada")
#Summary
summary(rare.PilouAreaModelCan)
#get ANOVA table
anova(rare.PilouAreaModelCan)

#Linear Regression
rare.PilouAreaModelTay <- lm(logPilou ~ logArea, data = rare.diversity.data, subset = Glacier == "Taylor")
#Summary
summary(rare.PilouAreaModelTay)
#get ANOVA table
anova(rare.PilouAreaModelTay)

myCol <- c('red', 'blue')
plot(logPilou ~ logArea, data = rare.diversity.data, col=myCol[Glacier], xlab='log Area', ylab = 'log Pilou')
abline(rare.PilouAreaModelCan, col='red')
abline(rare.PilouAreaModelTay, col='blue')
```

#Let's export our dataframes because they've got some useful stuff in them
```{r}
write.csv(tot.diversity.data,"../Outputs/Sommers2020/total-diversity-metadata.csv", row.names = FALSE)
write.csv(abun.diversity.data,"../Outputs/Sommers2020/abundant-diversity-metadata.csv", row.names = FALSE)
write.csv(rare.diversity.data,"../Outputs/Sommers2020/rare-diversity-metadata.csv", row.names = FALSE)
```