---
title: "Multiple Explanatory Variables"
author: "Amy Solman"
date: "23/04/2021"
output: html_document
---
#This script will follow the tutorials set out in the Multilingual Quantitative Biologist notebook
(https://nbviewer.jupyter.org/github/mhasoba/TheMulQuaBio/blob/master/content/notebooks/16-MulExpl.ipynb)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

######I WILL BE USING FAKE DATA FOR THIS ANALYSIS FOR DEMONSTRATION PURPOSES#####

#Here we're going to look at MULTI EXPLANATORY VARIABLE LINEAR MODELS
#How to fit a linear model when you have multiple explanatory variables

#Step One: Clear workspace and load packages
```{r}
rm(list=ls())

# library(dplyr)
# library(tibble)
# library(ggpubr)
# library(phyloseq)
# library(ape)
library(vegan)
# library(ggrepel)
# install.packages("repr")
# library(repr) #for plotting our pairs of data
library("picante")
library(gplots)
library(lattice)
```

#Step Two: Load Data and make into phyloseq object 
```{r}
load("results/cleaned_files/cleanedfiles_3.RData")
#hole area - calculate and add to dataframe
#Method as per Darcy et al (2018)
metadata$area <- pi*(as.numeric(metadata$diameter1)/2)*(as.numeric(metadata$diameter2)/2)
metadata$area <- metadata$area*10000

#Let's make our data into a phyloseq object
ASV = otu_table(data.frame(count_table_clean), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(taxonomy))
META	=	sample_data(data.frame(metadata, row.names = metadata$Run))
my.ps <- phyloseq(ASV, TAX, META, tree_clean)
```

#Step Three: Rarefy data and get relative abundances
```{r}
#set a seed so we get the same results each time we run this
set.seed(1961)
#rarefy the data to equal sampling depths
my.ps.rar <- rarefy_even_depth(my.ps)
#get relative abundances
my.ps.rar.rel <- transform_sample_counts(my.ps.rar, function(x) x/sum(x))
```

#Step Four: Calculate alpha diversity indices
```{r}
#Use vegan package to get shannon diversity index of each sample (remember to transpose the dataframe because it expects samples as rows and species as columns)
div_shan <- diversity(t(count_table_clean), index = "shannon")

#Use vegan package to get simpson diversity index of each sample
div_simp <- diversity(t(count_table_clean), index = "simpson")

#Use vegan package to get inverted simpson diversity index of each sample
div_inv_simp <- diversity(t(count_table_clean), index = "invsimpson")

#Faith's phylogenetic diversity 
asv_tab <- as.data.frame(my.ps@otu_table) #Hellinger transformed asv table

my_tree <- my.ps@phy_tree #get my phylogenetic tree

#Check if the tree is rooted
my.ps@phy_tree #Rooted, includes branch lengths

#It is a rooted tree so...
df.pd <- pd(t(asv_tab), my_tree, include.root = TRUE) #find faiths diversity metric for each sample.

#Calculate Pilou evenness
div_pilou <- diversity(t(count_table_clean), index = "simpson")/log(df.pd$SR) #Here we use the species count we got earlier

#bind alpha diversity indices into a dataframe
diversity <- cbind(df.pd, div_shan, div_simp, div_inv_simp, div_pilou)
colnames(diversity) <- c("Faith", "Richness", "Shannon", "Simpson", "Inv Simpson", "Pilou")

```

#Step Five: Bind alpha diversity indices with metadata 
```{r}
my.data <- cbind(metadata, diversity)

#remove columns with data we're not interested in
my.data.sub <- subset(my.data, select = c(1, 10, 16, 19, 34, 36, 48, 50, 51, 52, 53, 54, 55, 56))
rownames(my.data.sub) <- my.data.sub$Run
my.data.sub <- subset(my.data.sub, select = -c(1))
my.data.sub <- as.data.frame(lapply(my.data.sub, function(x) as.numeric(as.character(x))))
my.data.sub$Glacier <- metadata$Glacier
my.data.sub <- my.data.sub[complete.cases(my.data.sub), ]

```

#Step Six: Add another fake Glacier of data so we can compare three glaciers
```{r}
Chlorophyll <- sample(as.numeric(my.data.sub$Chlorophyll), size = 12)
Depth <- sample(as.numeric(my.data.sub$Depth), size = 12)
DNA_per_g <- sample(as.numeric(my.data.sub$DNA_per_g), size = 12)
org_matter <- sample(as.numeric(my.data.sub$org_matter), size = 12)
pH <- sample(as.numeric(my.data.sub$pH), size = 12)
water_content <- sample(as.numeric(my.data.sub$water_content), size = 12)
area <- sample(as.numeric(my.data.sub$area), size = 12)
Faith <- sample(as.numeric(my.data.sub$Faith), size = 12)
Richness <- sample(as.numeric(my.data.sub$Richness), size = 12)
Shannon <- sample(as.numeric(my.data.sub$Shannon), size = 12)
Simpson <- sample(as.numeric(my.data.sub$Simpson), size = 12)
Inv.Simpson <- sample(as.numeric(my.data.sub$Inv.Simpson), size = 12)
Pilou <- sample(as.numeric(my.data.sub$Pilou), size = 12)
Glacier <- rep("Solman", 12)

df <- cbind(Chlorophyll, Depth, DNA_per_g, org_matter, pH, water_content, area, Faith, Richness, Shannon, Simpson, Inv.Simpson, Pilou, Glacier)

my.data.sub.X <- rbind(my.data.sub, df)
my.data.sub.X.1 <- my.data.sub.X[,-c(14)] #remove glacier names

my.data.sub <- as.data.frame(lapply(my.data.sub.X.1, function(x) as.numeric(as.character(x))))
#re-add glacier names 
my.data.sub$Glacier <- as.factor(my.data.sub.X$Glacier)
```

#Step Seven: Add an addition factor for us to play with
```{r}
Substrate <- rep(c("Ice", "Snow", "Cryoconite"), 12) 
Substrate <- c(Substrate, rep("Cryoconite", 4))
my.data.sub$Substrate <- as.factor(Substrate)
```

#Step Eight: Boxplots within groups
```{r}
par(mfrow=c(1,3))
plot(Richness ~ Substrate, data = my.data.sub, subset = Glacier == "Canada", main = "Canada")
plot(Richness ~ Substrate, data = my.data.sub, subset = Glacier == "Taylor", main = "Taylor")
plot(Richness ~ Substrate, data = my.data.sub, subset = Glacier == "Solman", main = "Solman")

#we can do something similar with lattice plots
bwplot(Richness ~ Substrate | Glacier, data = my.data.sub)
```

#Step Nine: Barplots with standard errors
```{r}
groups <- list(my.data.sub$Substrate, my.data.sub$Glacier)
groupMeans <- tapply(my.data.sub$Richness, groups, FUN = mean)
print(groupMeans)

seMean <- function(x){
	# get rid of missing values
	x <- na.omit(x)
	# calculate the standard error
	se <- sqrt(var(x)/length(x))
	# tell the function to report the standard error
	return(se)
}

groups <- list(my.data.sub$Substrate, my.data.sub$Glacier)

groupMeans <- tapply(my.data.sub$Richness, groups, FUN=mean)
print(groupMeans)

groupSE <- tapply(my.data.sub$Richness, groups, FUN=seMean)
print(groupSE)


# get upper and lower standard error height
upperSE <- groupMeans + groupSE
lowerSE <- groupMeans - groupSE
# create barplot
barMids <- barplot(groupMeans, ylim=c(0, max(upperSE)), beside=TRUE, ylab= ' log C value (pg) ' , col=c( ' white ' , ' grey70 '))
arrows(barMids, upperSE, barMids, lowerSE, ang=90, code=3, len=0.05)
```

#Step Ten: An alternative to barplots 
#let's use gplots to plot group means and standard errors
```{r}
# Get plots of group means and standard errors
par(mfrow=c(1,3))
plotmeans(Richness ~ Substrate, data=my.data.sub, subset = Glacier == "Canada", p=0.95, connect=FALSE, main="Canada", ylim=c(0, 700))
plotmeans(Richness ~ Substrate, data=my.data.sub, subset = Glacier == "Taylor", p=0.95, connect=FALSE, main = "Taylor", ylim=c(0, 700))
plotmeans(Richness ~ Substrate, data=my.data.sub, subset = Glacier == "Solman", p=0.95, connect=FALSE, main = "Solman", ylim=c(0, 700))

```

#Step Eleven: Fitting the linear model
```{r}
model <- lm(Richness ~ Glacier + Substrate, data = my.data.sub)

#check diagnostic plot
par(mfrow=c(2,2))
plot(model)

#ANOVA of the model
anova(model) #here we see the sum of squares of the two variables are much smaller than the residuals - so there is a lot of unexplained variation. Substrate explained less vairation than glacier - this makes sense with our plots where there were big differences between the glacier plots but less variation within the glacier plots (where we compared the substrates)

#summary of the model
summary(model) #adjusted r = 0.03
#the intercept is the mean for Canada glacier cryoconite samples, we then have differences from this estimate for the different glaciers and different substrates
#glacier taylor makes a small difference to richness
```

#Step Twelve: Predicted values
#let's get our predicted values from the model
```{r}
# data frame of combinations of variables
gd <- rep(levels(my.data.sub$Substrate), times = 3)
print(gd)
tl <- rep(levels(my.data.sub$Glacier), each = 3)
print(tl)
predVals <- data.frame(Substrate = gd, Glacier = tl)

predVals$predict <- predict(model, newdata = predVals)
print(predVals)

barMids <- barplot(groupMeans, ylim=c(0, 500), ylab='Richness', beside=TRUE, col=c('white', 'grey70'))

arrows(barMids, upperSE, barMids, lowerSE, ang=90, code=3, len=0.1)
points(barMids, predVals$predict, col='red', pch=12)

#The red markers do not match the calculated means because the model only includes a single difference between substrates, which has to be the same for each glacier. So there is no interaction between glacier and substrate in this model.
```

#Step :
```{r}

```

#Step :
```{r}

```

#Step :
```{r}

```